

# install bartools
```{r}
# if (!requireNamespace("devtools", quietly = TRUE)) {
#   install.packages("devtools")
# }
# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
```

# load BARBIE functions
```{r}
# Specify the directory containing the R files
folder_path <- "/Users/feiliyang/Desktop/Barbie/R"

# List all files in the folder with a .R extension
r_files <- list.files(folder_path, pattern = "\\.R$", full.names = TRUE)

# Load each R file into the R session
for (file in r_files) {
  source(file)
}

```

# library packages

```{r}
# library(bartools)

# data manipulation
library(magrittr)
library(tidyverse)
library(dplyr)
library(kableExtra)

# visualization
library(ComplexHeatmap)
library(ggplot2)
library(patchwork)
library(gridtext)

# stats
library(logistf)
```

# read in preprocessed dataset and subset into individual treatments
```{r}
dge <- readRDS("dose2.rds")

# simplify barcode labels
rownames(dge) <- gsub("BFP_Barcode", "BC", rownames(dge))
```

## generate ibet and AraC specific datasets
```{r}
ibet <- dge[,which(dge$samples$Treatment == "IBET" | dge$samples$Treatment == "DMSO"| dge$samples$Treatment == "TP0")]
arac <- dge[,which(dge$samples$Treatment == "ARAC" | dge$samples$Treatment == "DMSO"| dge$samples$Treatment == "TP0")]
```

# split dataset into replicates
```{r}
dge.TR1 <- dge[,which(dge$samples$Technical_Replicate == "TR1")]
dge.TR2 <- dge[,which(dge$samples$Technical_Replicate == "TR2")]
```

# density
```{r}
plot(density(log2(dge$counts + 1)))
mean(dge$counts)
max(dge$counts)
```

# create BARBIE
```{r}
AML_bb <- createBarbie(
  object = dge$counts, 
  target = dge$samples,
  color_panel = NULL)

AML_bb <- getTopBar(Barbie = AML_bb, cut_off = 0.95)

sum(AML_bb$is_top)

AML_bb <- trimRow(Barbie = AML_bb, keep_rows = AML_bb$is_top)

dim(AML_bb$CPM)

View(AML_bb$metadata)

AML_bb$metadata %>% colnames()
```

useful variables: 

1) Treatment

2) Strategy

3) Technical_Replicate

4) Timepoint

# test between Tec Reps
```{r tech rep}
#trim out TP0
AML_TPx <- trimObjectByMetadata(
  Barbie = AML_bb, 
  condition = "Timepoint", specified = "TP0", 
  keep = FALSE) # exclude "immature" celltype from samples

# customize sample groups as you need.
Vector_customized <- AML_TPx$metadata$Technical_Replicate

# Generate tables and create the list
c_tables <- GetContingencyTable(AML_TPx, Vector_customized = Vector_customized)
#print the first 5 contingency table
lapply(c_tables[1:5], function(x) {knitr::kable(x)})

# Apply test, and get Bias group
AML_TPx <- GetOccBiasGroup(Barbie = AML_TPx, contingency_table_ls = c_tables)

kableExtra::kable(AML_TPx$Bias_Occ)
table(AML_TPx$Bias_Occ$group)

flag <- AML_TPx$Bias_Occ$group != "Unbiased"
failed_occ <-trimRow(Barbie = AML_TPx, keep_rows = flag)

PlotBiasVsRank(Barbie = AML_TPx, passing_data = "contribution")

PlotBiasVsRank(Barbie = AML_TPx, passing_data = "output")

PlotBiasVsRank(Barbie = AML_TPx, passing_data = "rank")

plot(density(AML_TPx$Bias_Occ$pvalue))
```

# Time Series

Could we fit a model about which barcodes are changing across time points (for IBET)? What I'm thinking is calling TP3 +TP5 late time point and TP1 early time point and then test for early vs late with a factor for the replicate. So the Fishers can't do the replicate and it would be good to do the logistic regression tests with and without the factor for replicate.

## Data Prep

```{r data prep}
# select IBET in Treatment, exclude T0
IBET_bb <- trimObjectByMetadata(
  Barbie = AML_bb,
  condition = "Treatment",
  specified = "IBET",
  keep = TRUE)

ARAC_bb <- trimObjectByMetadata(
  Barbie = AML_bb,
  condition = "Treatment",
  specified = "ARAC",
  keep = TRUE)

DMSO_bb <- trimObjectByMetadata(
  Barbie = AML_bb,
  condition = "Treatment",
  specified = "DMSO",
  keep = TRUE)

# split gradual / high_dose in Strategy 
IBET_gra <- trimObjectByMetadata(
  Barbie = IBET_bb,
  condition = "Strategy",
  specified = "gradual",
  keep = TRUE
)

IBET_high <- trimObjectByMetadata(
  Barbie = IBET_bb,
  condition = "Strategy",
  specified = "high_dose",
  keep = TRUE
)

IBET_bb$metadata %>% 
  select(Treatment, Strategy, Technical_Replicate, Timepoint) %>% 
  View()

IBET_gra$metadata %>% 
  select(Treatment, Strategy, Technical_Replicate, Timepoint) %>% 
  View()

IBET_high$metadata %>% 
  select(Treatment, Strategy, Technical_Replicate, Timepoint) %>% 
  View()
```

## recode target meta

```{r recode}
# transform variables into factors, recode Timepoint
# TP3-TP5: late; TP1: early
##go with gradient
meta <- IBET_bb$metadata %>% 
  select(Treatment, Strategy, Technical_Replicate, Timepoint) %>% 
  as.data.frame()

meta$Technical_Replicate <- as.factor(meta$Technical_Replicate)

meta$Time <- ifelse(
  meta$Timepoint == "TP1",
  "early",
  ifelse(
    meta$Timepoint %in% c("TP2"),
    "mid",
    ifelse(meta$Timepoint %in% c("TP3", "TP4", "TP5"),
           "late",
           NA)
    )
  )

meta$Time <- factor(meta$Time, levels = c("early", "mid", "late"))

IBET_bb$metadata$Time <- meta$Time

# recode TechRep nested under Strategy
meta <- meta %>%
  mutate(TR = paste(Strategy, Technical_Replicate, sep = "."))

IBET_bb$metadata$TR <- meta$TR
```


## fit model

```{r fit model}
meta %>% View()



# fit variables: Timepoint + Technical_Replicate
design <- meta %>%
  with(model.matrix(~ Time + TR))

# fit variables: Timepoint for comparison
design0 <- meta %>%
  with(model.matrix(~ Time))

  # check if mydesign matrix is full rank or not.
  is_full_rank <- qr(design)$rank == min(dim(design))
  print(paste("Design matrix full rank?", is_full_rank))

# make design matrix full rank
  # design2 <- WeightIt::make_full_rank(design, with.intercept = FALSE) # delete column with nested factors

IBET_bb <- GetLogitOcc(
  Barbie = IBET_bb, 
  penalized = F, 
  group_names = c("early", "late"),
  mydesign = design) 

table(IBET_bb$LR_Occ_classic$group)

IBET_bb <- GetLogitOcc(
  Barbie = IBET_bb, 
  penalized = T, 
  group_names = c("early", "late"),
  mydesign = design) 

table(IBET_bb$LR_Occ_penalized$group)

# fit design0
IBET_bb0 <- GetLogitOcc(
  Barbie = IBET_bb, 
  penalized = T, 
  group_names = c("early", "late"),
  mydesign = design0) 

table(IBET_bb0$LR_Occ_penalized$group)

# test one row
row <- IBET_bb$presence[1,]

ttt <- fit_logistic_model(Barbie = IBET_bb, row = row)
tt2 <- firth_model(Barbie = IBET_bb, row = row)

beta <- esti$coefficients["Timelate", "Estimate"]
p.value <- esti$coefficients["Timelate", "Pr(>|z|)"]


# apply Fisher's exact test
IBET_bb_2 <- trimObjectByMetadata(Barbie = IBET_bb, condition = "Time", specified = c("early", "late"), keep = T)

vec <- IBET_bb_2$metadata$Time %>% as.character()

ctable <- GetContingencyTable(Barbie = IBET_bb_2, Vector_customized = vec)

IBET_bb_2 <- GetOccBiasGroup(Barbie = IBET_bb_2, contingency_table_ls = ctable)

table(IBET_bb_2$Bias_Occ$group)

# linear model prop
IBET_bb <- getPropBiasGroup_t(
  Barbie = IBET_bb, 
  mydesign = design, 
  mytargets = meta)

table(IBET_bb$Bias_Prop$group)

# fit linear model prop design0
IBET_bb0 <- getPropBiasGroup_t(
  Barbie = IBET_bb, 
  mydesign = design0, 
  mytargets = meta)

table(IBET_bb0$Bias_Prop$group)
```

### details - model output

```{r output details, fig.width=9, fig.height=5}
flag <- IBET_bb$LR_Occ_penalized$group != "Unbiased"

diff_bb <- trimRow(Barbie = IBET_bb, keep_rows = flag)

set.seed(42)

right_ha <- diff_bb %>% with(
  rowAnnotation(
    Prop = Bias_Prop$group,
    Fishers = IBET_bb_2$Bias_Occ$group[flag],
    Logit_classic = LR_Occ_classic$group,
    Logit_penalized = LR_Occ_penalized$group,
    col = list(
      Prop = color_panel$bias_group,
      Fishers = color_panel$bias_group,
      Logit_classic = color_panel$bias_group,
      Logit_penalized = color_panel$bias_group
      ),
    show_legend = TRUE,
    show_annotation_name = TRUE,
    annotation_name_side = "top",
    annotation_name_gp = gpar(fontsize = 10)
    )
)
  
top_ha <- diff_bb %>% with(
  HeatmapAnnotation(
    Time = metadata$Time,    
    Timepoint = metadata$Timepoint,
    TechRep = metadata$Technical_Replicate,

    col = list(
      Time = c("early" = "#33AAFF", "late" = "#FF5959", "mid" = "#999999")
      ),
    show_annotation_name = T,
    annotation_name_side = "left",
    annotation_name_gp = gpar(fontsize = 10)
    )
)

PlotCpmHP(
  Barbie = diff_bb, 
  show_bias = F, 
  row_ha = right_ha,
  top_ha = top_ha, 
  # bottom_ha = bottom_ha, 
  split_by = diff_bb$metadata$Time, 
  col_title = "IBET_DMSO_gradual"
  ) +

PlotPreHP(
  Barbie = diff_bb, 
  show_bias = TRUE, 
  row_ha = right_ha, 
  top_ha = top_ha, 
  # bottom_ha = bottom_ha, 
  split_by = diff_bb$metadata$Time, 
  col_title = "%s"
  ) 
```


## ggplot

```{r dotplot}
# visualize

gg1 <- PlotBiasVsRank(
  Barbie = IBET_bb, passing_data = "contribution", # "output" "rank" "contribution"
  bias_group = IBET_bb$LR_Occ_classic$group,
  bias_pvalue = IBET_bb$LR_Occ_classic$p.values) + ggtitle("Logit_classic")

gg2 <- PlotBiasVsRank(
  Barbie = IBET_bb, passing_data = "contribution", 
  bias_group = IBET_bb$LR_Occ_penalized$group,
  bias_pvalue = IBET_bb$LR_Occ_penalized$p.values) + ggtitle("Logit_penalized")

gg3 <- PlotBiasVsRank(
  Barbie = IBET_bb, passing_data = "contribution", 
  bias_group = IBET_bb_2$Bias_Occ$group,
  bias_pvalue = IBET_bb_2$Bias_Occ$pvalue) + ggtitle("Fisher's")

gg4 <- PlotBiasVsRank(
  Barbie = IBET_bb, passing_data = "contribution", 
  bias_group = IBET_bb$Bias_Prop$group,
  bias_pvalue = IBET_bb$Bias_Prop$pvalue) + ggtitle("Linear_Reg")

ggsave(here::here("dose-escalation", "pvalues.png"),
       (gg1 | gg2) / (gg3 | gg4),
       width = unit(8,"cm"),
       height = unit(8,"cm")
       )

```


## FET vs GLM_pen - p.values

```{r p.values FET GLM}
FET <- IBET_bb_2$Bias_Occ %>% 
  select(pvalue, group) %>% 
  rename(FET_p.value = pvalue,
         FET_group = group)

GLM <- IBET_bb$LR_Occ_penalized %>% 
  rename(GLM_p.value = p.values, 
         GLM_group = group)

GLM0 <- IBET_bb0$LR_Occ_penalized %>% 
  rename(GLM0_p.value = p.values, 
         GLM0_group = group)

LM <- IBET_bb$Bias_Prop%>% 
  select(pvalue, group) %>% 
  rename(LM_p.value = pvalue, 
         LM_group = group)

LM0 <- IBET_bb0$Bias_Prop%>% 
  select(pvalue, group) %>% 
  rename(LM0_p.value = pvalue, 
         LM0_group = group)

df <- cbind(GLM, FET, GLM0, LM, LM0)

df$GLM_group[df$GLM_group == "Unbiased"] <- "n.s."
df$FET_group[df$FET_group == "Unbiased"] <- "n.s."
df$GLM0_group[df$GLM0_group == "Unbiased"] <- "n.s."
df$LM_group[df$LM_group == "Unbiased"] <- "n.s."
df$LM0_group[df$LM0_group == "Unbiased"] <- "n.s."


# Create a new column indicating if all columns have the same value
df <- df %>%
  rowwise() %>%
  mutate(
    GLM_p.values_transformed = -log10(GLM_p.value),
    GLM0_p.values_transformed = -log10(GLM0_p.value),
    FET_p.values_transformed = -log10(FET_p.value),
    LM_p.values_transformed = -log10(LM_p.value),
    LM0_p.values_transformed = -log10(LM0_p.value),
    overlap = ifelse(length(unique(c_across(c(LM_group, LM0_group)))) == 1, 
                     LM_group, 
                     paste("~ Time+TR:", LM_group, ",~ Time:", LM0_group)
                     ),
    match = ifelse(length(unique(c_across(c(LM_group, LM0_group)))) == 1, 
                     LM_group, 
                     "notmatch"
                     )
    )

head(df)

ggplot(data = df) +
  geom_point(
    aes(
      x = LM0_p.values_transformed, 
      y = LM_p.values_transformed, 
      col = overlap
      )
    ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "navy") +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed", color = "navy") +
  labs(
    y = "-log10(Linear_p.values ~ Time + TR)",
    x = "-log10(Linear_p.values ~ Time)"
    ) +
  theme_classic() +
    coord_fixed() 



```

### not match

```{r not match, fig.width=9, fig.height=5}
# for those not match
pointer <- df$match == "notmatch"
sum(pointer)

# pointer[1] <- T
nomatch_bb <- trimRow(Barbie = IBET_bb, keep_rows = pointer)

set.seed(42)

right_ha <- nomatch_bb %>% with(
  rowAnnotation(
    LM_Time_TR = Bias_Prop$group,
    LM_Time = IBET_bb0$Bias_Prop$group[pointer],
    # FET = IBET_bb_2$Bias_Occ$group[pointer],
    # Logit_classic = LR_Occ_classic$group,
    # Logit_pen_Time_TR = LR_Occ_penalized$group,
    # Logit_pen_Time = IBET_bb0$LR_Occ_penalized$group[pointer],
    col = list(
      LM_Time_TR = color_panel$bias_group,
      LM_Time = color_panel$bias_group,
      Logit_pen_Time_TR = color_panel$bias_group,
      Logit_pen_Time = color_panel$bias_group
      ),
    show_legend = F,
    show_annotation_name = TRUE,
    annotation_name_side = "top",
    annotation_name_gp = gpar(fontsize = 10)
    )
)
  
top_ha <- nomatch_bb %>% with(
  HeatmapAnnotation(
    Time = metadata$Time,    
    Timepoint = metadata$Timepoint,
    # Strategy = metadata$Strategy,
    TR = metadata$TR,

    col = list(
      Time = c("early" = "#33AAFF", "late" = "#FF5959", "mid" = "#999999")
      ),
    show_annotation_name = T,
    annotation_name_side = "left",
    annotation_name_gp = gpar(fontsize = 10)
    )
)

ht <- PlotCpmHP(
  Barbie = nomatch_bb, 
  show_bias = T, 
  row_ha = right_ha,
  top_ha = top_ha, 
  # bottom_ha = bottom_ha, 
  split_by = nomatch_bb$metadata$Time, 
  col_title = gt_render("IBET_DMSO<br> Barcode ~ Time + TR") 
  ) +

PlotPreHP(
  Barbie = nomatch_bb, 
  show_bias = TRUE, 
  row_ha = right_ha, 
  top_ha = top_ha, 
  # bottom_ha = bottom_ha, 
  split_by = nomatch_bb$metadata$Time, 
  col_title = "%s") 

# Custom legend
custom_legend <- Legend(
  labels = c("sig.to_early", "sig.to_late", "n.s."),
  legend_gp = gpar(fill = c("#33AAFF", "#FF5959", "#FFC000")),
  title = "Test: FET/Logitpen"
)

draw(
  ht, 
  annotation_legend_list = list(custom_legend)
  )

# decorate_heatmap_body("htname", {
#   grid.text("xxx")
#   })
```


## heatmap

```{r hp, fig.height=5, fig.width=9}
set.seed(42)

right_ha <- IBET_bb %>% with(
  rowAnnotation(
    Fishers = IBET_bb_2$Bias_Occ$group,
    Logit_pen_Time_TR = LR_Occ_penalized$group,
    Logit_pen_Time = IBET_bb0$LR_Occ_penalized$group,
    col = list(
      Fishers = color_panel$bias_group,
      Logit_pen_Time_TR = color_panel$bias_group,
      Logit_pen_Time = color_panel$bias_group
      ),
    show_legend = F,
    show_annotation_name = TRUE,
    annotation_name_side = "top",
    annotation_name_gp = gpar(fontsize = 10)
    )
)
  
top_ha <- IBET_bb %>% with(
  HeatmapAnnotation(
    Time = metadata$Time,    
    Timepoint = metadata$Timepoint,
    # Strategy = metadata$Strategy,
    TR = metadata$TR,

    col = list(
      Time = c("early" = "#33AAFF", "late" = "#FF5959", "mid" = "#999999")
      ),
    show_annotation_name = T,
    annotation_name_side = "left",
    annotation_name_gp = gpar(fontsize = 10)
    )
)

ht <- PlotCpmHP(
  Barbie = IBET_bb, 
  show_bias = F, 
  row_ha = right_ha,
  top_ha = top_ha, 
  # bottom_ha = bottom_ha, 
  split_by = IBET_bb$metadata$Time, 
  col_title = gt_render("IBET_DMSO<br>")) +

PlotPreHP(
  Barbie = IBET_bb, 
  show_bias = TRUE, 
  row_ha = right_ha, 
  top_ha = top_ha, 
  # bottom_ha = bottom_ha, 
  split_by = IBET_bb$metadata$Time, 
  col_title = "%s") 

# Custom legend
custom_legend <- Legend(
  labels = c("sig.to_early", "sig.to_late", "n.s."),
  legend_gp = gpar(fill = c("#33AAFF", "#FF5959", "#FFC000")),
  title = "Test: prop/occ/logit"
)

draw(ht, annotation_legend_list = list(custom_legend))
```

## functions - model

```{r function}
fit_logistic_model <- function(Barbie, row, mydesign = NULL) {
  dat <- cbind(meta, row)

  logiM <- glm(
  row ~ 0 + mydesign,
  data = dat,
  family = "binomial")
  
  return(invisible(summary(logiM)))
  
}

firth_model <- function(Barbie, row, mydesign = NULL) {
  dat <- cbind(meta, row)

  firthM <- logistf(
    row ~ 0 + mydesign, 
    data = dat)

  return(invisible(firthM))

}

GetLogitOcc <- function(Barbie, penalized = F, group_names, mydesign = NULL) {
  # get occurence data
  outcome <- Barbie$presence +1 -1
  # get group names for comparison
  if(is.null(group_names)){
    group_names <- unique(Barbie$metadata$group)
  }

  if(!penalized){ # fit classic logistic model

    results <- lapply(
      1:nrow(outcome),
      function(i)
        fit_logistic_model(Barbie, outcome[i, ], mydesign)
      )

    # View the summary of the first model as an example
    # results[[1]]$coefficients[2,4]

    p.values.logi <- lapply(results, function(x)
      x$coefficients["mydesignTimelate", "Pr(>|z|)"]) %>% unlist()

    beta.logi <- lapply(results, function(x)
      x$coefficients["mydesignTimelate", "Estimate"]) %>% unlist()

    signif.logi <- ifelse(p.values.logi >= 0.05,
                          "Unbiased",
                          ifelse(beta.logi > 0,
                                 group_names[2], group_names[1]))

    # get stat output
    LR_Occ <- data.frame(
      p.values = p.values.logi,
      group = signif.logi
    )

    Barbie$LR_Occ_classic <- LR_Occ

  } else { # fit penalized logistic model

    # Apply the function to each row of outcome and store the results
    results_firth <- lapply(
      1:nrow(outcome),
      function(i)
        invisible(firth_model(Barbie, outcome[i, ], mydesign))
      )

    p.values.firth <- lapply(results_firth, function(x) x$prob["mydesignTimelate"]) %>%
      unlist()

    beta.firth <- lapply(
      results_firth, function(x)
        x$coefficients["mydesignTimelate"]) %>% unlist()

    signif.firth <- ifelse(p.values.firth >= 0.05,
                           "Unbiased",
                           ifelse(beta.firth > 0,
                                  group_names[2], group_names[1]))

    # get stat output
    LR_Occ <- data.frame(
      p.values = p.values.firth,
      group = signif.firth
    )

    Barbie$LR_Occ_penalized <- LR_Occ

  }

  #return
  if(is.null(Barbie$color_panel$bias_group)) {
    Barbie$color_panel$bias_group <- c("C1" = "#33AAFF", "C2" = "#FF5959", "Unbiased" = "#FFC000")
    names(Barbie$color_panel$bias_group)[1:2] <- group_names
  }

  return(Barbie)
}


```

## modify prop function

```{r prop}
getPropBiasGroup_t <- function(Barbie, mydesign = NULL, group_vector = NULL, myblock = NULL, mytargets = NULL){

  # # targets
  # mytargets <- data.frame(Barbie$metadata, category = group_vector)
  #
  # # model design matrix
  # mydesign <- mytargets %>%
  #   with(model.matrix(~0 + category + treat + mouse + tissue))

  # make design matrix full rank
  mydesign2 <- WeightIt::make_full_rank(mydesign, with.intercept = FALSE) # delete column with nested factors

  # check if mydesign matrix is full rank or not.
  is_full_rank2 <- qr(mydesign2)$rank == min(dim(mydesign2))
  print(paste("Design matrix full rank?", is_full_rank2))

  # set up groups for correction of sample duplication
  # myblock <- mytargets %>%
  #   with(paste(treat, mouse, tissue, category))

  # arcsin squreroot transformation for data
  mydata <- Barbie$prop %>% sqrt() %>% asin()

  # perform duplication correction
  dup <- limma::duplicateCorrelation(object = mydata,
                                     design = mydesign2,
                                     block = myblock)
  print(paste("Consensus correlation: ", dup$consensus.correlation))


  # set up contrast
  mycontrasts <- mytargets %>%
    with(limma::makeContrasts(Bias = Timelate,
                              levels = colnames(mydesign2)))

  # fit limma model
  myfit <- limma::lmFit(object = mydata,
                        design = mydesign2,
                        block = myblock,
                        correlation = dup$consensus.correlation)

  myfit2 <- limma::contrasts.fit(fit = myfit, contrasts = mycontrasts)
  myfit3 <- limma::eBayes(myfit2)

  # apply test
  myresults <- limma::decideTests(myfit3)

  Group <- dplyr::recode(as.vector(myresults), "1" = "late", "-1" = "early", "0" = "Unbiased")
  names(Group) <- rownames(myresults)

  # Extract the moderated p-values
  moderated_pvalues <- myfit3$p.value %>% as.vector()

  Barbie$Bias_Prop <- data.frame(group = Group,
                                 pvalue = moderated_pvalues)

  # # Extract adjusted p-values using topTable
  top_bars <- limma::topTable(myfit3, number = Inf)  # Set number to Inf to get all barcodes
  adjusted_pvalues <- top_bars[rownames(Barbie$Bias_Prop),]$adj.P.Val

  Barbie$Bias_Prop <- data.frame(group = Group,
                                 pvalue = adjusted_pvalues)

  if(is.null(Barbie$color_panel$bias_group)) {
    Barbie$color_panel$bias_group <- c("late" = "#33AAFF", "early" = "#FF5959", "Unbiased" = "#FFC000")
  }

  return(Barbie)
}

```

