---
title: "neg_sim_DMSO_barbieQ"
author: "LF"
date: "2025-04-04"
output: workflowr::wflow_html
output_dir: ../../docs/bartools_dose_escalation
---

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)

library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
library(bartools)
```

# Install barbieQ - devel

```{r install barbieQ}
# devtools::install_github("Oshlack/barbieQ", force = TRUE)

library(barbieQ)
```

# Load rds

```{r save rds}
# rds saved in "../../analysis/bartools_dose_escalation/preprocessing_barbieQ.Rmd"

load("../../output/bartools_dose_escalation/AML_barbieQ.rda")
```

## t-SNE

```{r tsne}
# Set an appropriate perplexity value (e.g., num_samples / 3)
set.seed(12)
perplexity_value <- min(30, floor(ncol(AML_top) / 3))

tsne_results <- (assays(AML_top)$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  Rtsne::Rtsne(perplexit = perplexity_value)
tsne_coords <- tsne_results$Y
tsne_coords <- as.data.frame(tsne_coords)

ggplot(tsne_coords) +
  geom_point(aes(x = V1, y = V2, col = AML_top$sampleMetadata$Treatment)) +
  labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", col = "Mouse") +
  theme_classic() +
  theme(aspect.ratio = 1)

lapply(
  AML_top$sampleMetadata %>% ncol() %>% seq(), 
  function(i) {
    col_fac <- colnames(AML_top$sampleMetadata)[i]
    ggplot(tsne_coords) +
     geom_point(aes(x = V1, y = V2, col = AML_top$sampleMetadata[[i]])) +
     labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = col_fac) +
      theme_classic() +
     theme(aspect.ratio = 1)
  }
  )

pca_results <- (assays(AML_top)$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  prcomp( scale. = TRUE)

# Step 4: Create a data frame with PCA results and sample metadata
pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  Treatment = AML_top$sampleMetadata$Treatment
)

# Step 5: Plot
ggplot(pca_df, aes(x = PC1, y = PC2, color = Treatment)) +
  geom_point(size = 3) +
  labs(title = "PCA on logCPM", x = "PC1", y = "PC2") +
  theme_classic() +
  theme(aspect.ratio = 1)

```

## bartools

```{r bartools, fig.width=16, fig.height=10}
AML_dge <- DGEList(
  counts = assay(AML_top), 
  group = AML_top$sampleMetadata$Treatment)

bartools::plotBarcodeHistogram(AML_dge)
```


# Subet DMSO

```{r subset dmso}
DMSO <- AML_top[, AML_top$sampleMetadata$Treatment == "DMSO"]

myds <- AML_top$sampleMetadata %>%
    with(model.matrix(~0 + Treatment))

AML_top <- testBarcodeSignif(
  barbieQ = AML_top, designMatrix = myds, method = "diffOcc", contrastFormula = "TreatmentARAC - TreatmentDMSO")

AML_top <- testBarcodeSignif(
  barbieQ = AML_top, designMatrix = myds, contrastFormula = "TreatmentARAC - TreatmentDMSO")

bar_test <- rowData(AML_top)$testingBarcode
bar_test %>% as.data.frame() %>% View()
metadata(bar_test)$design
bar_test$direction %>% table()

plotBarcodeMA(AML_top)
plotBarcodePValue(AML_top)
plotSignifBarcodeHeatmap(AML_top)
plotSignifBarcodeProportion(AML_top)


```

## DMSO t-SNE

```{r DMSO t-SNE}
# Set an appropriate perplexity value (e.g., num_samples / 3)
perplexity_value <- min(30, floor(ncol(DMSO) / 3))

tsne_results <- (assays(DMSO)$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  Rtsne::Rtsne(perplexit = perplexity_value)
tsne_coords <- tsne_results$Y
tsne_coords <- as.data.frame(tsne_coords)

ggplot(tsne_coords) +
  geom_point(aes(x = V1, y = V2, col = DMSO$sampleMetadata$Treatment)) +
  labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", col = "Mouse") +
  theme(aspect.ratio = 1)

lapply(
  DMSO$sampleMetadata %>% ncol() %>% seq(), 
  function(i) {
    col_fac <- colnames(DMSO$sampleMetadata)[i]
    ggplot(tsne_coords) +
     geom_point(aes(x = V1, y = V2, col = DMSO$sampleMetadata[[i]])) +
     labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = col_fac) +
      theme_classic() +
     theme(aspect.ratio = 1)
  }
  )

```


# NegSim Functions


```{r random sampling at each size}

# Calculate the combos
for(i in 3:5) {
  print(choose(10,i) * choose(10-i,i))
}

DMSO <- AML_top[, AML_top$sampleMetadata$Treatment == "DMSO"]

assays(DMSO)$proportion %>% rowMeans() %>% summary()
barbieQ::plotBarcodeProportion(DMSO, absoluteProportion = T)

```




## NegSim_Time

```{r negsim dmso, fig.width=7, fig.height=5}
source("../../analysis/bartools_dose_escalation/NegSim.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/bartools_dose_escalation/AML_barbieQ.rda")

DMSO <- AML_top[, AML_top$sampleMetadata$Treatment == "DMSO"]

set.seed(2025)

results_negsi <- random_sampling(barbieQ = DMSO, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/bartools_dose_escalation/all_loops_time.rda")

# lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, DMSO")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, DMSO")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, DMSO")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, DMSO")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, DMSO")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, DMSO")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, DMSO")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, DMSO")

ggsave(
  "../../output/bartools_dose_escalation/test_time.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 30, height = 15,
  units = "cm"
  )

```

## Inspect NegSim

### load results

```{r extract results from random sampling}
load("../../output/bartools_dose_escalation/all_loops_time.rda")

  # extract results from the loops
  end_sampling <- 5
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```

### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


## flag out the FPR-high simulation
flag_highFPR <- which(df_FPR$Prop_asin > 0.05)
SimID_highFPR <- df_FPR[flag_highFPR,]$SimulationID

# df_Group_Vec[SimID_highFPR,]
```

### P.Val vs. Prop

```{r plot p.val}
stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
# stats_barcodes$Gmean_prop <- (2^stats_barcodes$Gmean_prop-0.5) / 1e6

ggplot(stats_barcodes) +
  geom_point(aes(y = -log10(Prop_asin), x = log10(Amean_prop + 1e-7)), color = "skyblue", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_logit), x = log10(Amean_prop + 1e-7)), color = "yellow", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_noTrans), x = log10(Amean_prop + 1e-7)), color = "pink", alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05))
```

### Binned P.Val vs. Prop

#### Function

```{r bin Amean}
plotP.Val_binnedAmean <- function(stats_barcodes) {
  stats_binned <- stats_barcodes %>%
    mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, 0, by = 1), include.lowest = TRUE)
  ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>%
    mutate(neglog10_p = -log10(pvalue + 1e-6))

ggplot(stats_binned, aes(x = bin, y = neglog10_p, fill = method)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ N_Samples) +
  labs(x = "log10(Amean_prop) bins", y = "-log10(p-value)", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r bin Amean, fig.width=10, fig.height=5}
plotP.Val_binnedAmean(stats_barcodes)
```

### High FPR simulation

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.2)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.2)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.2)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
# df_Group_Vec[flag_highFPR_asin,]
# df_Group_Vec[flag_highFPR_logit,]
# df_Group_Vec[flag_highFPR_noTrans,]

```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)

# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_logit))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_noTrans))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_asin < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_logit < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_noTrans < 0.05)))
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- DMSO
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## re-scale timepoint
  simData$sampleMetadata$Timepoint <- gsub("TP(\\d)", "\\1", simData$sampleMetadata$Timepoint) %>% as.numeric()
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

#### heatmap

```{r, fig.width=10, fig.height=6}
Heatmeap_list <- lapply(rownames(groups_highFPR), plotSimHeatmap)

# Set the resolution and output file for PNG
png("../../output/bartools_dose_escalation/heatmap_simulation_time.png", width = 45, height = 30, units = "in", res = 300)

# Set layout: 2 rows × 2 columns
n_rows <- 6
n_cols <- 5

grid.newpage()
pushViewport(viewport(layout = grid.layout(n_rows, n_cols)))

vplayout <- function(row, col) viewport(layout.pos.row = row, layout.pos.col = col)

for (i in seq_along(Heatmeap_list)) {
  row <- ceiling(i / n_cols)
  col <- (i - 1) %% n_cols + 1
  pushViewport(vplayout(row, col))
  draw(Heatmeap_list[[i]], newpage = FALSE)  # This renders the heatmap in the current viewport
  upViewport()
}

# Close the PNG device
dev.off()
```

## NegSim

```{r negsim dmso, fig.width=7, fig.height=5}
source("../../analysis/bartools_dose_escalation/NegSim.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/bartools_dose_escalation/AML_barbieQ.rda")

DMSO <- AML_top[, AML_top$sampleMetadata$Treatment == "DMSO"]

set.seed(2025)

results_negsi <- random_sampling(barbieQ = DMSO, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/bartools_dose_escalation/all_loops.rda")

# lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, DMSO")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, DMSO")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, DMSO")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, DMSO")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, DMSO")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, DMSO")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, DMSO")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, DMSO")

ggsave(
  "../../output/bartools_dose_escalation/test.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 30, height = 15,
  units = "cm"
  )

```

## Inspect NegSim

### load results

```{r extract results from random sampling}
load("../../output/bartools_dose_escalation/all_loops.rda")

  # extract results from the loops
  end_sampling <- 5
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```

### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


## flag out the FPR-high simulation
flag_highFPR <- which(df_FPR$Prop_asin > 0.05)
SimID_highFPR <- df_FPR[flag_highFPR,]$SimulationID

# df_Group_Vec[SimID_highFPR,]
```

### P.Val vs. Prop

```{r plot p.val}
stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
# stats_barcodes$Gmean_prop <- (2^stats_barcodes$Gmean_prop-0.5) / 1e6

ggplot(stats_barcodes) +
  geom_point(aes(y = -log10(Prop_asin), x = log10(Amean_prop + 1e-7)), color = "skyblue", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_logit), x = log10(Amean_prop + 1e-7)), color = "yellow", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_noTrans), x = log10(Amean_prop + 1e-7)), color = "pink", alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05))
```

### Binned P.Val vs. Prop

#### Function

```{r bin Amean}
plotP.Val_binnedAmean <- function(stats_barcodes) {
  stats_binned <- stats_barcodes %>%
    mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, 0, by = 1), include.lowest = TRUE)
  ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>%
    mutate(neglog10_p = -log10(pvalue + 1e-6))

ggplot(stats_binned, aes(x = bin, y = neglog10_p, fill = method)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ N_Samples) +
  labs(x = "log10(Amean_prop) bins", y = "-log10(p-value)", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r bin Amean, fig.width=10, fig.height=5}
plotP.Val_binnedAmean(stats_barcodes)
```

### High FPR simulation

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.4)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.4)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.4)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
# df_Group_Vec[flag_highFPR_asin,]
# df_Group_Vec[flag_highFPR_logit,]
# df_Group_Vec[flag_highFPR_noTrans,]

```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)

# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_logit))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_noTrans))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_asin < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_logit < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_noTrans < 0.05)))
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- DMSO
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## re-scale timepoint
  simData$sampleMetadata$Timepoint <- gsub("TP(\\d)", "\\1", simData$sampleMetadata$Timepoint) %>% as.numeric()
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

#### heatmap

```{r, fig.width=10, fig.height=6}
Heatmeap_list <- lapply(rownames(groups_highFPR), plotSimHeatmap)

# Set the resolution and output file for PNG
png("../../output/bartools_dose_escalation/heatmap_simulation.png", width = 45, height = 30, units = "in", res = 300)

# Set layout: 2 rows × 2 columns
n_rows <- 6
n_cols <- 5

grid.newpage()
pushViewport(viewport(layout = grid.layout(n_rows, n_cols)))

vplayout <- function(row, col) viewport(layout.pos.row = row, layout.pos.col = col)

for (i in seq_along(Heatmeap_list)) {
  row <- ceiling(i / n_cols)
  col <- (i - 1) %% n_cols + 1
  pushViewport(vplayout(row, col))
  draw(Heatmeap_list[[i]], newpage = FALSE)  # This renders the heatmap in the current viewport
  upViewport()
}

# Close the PNG device
dev.off()
```

## NegSim_late

```{r negsim dmso, fig.width=7, fig.height=5}
source("../../analysis/bartools_dose_escalation/NegSim.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/bartools_dose_escalation/AML_barbieQ.rda")

DMSO <- AML_top[, AML_top$sampleMetadata$Treatment == "DMSO"]

DMSO_late <- DMSO[,DMSO$sampleMetadata$Timepoint %in% c("TP3", "TP4", "TP5")]

set.seed(2025)

results_negsi <- random_sampling(barbieQ = DMSO_late, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/bartools_dose_escalation/all_loops_late.rda")

# lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, DMSO_late")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, DMSO_late")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, DMSO_late")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, DMSO_late")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, DMSO_late")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, DMSO_late")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, DMSO_late")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, DMSO_late")

ggsave(
  "../../output/bartools_dose_escalation/test_late.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 30, height = 15,
  units = "cm"
  )+
  

```

## Inspect NegSim

### load results

```{r extract results from random sampling}
load("../../output/bartools_dose_escalation/all_loops_late.rda")

  # extract results from the loops
  end_sampling <- 3
  
  ## extract FPR
  all_FPR <- lapply(seq_along(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq_along(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq_along(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq_along(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```

### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


## flag out the FPR-high simulation
flag_highFPR <- which(df_FPR$Prop_asin > 0.05)
SimID_highFPR <- df_FPR[flag_highFPR,]$SimulationID

# df_Group_Vec[SimID_highFPR,]
```

### P.Val vs. Prop

```{r plot p.val}
stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
# stats_barcodes$Gmean_prop <- (2^stats_barcodes$Gmean_prop-0.5) / 1e6

ggplot(stats_barcodes) +
  geom_point(aes(y = -log10(Prop_asin), x = log10(Amean_prop + 1e-7)), color = "skyblue", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_logit), x = log10(Amean_prop + 1e-7)), color = "yellow", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_noTrans), x = log10(Amean_prop + 1e-7)), color = "pink", alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05))
```

### Binned P.Val vs. Prop

#### Function

```{r bin Amean}
plotP.Val_binnedAmean <- function(stats_barcodes) {
  stats_binned <- stats_barcodes %>%
    mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, 0, by = 1), include.lowest = TRUE)
  ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>%
    mutate(neglog10_p = -log10(pvalue + 1e-6))

ggplot(stats_binned, aes(x = bin, y = neglog10_p, fill = method)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ N_Samples) +
  labs(x = "log10(Amean_prop) bins", y = "-log10(p-value)", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r bin Amean, fig.width=10, fig.height=5}
plotP.Val_binnedAmean(stats_barcodes)
```

### High FPR simulation

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.2)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.15)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.15)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
# df_Group_Vec[flag_highFPR_asin,]
# df_Group_Vec[flag_highFPR_logit,]
# df_Group_Vec[flag_highFPR_noTrans,]

```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)

# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_logit))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_noTrans))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_asin < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_logit < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_noTrans < 0.05)))
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- DMSO_late
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## re-scale timepoint
  simData$sampleMetadata$Timepoint <- gsub("TP(\\d)", "\\1", simData$sampleMetadata$Timepoint) %>% as.numeric()
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

#### heatmap

```{r, fig.width=10, fig.height=6}
Heatmeap_list <- lapply(rownames(groups_highFPR), plotSimHeatmap)

# Set the resolution and output file for PNG
png("../../output/bartools_dose_escalation/heatmap_simulation_late.png", width = 45, height = 30, units = "in", res = 300)

# Set layout: 2 rows × 2 columns
n_rows <- 6
n_cols <- 5

grid.newpage()
pushViewport(viewport(layout = grid.layout(n_rows, n_cols)))

vplayout <- function(row, col) viewport(layout.pos.row = row, layout.pos.col = col)

for (i in seq_along(Heatmeap_list)) {
  row <- ceiling(i / n_cols)
  col <- (i - 1) %% n_cols + 1
  pushViewport(vplayout(row, col))
  draw(Heatmeap_list[[i]], newpage = FALSE)  # This renders the heatmap in the current viewport
  upViewport()
}

# Close the PNG device
dev.off()
```

