---
title: "compare_original_analysis"
author: "LF"
date: "2024-12-30"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/tagged_bq.rda")
load("../../output/WuC/tagged_bq_min.rda")
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

 - 81 barcodes are selected by the paper, consisting 42% contribution out the of whole dataset based on my proportion data.

```{r paper selected barcode contri}
## tag these barcodes selected by the paper in the pre-filtering data.
## compare the collective proportion of these selected barcodes.
flag_paper <- rownames(ZG66_bq$assay) %in% rownames(ZG66_Fig2)

sum(flag_paper) / nrow(ZG66_bq$assay)

ZG66_bq$isTop$vec <- flag_paper

barbieQ::plotBarcodePareto(barbieQ = ZG66_bq)
barbieQ::plotBarcodeSankey(barbieQ = ZG66_bq)
```

 - My filtered data included 79 out of 81 paper-selected barcodes, missing 2 of them. The two didn't make the rule of being top 99% in at least 6 samples.

```{r overlap with N6_0.99}
ZG66_bq_min <- barbieQ::subsetBarcodes(
  barbieQ = ZG66_bq_min, retainedRows = ZG66_bq_min$isTop$vec)
venn2 <- eulerr::euler(
  list(
    N6_0.99 = rownames(ZG66_bq_min$assay), 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn2, fills = c("red", "blue"), alpha = 0.5, labels = TRUE, quantities = TRUE)
```

## NK-secific Clone Paper Selected

```{r paper barcode prop provided }
## calculate the total proportion of the paper selected barcodes based on the proportion data provided by the paper
NK_clone_contribution <- ZG66_Fig2 %>% colSums()
dat <- data.frame(
  contr = NK_clone_contribution,
  sample = c(rep("T", 6),
             rep("B", 6),
             rep("Gr", 6),
             rep("NK_CD56p_CD16n", 6),
             rep("NK_CD56n_CD16p", 6)))
dat$sample <- factor(dat$sample, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p"))

dat <- dat %>%
  mutate(id = row_number()) 
p11 <- ggplot(dat, aes(x = as.factor(id), y = contr, color = sample)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  ggtitle("Paper-selected clones, paper proportion")
```

```{r paper my prop, fig.width=10, fig.height=5}
## check out paper selected barcodes
flag_paper_filtered <- rownames(myObject$assay) %in% rownames(ZG66_Fig2)
paper_myObject <- barbieQ::subsetBarcodes(
  barbieQ = myObject, retainedRows = flag_paper_filtered)
paper_myObject$testBarcodes$diffProp_Celltype <- myObject$testBarcodes$diffProp_Celltype
paper_myObject$testBarcodes$diffProp_Celltype$results <-   paper_myObject$testBarcodes$diffProp_Celltype$results[flag_paper_filtered,]

## select barcodes selected by paper
paper_contribution <- paper_myObject$proportion[,1:30] %>% colSums()
dat_paper <- data.frame(
  contr = paper_contribution,
  sample = paper_myObject$metadata$Celltype[1:30])

dat_paper <- dat_paper %>% 
  mutate(id = row_number())

p12 <- ggplot(dat_paper, aes(x = as.factor(id), y = contr, color = sample)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  ggtitle("Paper-selected clones, my proportion")

p11 + p12 
```

# Model Design Matrix - filtered

- My selecting strategy to select `NK_CD56n_CD16p biased barcodes` by testing

 - [1] "(CelltypeNK_CD56n_CD16p)" v.s.                                      

 - [2] "(CelltypeB + CelltypeGr + CelltypeT + CelltypeNK_CD56p_CD16n)/4"

- Only keep the samples of "mid" phase, which is of timepoints between "6.5m to 42m".

## model

```{r model desing matrix 1}
## default Gr_1 and Gr_2 to Gr
## remove NKG2A_blah NK subtype, whice were collected in late timepoints

mytargets <-ZG66_bq_min$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6.5, "early", ifelse(Months >42, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))

myObject <- ZG66_bq_min
myObject$metadata <- mytargets
myObject <- barbieQ::subsetSamplesByMetadata(barbieQ = myObject, factor = "Phase", specifiedConditions = "mid")

table(myObject$metadata$Celltype)

```


```{r checkcout df after contrast}
mytargets <-ZG66_bq_min$metadata %>%
  select(Celltype, Months) %>%
  mutate(Phase = ifelse(Months < 6.5, "early", ifelse(Months >42, "late", "mid"))) %>%
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))
mid_flag <- mytargets$Phase == "mid"
mydata <- ZG66_bq_min$proportion[,mid_flag]
mydata <- asin(sqrt(mydata))
mytargets <- mytargets[mid_flag,]
mydesign <- model.matrix(~ 0 + Celltype, data = mytargets)

## complex contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4,
  levels = colnames(mydesign))
## fit design
fit1 <- limma::lmFit(mydata, design = mydesign)
fit2 <- limma::contrasts.fit(fit1, contrasts = mycontrast)
identical(fit1$df.residual, fit2$df.residual)

## simple contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p) - (CelltypeNK_CD56p_CD16n),
  levels = colnames(mydesign))
## fit design
fit1B <- limma::lmFit(mydata, design = mydesign)
fit2B <- limma::contrasts.fit(fit1, contrasts = mycontrast)
identical(fit1B$df.residual, fit2B$df.residual)

identical(fit1B$df.residual, fit1$df.residual)
```

```{r fit model 1}
## define matrix
## will consider Months factor later as each cell type may have different trend
mydesign <- model.matrix(~ 0 + Celltype, data = myObject$metadata)
View(mydesign)

## define contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4,
  levels = colnames(mydesign))

myObject$factorColors$diffProp_Celltype <- NULL
myObject <- barbieQ::testBarcodeBias(
  barbieQ = myObject,
  targets = myObject$metadata,
  designMatrix = mydesign,
  contrast = mycontrast
  )

myObject$testBarcodes$diffProp_Celltype$methods$groups
```

## visualization

```{r compare selects 1}
# barbieQ::plotBarcodeBiasHeatmap(barbieQ = myObject)

N6_0.99_select <- myObject$testBarcodes$diffProp_Celltype$results$direction == "group2"

venn3 <- eulerr::euler(
  list(
    N6_0.99_select = rownames(ZG66_bq_min$assay)[N6_0.99_select], 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn3, fills = c("red", "blue"), alpha = 0.5, labels = TRUE, quantities = TRUE)

## order the slices
myObject$metadata$Celltype <- factor(
  myObject$metadata$Celltype, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p"))
## visualize the sig.
NK_myObject <- barbieQ::subsetBarcodes(
  barbieQ = myObject, retainedRows = myObject$testBarcodes$diffProp_Celltype$results$direction == "group2")
NK_myObject$testBarcodes$diffProp_Celltype <- myObject$testBarcodes$diffProp_Celltype
NK_myObject$testBarcodes$diffProp_Celltype$results <- NK_myObject$testBarcodes$diffProp_Celltype$results[myObject$testBarcodes$diffProp_Celltype$results$direction == "group2",]

## check out my selected barcodes 
anno_sig <- rowAnnotation(
  N6_0.99 = NK_myObject$testBarcodes$diffProp_Celltype$results$direction,
  Paper = rownames(NK_myObject$assay) %in% rownames(ZG66_Fig2),
  col = list(
    N6_0.99 = NK_myObject$factorColors$diffProp_Celltype
    )
  )
set.seed(234)
hp1 <- barbieQ::plotBarcodeHeatmap(barbieQ = NK_myObject, barcodeAnnotation = anno_sig, sampleGroups = "Celltype", splitSamples = T)
hp1
```

```{r compare hp 1}
anno_sig <- rowAnnotation(
  N6_0.99 = myObject$testBarcodes$diffProp_Celltype$results$direction,
  Paper = rownames(myObject$assay) %in% rownames(ZG66_Fig2),
  col = list(
    N6_0.99 = myObject$factorColors$diffProp_Celltype
))

## visualize all barcodes tested
set.seed(234)
barbieQ::plotBarcodeHeatmap(barbieQ = myObject, barcodeAnnotation = anno_sig, sampleGroups = "Celltype", splitSamples = T)

anno_sig <- rowAnnotation(
  N6_0.99 = paper_myObject$testBarcodes$diffProp_Celltype$results$direction,
  Paper = rownames(paper_myObject$assay) %in% rownames(ZG66_Fig2),
  col = list(
    N6_0.99 = paper_myObject$factorColors$diffProp_Celltype
    )
  )
set.seed(234)
barbieQ::plotBarcodeHeatmap(barbieQ = paper_myObject, barcodeAnnotation = anno_sig, sampleGroups = "Celltype", splitSamples = T)

## visualize barcodes not exist in post-filtering data.
point_sup <- ! rownames(ZG66_Fig2) %in% rownames(myObject$assay)

barcode_sup <- rownames(ZG66_Fig2)[point_sup]

count_sup <- ZG66_bq$CPM[barcode_sup, colnames(myObject$assay)]
# count_sup2 <- ZG66_Fig2[barcode_sup, ]

kable(count_sup)

```

## validation

```{r function validation}
plot_validation <- function(NK_myObject) {
  
  NK_clone_contributionU <- NK_myObject$proportion %>% colSums()
  dat <- data.frame(
    contr = NK_clone_contributionU,
    Celltype = NK_myObject$metadata$Celltype,
    Months = NK_myObject$metadata$Months)
  ## sort samples by celltype and months
  dat <- dat %>%
    arrange(Celltype, Months) %>% 
    mutate(id = row_number()) 
  ## add dummy rows for spacing between celltypes
  dat <- dat %>%
    group_by(Celltype) %>%
    mutate(dummy = n() + 1) %>%    # Add a dummy count for spacing
    ungroup() %>%
    bind_rows(dat %>%             # Add dummy rows for spacing
                group_by(Celltype) %>%
                slice_tail(n = 1) %>%
                mutate(id = as.numeric(id) + 0.5, contr = NA, Months = NA)) %>%
    arrange(as.numeric(id))        # Arrange with dummy rows in place
  
  p <- ggplot(dat, aes(x = as.factor(id), y = contr, color = Celltype)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 0.5, size = 4),
        panel.grid.major.x = element_blank()) +
  scale_x_discrete(labels = ifelse(is.na(dat$Months), "", dat$Months)) + # Relabel with `Celltype`
  labs(x = "Samples at at different time points (m)", y = "Total prop. of sig. clones") +  # Update axis labels
  geom_tile(
    data = dat,
    aes(x = ifelse(is.na(Months), NA,order(id)), y = -0.02, fill = Celltype),
    height = 0.02, width = 1
  ) +
  geom_text(
    data = dat %>% 
      mutate(id = order(id)) %>% 
      group_by(Celltype) %>% 
      summarize(mid_point = mean(as.numeric(id))),
    aes(x = mid_point-0.5, y = -0.02, label = Celltype),
    size = 2, vjust = 0.5, color = "black"
  )
  
  return(p)
}
```

```{r NK my select prop}
p22 <- plot_validation(NK_myObject = NK_myObject) + ggtitle("barbieQ-selected, filtered by N6_0.99")

p22

p21 <- plot_validation(NK_myObject = paper_myObject) + ggtitle("paper-selected")

p21
```


# Model Design Matrix - unfiltered

## model

```{r model desing matrix 2}
## default Gr_1 and Gr_2 to Gr
## remove NKG2A_blah NK subtype, whice were collected in late timepoints

mytargetsU <-ZG66_bq$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))

myObjectU <- ZG66_bq
myObjectU$metadata <- mytargetsU
myObjectU <- barbieQ::subsetSamplesByMetadata(barbieQ = myObjectU, factor = "Phase", specifiedConditions = "mid")

table(myObjectU$metadata$Celltype)
table(myObjectU$metadata$Months)
```

```{r fit model 2}
## define matrix
## will consider Months factor later as each cell type may have different trend
mydesignU <- model.matrix(~ 0 + Celltype, data = myObjectU$metadata)
# View(mydesignU)

## define contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4,
  levels = colnames(mydesignU))

myObjectU$factorColors$diffProp_Celltype <- NULL
myObjectU <- barbieQ::testBarcodeBias(
  barbieQ = myObjectU,
  targets = myObjectU$metadata,
  designMatrix = mydesignU,
  contrast = mycontrast
  )

myObjectU$testBarcodes$diffProp_Celltype$methods$groups
```

## visualization

```{r compare selects 2}
# barbieQ::plotBarcodeBiasHeatmap(barbieQ = myObject)

unfiltered_select <- myObjectU$testBarcodes$diffProp_Celltype$results$direction == "group2"

venn4 <- eulerr::euler(
  list(
    unfiltered_select = rownames(ZG66_bq$assay)[unfiltered_select], 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn4, fills = c("red", "blue"), alpha = 0.5, labels = TRUE, quantities = TRUE)

## order the slices
myObjectU$metadata$Celltype <- factor(
  myObjectU$metadata$Celltype, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p"))
## visualize the sig.
NK_myObjectU <- barbieQ::subsetBarcodes(
  barbieQ = myObjectU, retainedRows = unfiltered_select)
NK_myObjectU$testBarcodes$diffProp_Celltype <- myObjectU$testBarcodes$diffProp_Celltype
NK_myObjectU$testBarcodes$diffProp_Celltype$results <- NK_myObjectU$testBarcodes$diffProp_Celltype$results[unfiltered_select,]

## check out my selected barcodes 
anno_sig <- rowAnnotation(
  Unfiltered = NK_myObjectU$testBarcodes$diffProp_Celltype$results$direction,
  Paper = rownames(NK_myObjectU$assay) %in% rownames(ZG66_Fig2),
  col = list(
    Unfiltered = NK_myObjectU$factorColors$diffProp_Celltype
    )
  )
set.seed(234)
hp2 <- barbieQ::plotBarcodeHeatmap(barbieQ = NK_myObjectU, barcodeAnnotation = anno_sig, sampleGroups = "Celltype", splitSamples = T)
hp2
```

```{r compare hp 2}
## visualize all barcodes tested
## over 2000 rows; don't run
# anno_sig <- rowAnnotation(
#   Unfiltered = myObjectU$testBarcodes$diffProp_Celltype$results$direction,
#   Paper = rownames(myObjectU$assay) %in% rownames(ZG66_Fig2),
#   col = list(
#     Unfiltered = myObjectU$factorColors$diffProp_Celltype
# ))
# set.seed(234)
# barbieQ::plotBarcodeHeatmap(barbieQ = myObjectU, barcodeAnnotation = anno_sig, sampleGroups = "Celltype", splitSamples = T)

## check out paper selected barcodes
flag_paper_unfiltered <- rownames(myObjectU$assay) %in% rownames(ZG66_Fig2)
paper_myObjectU <- barbieQ::subsetBarcodes(
  barbieQ = myObjectU, retainedRows = flag_paper_unfiltered)
paper_myObjectU$testBarcodes$diffProp_Celltype <- myObjectU$testBarcodes$diffProp_Celltype
paper_myObjectU$testBarcodes$diffProp_Celltype$results <-   paper_myObjectU$testBarcodes$diffProp_Celltype$results[flag_paper_unfiltered,]

anno_sig <- rowAnnotation(
  Unfiltered = paper_myObjectU$testBarcodes$diffProp_Celltype$results$direction,
  Paper = rownames(paper_myObjectU$assay) %in% rownames(ZG66_Fig2),
  col = list(
    Unfiltered = paper_myObjectU$factorColors$diffProp_Celltype
    )
  )

set.seed(234)
barbieQ::plotBarcodeHeatmap(barbieQ = paper_myObjectU, barcodeAnnotation = anno_sig, sampleGroups = "Celltype", splitSamples = T)

```

## validation

```{r NK my select prop 2}
p23 <- plot_validation(NK_myObject = NK_myObjectU) + ggtitle("barbieQ selected, unfiltered data")
p23

```

::: scroll-plot
```{r compare validation, fig.width=17, fig.height=5}
(p21 + p22 + p23)
```
:::


