---
title: "filtering_explore"
author: "LF"
date: "2024-12-24"
output: 
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_sections: true
---

```{=html}
<style>
.scroll-plot {
    width: 1000px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: scroll;
}
</style>
```

```{r dummy}
# dummy data
data <- data.frame(
  METRIC = paste('var', 1:20),
  VAL = runif(10),
  stringsAsFactors = FALSE
)
```


```{r results = 'asis'}
# automatically generate <style> block of css for each tab
tab_number <- 1:nrow(data)
tab_color <- "#CCCCCC"
css <- sprintf(".color-tabs>.nav-pills>li:nth-child(%d){background:%s;}", tab_number, tab_color)
css <- paste(css, collapse = "\n")
css <- paste("<style>", css, "</style>")
cat(css)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(ggplot2)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
# library(circlize)
library(limma)
```

# Load Processed Data

```{r load rda}
load("tagged_bq.rda")
```

# Filter Barcodes

## 0.99 / minSize

 - Find minimum group size.

```{r clean up}
ZH33_bq$metadata$Celltype %>% table()
groups <- ZG66_bq$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype)) %>% 
  filter(Phase == "mid")
```

 - Tag top barcodes.

```{r tag 0.99 minSize}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ## 4
# ZJ31_bq_min <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq, proportionThreshold = 0.99, minTopFrequency = 4)
# ### 8
# ZH33_bq_min <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq, proportionThreshold = 0.99, minTopFrequency = 8)
# ##6
# ZG66_bq_min <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq, proportionThreshold = 0.99, minTopFrequency = 6)
# 
# save(ZJ31_bq_min, ZH33_bq_min, ZG66_bq_min, file = "tagged_bq_min.rda")

load("tagged_bq_min.rda")
```

 - Visualize top vs. bottom barcodes

```{r vis tagging}
barbieQ::plotBarcodeSankey(barbieQ = ZJ31_bq_min)
barbieQ::plotBarcodeSankey(barbieQ = ZH33_bq_min)
barbieQ::plotBarcodeSankey(barbieQ = ZG66_bq_min)
```

```{r vis pareto}
# taking too long
# barbieQ::plotBarcodePareto(barbieQ = ZJ31_bq_min)
# barbieQ::plotBarcodePareto(barbieQ = ZH33_bq_min)
# barbieQ::plotBarcodePareto(barbieQ = ZG66_bq_min)
```

 - Subset barcodes

```{r subset barcodes}

ZG66_bq_min$isTop$vec %>% table()
ZG66_bq_min <- barbieQ::subsetBarcodes(barbieQ = ZG66_bq_min, retainedRows = ZG66_bq_min$isTop$vec)
```

```{r heatmap min}
set.seed(99)
barbieQ::plotBarcodeHeatmap(ZG66_bq_min)
```

## 0.9 / 2

```{r load 0.9}
load("tagged_bq_0.9.rda")
ZG66_bq_0.9$isTop$vec %>% table()

## let n = 2
new_tops <- rowSums(ZG66_bq_0.9$isTop$mat) >= 2
table(new_tops)
ZG66_bq_0.9$isTop$vec <- new_tops

ZG66_bq_0.9 <- barbieQ::subsetBarcodes(barbieQ = ZG66_bq_0.9, retainedRows = new_tops)
```

```{r heatmap 0.9}
set.seed(99)
# barbieQ::plotBarcodeHeatmap(ZG66_bq_0.9)
```

## overlaps

 - Similar number of barcodes were retained through the two filtering strategy, and most of them overlap. However, approximately 20% from each are unique.

```{r filtering overlaps}
top_by_min <- ZG66_bq_min$assay %>% rownames()
top_by_0.9 <- ZG66_bq_0.9$assay %>% rownames()

venn1 <- eulerr::euler(
  list(
    N6_0.99 = top_by_min, 
    N2_0.9 = top_by_0.9))

plot(venn1, fills = c("red", "blue"), alpha = 0.5, labels = TRUE, quantities = TRUE)

# ggVennDiagram::ggVennDiagram(
#   x = list(top_by_min, top_by_0.9),
#   label = "both",
#   category.names = c("N6_0.99", "N2_0.9"),
#   label_size = 3,
#   set_size = 3,
#   order.set.by = "name"
# ) + 
#   scale_x_continuous(expand = expansion(mult = .2)) +
#   labs(
#     title = "Venn Diagram of Barcodes Sets Filtered by Two Strategies",
#     subtitle = "Strategy 1: top 90% in 2 samples at least; \nStrategy 2: top 99% in 6 samples as minimum group size.",
#     caption = Sys.Date())

ggVennDiagram::ggVennDiagram(
  x = list(top_by_min, top_by_0.9),
  label = "both",
  category.names = c("N6_0.99", "N2_0.9"),
  label_size = 3,
  set_size = 3,
  force_upset = T
) 
```


# Model Sample Design

 - Model experimental design

```{r model desing matrix}
## default Gr_1 and Gr_2 to Gr
## remove NKG2A_blah NK subtype, which were collected in late timepoints

mytargets <-ZG66_bq_min$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))

## trim samples - exclude early and late time samples, keeping only mid samples to test
## First filtering strategy
ZG66_test_min <- ZG66_bq_min
ZG66_test_min$metadata <- mytargets
ZG66_test_min <- barbieQ::subsetSamplesByMetadata(barbieQ = ZG66_test_min, factor = "Phase", specifiedConditions = "mid")
table(ZG66_test_min$metadata$Celltype)
## Second filtering strategy
ZG66_test_0.9 <- ZG66_bq_0.9
ZG66_test_0.9$metadata <- mytargets
ZG66_test_0.9 <- barbieQ::subsetSamplesByMetadata(barbieQ = ZG66_test_0.9, factor = "Phase", specifiedConditions = "mid")
table(ZG66_test_0.9$metadata$Celltype)

## load pre-filtering data
load("ZJ31_ZH33_ZG66_bq.rda")
ZG66_test <- ZG66_bq
ZG66_test$metadata <- mytargets
ZG66_test <- barbieQ::subsetSamplesByMetadata(barbieQ = ZG66_test, factor = "Phase", specifiedConditions = "mid")
table(ZG66_test$metadata$Celltype)
```

 - Fit model
 
```{r fit model}
## define matrix
## will consider Months factor later as each cell type may have different trend
mydesign <- model.matrix(~ 0 + Celltype, data = ZG66_test_min$metadata)
# View(mydesign)

## define contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p+CelltypeNK_CD56p_CD16n)/2 - (CelltypeB+CelltypeGr+CelltypeT)/3,
  levels = colnames(mydesign))
```
 
```{r test}
ZG66_test_min <- barbieQ::testBarcodeBias(
  barbieQ = ZG66_test_min,
  targets = ZG66_test_min$metadata,
  designMatrix = mydesign,
  contrast = mycontrast
  )

ZG66_test_min$testBarcodes$diffProp_Celltype$methods$groups


ZG66_test_0.9 <- barbieQ::testBarcodeBias(
  barbieQ = ZG66_test_0.9,
  targets = ZG66_test_0.9$metadata,
  designMatrix = mydesign,
  contrast = mycontrast
  )

ZG66_test_0.9$testBarcodes$diffProp_Celltype$methods$groups

ZG66_test <- barbieQ::testBarcodeBias(
  barbieQ = ZG66_test,
  designMatrix = mydesign,
  contrast = mycontrast
)
ZG66_test$testBarcodes$diffProp_Celltype$methods$groups
```

# Compare Test Results

 - Heatmap visualization 
 
 - More significant results after filtering, while less barcodes are tested during multiple testing. The multiple testing size of pre-filtering data is ten times the post-filtering data, leading to stronger FDR correction and resulting in less significant results.
 
```{r test category}
ZG66_test_min$testBarcodes$diffProp_Celltype$results$direction %>% table()

ZG66_test_0.9$testBarcodes$diffProp_Celltype$results$direction %>% table()

ZG66_test$testBarcodes$diffProp_Celltype$results$direction %>% table()
```

```{r vis test}
print(nrow(ZG66_test_min$CPM))
print(nrow(ZG66_test_0.9$CPM))
# barbieQ::plotBarcodeBiasHeatmap(ZG66_test_min)
# barbieQ::plotBarcodeBiasHeatmap(ZG66_test_0.9)
```

```{r group1 compare}
plot(
  eulerr::euler(list(
    Pre = rownames(ZG66_test$assay)[ZG66_test$testBarcodes$diffProp_Celltype$results$direction == "group1"],
    N6_0.99 = rownames(ZG66_test_min$assay)[ZG66_test_min$testBarcodes$diffProp_Celltype$results$direction == "group1"],
    N2_0.9 = rownames(ZG66_test_0.9$assay)[ZG66_test_0.9$testBarcodes$diffProp_Celltype$results$direction == "group1"])),
  fills = c("red", "blue", "green"), alpha = 0.5, labels = TRUE, quantities = TRUE
  )
```

```{r group2 compare}
plot(
  eulerr::euler(list(
    Pre = rownames(ZG66_test$assay)[ZG66_test$testBarcodes$diffProp_Celltype$results$direction == "group2"],
    N6_0.99 = rownames(ZG66_test_min$assay)[ZG66_test_min$testBarcodes$diffProp_Celltype$results$direction == "group2"],
    N2_0.9 = rownames(ZG66_test_0.9$assay)[ZG66_test_0.9$testBarcodes$diffProp_Celltype$results$direction == "group2"])),
  fills = c("red", "blue", "green"), alpha = 0.5, labels = TRUE, quantities = TRUE
  )
```

```{r n.s. compare}
plot(
  eulerr::euler(list(
    Pre = rownames(ZG66_test$assay)[ZG66_test$testBarcodes$diffProp_Celltype$results$direction == "n.s."],
    N6_0.99 = rownames(ZG66_test_min$assay)[ZG66_test_min$testBarcodes$diffProp_Celltype$results$direction == "n.s."],
    N2_0.9 = rownames(ZG66_test_0.9$assay)[ZG66_test_0.9$testBarcodes$diffProp_Celltype$results$direction == "n.s."])),
  fills = c("red", "blue", "green"), alpha = 0.5, labels = TRUE, quantities = TRUE
  )
```

```{r sig. compare}
plot(
  eulerr::euler(list(
    Pre = rownames(ZG66_test$assay)[ZG66_test$testBarcodes$diffProp_Celltype$results$direction != "n.s."],
    N6_0.99 = rownames(ZG66_test_min$assay)[ZG66_test_min$testBarcodes$diffProp_Celltype$results$direction != "n.s."],
    N2_0.9 = rownames(ZG66_test_0.9$assay)[ZG66_test_0.9$testBarcodes$diffProp_Celltype$results$direction != "n.s."])),
  fills = c("red", "blue", "green"), alpha = 0.5, labels = TRUE, quantities = TRUE
  )
```

# All Relevant Barcodes {.color-tabs .tabset .tabset-fade .tabset-pills}

## Sets of significant barcodes

 - Retain barcodes being significant in pre-filtering data, or being included in either the filtered data.
 
 - Pointing based on the pre-filtering data.

```{r keep all}
pointer <- ZG66_test$testBarcodes$diffProp_Celltype$results$direction != "n.s." |
  rownames(ZG66_test$assay) %in% rownames(ZG66_test_min$assay) |
  rownames(ZG66_test$assay) %in% rownames(ZG66_test_0.9$assay)

sum(pointer)

ZG66_all <- barbieQ::subsetBarcodes(barbieQ = ZG66_bq, retainedRows = pointer)

## make a big data.frame including all results
## extract test results from each
results_N6_0.99 <- ZG66_test_min$testBarcodes$diffProp_Celltype$results
colnames(results_N6_0.99) <- paste0("N6_0.99_", colnames(results_N6_0.99))
results_N2_0.9 <- ZG66_test_0.9$testBarcodes$diffProp_Celltype$results
colnames(results_N2_0.9) <- paste0("N2_0.9_", colnames(results_N2_0.9))
results_Pre <- ZG66_test$testBarcodes$diffProp_Celltype$results[pointer,]
colnames(results_Pre) <- paste0("Pre_", colnames(results_Pre))

## merge the results by rownames
results_all <- merge(results_Pre, results_N6_0.99, by = "row.names", all = TRUE)
rownames(results_all) <- results_all$Row.names
results_all <- results_all[, -1] # Remove the extra row names column

results_all <- merge(results_all, results_N2_0.9, by = "row.names", all = TRUE)
rownames(results_all) <- results_all$Row.names
results_all <- results_all[, -1] # Remove the extra row names column
## Order barcode in line with ZG66_all
results_all <- results_all[rownames(ZG66_all$assay),]
```

 - In total, 799 barcodes are being tested as significant in either pre-filtering data, or post-filtering data out of two strategies. Here we discuss significance based on **adjusted p.values**.
 
 - From those, nearly 50% are consistently significant in the three data; 
 
  - 40% are significant in both the post-filtering data but n.s in the pre-filtering data; 
  
  - the rest barcodes are either uniquely sig. in one data or sig. in other combos.

 - For barcodes with p.values (**non-adjusted**) < 0.05 in the filtered data but >= 0.05 in the pre-filtering data, this is unexpected in normal testing because each barcodes are independently tested. But here this phenomenon is because of the moderated t-test implemented in the limma model that the estimated variance is dependant on across the barcodes.

```{r venn of all}
ggVennDiagram::ggVennDiagram(
  x = list(
    Pre_sig = rownames(results_all)[!is.na(results_all$Pre_direction) & results_all$Pre_direction != "n.s."],
    N6_0.99_sig = rownames(results_all)[!is.na(results_all$N6_0.99_direction) & results_all$N6_0.99_direction != "n.s."],
    N2_0.9_sig = rownames(results_all)[!is.na(results_all$N2_0.9_direction) & results_all$N2_0.9_direction != "n.s."]),
  label = "both",
  # category.names = c("Pre", "N6_0.99", "N2_0.9"),
  label_size = 3,
  set_size = 3,
  force_upset = T
) 

ggVennDiagram::ggVennDiagram(
  x = list(
    Pre_sig = rownames(results_all)[!is.na(results_all$Pre_p.value) & results_all$Pre_p.value < 0.05],
    N6_0.99_sig = rownames(results_all)[!is.na(results_all$N6_0.99_p.value) & results_all$N6_0.99_p.value < 0.05],
    N2_0.9_sig = rownames(results_all)[!is.na(results_all$N2_0.9_p.value) & results_all$N2_0.9_p.value < 0.05]),
  label = "both",
  # category.names = c("Pre", "N6_0.99", "N2_0.9"),
  label_size = 3,
  set_size = 3,
  force_upset = T
) 

## check the unexpected set.
# ttt <- ((!is.na(results_all$N6_0.99_p.value) & results_all$N6_0.99_p.value < 0.05) & (!is.na(results_all$Pre_p.value) & results_all$Pre_p.value >= 0.05) )
# View(results_all[ttt,])

ggVennDiagram::ggVennDiagram(
  x = list(
    Pre_Other = rownames(results_all)[!is.na(results_all$Pre_direction) & results_all$Pre_direction == "group1"],
    Pre_NK = rownames(results_all)[!is.na(results_all$Pre_direction) & results_all$Pre_direction == "group2"],
    
    N6_0.99_Other = rownames(results_all)[!is.na(results_all$N6_0.99_direction) & results_all$N6_0.99_direction == "group1"],
    N6_0.99_NK = rownames(results_all)[!is.na(results_all$N6_0.99_direction) & results_all$N6_0.99_direction == "group2"],
    
    N2_0.9_Other = rownames(results_all)[!is.na(results_all$N2_0.9_direction) & results_all$N2_0.9_direction == "group1"],
    N2_0.9_NK = rownames(results_all)[!is.na(results_all$N2_0.9_direction) & results_all$N2_0.9_direction == "group2"]),
  label = "both",
  # category.names = c("Pre", "N6_0.99", "N2_0.9"),
  label_size = 3,
  set_size = 3,
  force_upset = T
) 
```

## P.values of each set

```{r barcode set}
# Process the data for Venn diagram
cate_adj <- ggVennDiagram::process_data(
  Venn(
    list(
    Pre_sig = rownames(results_all)[!is.na(results_all$Pre_direction) & results_all$Pre_direction != "n.s."],
    N6_0.99_sig = rownames(results_all)[!is.na(results_all$N6_0.99_direction) & results_all$N6_0.99_direction != "n.s."],
    N2_0.9_sig = rownames(results_all)[!is.na(results_all$N2_0.9_direction) & results_all$N2_0.9_direction != "n.s."])
))
# Extract set (region) information (elements for each region)
region_info <- cate_adj$regionData
region_info$item[[1]] %>% class()

long_region <- region_info %>%
  select(item, name) %>%         # Select relevant columns: items and region names
  unnest(item)              # Expand list-column `item` into rows
# rownames(long_region) <- long_region$item
# long_region <- long_region[,2, drop=FALSE]

## Merge sregion information with test results
results_all <- merge(results_all, long_region, by.x = "row.names", by.y = "item", all = TRUE)
colnames(results_all)[17] <- "region"
rownames(results_all) <- results_all$Row.names
results_all <- results_all[, -1]
## order results_all by ZG66_all
results_all <- results_all[rownames(ZG66_all$assay),]

```

 - The most significant barcodes (with relatively low p.values) tend to be defined as "sig" in all pre-filtering and post-filtering data. 

```{r vis barcode region}
## visualize adj.pvalues of different category
ggplot(results_all) +
  geom_point(aes(x = rowMeans(ZG66_all$rank), y = Pre_adj.p.value, shape = Pre_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 

## visualize adj.pvalues of different category
ggplot(results_all) +
  geom_point(aes(x = rowMeans(ZG66_all$rank), y = Pre_p.value, shape = Pre_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 

## visualize adj.pvalues of different category
ggplot(results_all) +
  geom_point(aes(x = rowMeans(ZG66_all$rank), y = N6_0.99_adj.p.value, shape = N6_0.99_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 

## visualize adj.pvalues of different category
ggplot(results_all) +
  geom_point(aes(x = rowMeans(ZG66_all$rank), y = N6_0.99_p.value, shape = N6_0.99_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 

## visualize adj.pvalues of different category
ggplot(results_all) +
  geom_point(aes(x = rowMeans(ZG66_all$rank), y = N2_0.9_adj.p.value, shape = N2_0.9_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 

## visualize adj.pvalues of different category
ggplot(results_all) +
  geom_point(aes(x = rowMeans(ZG66_all$rank), y = N2_0.9_p.value, shape = N2_0.9_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 
```


```{r subset sig}
## flag barcodes that belongs to regions other than NA - which means being significant in at least pre/filter1/filter2 data
flag_sig <- !is.na(results_all$region)
sum(flag_sig)

identical(rownames(results_all), rownames(ZG66_all$assay))

## subset object by the flagged barcodes
ZG66_sig <- barbieQ::subsetBarcodes(barbieQ = ZG66_all, retainedRows = flag_sig)
results_sig <- results_all[flag_sig,]
ggplot(results_sig) +
  geom_point(aes(x = rowMeans(ZG66_sig$rank), y = Pre_adj.p.value, shape = Pre_direction, color = region)) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed", linewidth = 1) 
```

## CPM of significant barcodes

 - Linear model on barcode proportion is based on only the right two slices in the heatmap (midgroup1 and midgroup2).

 - group1 is B, T, Gr cells; group2 is NK cells inclduing subtypes.

 - Statistical test results on the three pre- or post- filtering data are visualized by the row annotation in the right side of the heatmap. 
 
 - Mostly, results on the two post-filtering data of different filtering strategies mostly agree with each other unless barcodes are filtered out in one data.
 
 - For barcodes that are decided as significant in the post-filtering data but n.s. in pre-filtering data, they still look interesting and do show variability based on visual vibe.
 
 - For barcodes being significant in the pre-filtering data only, they are not included in the post-filtering data, but do look interesting and do show variability based on visual vibe. So we've risked filtering our some informative barcodes using the current two strategies - but that's less than 10%.

```{r vis subset sig, fig.width=9, fig.height=5}
barbieQ::plotBarcodeHeatmap(barbieQ = ZG66_sig)

anno_sig <- rowAnnotation(
  Pre = results_sig$Pre_direction,
  N6_0.99 = results_sig$N6_0.99_direction,
  N2_0.9 = results_sig$N2_0.9_direction,
  col = list(
    Pre = ZG66_test$factorColors$diffProp_Celltype,
    N6_0.99 = ZG66_test$factorColors$diffProp_Celltype,
    N2_0.9 = ZG66_test$factorColors$diffProp_Celltype)
)



ZG66_sig$metadata <-ZG66_sig$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  # mutate(Phase = factor(Phase, levels = c("early", "mid", "late"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype)) %>% 
  mutate(Group = ifelse(Celltype %in% c("T", "B", "Gr"), "group1", "group2")) %>% 
  mutate(Slice = paste0(Phase, Group))

set.seed(234)
barbieQ::plotBarcodeHeatmap(barbieQ = ZG66_sig, barcodeAnnotation = anno_sig, sampleGroups = "Slice", splitSamples = T)
```

