---
title: "filtered v.s. unfiltered"
author: "LF"
date: "2025-02-17"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
## ../../ moves two levels up; ../ moves one level up; ./ refers to current directory
## this processed rda file is conducted in the ./LF_start_mat.Rmd
load("../../output/WuC/tagged_bq.rda")
```

# Filter Barcodes

## 0.99 / minSize

 - Find minimum group size.

```{r clean up}
ZG66_bq$metadata$Celltype %>% table()
groups <- ZG66_bq$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype)) %>% 
  filter(Phase == "mid")
groups$Celltype %>% table()
```

 - Tag top barcodes.

```{r tag 0.99 minSize}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ## 4
# ZJ31_bq_min <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq, proportionThreshold = 0.99, minTopFrequency = 4)
# ## 8
# ZH33_bq_min <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq, proportionThreshold = 0.99, minTopFrequency = 8)
# ## 6
# ZG66_bq_min <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq, proportionThreshold = 0.99, minTopFrequency = 6)
# 
# save(ZJ31_bq_min, ZH33_bq_min, ZG66_bq_min, file = "../../output/WuC/tagged_bq_min.rda")

load("../../output/WuC/tagged_bq_min.rda")
```

 - Visualize top vs. bottom barcodes

```{r vis tagging}
barbieQ::plotBarcodeSankey(barbieQ = ZJ31_bq_min)
barbieQ::plotBarcodeSankey(barbieQ = ZH33_bq_min)
barbieQ::plotBarcodeSankey(barbieQ = ZG66_bq_min)
```

 - Subset barcodes

```{r subset barcodes}
ZG66_bq_min$isTop$vec %>% table()
ZG66_bq_min <- barbieQ::subsetBarcodes(barbieQ = ZG66_bq_min, retainedRows = ZG66_bq_min$isTop$vec)
```

::: scroll-plot
```{r heatmap min, fig.width=9, fig.height=5}
set.seed(99)
barbieQ::plotBarcodeHeatmap(ZG66_bq_min)
```
:::

# Model Sample Design

 - Model experimental design

```{r model desing matrix}
## default Gr_1 and Gr_2 to Gr
## remove NKG2A_blah NK subtype, which were collected in late timepoints
mytargets <-ZG66_bq_min$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))

## trim samples - exclude early and late time samples, keeping only mid samples to test
## filtering strategy
ZG66_test_min <- ZG66_bq_min
ZG66_test_min$metadata <- mytargets
ZG66_test_min <- barbieQ::subsetSamplesByMetadata(barbieQ = ZG66_test_min, factor = "Phase", specifiedConditions = "mid")
table(ZG66_test_min$metadata$Celltype)

## load Unfiltered data
load("../../output/WuC/ZJ31_ZH33_ZG66_bq.rda")
ZG66_test <- ZG66_bq
ZG66_test$metadata <- mytargets
ZG66_test <- barbieQ::subsetSamplesByMetadata(barbieQ = ZG66_test, factor = "Phase", specifiedConditions = "mid")
table(ZG66_test$metadata$Celltype)
```

 - Fit model
 
```{r fit model}
## define matrix
## will consider Months factor later as each cell type may have different trend
mydesign <- model.matrix(~ 0 + Celltype, data = ZG66_test_min$metadata)
# View(mydesign)

## define contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4,
  levels = colnames(mydesign))
```
 
```{r test}
ZG66_test_min <- barbieQ::testBarcodeBias(
  barbieQ = ZG66_test_min,
  # targets = ZG66_test_min$metadata,
  designMatrix = mydesign,
  contrast = mycontrast
  )
ZG66_test_min$testBarcodes$diffProp_Celltype$methods$groups
barbieQ::plotBarcodeBiasHeatmap(ZG66_test_min)


ZG66_test <- barbieQ::testBarcodeBias(
  barbieQ = ZG66_test,
  designMatrix = mydesign,
  contrast = mycontrast
)
ZG66_test$testBarcodes$diffProp_Celltype$methods$groups

```

# Question2: Should we filter at all?

## Compare Testing Results from Filtered and Unfilterd Data

Overlaps

```{r filtered vs un}
ggVennDiagram::ggVennDiagram(
  x = list(
    Unfiltered_sig = rownames(ZG66_test$assay)[ZG66_test$testBarcodes$diffProp_Celltype$results$direction != "n.s."],
    N6_0.99_sig = rownames(ZG66_test_min$assay)[ZG66_test_min$testBarcodes$diffProp_Celltype$results$direction != "n.s."]),
  label = "both",
  # category.names = c("Unfiltered", "N6_0.99"),
  label_size = 3,
  set_size = 3,
  force_upset = T
) 
```

## All Relevant Barcodes {.color-tabs .tabset .tabset-fade .tabset-pills}

### Subset of significant barcodes

 - Retain barcodes being significant in unfiltered data, or being included in the filtered data.
 
 - Pointing based on the unfiltered data.

```{r keep all 2}
pointer <- ZG66_test$testBarcodes$diffProp_Celltype$results$direction != "n.s." |
  rownames(ZG66_test$assay) %in% rownames(ZG66_test_min$assay) 

sum(pointer)

ZG66_two <- barbieQ::subsetBarcodes(barbieQ = ZG66_bq, retainedRows = pointer)

## make a big data.frame including all results
## extract test results from each
results_N6_0.99 <- ZG66_test_min$testBarcodes$diffProp_Celltype$results
colnames(results_N6_0.99) <- paste0("N6_0.99_", colnames(results_N6_0.99))
results_Unfiltered <- ZG66_test$testBarcodes$diffProp_Celltype$results[pointer,]
colnames(results_Unfiltered) <- paste0("Unfiltered_", colnames(results_Unfiltered))

## merge the results by rownames
results_two <- merge(results_Unfiltered, results_N6_0.99, by = "row.names", all = TRUE)
rownames(results_two) <- results_two$Row.names
results_two <- results_two[, -1] # Remove the extra row names column

## Order barcode in line with ZG66_two
results_two <- results_two[rownames(ZG66_two$assay),]
```

### Regions of sig. status

```{r barcode set}
# Process the data for Venn diagram
cate_adj <- ggVennDiagram::process_data(
  Venn(
    list(
    Unfiltered_sig = rownames(results_two)[!is.na(results_two$Unfiltered_direction) & results_two$Unfiltered_direction != "n.s."],
    N6_0.99_sig = rownames(results_two)[!is.na(results_two$N6_0.99_direction) & results_two$N6_0.99_direction != "n.s."])
))
# Extract set (region) information (elements for each region)
region_info <- cate_adj$regionData
region_info$item[[1]] %>% class()

long_region <- region_info %>%
  select(item, name) %>%         # Select relevant columns: items and region names
  unnest(item)              # Expand list-column `item` into rows
# rownames(long_region) <- long_region$item
# long_region <- long_region[,2, drop=FALSE]

## Merge sregion information with test results
results_two <- merge(results_two, long_region, by.x = "row.names", by.y = "item", all = TRUE)
colnames(results_two)[12] <- "region"
rownames(results_two) <- results_two$Row.names
results_two <- results_two[, -1]
## order results_two by ZG66_two
results_two <- results_two[rownames(ZG66_two$assay),]

## reorder levels in region
results_two$region <- factor(results_two$region, 
                             levels = c("NA", "Unfiltered_sig/N6_0.99_sig", "N6_0.99_sig", "Unfiltered_sig")) 
results_two$Unfiltered_direction <- factor(results_two$Unfiltered_direction,
                                           levels = c("group2", "group1", "n.s."),
                                           labels = c("Higher prop. in NKnp", "Higher prop. in others", "n.s."))
results_two$N6_0.99_direction <- factor(results_two$N6_0.99_direction,
                                           levels = c("group2", "group1", "n.s."),
                                           labels = c("Higher prop. in NKnp", "Higher prop. in others", "n.s."))

```

### logFC

```{r logFC}
## the logFC in the testing results is actually the minus of two group mean
colnames(results_two)[4] <- "Unfiltered_minus"
colnames(results_two)[9] <- "N6_0.99_minus"

# calculate the actual logFC of the proportion
ZG66_two$metadata <- ZG66_two$metadata %>% 
  mutate(group = ifelse(Celltype == "NK_CD56n_CD16p", "NKnp", "others"))
group_mean <- sapply(
  unique(ZG66_two$metadata$group), 
  function(g) rowMeans(ZG66_two$proportion[, ZG66_two$metadata$group == g, drop = FALSE]))

group_mean <- as.data.frame(group_mean)

## by defination, MA plot should be LFC v.s. average log data
group_mean <- group_mean %>% 
  mutate(LFC = log2((NKnp + 0.0000001) / (others + 0.0000001))) %>% 
  mutate(LogAvg = log2(rowMeans(ZG66_two$proportion))) %>% 
  mutate(AvgLog = rowMeans(log2(ZG66_two$proportion + 0.0000001)))

## compute MA measurements for transformed data
group_mean_tf <- sapply(
  unique(ZG66_two$metadata$group), 
  function(g) rowMeans(asin(sqrt(ZG66_two$proportion))[, ZG66_two$metadata$group == g, drop = FALSE]))

group_mean_tf <- as.data.frame(group_mean_tf)

## by defination, MA plot should be LFC v.s. average log data
group_mean_tf <- group_mean_tf %>% 
  mutate(LFC = log2((NKnp + 0.0000001) / (others + 0.0000001))) %>% 
  mutate(LogAvg = log2(rowMeans(asin(sqrt(ZG66_two$proportion))))) %>% 
  mutate(AvgLog = rowMeans(log2(asin(sqrt(ZG66_two$proportion)) + 0.0000001)))

identical(rownames(group_mean_tf), rownames(results_two))

group_mean$region <- results_two$region
group_mean_tf$region <- results_two$region
```

### MA plot: proportion

```{r MA proportion}
## visualize logFC v.s. Avg. log prop
ggplot(group_mean) +
  geom_point(aes(x = AvgLog, y = LFC, color = region), shape = 1) +
  ggtitle("MA plot of Barcode proportion") +
  labs(y = "Log fold change of barcode proportion", x = "Average log barcode proportion") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig"))) 
  ## set up shape, specify shape by direction
  # scale_shape_manual(values = setNames(c(2,6,1), c("Higher prop. in NKnp", "Higher prop. in others", "n.s.")))

## visualize logFC v.s. log avg. prop
ggplot(group_mean) +
  geom_point(aes(x = LogAvg, y = LFC, color = region), shape = 1) +
  ggtitle("MA plot of Barcode proportion") +
  labs(y = "Log fold change of barcode proportion", x = "Log average barcode proportion") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))
```

### MA plot: asin-sqrt proportion

```{r MA asin-sqrt proportion}
## visualize logFC v.s. Avg. log prop
ggplot(group_mean_tf) +
  geom_point(aes(x = AvgLog, y = LFC, color = region), shape = 1) +
  ggtitle("MA plot of transformed barcode proportion") +
  labs(y = "Log fold change of barcode asin-sqrt proportion", x = "Average log barcode asin-sqrt proportion") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig"))) 
  ## set up shape, specify shape by direction
  # scale_shape_manual(values = setNames(c(2,6,1), c("Higher prop. in NKnp", "Higher prop. in others", "n.s.")))

## visualize logFC v.s. log avg. prop
ggplot(group_mean_tf) +
  geom_point(aes(x = LogAvg, y = LFC, color = region), shape = 1) +
  ggtitle("MA plot of transformed barcode proportion") +
  labs(y = "Log fold change of barcode asin-sqrt proportion", x = "Log average barcode asin-sqrt proportion") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))
```

### Volcano plot

```{r volcano}
## visualize logFC v.s. Avg. log prop
ggplot(group_mean) +
  geom_point(aes(x = -log10(results_two$Unfiltered_adj.p.value), y = LFC, color = region), shape = 1) +
  ggtitle("MA plot of Barcode proportion") +
  labs(y = "Log fold change of barcode proportion", x = "-Log adjusted p.value") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig"))) 
```


### P.values

 - The multiple testing correction in the unfiltered data should be more intense than that in the filtered data. So, we'd expect adjusted p.values on the x axis being increased more than that on the y axis. But why is the curve not smooth? At some points, y axis corrected (enhanced) the p.values more intense than we expect. My hypothesis is that, those points are dense with barcodes got filtered out - resulting in the retained p.values' ranking of significant level much lower than their ranking in the unfiltered data.

```{r plot p.values two data}
sum(pointer)
missing <- ZG66_test$testBarcodes$diffProp_Celltype$results[!pointer,]

ggplot(results_two) +
  geom_point(aes(x = Unfiltered_adj.p.value, y = N6_0.99_adj.p.value, color = region), shape = 1) +
  # geom_point(data = missing, 
  #            aes(x = adj.p.value, y = p.value), shape = 2, color = "yellow") +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, color = "purple", linetype = "dashed") +
  theme(aspect.ratio = 1) +
  labs(x = "Adjusted p.values from unfiltered data", y = "Adjusted p.values from data filtered by N6_0.99") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))

ggplot(results_two) +
  geom_point(aes(x = Unfiltered_p.value, y = N6_0.99_p.value, color = region), shape = 1) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, color = "purple", linetype = "dashed") +
  theme(aspect.ratio = 1) +
  labs(x = "p.values from unfiltered data", y = "p.values from data filtered by N6_0.99") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))

ggplot(results_two) +
  geom_point(aes(x = N6_0.99_p.value, y = N6_0.99_adj.p.value, color = region), shape = 1) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, color = "purple", linetype = "dashed") +
  theme(aspect.ratio = 1) +
  labs(x = "p.values from filtered data", y = "Adjusted p.values from data filtered by N6_0.99") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))

ggplot(results_two) +
  geom_point(aes(x = Unfiltered_p.value, y = Unfiltered_adj.p.value, color = region), shape = 1) +
  geom_hline(yintercept = 0.05, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, color = "purple", linetype = "dashed") +
  theme(aspect.ratio = 1) +
  labs(x = "p.values from unfiltered data", y = "Adjusted p.values from unfiltered data") +
  theme_classic() +
  scale_color_manual(
    values = setNames(scales::hue_pal()(3), 
                      c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))

# ggplot(results_two) +
#   geom_point(aes(x = -log10(Unfiltered_adj.p.value), y = -log10(N6_0.99_adj.p.value), color = region), shape = 1) +
#   geom_hline(yintercept = -log10(0.05), color = "red", linetype = "dashed") +
#   geom_vline(xintercept = -log10(0.05), color = "red", linetype = "dashed") +
#   geom_abline(slope = 1, intercept = 0, color = "purple", linetype = "dashed") +
#   theme(aspect.ratio = 1) +
#   labs(x = "-log10 (adjusted p.values) from unfiltered data", y = "-log10 (adjusted p.values) from data filtered by N6_0.99") +
#   theme_classic() +
#   scale_color_manual(
#     values = setNames(scales::hue_pal()(3), 
#                       c("N6_0.99_sig", "Unfiltered_sig", "Unfiltered_sig/N6_0.99_sig")))
```

