---
title: "LF_start_mat"
author: "LF"
date: "2024-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(ggplot2)
library(scales)
library(ComplexHeatmap)
library(limma)
```

# Read Data

## wu_subset rda

```{r read data}
getwd()

load("wu_subset.rda")

# metadata
wu_subset$SAMPLENAME
wu_subset$monkey
wu_subset$timepoint
wu_subset$months
wu_subset$celltype

# count data
wu_count <- assay(wu_subset, "counts") 

dim(wu_subset)

```

215 barcodes x 20 samples.

## txt data ZJ31

```{r read ZJ31 counts}
getwd()
## data is separated by "\t", so use read.table()
ZJ31 <- read.table("monkey_ZJ31.txt")
## check data dimension
print("Number of rows and columns: ")
dim(ZJ31)
## observing around 60k barcodes in rows, check their names
barcodes <- rownames(ZJ31) 
## barcodes names by sequence - each barcode is unique
unique(barcodes) %>% length()
## check if barcodes sequence only contains ATGC letters
grepl("^[ATGC]+$", barcodes[1])
is.seq <- lapply(barcodes, function(x) grepl("^[ATGC]+$", x)) %>% unlist()
table(is.seq)
## observing only 2 barcodes not following this pattern, print it out
barcodes[!is.seq]
grepl("^[ATGC]+$", "GTAGCCCATCTTACGCCGGATGACGTCATAAGTGTATCGANGATCGGAAG")
grepl("^[ATGC]+$", "GTAGCCTGGGTTTTGTCCGACATAGAGTCTGTACTTTCGCNAGATCGGAA")
## "N" is included in these barcodes' sequence - seem to be unknown nucleotide

## check numeric of dataset
is.numeric(as.matrix(ZJ31))
## plot density of dataset
plot(density(as.matrix(ZJ31)))
plot(density(as.matrix(log2(ZJ31 + 1))))
## plot the library size
lib_size <- colSums(ZJ31)
hist(lib_size)
hist(log10(lib_size +1))
## most samples' library size are at million's level
any(lib_size < 1e5)
## All samples' library size are greater than 0.1M.
```

??? Should I exclude those barcodes with unknown nucleotides in the subsequent analysis???

```{r read ZJ31 target}
## read target file of ZJ31 dataset
ZJ31Target <- read.table("monkey_ZJ31_metadata.txt", header = 1, row.names = 1)
## checkout target factors
colnames(ZJ31Target)
lapply(ZJ31Target, class)
```

## txt data ZH33

```{r read ZH33 counts}
getwd()
## data is separated by "\t", so use read.table()
ZH33 <- read.table("monkey_ZH33.txt")
## check data dimension
print("Number of rows and columns: ")
dim(ZH33)
## observing around 26k barcodes in rows, check their names
barcodes <- rownames(ZH33) 
## barcodes names by sequence - each barcode is unique
unique(barcodes) %>% length()
## check if barcodes sequence only contains ATGC letters
grepl("^[ATGC]+$", barcodes[1])
is.seq <- lapply(barcodes, function(x) grepl("^[ATGC]+$", x)) %>% unlist()
table(is.seq)
## all barcodes' sequence are consist of only  A T G C 
barcodes[!is.seq]

## check numeric of dataset
is.numeric(as.matrix(ZH33))
## plot density of dataset
plot(density(as.matrix(ZH33)))
plot(density(as.matrix(log2(ZH33 + 1))))
## plot the library size
lib_size <- colSums(ZH33)
hist(lib_size)
hist(log10(lib_size +1))
## most samples' library size are at million's level
any(lib_size < 1e5)
## All samples' library size are greater than 0.1M.
```

```{r read ZH33 target}
## read target file of ZH33 dataset
ZH33Target <- read.table("monkey_ZH33_metadata.txt", header = 1, row.names = 1)
## checkout target factors
colnames(ZH33Target)
lapply(ZH33Target, class)
```

## txt data ZG66

```{r read ZG66 counts}
getwd()
## data is separated by "\t", so use read.table()
ZG66 <- read.table("monkey_ZG66.txt")
## check data dimension
print("Number of rows and columns: ")
dim(ZG66)
## observing around 60k barcodes in rows, check their names
barcodes <- rownames(ZG66) 
## barcodes names by sequence - each barcode is unique
unique(barcodes) %>% length()
## check if barcodes sequence only contains ATGC letters
grepl("^[ATGC]+$", barcodes[1])
is.seq <- lapply(barcodes, function(x) grepl("^[ATGC]+$", x)) %>% unlist()
table(is.seq)
## observing only 6 barcodes not following this pattern, print it out
barcodes[!is.seq]
grepl("^[ATGC]+$", "GTAGCCACGTTTACGTATCGTTCTAATAGCTGGTGCCTTGGAGATCGGAN")
grepl("^[ATGC]+$", "GTAGCCGTCGGCACTTTGTTTATAAAGTCTTGCGTCTTCGTAGATCGGAN")
grepl("^[ATGC]+$", "GTAGCCTCGCGCCGTTAGCTTATGGGATAAACTCAATATGAAGATCGGAN")
grepl("^[ATGC]+$", "GTAGCCTGCGAGTGGAGCCCGTTTTAGGGATAGTGACGTGTAGATCGGAN")
grepl("^[ATGC]+$", "GTAGCCTTTATTGTCTCGATCTAAGCAGCTCAGTAGCAAAGAGATCGGAN")
grepl("^[ATGC]+$", "GTAGCCTTTGGCTTGCCGGGGGTTGGGGTCTTGTTCTGTCGAGATCGGAN")
## "N" is included in these barcodes' sequence - seem to be unknown nucleotide

## check numeric of dataset
is.numeric(as.matrix(ZG66))
## plot density of dataset
plot(density(as.matrix(ZG66)))
plot(density(as.matrix(log2(ZG66 + 1))))
## plot the library size
lib_size <- colSums(ZG66)
hist(lib_size)
hist(log10(lib_size +1))
## most samples' library size are at million's level
any(lib_size < 1e5)
lib_size[lib_size < 1e5]
## only one sample's library size is less than 0.1M, around 59k is fine.
```

??? Should I exclude those barcodes with unknown nucleotides in the subsequent analysis???

```{r read ZG66 target}
## read target file of ZG66 dataset
ZG66Target <- read.table("monkey_ZG66_metadata.txt", header = 1, row.names = 1)
## checkout target factors
colnames(ZG66Target)
lapply(ZG66Target, class)
```

## checkout sample targets: factors

The three monkeys can be considered as biological replicates, as they share identical sample factors. But I need to check specific conditions within each factor.

```{r clean target}
## tidy up colnames of the three targets
colnames(ZJ31Target)
colnames(ZH33Target)
colnames(ZG66Target)

## update the ZH33 factor names following camelCase
colnames(ZH33Target)[4] <- "Celltype"
colnames(ZH33Target)
## update ZJ31 and ZG66's factor names by ZH33's
colnames(ZJ31Target) <- colnames(ZH33Target)
colnames(ZG66Target) <- colnames(ZH33Target)

## checkout
colnames(ZJ31Target)
colnames(ZH33Target)
colnames(ZG66Target)
lapply(ZJ31Target, class) %>% unlist()
lapply(ZH33Target, class) %>% unlist()
lapply(ZG66Target, class) %>% unlist()
```

## checkout condition settings

There are two main factors in the sample targets: Months and Celltype.

 - Months: 
 Most of the time, blood samples from the monkeys were collected separately at independent time points. 
There are only 5 time points (months) when samples are collected from all individual monkeys - and we don't know whether these samples were collected at exactly the same time or if this alignment is due to the recording resolution (accuracy), as 0.5 months appears to be the minimum interval. But this likely doesn't matter, as I believe the researchers must have chosen a reasonable time measurement to reveal the real time-related changes in monkeys. So, I assume the timepoint alignment represents actual controlled time points.

 - Celltype: 
 Only T, B Gr celltypes were collected from all money individuals.
 
 - Combo of Months + Celltype:
 
 Only 9 homogeneous settings across three monkeys;
 
 13 homogeneous settign between two monkeys;
 
 128 samples of unique settings.
 
 No technique replicates. Each condition (Monkey_Celltype_Timepoint) has only one sample.
 

```{r check conditions months}
Targets <- rbind(ZJ31Target, ZH33Target, ZG66Target)
head(Targets)

ggplot(data = Targets) +
  geom_histogram(aes(x = Months, color = Monkey), fill = NA, binwidth = 0.5) +
  scale_x_continuous(breaks =seq(0, 70, by = 5)) +
  theme_minimal()
```

```{r celltype alignment}
ggplot(data = Targets) +
  geom_bar(aes(x = as.factor(Celltype), color = Monkey), fill = NA) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r condition combo, fig.height=5, fig.width=5}

Targets2 <- Targets %>%
  group_by(Celltype, Months) %>%
  mutate(count = n()) %>% 
  ungroup()
ggplot(data = Targets2) +
  geom_bar(aes(x = as.factor(paste0(Celltype, Months)), color = Monkey), fill = NA) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1, size = 3)) +
  geom_text(
    aes(x = as.factor(paste0(Celltype, Months)), 
        label = ifelse(count == 3, Celltype, "")),
    stat = "count",
    position = position_jitter(width = 0.1, height = 0.2), 
    y = 3.1, angle = 45, size = 2) +
  geom_text(
    aes(x = as.factor(paste0(Celltype, Months)), 
        label = ifelse(count == 3, Timepoint, "")),
    stat = "count",
    position = position_jitter(width = 0.1, height = 0.2), 
    y = 2.9, angle = 90, size = 2) +
  scale_x_discrete(expand = expansion(add = c(5, 5)))

## checkout homogeneous settings
table(Targets2$count)

## checkout sample replicates
table(paste0(Targets$Monkey, Targets$Celltype, Targets$Months)) %>% table()
```

## checkout barcodes

Observing three independent barcode libraries, each injected into three monkey individuals corresponding to three datasets.

```{r barcode overlaps}
## check if the barcodes injected into three monkeys overlap
## Euler/Venn diagram
fit <- euler(
  list(
    ZJ31 = rownames(ZJ31Target), 
    ZH33 = rownames(ZH33Target), 
    ZG66 = rownames(ZG66Target)))
plot(fit, fills = c("red", "blue", "green"), alpha = 0.5, labels = TRUE)
```

# Load barbieQ

```{r load barbieQ}
install.packages("/Users/feiliyang/Desktop/barbieQ/", repos = NULL, type = "source")
library(barbieQ)
```

```{r merge datasets}
## combine three targets
Targets <- rbind(ZJ31Target, ZH33Target, ZG66Target)
## combine three count matrix
# Counts <- merge(ZJ31, ZH33, by = "row.names", all = TRUE)
# rownames(Counts) <- Counts$Row.names
# Counts <- Counts[,-1]
# Counts <- merge(Counts, ZG66, by = "row.names", all = TRUE)
# rownames(Counts) <- Counts$Row.names
# Counts <- Counts[,-1]
# 
# ## this doesn;t look right
# sum(nrow(ZJ31), nrow(ZH33), nrow(ZG66))
# nrow(Counts)
# identical(colnames(Counts), rownames(Targets))
```

# Creare barbieQ Object

```{r read by barbieQ}
ZJ31_bq <- barbieQ::createBarbieQ(object = ZJ31, target = ZJ31Target)
ZH33_bq <- barbieQ::createBarbieQ(object = ZH33, target = ZH33Target)
ZG66_bq <- barbieQ::createBarbieQ(object = ZG66, target = ZG66Target)

save(ZJ31_bq, ZH33_bq, ZG66_bq, file = "ZJ31_ZH33_ZG66_bq.rda")
```

```{r distribution}
long_data <- as.data.frame(ZG66_bq$assay) %>%
  rownames_to_column(var = "RowName") %>%   # Add row names as a column named "RowName"
  pivot_longer(
    cols = -RowName,
    names_to = "Sample",                # Create a "ColumnName" column
    values_to = "Value"                     # Create a "Value" column
  )

ggplot(data = long_data) +
  geom_histogram(aes(x = log2(Value +1), color = Sample), fill = NA, binwidth = 1) +
  theme(legend.position = "none")
```


# Filter

## Round 1

```{r tag}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ZJ31_bq <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq)
# ZH33_bq <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq)
# ZG66_bq <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq)
# 
# save(ZJ31_bq, ZH33_bq, ZG66_bq, file = "tagged_bq.rda")

load("tagged_bq.rda")
```

```{r vis tags}
## visualize tags
barbieQ::plotBarcodeSankey(ZJ31_bq)
barbieQ::plotBarcodeSankey(ZH33_bq)
barbieQ::plotBarcodeSankey(ZG66_bq)
```

0.99 cut-off didn’t filter out most of the barcodes.

```{r prop vis}
barbieQ::plotBarcodeProportion(barbieQ = ZJ31_bq)
barbieQ::plotBarcodeProportion(barbieQ = ZH33_bq)
barbieQ::plotBarcodeProportion(barbieQ = ZG66_bq)
```

## PCA

```{r PCA}
pcZJ31 <- prcomp(t(ZJ31_bq$proportion))
PCscore <- pcZJ31$x

ggplot(PCscore) +
  geom_point(aes(x = PC1, y = PC2, color = ZJ31_bq$metadata$Celltype))

```


## Round 2 - cut off

```{r tag 2 0.9}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ZJ31_bq_0.9 <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq, proportionThreshold = 0.9)
# ZH33_bq_0.9 <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq, proportionThreshold = 0.9)
# ZG66_bq_0.9 <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq, proportionThreshold = 0.9)
# 
# save(ZJ31_bq_0.9, ZH33_bq_0.9, ZG66_bq_0.9, file = "tagged_bq_0.9.rda")

load("tagged_bq_0.9.rda")
```

```{r tag 2 0.8}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ZJ31_bq_0.8 <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq, proportionThreshold = 0.8)
# ZH33_bq_0.8 <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq, proportionThreshold = 0.8)
# ZG66_bq_0.8 <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq, proportionThreshold = 0.8)
# 
# save(ZJ31_bq_0.8, ZH33_bq_0.8, ZG66_bq_0.8, file = "tagged_bq_0.8.rda")

load("tagged_bq_0.8.rda")
```

```{r tag 2 0.7}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ZJ31_bq_0.7 <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq, proportionThreshold = 0.7)
# ZH33_bq_0.7 <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq, proportionThreshold = 0.7)
# ZG66_bq_0.7 <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq, proportionThreshold = 0.7)
# 
# save(ZJ31_bq_0.7, ZH33_bq_0.7, ZG66_bq_0.7, file = "tagged_bq_0.7.rda")

load("tagged_bq_0.7.rda")
```

```{r vis tags 2}
## visualize tags
barbieQ::plotBarcodeSankey(ZJ31_bq_0.9)
barbieQ::plotBarcodeSankey(ZH33_bq_0.9)
barbieQ::plotBarcodeSankey(ZG66_bq_0.9)
```

```{r vis tags 2 0.8}
## visualize tags
barbieQ::plotBarcodeSankey(ZJ31_bq_0.8)
barbieQ::plotBarcodeSankey(ZH33_bq_0.8)
barbieQ::plotBarcodeSankey(ZG66_bq_0.8)
```

```{r vis tags 2 0.7}
## visualize tags
barbieQ::plotBarcodeSankey(ZJ31_bq_0.7)
barbieQ::plotBarcodeSankey(ZH33_bq_0.7)
barbieQ::plotBarcodeSankey(ZG66_bq_0.7)
```

```{r prop vis 2}
barbieQ::plotBarcodeProportion(barbieQ = ZJ31_bq_0.9)
barbieQ::plotBarcodeProportion(barbieQ = ZH33_bq_0.9)
barbieQ::plotBarcodeProportion(barbieQ = ZG66_bq_0.9)

barbieQ::plotBarcodePareto(barbieQ = ZG66_bq_0.9)
```

## Round 3 - exclude starting point

Any starting point sample included? 

 - Month 1.0 seems to be the starting point and may contain too many equally contributing barcodes. So, exclude these samples when tagging top/bottom.

```{r Months}
ggplot(data = Targets) +
  geom_histogram(aes(x = Months), binwidth = 0.5) +
  scale_x_continuous(trans = log_trans(base = 2))
min(Targets$Months)
head(sort(Targets$Months), 10)
sort(Targets$Months)
```

```{r exclude month1}
ZJ31_bq_ex <- barbieQ::subsetSamplesByMetadata(
  barbieQ = ZJ31_bq, factor = "Months", specifiedConditions = 1, keep = FALSE)

ZH33_bq_ex <- barbieQ::subsetSamplesByMetadata(
  barbieQ = ZH33_bq, factor = "Months", specifiedConditions = 1, keep = FALSE)

ZG66_bq_ex <- barbieQ::subsetSamplesByMetadata(
  barbieQ = ZG66_bq, factor = "Months", specifiedConditions = 1, keep = FALSE)
```

```{r tag 3}
## This step is very time-consuming, don't run again
## Just use the temporary results saved from the first run
# ZJ31_bq_ex <- barbieQ::tagTopBarcodes(barbieQ = ZJ31_bq_ex)
# ZH33_bq_ex <- barbieQ::tagTopBarcodes(barbieQ = ZH33_bq_ex)
# ZG66_bq_ex <- barbieQ::tagTopBarcodes(barbieQ = ZG66_bq_ex)
# 
# save(ZJ31_bq_ex, ZH33_bq_ex, ZG66_bq_ex, file = "tagged_bq_ex.rda")

load("tagged_bq_ex.rda")
```

```{r vis tags 3}
## visualize tags
barbieQ::plotBarcodeSankey(ZJ31_bq_ex)
barbieQ::plotBarcodeSankey(ZH33_bq_ex)
barbieQ::plotBarcodeSankey(ZG66_bq_ex)
```

```{r prop vis 2}
barbieQ::plotBarcodeProportion(barbieQ = ZJ31_bq_ex)
barbieQ::plotBarcodeProportion(barbieQ = ZH33_bq_ex)
barbieQ::plotBarcodeProportion(barbieQ = ZG66_bq_ex)
```

```{r ZG66 exclude early timepoints}
ZG66_bq$metadata$Months %>% table()
## exclude months less than 6 when considering top barcodes
pointer <- ZG66_bq$metadata$Months < 6
table(pointer)
subset_top_table <- ZG66_bq$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 1
table(new_tops)

subset_top_table <- ZG66_bq_0.9$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 1
table(new_tops)

subset_top_table <- ZG66_bq_0.8$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 1
table(new_tops)

subset_top_table <- ZG66_bq_0.7$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 1
table(new_tops)



subset_top_table <- ZG66_bq$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 2
table(new_tops)

subset_top_table <- ZG66_bq_0.9$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 2
table(new_tops)

subset_top_table <- ZG66_bq_0.8$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 2
table(new_tops)

subset_top_table <- ZG66_bq_0.7$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 2
table(new_tops)
```

# Subset Barcodes

Filtering parameters:

 - Being top 90% in at least 2 samples.

```{r subset}
subset_top_table <- ZG66_bq_0.9$isTop$mat[,!pointer]
new_tops <- rowSums(subset_top_table) >= 2
table(new_tops)
ZG66_bq_0.9$isTop$vec <- new_tops

ZG66_bq_0.9 <- barbieQ::subsetBarcodes(barbieQ = ZG66_bq_0.9, retainedRows = new_tops)
```

# Heatmap

```{r heatmap}
set.seed(99)
barbieQ::plotBarcodeHeatmap(ZG66_bq_0.9)
```


# Barcode Correlation

Too many barcodes! Will take forever to run the barcode pair wise correlation, so send to background job.

```{r barcode cor}
p_cor_ZG <- barbieQ::plotBarcodePairCorrelation(barbieQ = ZG66_bq_0.9)
ggplot2::ggsave("p_cor_ZG.pdf", plot = p_cor_ZG, width = 8, height = 6)
```


# Sample Correlation

I'll continue with `ZG66_bq_0.9`, as it contains only 6k barcodes after filtering.

Samples of Month 1 is so different.

```{r sample cor, ZG66, fig.width=8, fig.height=4}
hp_targets <- ZG66_bq_0.9$metadata %>% 
  select(Months, Celltype, Monkey) %>% 
  mutate(Months = log2(Months))

set.seed(88)
barbieQ::plotSamplePairCorrelation(
  barbieQ = ZG66_bq_0.9, targets = hp_targets)

hp_targets <- ZG66_bq_0.9$metadata %>% 
  select(Months, Celltype, Monkey) %>% 
  mutate(Months = as.factor(Months))

barbieQ::plotSamplePairCorrelation(
  barbieQ = ZG66_bq_0.9, targets = hp_targets[,c(2,1,3)])

table(ZG66_bq_0.9$metadata$Months)

## check sample repilcates
table(paste0(hp_targets$Celltype, hp_targets$Months)) %>% summary()

```

# Sample PCA

NK cells are so different from other cell types and so different between NK subsets.

```{r PCA ZG66}
pcZG66 <- prcomp(t(ZG66_bq_0.9$proportion))
PCscore <- pcZG66$x

ggplot(PCscore) +
  geom_point(aes(x = PC1, y = PC2, color = ZG66_bq_0.9$metadata$Celltype))
ggplot(PCscore) +
  geom_point(aes(x = PC1, y = PC2, color = ZG66_bq_0.9$metadata$Months))
ggplot(PCscore) +
  geom_point(aes(x = PC1, y = PC3, color = ZG66_bq_0.9$metadata$Celltype))
ggplot(PCscore) +
  geom_point(aes(x = PC1, y = PC3, color = ZG66_bq_0.9$metadata$Months))
ggplot(PCscore) +
  geom_point(aes(x = PC4, y = PC3, color = ZG66_bq_0.9$metadata$Celltype))
ggplot(PCscore) +
  geom_point(aes(x = PC4, y = PC3, color = ZG66_bq_0.9$metadata$Months))

```

# Model Design Matrix

```{r model desing matrix}
## default Gr_1 and Gr_2 to Gr
## remove NKG2A_blah NK subtype, whice were collected in late timepoints

mytargets <-ZG66_bq_0.9$metadata %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))

myObject <- ZG66_bq_0.9
myObject$metadata <- mytargets
myObject <- barbieQ::subsetSamplesByMetadata(barbieQ = myObject, factor = "Phase", specifiedConditions = "mid")

table(myObject$metadata$Celltype)

```

```{r fit model}
## define matrix
## will consider Months factor later as each cell type may have different trend
mydesign <- model.matrix(~ 0 + Celltype, data = myObject$metadata)
View(mydesign)

## define contrast
mycontrast <- limma::makeContrasts(
  (CelltypeNK_CD56n_CD16p+CelltypeNK_CD56p_CD16n)/2 - (CelltypeB+CelltypeGr+CelltypeT)/3,
  levels = colnames(mydesign))

myObject$factorColors$diffProp_Celltype <- NULL
myObject <- barbieQ::testBarcodeBias(
  barbieQ = myObject,
  targets = myObject$metadata,
  designMatrix = mydesign,
  contrast = mycontrast
  )

myObject$testBarcodes$diffProp_Celltype$methods$groups
```

```{r vis test hp}
barbieQ::plotBarcodeBiasHeatmap(barbieQ = myObject)

```

```{r vis test dot}

barbieQ::plotBarcodeBiasScatterPlot(
  barbieQ = myObject, reorderRank = TRUE, pValuesAdjusted = T, xAxis = "avgLogCPM")
```


## NK-secific Clone

```{r NK }
point_NK <- myObject$testBarcodes$diffProp_Celltype$results$direction == "group2"
myNK <- barbieQ::subsetBarcodes(barbieQ = myObject, retainedRows = point_NK)
myNK$testBarcodes <- myObject$testBarcodes
myNK$testBarcodes$diffProp_Celltype$results <- myNK$testBarcodes$diffProp_Celltype$results[point_NK,]
barbieQ::plotBarcodeBiasHeatmap(barbieQ = myNK)

NK_clone_contribution <- myNK$proportion %>% colSums()
dat <- data.frame(
  contr = NK_clone_contribution,
  sample = myNK$metadata$Celltype)

dat <- dat %>%
  mutate(id = row_number()) 
ggplot(dat, aes(x = as.factor(id), y = contr, color = sample)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal()
```

