---
title: "case_study"
author: "LF"
date: "2025-06-17"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(circlize)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_barbieQ.rda")
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

# Compare filtering

```{r filtering original}
tagTop10 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 10
  # summary(isTop)
  return(isTop)
}

colTags <- lapply(as.data.frame(assays(monkeyHSPC_agg)$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC_agg)
dim(colTagMat)
colSums(colTagMat)

top10each <- monkeyHSPC_agg
rowData(top10each)$isTopBarcode$isTop <- colTagVec
# assays(top10each)$isTopAssay <- colTagMat
barbieQ::plotBarcodeSankey(top10each)
ggsave("../../output/WuC/sankey_top10each.png", dpi = 300, width = 4, height = 4, 
       barbieQ::plotBarcodeSankey(top10each))

tagTop20 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 20
  # summary(isTop)
  return(isTop)
}
colTags2 <- lapply(as.data.frame(assays(monkeyHSPC_agg)$proportion), 
                  function(col) tagTop20(col))
colTagMat2 <- do.call(cbind, colTags2)
colTagVec2 <- rowSums(colTagMat2) >= 1

top20each <- monkeyHSPC_agg
rowData(top20each)$isTopBarcode$isTop <- colTagVec2
# assays(top10each)$isTopAssay <- colTagMat2
barbieQ::plotBarcodeSankey(top20each)
ggsave("../../output/WuC/sankey_top20each.png", dpi = 300, width = 4, height = 4, 
       barbieQ::plotBarcodeSankey(top20each))

venn1 <- eulerr::euler(
  list(
    barbieQ = rownames(monkeyHSPC_top), 
    top10_per_sample = rownames(monkeyHSPC_agg)[colTagVec],
    top20_per_sample = rownames(monkeyHSPC_agg)[colTagVec2])
  )

plot(venn1, fills = c("white", "grey", "black"), alpha = 0.5, labels = TRUE, quantities = TRUE)
ggsave("../../output/WuC/venn_filtering.png", dpi = 300, width = 6, height = 4, 
       plot(venn1, fills = c("white", "grey", "black"), alpha = 0.5, labels = TRUE, quantities = TRUE))
```

# Clone Paper Selected (top10)

Selected barcodes downloaded from the table in the supplementary materials of the original paper differ from the set of barcodes selected using "top10each" from the downloaded whole ZG66 dataset. I think it's because the downloaded table doesn't include the later time points. But I cannot find the corresponding table from the paper's materials. For the sake of consistency, I'll use the set of barcodes selected by "top10each" from the downloaded dataset that processed by myself.

```{r overlap}
venn2 <- eulerr::euler(
  list(
    barbieQ = rownames(monkeyHSPC_top), 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn2, fills = c("white", "grey"), alpha = 0.5, labels = TRUE, quantities = TRUE)
```

```{r paper barcode prop provided }
## calculate the total proportion of the paper selected barcodes based on the proportion data provided by the paper
NK_clone_contribution <- ZG66_Fig2 %>% colSums()
dat <- data.frame(
  contr = NK_clone_contribution,
  sample = c(rep("T", 6),
             rep("B", 6),
             rep("Gr", 6),
             rep("NK_CD56p_CD16n", 6),
             rep("NK_CD56n_CD16p", 6)))
dat$sample <- factor(dat$sample, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p"))

dat <- dat %>%
  mutate(id = row_number()) 
p11 <- ggplot(dat, aes(x = as.factor(id), y = contr, color = sample)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  ggtitle("Paper - identified clones") +
  labs(y = "Prop.", x = "Sample (cell type)")
p11
```


# Reanalysis - filtered data

- My selecting strategy to select `NK_CD56n_CD16p biased barcodes` by testing

 - [1] "(CelltypeNK_CD56n_CD16p)" v.s.                                      

 - [2] "(CelltypeB + CelltypeGr + CelltypeT + CelltypeNK_CD56p_CD16n)/4"

- Only keep the samples of timepoints at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m. (to be consistent with the paper's analysis)

## model: monkeyHSPC_top (at time points when all cell types colleted)

```{r checkcout df after contrast}
## For time points at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m, each collecting all cell types.
## These samples refer to the first 30 samples in the object, which is also in line with the samples in the table downloaded from the paper's materials.
whenAllCell <- monkeyHSPC_top[,1:30]
dim(whenAllCell)
whenAllCell$sampleMetadata %>% as.data.frame() %>% View()

##need to take Time into account
designs <- whenAllCell$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

asinTrans <- testBarcodeSignif(
  barbieQ = whenAllCell, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
# check design
asinTrans@elementMetadata$testingBarcode@metadata$design
asinTrans@elementMetadata$testingBarcode@metadata$contrasts



## test Barcode differential occurrence between sample groups
## set up the targets (sample conditions)
targets <- monkeyHSPC_top$sampleMetadata %>% 
  as.data.frame() %>% 
  mutate(Group = ifelse(
    Celltype == "NK_CD56n_CD16p",
    "NK_CD56n_CD16p",
    "B.Gr.T.NK_CD56p_CD16n"))

occTest <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top, 
  sampleMetadata = targets[,"Group", drop=FALSE], 
  method = "diffOcc"
)
```

```{r, fig.width=14, fig.height=5}
plotSignifBarcodeHeatmap(asinTrans)
```

```{r}
plotBarcodeMA(asinTrans)
```


## model paper table

```{r}
paper_bb <- monkeyHSPC[rownames(monkeyHSPC) %in% rownames(ZG66_Fig2), 1:30]
```

```{r}
# paper_bb$sampleMetadata$Months <- as.factor(paper_bb$sampleMetadata$Months)

##need to take Time into account
designs <- paper_bb$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

asinTrans_paper <- testBarcodeSignif(
  barbieQ = paper_bb, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

# check design
asinTrans@elementMetadata$testingBarcode@metadata$design %>% View()
asinTrans@elementMetadata$testingBarcode@metadata$contrasts
```

## model paper table all time

```{r}
paper_bb_allTime <- monkeyHSPC[rownames(monkeyHSPC) %in% rownames(ZG66_Fig2)]

##need to take Time into account
designs <- paper_bb_allTime$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)


asinTrans_paper_allTime <- testBarcodeSignif(
  barbieQ = paper_bb_allTime, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
```


## validation function

```{r function validation}
plot_validation <- function(myObject) {
  
  flag <- rowData(myObject)$testingBarcode$direction == 1
  myObject <- myObject[flag,]
  
  NK_clone_contributionU <- assays(myObject)$proportion %>% colSums()
  dat <- data.frame(
    contr = NK_clone_contributionU,
    Celltype = factor(myObject$sampleMetadata$Celltype,
                      levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p")),
    Months = myObject$sampleMetadata$Months)
  ## sort samples by celltype and months
  dat <- dat %>%
    arrange(Celltype, Months) %>% 
    mutate(id = row_number()) 
  ## add dummy rows for spacing between celltypes
  dat <- dat %>%
    group_by(Celltype) %>%
    mutate(dummy = n() + 1) %>%    # Add a dummy count for spacing
    ungroup() %>%
    bind_rows(dat %>%             # Add dummy rows for spacing
                group_by(Celltype) %>%
                slice_tail(n = 1) %>%
                mutate(id = as.numeric(id) + 0.5, contr = NA, Months = NA)) %>%
    arrange(as.numeric(id))        # Arrange with dummy rows in place
  
  celltype_colors <- c(
  "T" = "yellowgreen",
  "B" = "yellowgreen",
  "Gr" = "yellowgreen",
  "NK_CD56p_CD16n" = "yellowgreen",
  "NK_CD56n_CD16p" = "slateblue1" # "orchid"
)
  
  p <- ggplot(dat, aes(x = as.factor(id), y = contr, color = Celltype)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 0.5, size = 10),
        axis.title.x = element_text(size = 14), 
        axis.title.y = element_text(size = 14), 
        panel.grid.major.x = element_blank()) +
  scale_x_discrete(labels = ifelse(is.na(dat$Months), "", dat$Months)) + # Relabel with `Celltype`
  scale_color_manual(values = celltype_colors) +  # <- set bar outline color
  scale_fill_manual(values = celltype_colors) +   # <- set tile fill color
  scale_y_continuous(limits = c(-0.09, 0.8)) +
  labs(x = "Samples of cell types at different months", y = "Total prop. of selected barcodes") +  # Update axis labels
  geom_tile(
    data = dat,
    aes(x = ifelse(is.na(Months), NA,order(id)), y = -0.03, fill = Celltype),
    height = 0.03, width = 1
  ) +
  geom_text(
    data = dat %>%
      mutate(id = order(id)) %>%
      group_by(Celltype) %>%
      mutate(Celltype = recode(Celltype, "NK_CD56n_CD16p" = "NK.CD56-CD16+")) %>% 
      mutate(Celltype = recode(Celltype, "NK_CD56p_CD16n" = "NK.CD56+CD16-")) %>% 
      summarize(mid_point = mean(as.numeric(id))),
    aes(x = mid_point-0.5, y = -0.03, label = Celltype),
    size = 3, vjust = 0.5, color = "black"
  )
  
  return(p)
}
```

### barbieQ test on monkeyHSPC_top and on paper table

```{r NK my select prop, fig.width=6, fig.height=5}
plot_validation(asinTrans) 
plot_validation(asinTrans_paper)
plot_validation(asinTrans_paper_allTime)

# ggsave("../../output/WuC/validation_barbieQ.png", 
#        dpi = 300, width = 6, height = 5, plot_validation(asinTrans))


plotSignifBarcodeProportion(asinTrans)
```

### Bias approaches for paper table

(i) >1% contribution to CD56−CD16+ cells and 

(ii) >10-fold abundance in contribution to CD56−CD16+ cells compared with T, B, Gr, and CD56+CD16− cells.

```{r paper nk, fig.width=6, fig.height=5}
colnames(ZG66_Fig2)
colnames(monkeyHSPC_top)
ZG66_Fig2_bb <- createBarbieQ(ZG66_Fig2)
colData(ZG66_Fig2_bb)$sampleMetadata <- monkeyHSPC_top$sampleMetadata[1:30,]
assays(ZG66_Fig2_bb)$proportion <- ZG66_Fig2
rowData(ZG66_Fig2_bb)$testingBarcode$direction = 1
plot_validation(ZG66_Fig2_bb)

# ggsave("../../output/WuC/validation_original.png", 
#        dpi = 300, width = 6, height = 5, plot_validation(ZG66_Fig2_bb))

```

#### Calc LFC

```{r}
## extract props. and sample conditions
prop. <- paper_data[,-1] 
cond. <- monkeyHSPC$sampleMetadata[1:30,] %>% as.data.frame()
## group samples by NKnp or others
cond. <- cond. %>% 
  mutate(isNK = ifelse(Celltype == "NK_CD56n_CD16p", "NKnp", "others"))
cond.$Celltype %>% table()

## calc mean per group at each month
group_labels <- paste(cond.$Months, cond.$isNK, sep = "__")
group_means <- sapply(unique(group_labels), function(g) {
  idx <- which(group_labels == g)
  rowMeans(prop.[, idx, drop = FALSE])
})

## calc group LFC at each month
colnames(group_means)
months <- gsub("(\\1)__.*", "\\1",colnames(group_means))

LFC_months <- sapply(unique(months), function(m) {
  idx <- which(months == m)
  sub <- group_means[, idx, drop = FALSE]
  group <- gsub(".*__()", "\\1",colnames(sub))
  NK <- sub[, which(group == "NKnp"), drop = F]
  others <- sub[, which(group=="others"),drop = F]
  NK / others 
})

## decide if each barcode LFC is high at each month
is_LFC_high <- LFC_months > 10
colnames(is_LFC_high)

## calc if each barcode NKnp  prop > 0.01 at each month
NK_prop <- prop.[,cond.$isNK == "NKnp", drop=F]
is_NKnp_high <- NK_prop > 0.01
colnames(is_NKnp_high)

## barcodes meet both metric at each month
select_months <- (is_LFC_high & is_NKnp_high)

## select barcodes at each month
select_barcodes_months <- lapply(colnames(select_months), function(j) 
  rownames(prop.)[select_months[,j]])

names(select_barcodes_months) <- paste0("heuristic at ", colnames(is_LFC_high), "m")

```

#### Venn: LFC at each month vs barbieQ test

```{r}

## check overlaps between months
venn0 <- eulerr::euler(select_barcodes_months)
plot(venn0, fills = c("white", "grey"), alpha = 0.5, labels = T, quantities = T)

## check overlaps
compare_list <- select_barcodes_months
compare_list[["barbieQ_select"]] <- rownames(asinTrans_paper)[rowData(asinTrans_paper)$testingBarcode$direction == 1]
compare_list[["barbieQ_select_allTime"]] <- rownames(asinTrans_paper_allTime)[rowData(asinTrans_paper_allTime)$testingBarcode$direction == 1]

venn00 <- eulerr::euler(compare_list)

plot(venn00, fills = c("white", "grey"), alpha = 0.5, labels = T, quantities = T)
```


#### Validation: LFC at each month

```{r}
ZG66_Fig2_bb <- createBarbieQ(ZG66_Fig2)
colData(ZG66_Fig2_bb)$sampleMetadata <- monkeyHSPC_top$sampleMetadata[1:30,]
assays(ZG66_Fig2_bb)$proportion <- ZG66_Fig2
rowData(ZG66_Fig2_bb)$testingBarcode$direction = 1

lapply(select_barcodes_months, function(m) plot_validation(ZG66_Fig2_bb[m,]))

```

#### Venn and validation: union high LFC barcodes at all months vs. barbieQ test

```{r}
union_all_months <- unlist(select_barcodes_months) %>% unique()

venn000 <- eulerr::euler(
  list(
    barbieQ_select = rownames(asinTrans_paper)[rowData(asinTrans_paper)$testingBarcode$direction == 1],
    barbieQ_select_allTime = rownames(asinTrans_paper_allTime)[rowData(asinTrans_paper_allTime)$testingBarcode$direction == 1],
    union_heuristic = union_all_months)
)

plot(venn000, fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)


plot_validation(ZG66_Fig2_bb[union_all_months,])

plot_validation(asinTrans_paper)
plot_validation(asinTrans_paper_allTime)
```

## Validation: change across time

```{r}
heristic_allTime <- monkeyHSPC[union_all_months,]
rowData(heristic_allTime)$testingBarcode$direction = 1
plot_validation(heristic_allTime)

```

```{r}
test_allTime <- monkeyHSPC[rownames(asinTrans_paper)[rowData(asinTrans_paper)$testingBarcode$direction == 1],]
rowData(test_allTime)$testingBarcode$direction = 1
plot_validation(test_allTime)
```


### Bias approaches for monkeyHSPC_top

```{r}
venn3 <- eulerr::euler(
  list(
    barbieQ_select = rownames(monkeyHSPC_top)[rowData(asinTrans)$testingBarcode$direction == 1], 
    Heuristic_LFC = union_all_months))

plot(venn3, fills = c("white", "grey"), alpha = 0.5, labels = T, quantities = T)
```


```{r prop}

ps.asinTrans <- asinTrans
SummarizedExperiment::rowData(ps.asinTrans)$testingBarcode$tendencyTo <- rowData(asinTrans)$testingBarcode$tendencyTo
ps.logitTrans <- noTrans
SummarizedExperiment::rowData(ps.logitTrans)$testingBarcode$tendencyTo <- rowData(logitTrans)$testingBarcode$tendencyTo

(plotBarcodeMA(ps.asinTrans) + coord_trans(x = "log10"))|> plot()
(plotBarcodeMA(ps.logitTrans) + coord_trans(x = "log10"))|> plot()
(plotBarcodeMA(noTrans) + coord_trans(x = "log10"))|> plot()

## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../barcoding/public_barcode_count/data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]
ps.paper <- monkeyHSPC[rownames(monkeyHSPC) %in% rownames(ZG66_Fig2) | rownames(monkeyHSPC) %in% rownames(noTrans),]


ps.paper <- testBarcodeSignif(
  barbieQ = paper_bb,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "none"
)
SummarizedExperiment::rowData(ps.paper)$testingBarcode$tendencyTo <- rep("n.s.", nrow(ps.paper))
SummarizedExperiment::rowData(ps.paper)$testingBarcode$tendencyTo[rownames(ps.paper) %in% union_all_months] <- "CelltypeNK_CD56n_CD16p"
(plotBarcodeMA(ps.paper) + coord_trans(x = "log10"))|> plot()

```


```{r occ, fig.width = 6, fig.height = 4}

ps.occ <- noTrans
SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo <- rowData(monkeyHSPC_top)$testingBarcode$tendencyTo
SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo[SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo == "GroupNK_CD56n_CD16p"] <- "CelltypeNK_CD56n_CD16p"
SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo[SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo == "GroupB.Gr.T.NK_CD56p_CD16n"] <- "CelltypeB + CelltypeGr + CelltypeNK_CD56p_CD16n + CelltypeT"
(plotBarcodeMA(ps.occ) + coord_trans(x = "log10"))|> plot()


plotBarcodeMA(ps.asinTrans) + 
  theme(legend.position = "none") + 
  labs(title = "barbieQ selected, red: NK-specific") |
plotBarcodeMA(ps.paper) + 
  theme(legend.position = "none") +
  labs(title = "paper selected, red: NK-specific")
```


# Reanalysis - time dynamics

## HP clustering - paper analysis

subset NKnp samples from 7 timepoints (all) and keep top10each

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

colTags_NKnp <- lapply(as.data.frame(assays(NKnp)$proportion), 
                  function(col) tagTop10(col))
colTagMat_NKnp <- do.call(cbind, colTags_NKnp)
colTagVec_NKnp <- rowSums(colTagMat_NKnp) >= 1
sum(colTagVec_NKnp) 
colSums(colTagMat_NKnp)

NKnp <- NKnp[colTagVec_NKnp,]
```

hierarchically clustering barcodes and extract the clusters

```{r}
NKnp_slot <- NKnp@assays@data$proportion
NKnp_hp <- Heatmap(log10(NKnp_slot+1e-7), show_row_names = F, 
        column_order = order(NKnp$sampleMetadata$Months),
        col = colorRamp2(
    c(0, -2, -4.602),                     # log10 values for 1, 0.01, 0.000025
    c("red", "white", "blue")             # choose any colors you prefer
  ))

NKnp_hp
```

```{r}
# 3. Extract row dendrogram and cut into 3 clusters
NKnp_row_dend <- row_dend(NKnp_hp)
NKnp_clusters <- cutree(as.hclust(NKnp_row_dend), k = 4)
```

```{r}

# 5. Create row annotation based on clusters
row_ha <- rowAnnotation(
  NKnp_clusters = as.factor(NKnp_clusters),
  col = list(NKnp_clusters = structure(
    RColorBrewer::brewer.pal(4, "Set1"),
    names = as.character(1:4)
  )),
  show_annotation_name = TRUE
)

# 6. Plot heatmap with annotation
ht_with_anno <- Heatmap(
  log10(NKnp_slot + 1e-7),
  show_row_names = FALSE, column_labels = NKnp$sampleMetadata$Months,
  column_order = order(NKnp$sampleMetadata$Months),
  col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  left_annotation = row_ha
)

ht_with_anno
```


## model all celltypes

### monkeyHSPC_top

```{r}

##need to take Time into account
designs <- monkeyHSPC_top$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


asinTrans_allCell_testTime <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top, designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

### NKnp top10each

```{r}
NKnp_top10each <- monkeyHSPC[rownames(NKnp),]
##need to take Time into account
designs <- NKnp_top10each$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


asinTrans_allCell_NKnp_top10each_testTime <- testBarcodeSignif(
  barbieQ = NKnp_top10each, designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## model NKnp - Need to fix pkg here!!!

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC_top$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

designs <- NKnp$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Months)
)

asinTrans_NKnp_testTime <- testBarcodeSignif(
  barbieQ = NKnp, ,sampleMetadata = NKnp$sampleMetadata, sampleGroup = "Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

##Validation: vis

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

cluster_selected_list <- lapply(1:4, function(i) names(NKnp_clusters)[NKnp_clusters == i])
names(cluster_selected_list) <- paste("cluster_", 1:4, sep = "_")

seletced_barbieQ_NKnp_up <- rowData(asinTrans_allCell_NKnp_top10each_testTime)$testingBarcode$direction == 1
seletced_barbieQ_NKnp_down <- rowData(asinTrans_allCell_NKnp_top10each_testTime)$testingBarcode$direction == -1

cluster_selected_list[["barbieQ_NKnp_up"]] <- rownames(asinTrans_allCell_NKnp_top10each_testTime)[seletced_barbieQ_NKnp_up]
cluster_selected_list[["barbieQ_NKnp_down"]] <- rownames(asinTrans_allCell_NKnp_top10each_testTime)[seletced_barbieQ_NKnp_down]

venn99 <- eulerr::euler(
  cluster_selected_list
)

plot(venn99, fills = c("white", "white","white","white","grey","grey"), alpha = 0.5, labels = T, quantities = T)


```


# Need to get data at early timepoint: 4.5m
