---
title: "case_study"
author: "LF"
date: "2025-04-22"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_barbieQ.rda")
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

# Compare filtering

```{r filtering original}
tagTop10 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 10
  # summary(isTop)
  return(isTop)
}

colTags <- lapply(as.data.frame(assays(monkeyHSPC_agg)$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC_agg)
dim(colTagMat)
colSums(colTagMat)

top10each <- monkeyHSPC_agg
rowData(top10each)$isTopBarcode$isTop <- colTagVec
# assays(top10each)$isTopAssay <- colTagMat
barbieQ::plotBarcodeSankey(top10each)
ggsave("../../output/WuC/sankey_top10each.png", dpi = 300, width = 4, height = 4, 
       barbieQ::plotBarcodeSankey(top10each))

tagTop20 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 20
  # summary(isTop)
  return(isTop)
}
colTags2 <- lapply(as.data.frame(assays(monkeyHSPC_agg)$proportion), 
                  function(col) tagTop20(col))
colTagMat2 <- do.call(cbind, colTags2)
colTagVec2 <- rowSums(colTagMat2) >= 1

top20each <- monkeyHSPC_agg
rowData(top20each)$isTopBarcode$isTop <- colTagVec2
# assays(top10each)$isTopAssay <- colTagMat2
barbieQ::plotBarcodeSankey(top20each)
ggsave("../../output/WuC/sankey_top20each.png", dpi = 300, width = 4, height = 4, 
       barbieQ::plotBarcodeSankey(top20each))

venn1 <- eulerr::euler(
  list(
    barbieQ = rownames(monkeyHSPC_top), 
    top10_per_sample = rownames(monkeyHSPC_agg)[colTagVec],
    top20_per_sample = rownames(monkeyHSPC_agg)[colTagVec2])
  )

plot(venn1, fills = c("white", "grey", "black"), alpha = 0.5, labels = TRUE, quantities = TRUE)
ggsave("../../output/WuC/venn_filtering.png", dpi = 300, width = 6, height = 4, 
       plot(venn1, fills = c("white", "grey", "black"), alpha = 0.5, labels = TRUE, quantities = TRUE))
```


```{r overlap}
venn2 <- eulerr::euler(
  list(
    barbieQ = rownames(monkeyHSPC_top), 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn2, fills = c("red", "blue"), alpha = 0.5, labels = TRUE, quantities = TRUE)
```

## NK-secific Clone Paper Selected

```{r paper barcode prop provided }
## calculate the total proportion of the paper selected barcodes based on the proportion data provided by the paper
NK_clone_contribution <- ZG66_Fig2 %>% colSums()
dat <- data.frame(
  contr = NK_clone_contribution,
  sample = c(rep("T", 6),
             rep("B", 6),
             rep("Gr", 6),
             rep("NK_CD56p_CD16n", 6),
             rep("NK_CD56n_CD16p", 6)))
dat$sample <- factor(dat$sample, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p"))

dat <- dat %>%
  mutate(id = row_number()) 
p11 <- ggplot(dat, aes(x = as.factor(id), y = contr, color = sample)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  ggtitle("Paper - identified clones") +
  labs(y = "Prop.", x = "Sample (cell type)")
p11
```


## bartools

```{r bartools, fig.width=10, fig.height=5}
temp <- assay(monkeyHSPC_top)[,1:30]
tp_condi <- monkeyHSPC_top$sampleMetadata[1:30,]
colnames(temp)
paste(tp_condi$Celltype, tp_condi$Months, sep = "_")
od_smp <- factor(colnames(temp), levels = unique(colnames(temp)))
colnames(temp) <- od_smp
monkey_dge <- DGEList(counts = temp)
bartools::plotBarcodeHistogram(monkey_dge)
cust_histogram(monkey_dge)


cust_histogram <- function (dgeObject, topN = 10, seedColors = 1, samples = NULL, 
    orderSamples = NULL, alphaLowFreq = 1) 
{
    # inputChecks(dgeObject, samples = c(samples, orderSamples))
    if (topN > 74) {
        message("Warning: number of labelled barcodes larger than number of colors in palette (74)")
    }
    counts <- as.data.frame(dgeObject$counts)
    if (is.null(samples)) {
        samples <- colnames(counts)
    }
    counts <- counts[, samples]
    if (!is.null(orderSamples)) {
        if (!all(orderSamples %in% samples)) {
            stop("samples to order by are not present or among selected samples")
        }
    }
    else {
        orderSamples <- samples
    }
    barcode_freqs <- counts %>% as.data.frame() %>% tibble::rownames_to_column(var = "barcode") %>% 
        tidyr::pivot_longer(-.data$barcode) %>% dplyr::rename(count = .data$value, 
        sample = .data$name) %>% dplyr::group_by(.data$sample) %>% 
        dplyr::mutate(freq = .data$count/sum(.data$count)) %>% 
        dplyr::ungroup()
    barcode_freqs$sample <- factor(barcode_freqs$sample, levels = unique(barcode_freqs$sample))
    barcode_order <- barcode_freqs %>% dplyr::filter(.data$sample %in% 
        orderSamples) %>% dplyr::select(.data$barcode, .data$freq) %>% 
        dplyr::group_by(.data$barcode) %>% dplyr::slice_max(order_by = .data$freq, 
        n = 1, with_ties = F) %>% dplyr::arrange(.data$freq)
    barcode_freqs$barcode <- factor(barcode_freqs$barcode, levels = barcode_order$barcode)
    qual_col_pals = RColorBrewer::brewer.pal.info[RColorBrewer::brewer.pal.info$category == 
        "qual", ]
    col_vector = unlist(mapply(RColorBrewer::brewer.pal, qual_col_pals$maxcolors, 
        rownames(qual_col_pals)))
    n <- length(unique(barcode_freqs$barcode))
    set.seed(seedColors)
    cols_vector <- sample(col_vector, n, replace = T)
    bc_labels <- barcode_freqs %>% dplyr::arrange(dplyr::desc(.data$freq)) %>% 
        dplyr::pull(.data$barcode) %>% unique() %>% utils::head(topN) %>% 
        as.character()
    label_values <- stats::setNames(as.list(cols_vector), unique(barcode_freqs$barcode))
    p <- barcode_freqs %>% ggplot2::ggplot(aes(fill = .data$barcode, 
        x = .data$sample, y = .data$freq, alpha = ifelse(.data$barcode %in% 
            bc_labels, 1, 0.5))) + ggplot2::geom_bar(position = "stack", 
        stat = "identity") + ggplot2::scale_alpha_continuous(guide = FALSE, 
        range = c(alphaLowFreq, 1)) + ggplot2::scale_fill_manual(values = label_values, 
        breaks = bc_labels) + ggplot2::theme_bw() + ggplot2::labs(fill = paste0("Top ", 
        topN, " barcodes")) + ggplot2::ylab("Proportion") + ggplot2::theme(legend.text = element_text(size = 8), 
        axis.text.x = element_text(angle = 45, hjust = 1))
    return(p)
}

```

```{r fig.width=10, fig.height=5}
cust_histogram(monkey_dge)
ggsave("../../output/WuC/histo_all.png", 
       width = 10, height = 5, dpi = 300, cust_histogram(monkey_dge))
```


# Model Design Matrix - filtered

- My selecting strategy to select `NK_CD56n_CD16p biased barcodes` by testing

 - [1] "(CelltypeNK_CD56n_CD16p)" v.s.                                      

 - [2] "(CelltypeB + CelltypeGr + CelltypeT + CelltypeNK_CD56p_CD16n)/4"

- Only keep the samples of "mid" phase, which is of timepoints between "6.5m to 42m".

## model

```{r checkcout df after contrast}
asinTrans <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top[,1:30],
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

## Alternatively: using logit transformation
logitTrans <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "logit"
)

## Alternatively: no transformation
ps.monkeyHSPC_top <- monkeyHSPC_top
assays(ps.monkeyHSPC_top)$proportion <- log2(assays(ps.monkeyHSPC_top)$proportion+1e-7)
noTrans <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "none"
)

## test Barcode differential occurrence between sample groups
## set up the targets (sample conditions)
targets <- monkeyHSPC_top$sampleMetadata %>% 
  as.data.frame() %>% 
  mutate(Group = ifelse(
    Celltype == "NK_CD56n_CD16p",
    "NK_CD56n_CD16p",
    "B.Gr.T.NK_CD56p_CD16n"))

monkeyHSPC_top <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top, 
  sampleMetadata = targets[,"Group", drop=FALSE], 
  method = "diffOcc"
)
```

## validation

```{r function validation}
plot_validation <- function(myObject) {
  
  flag <- rowData(myObject)$testingBarcode$direction == 1
  myObject <- myObject[flag,]
  
  NK_clone_contributionU <- assays(myObject)$proportion %>% colSums()
  dat <- data.frame(
    contr = NK_clone_contributionU,
    Celltype = factor(myObject$sampleMetadata$Celltype,
                      levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p")),
    Months = myObject$sampleMetadata$Months)
  ## sort samples by celltype and months
  dat <- dat %>%
    arrange(Celltype, Months) %>% 
    mutate(id = row_number()) 
  ## add dummy rows for spacing between celltypes
  dat <- dat %>%
    group_by(Celltype) %>%
    mutate(dummy = n() + 1) %>%    # Add a dummy count for spacing
    ungroup() %>%
    bind_rows(dat %>%             # Add dummy rows for spacing
                group_by(Celltype) %>%
                slice_tail(n = 1) %>%
                mutate(id = as.numeric(id) + 0.5, contr = NA, Months = NA)) %>%
    arrange(as.numeric(id))        # Arrange with dummy rows in place
  
  celltype_colors <- c(
  "T" = "yellowgreen",
  "B" = "yellowgreen",
  "Gr" = "yellowgreen",
  "NK_CD56p_CD16n" = "yellowgreen",
  "NK_CD56n_CD16p" = "slateblue1" # "orchid"
)
  
  p <- ggplot(dat, aes(x = as.factor(id), y = contr, color = Celltype)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 0.5, size = 10),
        axis.title.x = element_text(size = 14), 
        axis.title.y = element_text(size = 14), 
        panel.grid.major.x = element_blank()) +
  scale_x_discrete(labels = ifelse(is.na(dat$Months), "", dat$Months)) + # Relabel with `Celltype`
  scale_color_manual(values = celltype_colors) +  # <- set bar outline color
  scale_fill_manual(values = celltype_colors) +   # <- set tile fill color
  scale_y_continuous(limits = c(-0.09, 0.8)) +
  labs(x = "Samples of cell types at different months", y = "Total prop. of selected barcodes") +  # Update axis labels
  geom_tile(
    data = dat,
    aes(x = ifelse(is.na(Months), NA,order(id)), y = -0.03, fill = Celltype),
    height = 0.03, width = 1
  ) +
  geom_text(
    data = dat %>%
      mutate(id = order(id)) %>%
      group_by(Celltype) %>%
      mutate(Celltype = recode(Celltype, "NK_CD56n_CD16p" = "NK.CD56-CD16+")) %>% 
      mutate(Celltype = recode(Celltype, "NK_CD56p_CD16n" = "NK.CD56+CD16-")) %>% 
      summarize(mid_point = mean(as.numeric(id))),
    aes(x = mid_point-0.5, y = -0.03, label = Celltype),
    size = 3, vjust = 0.5, color = "black"
  )
  
  return(p)
}
```

```{r NK my select prop, fig.width=6, fig.height=5}
plot_validation(asinTrans) 
ggsave("../../output/WuC/validation_barbieQ.png", 
       dpi = 300, width = 6, height = 5, plot_validation(asinTrans))
plot_validation(logitTrans) 
plot_validation(noTrans) 

# plot_validation(myObject = paper_myObject) 

plotSignifBarcodeProportion(asinTrans)
```

```{r paper nk, fig.width=6, fig.height=5}
colnames(ZG66_Fig2)
colnames(monkeyHSPC_top)
ZG66_Fig2_bb <- createBarbieQ(ZG66_Fig2)
colData(ZG66_Fig2_bb)$sampleMetadata <- monkeyHSPC_top$sampleMetadata[1:30,]
assays(ZG66_Fig2_bb)$proportion <- ZG66_Fig2
rowData(ZG66_Fig2_bb)$testingBarcode$direction = 1
plot_validation(ZG66_Fig2_bb)

ggsave("../../output/WuC/validation_original.png", 
       dpi = 300, width = 6, height = 5, plot_validation(ZG66_Fig2_bb))
```

```{r}
venn3 <- eulerr::euler(
  list(
    barbieQ_select = rownames(monkeyHSPC_top)[rowData(asinTrans)$testingBarcode$direction == 1], 
    Heuristic_LFC = rownames(ZG66_Fig2)))

plot(venn3, fills = c("white", "grey"), alpha = 0.5, labels = F, quantities = F)
```


```{r prop}

ps.asinTrans <- noTrans
SummarizedExperiment::rowData(ps.asinTrans)$testingBarcode$tendencyTo <- rowData(asinTrans)$testingBarcode$tendencyTo
ps.logitTrans <- noTrans
SummarizedExperiment::rowData(ps.logitTrans)$testingBarcode$tendencyTo <- rowData(logitTrans)$testingBarcode$tendencyTo

(plotBarcodeMA(ps.asinTrans) + coord_trans(x = "log10"))|> plot()
(plotBarcodeMA(ps.logitTrans) + coord_trans(x = "log10"))|> plot()
(plotBarcodeMA(noTrans) + coord_trans(x = "log10"))|> plot()

## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../barcoding/public_barcode_count/data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]
ps.paper <- monkeyHSPC[rownames(monkeyHSPC) %in% rownames(ZG66_Fig2) | rownames(monkeyHSPC) %in% rownames(noTrans),]
ps.paper <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "none"
)
SummarizedExperiment::rowData(ps.paper)$testingBarcode$tendencyTo <- rep("n.s.", nrow(ps.paper))
SummarizedExperiment::rowData(ps.paper)$testingBarcode$tendencyTo[rownames(ps.paper) %in% rownames(ZG66_Fig2)] <- "CelltypeNK_CD56n_CD16p"
(plotBarcodeMA(ps.paper) + coord_trans(x = "log10"))|> plot()

```


```{r occ, fig.width = 6, fig.height = 4}

ps.occ <- noTrans
SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo <- rowData(monkeyHSPC_top)$testingBarcode$tendencyTo
SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo[SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo == "GroupNK_CD56n_CD16p"] <- "CelltypeNK_CD56n_CD16p"
SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo[SummarizedExperiment::rowData(ps.occ)$testingBarcode$tendencyTo == "GroupB.Gr.T.NK_CD56p_CD16n"] <- "CelltypeB + CelltypeGr + CelltypeNK_CD56p_CD16n + CelltypeT"
(plotBarcodeMA(ps.occ) + coord_trans(x = "log10"))|> plot()


plotBarcodeMA(ps.asinTrans) +theme(legend.position = "none")
plotBarcodeMA(ps.paper)+theme(legend.position = "none")
```




