---
title: "case_study"
author: "LF"
date: "2025-06-19"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
suppressMessages({
  library(readxl)
  library(magrittr)
  library(dplyr)
  library(tidyr) # for pivot_longer
  library(tibble) # for rownames_to _column
  library(knitr) # for kable()
  library(ggplot2)
  library(patchwork)
  library(scales)
  library(ggVennDiagram)
  library(ComplexHeatmap)
  library(proxy) ## for dist
  library(circlize)
  library(limma)
  
  library(SummarizedExperiment)
  library(S4Vectors)
  library(barbieQ)
  library(barcodetrackR)
})
  
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_ZG66_allPhases_barbieQ.rda")
```


```{r eval=FALSE, include=FALSE}
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "early",]
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "late",]
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx, eval=FALSE, include=FALSE}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

# Source FUN

```{r}
getwd()
source("FUN_validation.R")
```


# Data

Same 81 BC in paper Fig2

```{r}
# first 30 samples are included in the paper table
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,1:30])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_30 <- monkeyHSPC
rowData(top10each_30)$isTopBarcode$isTop <- colTagVec
# assays(top10each_30)$isTopAssay <- colTagMat
top10each_30 <- top10each_30[colTagVec,]

```

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,flag_early_mid])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_early_mid <- monkeyHSPC
rowData(top10each_early_mid)$isTopBarcode$isTop <- colTagVec
# assays(top10each_early_mid)$isTopAssay <- colTagMat
top10each_early_mid <- top10each_early_mid[colTagVec,]

```

subset NKnp samples of all 8 timepoints

Choose BC by top10each (suggested by paper): 39 BCs

Same 39 BC in paper Fig3

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

colTags_NKnp <- lapply(as.data.frame(assays(NKnp)$proportion), 
                  function(col) tagTop10(col, n = 10))
colTagMat_NKnp <- do.call(cbind, colTags_NKnp)
colTagVec_NKnp <- rowSums(colTagMat_NKnp) >= 1
sum(colTagVec_NKnp) 
all(colSums(colTagMat_NKnp) == 15)

top10each_NKnp <- monkeyHSPC[colTagVec_NKnp,]

## same data used by paper
NKnp <- NKnp[colTagVec_NKnp,]
```

monkeyHSPC_top99_early_mid saved in the rda object - no need to run again.

```{r eval=FALSE, include=FALSE}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
# tag top
monkeyHSPC <- tagTopBarcodes(barbieQ = monkeyHSPC, activeSamples = flag_early_mid, nSampleThreshold = 6)
flag_top <- monkeyHSPC@elementMetadata$isTopBarcode$isTop
table(flag_top)
monkeyHSPC_top99_early_mid <- monkeyHSPC[flag_top,]

```

## Save to rda

```{r}

save(
  monkeyHSPC, monkeyHSPC_top99_early_mid, file = "../../output/WuC/cluster_pca_facet.pngmonkeyHSPC_ZG66_allPhases_barbieQ.rda"
)

```


# Top10BC, 39BC from 8 NKnp samples

## HP clustering - paper analysis

### clustering approach


hierarchically clustering barcodes and extract the clusters

```{r}
NKnp_slot <- NKnp@assays@data$proportion
NKnp_hp <- Heatmap(log(NKnp_slot+1e-6), show_row_names = F, 
        column_order = order(NKnp$sampleMetadata$Months),
        col = colorRamp2(
    c(log(1+1e-6), log(0.01+1e-6), log(1e-6)),                     # log10 values for 1, 0.01, 0.000025
    c("red", "white", "blue")             # choose any colors you prefer
  ))

# NKnp_hp
```

```{r}
# 3. Extract row dendrogram and cut into 3 clusters
NKnp_row_dend <- row_dend(NKnp_hp)
NKnp_clusters <- cutree(as.hclust(NKnp_row_dend), k = 5)

# 5. Create row annotation based on clusters
row_ha <- rowAnnotation(
  NKnp_clusters = as.factor(NKnp_clusters),
  col = list(NKnp_clusters = structure(
    RColorBrewer::brewer.pal(5, "Set1"),
    names = as.character(1:5)
  )),
  show_annotation_name = TRUE
)

# 6. Plot heatmap with annotation
ht_with_anno <- Heatmap(
  log(NKnp_slot + 1e-6),
  show_row_names = FALSE, column_labels = NKnp$sampleMetadata$Months,
  column_order = order(NKnp$sampleMetadata$Months),
  # col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  left_annotation = row_ha,
  row_split = as.factor(NKnp_clusters)
)

ht_with_anno
```

### barcodetrackR source code

!!! this equals to the default clustering and algorithms in ComplexHeatmap

clustering_data <- SummarizedExperiment::assays(your_SE)[["logs"]]

log(NKnp_slot+1e-6)

clustering_data.dist <- proxy::dist(clustering_data, method = distance_method, p = minkowski_power)
hclustering <- hclust(clustering_data.dist, method = hclust_linkage)
clusters = cutree(hclustering, clusters)

```{r}
# Compute matrix and clustering
NKnp_slot <- NKnp@assays@data$proportion
log_mat <- log(NKnp_slot + 1e-6)

# Your existing clustering:
clustering_data.dist <- proxy::dist(log_mat, method = "euclidean", p = 2)
hclustering <- hclust(clustering_data.dist, method = "complete")
clusters <- cutree(hclustering, 5)
table(clusters)

# Make sure cluster vector is a factor in the same order as the matrix rows
row_split <- factor(clusters[rownames(log_mat)], levels = 1:5)

# Plot the heatmap with row_split
Heatmap(
  log_mat,
  show_row_names = FALSE,
  column_labels = NKnp$sampleMetadata$Months,
  column_order = order(NKnp$sampleMetadata$Months),
  # col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  row_split = row_split
)
```

## Test NKnp:Time, early_mid, top10each_NKnp

```{r}
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")


##need to take Time into account
designs <- top10each_NKnp[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


top10each_NKnp_test_time_early_mid <- testBarcodeSignif(
  barbieQ = top10each_NKnp[,flag_early_mid], designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## Test NKnp:Time, NKnp only, top10each_NKnp - Need to fix pkg here!!!

```{r eval=FALSE, include=FALSE}

designs <- NKnp$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Months)
)

top10each_NKnp_test_time_NKnp <- testBarcodeSignif(
  barbieQ = NKnp, ,sampleMetadata = NKnp$sampleMetadata, sampleGroup = "Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## HP Vis testing

```{r}
BC_kinetic <- top10each_NKnp_test_time_NKnp@elementMetadata$testingBarcode$tendencyTo

pch_fun <- ifelse(top10each_NKnp_test_time_NKnp@elementMetadata$testingBarcode$direction != 0, "*", NA)

row_ha <- rowAnnotation(
  NKnp_clusters = as.factor(NKnp_clusters),
  col = list(NKnp_clusters = structure(
    RColorBrewer::brewer.pal(5, "Set1"),
    names = as.character(1:5)
  )),
  show_annotation_name = TRUE
)


kinetic_ha <- rowAnnotation(
  Kinetic = anno_simple(
    BC_kinetic,
    col = top10each_NKnp_test_time_NKnp@metadata$factorColors$testingBarcode,
    pch = pch_fun),
  show_annotation_name = TRUE,
  annotation_name_side = "top",
  annotation_name_rot = 45 
)

HP_anno_kinetic <- Heatmap(
  log10(NKnp_slot + 1e-6), 
  show_row_names = FALSE, column_labels = NKnp$sampleMetadata$Months,
  column_title = "Months", column_names_side = "top", column_names_rot = 45,
  row_title = "Barcodes",
  column_order = order(NKnp$sampleMetadata$Months),
  # col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  # left_annotation = row_ha,
  right_annotation = kinetic_ha,
  row_split = as.factor(NKnp_clusters),
  heatmap_legend_param = list(
        title = "Prop.", at = c(-6, -4, -2, 0), 
        labels = c(0.000001, 0.0001, 0.01, 0)
    )
)

lgd_sig = Legend(pch = "*", type = "points", labels = "signif.")

lgd_kinetic = Legend(labels = c("up", "down", "n.s."), title = "Kinetic", 
             legend_gp = gpar(fill = c("#FF5959", "#33AAFF", "#FFC000"))
             )

```

```{r fig.width=4, fig.height=6}
draw(HP_anno_kinetic, annotation_legend_list = list(lgd_kinetic, lgd_sig))

png("../../output/WuC/kinetic_HP.png", width = 1200, height = 1200, res = 300)
draw(HP_anno_kinetic, annotation_legend_list = list(lgd_kinetic, lgd_sig))
dev.off()
```



## Compare

### Venn

```{r}
cluster_selected_list <- lapply(1:5, function(i) names(NKnp_clusters)[NKnp_clusters == i])
names(cluster_selected_list) <- paste("cluster_", 1:5, sep = "_")

seletced_barbieQ_NKnp_up <- rowData(top10each_NKnp_test_time_early_mid)$testingBarcode$direction == 1
seletced_barbieQ_NKnp_down <- rowData(top10each_NKnp_test_time_early_mid)$testingBarcode$direction == -1


cluster_selected_list[["test_NKnp_up"]] <- rownames(top10each_NKnp_test_time_NKnp)[seletced_barbieQ_NKnp_up]
cluster_selected_list[["test_NKnp_down"]] <- rownames(top10each_NKnp_test_time_NKnp)[seletced_barbieQ_NKnp_down]


cluster_selected_list[["test_all_up"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_up]
cluster_selected_list[["test_all_down"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_down]

venn99 <- eulerr::euler(
  cluster_selected_list[1:7]
)

plot(venn99, fills = c("white", "white","white","white","white","grey","grey"), alpha = 0.5, labels = T, quantities = T)


```

```{r}
venn99 <- eulerr::euler(
  cluster_selected_list
)

plot(venn99, fills = c("white", "white","white","white","white", "hotpink","royalblue","white","white","white"), alpha =0.8, labels = T, quantities = T)
```


### Line plot

```{r}

NK_long_time <- plot_line_validation(
  cluster_selected_list[c(1,2,3,5,7)], barbieQ = top10each_NKnp[,top10each_NKnp$sampleMetadata$Celltype == "NK_CD56n_CD16p"], get_long_data = T)

# Step 1: Summarise mean and SEM
NK_summary <- NK_long_time %>%
  filter(!str_detect(selectedBy, "none")) %>%
  group_by(selectedBy, Months) %>%
  summarise(
    mean_prop = mean(Proportion, na.rm = TRUE),
    sem = sd(Proportion, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) 

# Step 2: Plot mean with SEM ribbon
ggplot(NK_summary, aes(x = Months, y = mean_prop, color = selectedBy, fill = selectedBy)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_prop - sem, ymax = mean_prop + sem), alpha = 0.2, color = NA) +
  theme_minimal() +
  scale_color_manual(values = c("cluster__1,test_NKnp_down" = "limegreen", 
                                "cluster__2,test_NKnp_down" = "limegreen", 
                                "cluster__3,test_NKnp_down" = "limegreen", 
                                "cluster__5,test_NKnp_down" = "limegreen", 
                                "cluster__2" = "black")) +
  scale_fill_manual(values = c("cluster__1,test_NKnp_down" = "yellowgreen", 
                                "cluster__2,test_NKnp_down" = "yellowgreen", 
                                "cluster__3,test_NKnp_down" = "yellowgreen", 
                                "cluster__5,test_NKnp_down" = "yellowgreen", 
                                "cluster__2" = "grey")) 


```

```{r}
NK_long_time <- plot_line_validation(
  cluster_selected_list[c(4,6)], barbieQ = top10each_NKnp[,top10each_NKnp$sampleMetadata$Celltype == "NK_CD56n_CD16p"], get_long_data = T)

# Step 1: Summarise mean and SEM
NK_summary <- NK_long_time %>%
  filter(!str_detect(selectedBy, "none")) %>%
  group_by(selectedBy, Months) %>%
  summarise(
    mean_prop = mean(Proportion, na.rm = TRUE),
    sem = sd(Proportion, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Step 2: Plot mean with SEM ribbon
ggplot(NK_summary, aes(x = Months, y = mean_prop, color = selectedBy, fill = selectedBy)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_prop - sem, ymax = mean_prop + sem), alpha = 0.2, color = NA) +
  theme_minimal() +
  scale_color_manual(values = c("cluster__4,test_NKnp_up" = "red", 
                                "cluster__4" = "black")) +
  scale_fill_manual(values = c("cluster__4,test_NKnp_up" = "pink", 
                                "cluster__4" = "grey")) 

```

```{r}
NK_long_time_c1235 <- plot_line_validation(
  cluster_selected_list[c(1,2,3,4,5)], barbieQ = top10each_NKnp[,top10each_NKnp$sampleMetadata$Celltype == "NK_CD56n_CD16p"], get_long_data = T)

NK_long_time_down <- plot_line_validation(
  cluster_selected_list[c(6,7)], barbieQ = top10each_NKnp[,top10each_NKnp$sampleMetadata$Celltype == "NK_CD56n_CD16p"], get_long_data = T)

ggplot(NK_long_time_c1235 %>% mutate(testing_time = NK_long_time_down$selectedBy)) +
    geom_line(aes(x = Months, y = Proportion, group = Barcode, color = testing_time)) +
    facet_wrap(~selectedBy, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "top") +
  scale_color_manual(values = c("none" = "gold",
                                "test_NKnp_down" = "royalblue",
                                "test_NKnp_up" = "hotpink"))
```


```{r, fig.width=10, fig.height=12}
plot_line_validation(cluster_selected_list[c(1,2,3,5,7)], top10each_NKnp[,flag_early_mid])
```

```{r, fig.width=10, fig.height=12}
plot_line_validation(cluster_selected_list[c(1,2,3,5,7)], top10each_NKnp)
```

```{r, fig.width=10, fig.height=6}
plot_line_validation(cluster_selected_list[c(4,6)], top10each_NKnp[,flag_early_mid])
```

```{r, fig.width=10, fig.height=6}
plot_line_validation(cluster_selected_list[c(4,6)], top10each_NKnp)
```

# More BC

## Test NKnp, early_mid, 132 BC from 53 samples of all TP all celltype

```{r}
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")

##need to take Time into account
designs <- top10each_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


top10each_early_mid_test_time_early_mid <- testBarcodeSignif(
  barbieQ = top10each_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## Test, 1.6k BC

```{r}
dim(monkeyHSPC_top99_early_mid)
##need to take Time into account
designs <- monkeyHSPC_top99_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


top99n6_early_mid_test_time_early_mid <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top99_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
dim(top99n6_early_mid_test_time_early_mid)
```


## Compare

### Venn

```{r}
# cluster_selected_list <- lapply(1:5, function(i) names(NKnp_clusters)[NKnp_clusters == i])
# names(cluster_selected_list) <- paste("cluster_", 1:5, sep = "_")

NKnp_up_132BC <- top10each_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_132BC <- top10each_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == -1

NKnp_up_39BC <- top10each_NKnp_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_39BC <- top10each_NKnp_test_time_early_mid@elementMetadata$testingBarcode$direction == -1

NKnp_up_1.6kBC <- top99n6_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_1.6kBC <- top99n6_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == -1


cluster_selected_list[["barbieQ_NKnp_up"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_up]
cluster_selected_list[["barbieQ_NKnp_down"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_down]

list_test_NK_early_mid <- list(
    NKnp_up_132BC = rownames(top10each_early_mid_test_time_early_mid)[NKnp_up_132BC],
    NKnp_down_132BC = rownames(top10each_early_mid_test_time_early_mid)[NKnp_down_132BC],
    NKnp_up_39BC = rownames(top10each_NKnp_test_time_early_mid)[NKnp_up_39BC],
    NKnp_down_39BC = rownames(top10each_NKnp_test_time_early_mid)[NKnp_down_39BC],
    NKnp_up_1.6kBC = rownames(top99n6_early_mid_test_time_early_mid)[NKnp_up_1.6kBC],
    NKnp_down_1.6kBC = rownames(top99n6_early_mid_test_time_early_mid)[NKnp_down_1.6kBC]
  )

venn_test_NK_early_mid <- eulerr::euler(
  list_test_NK_early_mid[1:4]
)

plot(venn_test_NK_early_mid, fills = c("white", "white","grey","grey"), alpha = 0.5, labels = T, quantities = T)


```
```{r}
venn_test_NK_early_mid <- eulerr::euler(
  list_test_NK_early_mid
)

plot(venn_test_NK_early_mid, fills = c("white", "white","grey","grey", "grey2", "grey2"), alpha = 0.5, labels = T, quantities = T)
```

### Line plot

```{r}
plot_line_validation(list_test_NK_early_mid[c(1,3)], top10each_early_mid[,flag_early_mid])
plot_line_validation(list_test_NK_early_mid[c(2,4)], top10each_early_mid[,flag_early_mid])
```

```{r}
plot_line_validation(list_test_NK_early_mid[c(1,3)], top10each_early_mid)
plot_line_validation(list_test_NK_early_mid[c(2,4)], top10each_early_mid)
```

```{r, fig.width=10, fig.height=6}
plot_line_validation(list_test_NK_early_mid[c(1,3,5)], monkeyHSPC_top99_early_mid[,flag_early_mid])
plot_line_validation(list_test_NK_early_mid[c(2,4,6)], monkeyHSPC_top99_early_mid[,flag_early_mid])
```

```{r, fig.width=10, fig.height=6}
plot_line_validation(list_test_NK_early_mid[c(1,3,5)], monkeyHSPC_top99_early_mid)
plot_line_validation(list_test_NK_early_mid[c(2,4,6)], monkeyHSPC_top99_early_mid)
```

# Two figure, 132 BC from all celltype all times

## Test Celltype, early_mid, 132 BC from 53 samples of all TP all celltype

```{r}
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")

##need to take Time into account
designs <- top10each_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_early_mid_test_celltype_ealy_mid <- testBarcodeSignif(
  barbieQ = top10each_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

top10each_early_mid_test_celltype_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```

## Compare

### Venn

```{r}
NKnp_up_132BC <- top10each_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_132BC <- top10each_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == -1

NKnp_more <- top10each_early_mid_test_celltype_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"

NKnp_less <- top10each_early_mid_test_celltype_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeB + CelltypeGr + CelltypeNK_CD56p_CD16n + CelltypeT"

list_test_two_figure_early_mid <- list(
    NKnp_up = rownames(top10each_early_mid_test_time_early_mid)[NKnp_up_132BC],
    NKnp_down = rownames(top10each_early_mid_test_time_early_mid)[NKnp_down_132BC],
    NKnp_more = rownames(top10each_early_mid_test_celltype_ealy_mid)[NKnp_more],
    NKnp_less = rownames(top10each_early_mid_test_celltype_ealy_mid)[NKnp_less]
    
)

venn_test_two_figure_early_mid <- eulerr::euler(
  list_test_two_figure_early_mid
)

plot(venn_test_two_figure_early_mid, fills = c("white", "white","grey","grey"), alpha = 0.5, labels = T, quantities = T)
```

### Line plot

```{r}
plot_line_validation(list_test_two_figure_early_mid[1:3], top10each_early_mid[,flag_early_mid])
```

```{r}
plot_line_validation(list_test_two_figure_early_mid[1:3], top10each_early_mid)
```

# Two figure, 1.6k BC from all celltype all times

## Test Celltype, early_mid, 1.6k BC from 53 samples of all TP all celltype

```{r}
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")

##need to take Time into account
designs <- monkeyHSPC_top99_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top99n6_early_mid_test_celltype_ealy_mid <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top99_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## Compare

### Venn

```{r}
NKnp_up_132BC <- top99n6_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_132BC <- top99n6_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == -1

NKnp_more <- top99n6_early_mid_test_celltype_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"

NKnp_less <- top99n6_early_mid_test_celltype_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeB + CelltypeGr + CelltypeNK_CD56p_CD16n + CelltypeT"

list_test_two_figure_early_mid_top99n6 <- list(
    NKnp_up = rownames(top99n6_early_mid_test_time_early_mid)[NKnp_up_132BC],
    NKnp_down = rownames(top99n6_early_mid_test_time_early_mid)[NKnp_down_132BC],
    NKnp_more = rownames(top99n6_early_mid_test_celltype_ealy_mid)[NKnp_more],
    NKnp_less = rownames(top99n6_early_mid_test_celltype_ealy_mid)[NKnp_less]
    
)

venn_test_two_figure_early_mid_top99n6 <- eulerr::euler(
  list_test_two_figure_early_mid_top99n6 
)

plot(venn_test_two_figure_early_mid_top99n6, fills = c("white", "white","grey","grey"), alpha = 0.5, labels = T, quantities = T)
```

### Line plot

```{r, fig.width=8, fig.height=8}
plot_line_validation(list_test_two_figure_early_mid_top99n6, monkeyHSPC_top99_early_mid[,flag_early_mid])
```

```{r, fig.width=12, fig.height=8}
plot_line_validation(list_test_two_figure_early_mid_top99n6, monkeyHSPC_top99_early_mid)
```


# Compare filtering

## Venn 

```{r}
list_compare_filtering <- vector("list", length = 8)
names(list_compare_filtering) <- c(
  paste("pp", gsub("NKnp_(.*)", "\\1",names(list_test_two_figure_early_mid))),
  paste("bb", gsub("NKnp_(.*)", "\\1",names(list_test_two_figure_early_mid_top99n6))))

list_compare_filtering[1:4] <- list_test_two_figure_early_mid
list_compare_filtering[5:8] <- list_test_two_figure_early_mid_top99n6


venn_omg <- eulerr::euler(
  list_compare_filtering[c(3, 5:7)]
)

plot(venn_omg, fills = c("grey","white","white","white"), alpha = 0.5, labels = T, quantities = T)

venn_omg <- eulerr::euler(
  list_compare_filtering[c(1:3, 5:7)]
)

plot(venn_omg, fills = c("grey","grey","grey","white", "white","white"), alpha = 0.5, labels = T, quantities = T)
```

```{r}
cluster_selected_list[["LFC"]] <- union_LFC_top10each_30

venn_paper <- eulerr::euler(
  cluster_selected_list[c(6, 1:5)]
)

plot(venn_paper, fills = c("white", "white","white", "white","white", "grey"), alpha = 0.5, labels = T, quantities = T)

```

