---
title: "case_study"
author: "LF"
date: "2025-06-18"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(circlize)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_ZG66_allPhases_barbieQ.rda")
```


```{r}
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "early",]
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "late",]
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

# Top10BC per sample

```{r filtering original}
tagTop10 <- function(col, n=10) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= n
  # summary(isTop)
  return(isTop)
}

colTags <- lapply(as.data.frame(assays(monkeyHSPC)$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each <- monkeyHSPC
rowData(top10each)$isTopBarcode$isTop <- colTagVec
# assays(top10each)$isTopAssay <- colTagMat
barbieQ::plotBarcodeSankey(top10each)
# ggsave("../../output/WuC/sankey_top10each.png", dpi = 300, width = 4, height = 4, 
#        barbieQ::plotBarcodeSankey(top10each))
top10each <- top10each[colTagVec,]

tagTop20 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 20
  # summary(isTop)
  return(isTop)
}
colTags2 <- lapply(as.data.frame(assays(monkeyHSPC)$proportion), 
                  function(col) tagTop20(col))
colTagMat2 <- do.call(cbind, colTags2)
colTagVec2 <- rowSums(colTagMat2) >= 1

top20each <- monkeyHSPC
rowData(top20each)$isTopBarcode$isTop <- colTagVec2
# assays(top10each)$isTopAssay <- colTagMat2
barbieQ::plotBarcodeSankey(top20each)
# ggsave("../../output/WuC/sankey_top20each.png", dpi = 300, width = 4, height = 4, 
#        barbieQ::plotBarcodeSankey(top20each))
top20each <- top20each[colTagVec2,]

venn1 <- eulerr::euler(
  list(
    barbieQ = rownames(monkeyHSPC)[monkeyHSPC@elementMetadata$isTopBarcode$isTop], 
    top10_per_sample = rownames(top10each),
    top20_per_sample = rownames(top20each)
  )
)

plot(venn1, fills = c("white", "grey", "black"), alpha = 0.5, labels = TRUE, quantities = TRUE)
# ggsave("../../output/WuC/venn_filtering.png", dpi = 300, width = 6, height = 4, 
       # plot(venn1, fills = c("white", "grey", "black"), alpha = 0.5, labels = TRUE, quantities = TRUE))
```

## Paper table

Downloaded table from the original paper contains 81 barcodes and 30 samples. Barcodes selected using the "top10each" approach based on the 30 samples from the original ZG66 dataset, monkeyHSPC, perfectly overlap with the 81 barcodes in the downloaded table.

However, when samples of earlier or later months are included, more barcodes will be selected.

```{r}
# first 30 samples are included in the paper table
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,1:30])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_30 <- monkeyHSPC
rowData(top10each_30)$isTopBarcode$isTop <- colTagVec
# assays(top10each_30)$isTopAssay <- colTagMat
top10each_30 <- top10each_30[colTagVec,]

```

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,flag_early_mid])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_early_mid <- monkeyHSPC
rowData(top10each_early_mid)$isTopBarcode$isTop <- colTagVec
# assays(top10each_early_mid)$isTopAssay <- colTagMat
top10each_early_mid <- top10each_early_mid[colTagVec,]

```


```{r overlap}
venn2 <- eulerr::euler(
  list(
    top10each = rownames(top10each), 
    top10each_early_mid = rownames(top10each_early_mid), 
    top10each_30 = rownames(top10each_30), 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn2, fills = c("white", "grey"), alpha = 0.5, labels = TRUE, quantities = TRUE)
```

```{r paper barcode prop provided }
## calculate the total proportion of the paper selected barcodes based on the proportion data provided by the paper
NK_clone_contribution <- ZG66_Fig2 %>% colSums()
dat <- data.frame(
  contr = NK_clone_contribution,
  sample = c(rep("T", 6),
             rep("B", 6),
             rep("Gr", 6),
             rep("NK_CD56p_CD16n", 6),
             rep("NK_CD56n_CD16p", 6)))
dat$sample <- factor(dat$sample, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p"))

dat <- dat %>%
  mutate(id = row_number()) 
p11 <- ggplot(dat, aes(x = as.factor(id), y = contr, color = sample)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  ggtitle("Paper - identified clones") +
  labs(y = "Prop.", x = "Sample (cell type)")
p11
```

## Heatmap by Phase

Quite a few shared barcodes between early+mid phases and late phase where samples are mature NK subtypes.

```{r}
plotBarcodeHeatmap(top10each, splitSamples = T, sampleGroup = "Phase")
```

# Function: Validation 

```{r function validation}
plot_validation <- function(myObject) {
  
  flag <- rowData(myObject)$testingBarcode$direction == 1
  myObject <- myObject[flag,]
  
  NK_clone_contributionU <- assays(myObject)$proportion %>% colSums()
  dat <- data.frame(
    contr = NK_clone_contributionU,
    Celltype = factor(myObject$sampleMetadata$Celltype,
                      levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p")),
    Months = myObject$sampleMetadata$Months)
  ## sort samples by celltype and months
  dat <- dat %>%
    arrange(Celltype, Months) %>% 
    mutate(id = row_number()) 
  ## add dummy rows for spacing between celltypes
  dat <- dat %>%
    group_by(Celltype) %>%
    mutate(dummy = n() + 1) %>%    # Add a dummy count for spacing
    ungroup() %>%
    bind_rows(dat %>%             # Add dummy rows for spacing
                group_by(Celltype) %>%
                slice_tail(n = 1) %>%
                mutate(id = as.numeric(id) + 0.5, contr = NA, Months = NA)) %>%
    arrange(as.numeric(id))        # Arrange with dummy rows in place
  
  celltype_colors <- c(
  "T" = "yellowgreen",
  "B" = "yellowgreen",
  "Gr" = "yellowgreen",
  "NK_CD56p_CD16n" = "yellowgreen",
  "NK_CD56n_CD16p" = "slateblue1" # "orchid"
)
  
  p <- ggplot(dat, aes(x = as.factor(id), y = contr, color = Celltype)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 0.5, size = 10),
        axis.title.x = element_text(size = 14), 
        axis.title.y = element_text(size = 14), 
        panel.grid.major.x = element_blank()) +
  scale_x_discrete(labels = ifelse(is.na(dat$Months), "", dat$Months)) + # Relabel with `Celltype`
  scale_color_manual(values = celltype_colors) +  # <- set bar outline color
  scale_fill_manual(values = celltype_colors) +   # <- set tile fill color
  scale_y_continuous(limits = c(-0.09, 0.8)) +
  labs(x = "Samples of cell types at different months", y = "Total prop. of selected barcodes") +  # Update axis labels
  geom_tile(
    data = dat,
    aes(x = ifelse(is.na(Months), NA,order(id)), y = -0.03, fill = Celltype),
    height = 0.03, width = 1
  ) +
  geom_text(
    data = dat %>%
      mutate(id = order(id)) %>%
      group_by(Celltype) %>%
      mutate(Celltype = recode(Celltype, "NK_CD56n_CD16p" = "NK.CD56-CD16+")) %>% 
      mutate(Celltype = recode(Celltype, "NK_CD56p_CD16n" = "NK.CD56+CD16-")) %>% 
      summarize(mid_point = mean(as.numeric(id))),
    aes(x = mid_point-0.5, y = -0.03, label = Celltype),
    size = 3, vjust = 0.5, color = "black"
  )
  
  return(p)
}
```

# Function: line validation

```{r}
plot_line_validation <- function(barcode_set_list, barbieQ) {
  ## Example for barcode set list
  # barcode_set_list <- list(
  #   test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  #   test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  #   heuristic = union_LFC_top10each_30
  # )
  
  ## check barcode set list and barbieQ contents
  print(names(barcode_set_list))
  print(dim(barbieQ))
  
  # Create named vector with default value "none"
  selectedBy <- setNames(rep("none", nrow(barbieQ)), rownames(barbieQ))
  
  # Assign sources
  for (source_name in names(barcode_set_list)) {
    matching_barcodes <- barcode_set_list[[source_name]]
    selectedBy[matching_barcodes] <- ifelse(
      selectedBy[matching_barcodes] == "none",
      source_name,
      paste(selectedBy[matching_barcodes], source_name, sep = ",")
    )
  }
  
  ## extract props. and sample conditions
  prop. <- assays(barbieQ)$proportion
  cond. <- barbieQ$sampleMetadata %>% as.data.frame()
  
  cond.$Celltype <- factor(cond.$Celltype,
                      levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p", 
                                 "NK_NKG2Ap_CD16p", "NK_NKG2Ap_CD16p_KIR3DL01n",
                                 "NK_NKG2Ap_CD16p_KIR3DL01p"))

  ## combine slots
  prop_df <- data.frame(cond., t(prop.))
  ## keep only early and mid samples
  prop_df <- prop_df %>% 
    # dplyr::filter(Phase %in% c("early", "mid")) %>% 
    select(-Phase)
  ## classify BC by selected by test or LFC
  
  ## picot to long data
  prop_long <- prop_df %>%
    pivot_longer(
      cols = -c(Celltype, Months),     # keep Celltype and Months fixed
      names_to = "Barcode",
      values_to = "Proportion"
    )
  ## assign to barcode selectedBy status
  prop_long$selectedBy <- selectedBy[as.character(prop_long$Barcode)]
  
  ## plot
  p <- ggplot(prop_long) +
    geom_line(aes(x = Months, y = Proportion, group = Barcode, color = selectedBy)) +
    facet_grid(selectedBy ~ Celltype, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "top")
  
  return(p)
  
}
```


# Function: Calc LFC

```{r}
calc_bias_LFC <- function(barbieQ) {
  ## extract props. and sample conditions
  prop. <- assays(barbieQ)$proportion
  cond. <- barbieQ$sampleMetadata %>% as.data.frame()
  ## group samples by NKnp or others
  cond. <- cond. %>% 
    mutate(isNK = ifelse(Celltype == "NK_CD56n_CD16p", "NKnp", "others"))
  cond.$Celltype %>% table()
  
  ## calc mean per group at each month
  group_labels <- paste(cond.$Months, cond.$isNK, sep = "__")
  group_means <- sapply(unique(group_labels), function(g) {
    idx <- which(group_labels == g)
    rowMeans(prop.[, idx, drop = FALSE])
  })
  
  ## calc group LFC at each month
  colnames(group_means)
  months <- gsub("(\\1)__.*", "\\1",colnames(group_means))
  
  LFC_months <- sapply(unique(months), function(m) {
    idx <- which(months == m)
    sub <- group_means[, idx, drop = FALSE]
    group <- gsub(".*__()", "\\1",colnames(sub))
    NK <- sub[, which(group == "NKnp"), drop = F]
    others <- sub[, which(group=="others"),drop = F]
    NK / others 
  })
  
  ## decide if each barcode LFC is high at each month
  is_LFC_high <- LFC_months > 10
  colnames(is_LFC_high)
  
  ## calc if each barcode NKnp  prop > 0.01 at each month
  NK_prop <- prop.[,cond.$isNK == "NKnp", drop=F]
  is_NKnp_high <- NK_prop > 0.01
  colnames(is_NKnp_high)
  
  ## barcodes meet both metric at each month
  select_months <- (is_LFC_high & is_NKnp_high)
  
  ## select barcodes at each month
  select_barcodes_months <- lapply(colnames(select_months), function(j) 
    rownames(prop.)[select_months[,j]])
  
  names(select_barcodes_months) <- paste0("heuristic at ", colnames(is_LFC_high), "m")
  
  return(select_barcodes_months)
}

```

# Paper 81 barcodes (top10each from 30 samples)

- My selecting strategy to select `NK_CD56n_CD16p biased barcodes` by testing

 - [1] "(CelltypeNK_CD56n_CD16p)" v.s.                                      

 - [2] "(CelltypeB + CelltypeGr + CelltypeT + CelltypeNK_CD56p_CD16n)/4"

- Only keep the samples of timepoints at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m. (to be consistent with the paper's analysis)

## Test 30 samples

```{r checkcout df after contrast}
## For time points at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m, each collecting all cell types.
## These samples refer to the first 30 samples in the object, which is also in line with the samples in the table downloaded from the paper's materials.
dim(top10each_30)

##need to take Time into account
designs <- top10each_30[,1:30]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_30_test30 <- testBarcodeSignif(
  barbieQ = top10each_30[,1:30], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
# check design
# top10each_30_test30@elementMetadata$testingBarcode@metadata$design
top10each_30_test30@elementMetadata$testingBarcode@metadata$contrasts
dim(top10each_30_test30)



## test Barcode differential occurrence between sample groups
## set up the targets (sample conditions)
# targets <- top10each_30$sampleMetadata %>% 
#   as.data.frame() %>% 
#   mutate(Group = ifelse(
#     Celltype == "NK_CD56n_CD16p",
#     "NK_CD56n_CD16p",
#     "B.Gr.T.NK_CD56p_CD16n"))
# 
# occTest <- testBarcodeSignif(
#   barbieQ = top10each_30, 
#   sampleMetadata = targets[,"Group", drop=FALSE], 
#   method = "diffOcc"
# )
```


## Test early+mid

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
dim(top10each_30)

##need to take Time into account
designs <- top10each_30[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_30_test_ealy_mid <- testBarcodeSignif(
  barbieQ = top10each_30[,flag_early_mid], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top10each_30_test_ealy_mid)
# check design
# top10each_30_test_ealy_mid@elementMetadata$testingBarcode@metadata$design
top10each_30_test_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```

## Compare

```{r}
plot_validation(top10each_30_test30)
plot_validation(top10each_30_test_ealy_mid)
```

```{r}

venn_top10each_30 <- eulerr::euler(
  list(
    test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"])
)

plot(venn_top10each_30 , fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)
```


## LFC approach

(i) >1% contribution to CD56−CD16+ cells and 

(ii) >10-fold abundance in contribution to CD56−CD16+ cells compared with T, B, Gr, and CD56+CD16− cells.

```{r}
list_BC_top10each_30_LFC <- calc_bias_LFC(top10each_30_test30)
```

### Venn: LFC at each month

```{r}
## check overlaps between months
venn0 <- eulerr::euler(list_BC_top10each_30_LFC)
plot(venn0, fills = c("white"), alpha = 0.5, labels = T, quantities = T)
```

## Compare

```{r}
## check overlaps
compare_list <- list_BC_top10each_30_LFC 
compare_list[["test_30sample"]] <- rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"]
compare_list[["test_early_mid"]] <- rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"]

venn00 <- eulerr::euler(compare_list)

plot(venn00, fills = c("white"), alpha = 0.5, labels = T, quantities = T)
```


## Validation: LFC at each month

```{r}
plot_list_LFC_top10each_30 <- lapply(list_BC_top10each_30_LFC, function(m) plot_validation(top10each_30_test30[m,]))

plot_list_LFC_top10each_early_mid <- lapply(list_BC_top10each_30_LFC, function(m) plot_validation(top10each_30_test_ealy_mid[m,]))
```


```{r, fig.width=12, fig.height=8}
suppressWarnings(wrap_plots(plot_list_LFC_top10each_30, ncol = 3))

```

```{r, fig.width=18, fig.height=8}
suppressWarnings(wrap_plots(plot_list_LFC_top10each_early_mid, ncol = 3))

```

#### LFC vs two tests

```{r}
union_LFC_top10each_30 <- unlist(list_BC_top10each_30_LFC) %>% unique()

venn_top10each_30 <- eulerr::euler(
  list(
    test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    union_heuristic = union_LFC_top10each_30)
)

plot(venn_top10each_30 , fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)
```

```{r}
top10each_30_LFC <- top10each_30[union_LFC_top10each_30, 1:30]
dim(top10each_30_LFC)
rowData(top10each_30_LFC)$testingBarcode$direction = 1

top10each_early_mid_LFC <- top10each_early_mid[union_LFC_top10each_30, flag_early_mid]
rowData(top10each_early_mid_LFC)$testingBarcode$direction = 1

(plot_validation(top10each_30_test30) + plot_validation(top10each_30_test_ealy_mid)) /
  (plot_validation(top10each_30_LFC) + plot_validation(top10each_early_mid_LFC))

```

## Line plot: test early+mid vs LFC

indicate each barcode is selected by test_30sample, test_early-mid, or LFC



```{r}
# Define barcode sets
barcode_list <- list(
  test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  heuristic = union_LFC_top10each_30
)

dim(top10each_early_mid)
```


```{r, fig.width=8, fig.height=7}
plot_line_validation(
  barcode_set_list = barcode_list, 
  barbieQ = top10each_30[,top10each_30$sampleMetadata$Phase != "late"])

plot_line_validation(
  barcode_set_list = barcode_list, 
  barbieQ = top10each_early_mid[,top10each_early_mid$sampleMetadata$Phase != "late"])
```

# More BC, using same 53 samples of all TP

## Test: 132 BC (top10each from 53 samples)

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- top10each_early_mid$sampleMetadata$Phase %in% c("early", "mid")
dim(top10each_early_mid)

##need to take Time into account
designs <- top10each_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_early_mid_test_ealy_mid <- testBarcodeSignif(
  barbieQ = top10each_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top10each_early_mid_test_ealy_mid)
# check design
# top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$design
top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```

## Test bbQ BC (99%, n=6, from 53 samples)

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")

# tag top
monkeyHSPC_top99_early_mid_0 <- tagTopBarcodes(barbieQ = monkeyHSPC[,flag_early_mid], nSampleThreshold = 6)
flag_top <- monkeyHSPC_top99_early_mid_0@elementMetadata$isTopBarcode$isTop
table(flag_top)
monkeyHSPC_top99_early_mid <- monkeyHSPC_top99_early_mid_0[flag_top,]

##need to take Time into account
designs <- monkeyHSPC_top99_early_mid$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>%
  with(
    model.matrix(~0+ Celltype + Months)
)
# colnames(designs) <- gsub(":", ".", colnames(designs))
# Celltype + Celltype:Months didn't work well

top99n6_early_mid_test_ealy_mid <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top99_early_mid, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top99n6_early_mid_test_ealy_mid)
# check design
# top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$design
top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```


## Compare

### Total prop

```{r}
plot_validation(top10each_30_test_ealy_mid)
plot_validation(top10each_early_mid_test_ealy_mid)
plot_validation(top99n6_early_mid_test_ealy_mid)
```

### Venn

```{r}

venn_test_early_mid <- eulerr::euler(
  list(
    top10each_30 = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    top10each_early_mid = rownames(top10each_early_mid_test_ealy_mid)[top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    top99n6 = rownames(top99n6_early_mid_test_ealy_mid)[top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"]
    ## to chech if the missing one is filtered out: no, it's included
    # all_top99n6 = rownames(top99n6_early_mid_test_ealy_mid)
    )
)

plot(venn_test_early_mid, fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)
```

### Line plot

```{r, fig.width=8, fig.height=8}
BC_list <- list(
    BC81 = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    BC132 = rownames(top10each_early_mid_test_ealy_mid)[top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    BC1.6K = rownames(top99n6_early_mid_test_ealy_mid)[top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"])



plot_line_validation(BC_list, monkeyHSPC_top99_early_mid)
plot_line_validation(BC_list, monkeyHSPC)
```


-------------

