---
title: "case_study"
author: "LF"
date: "2025-07-16"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(circlize)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_ZG66_allPhases_barbieQ.rda")
```



```{r}
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "early",]
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "late",]
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```


# Source FUN

```{r}
getwd()
source("FUN_validation.R")
```

# Data

Same 81 BC in paper Fig2

```{r}
# first 30 samples are included in the paper table
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,1:30])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_30 <- monkeyHSPC
rowData(top10each_30)$isTopBarcode$isTop <- colTagVec
# assays(top10each_30)$isTopAssay <- colTagMat
top10each_30 <- top10each_30[colTagVec,]

```

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,flag_early_mid])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_early_mid <- monkeyHSPC
rowData(top10each_early_mid)$isTopBarcode$isTop <- colTagVec
# assays(top10each_early_mid)$isTopAssay <- colTagMat
top10each_early_mid <- top10each_early_mid[colTagVec,]

```

subset NKnp samples of all 8 timepoints

Choose BC by top10each (suggested by paper): 39 BCs

Same 39 BC in paper Fig3

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

colTags_NKnp <- lapply(as.data.frame(assays(NKnp)$proportion), 
                  function(col) tagTop10(col, n = 10))
colTagMat_NKnp <- do.call(cbind, colTags_NKnp)
colTagVec_NKnp <- rowSums(colTagMat_NKnp) >= 1
sum(colTagVec_NKnp) 
all(colSums(colTagMat_NKnp) == 15)

top10each_NKnp <- monkeyHSPC[colTagVec_NKnp,]

## same data used by paper
NKnp <- NKnp[colTagVec_NKnp,]
```

monkeyHSPC_top99_early_mid saved in the rda object - no need to run again.

```{r eval=FALSE, include=FALSE}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
# tag top
## no need to run again
# monkeyHSPC <- tagTopBarcodes(barbieQ = monkeyHSPC, activeSamples = flag_early_mid, nSampleThreshold = 6)
## top BC from 59 samples
flag_top <- monkeyHSPC@elementMetadata$isTopBarcode$isTop
table(flag_top)
monkeyHSPC_top99_early_mid <- monkeyHSPC[flag_top,]
## top BC from first 30 sampels
flag_top_2 <- monkeyHSPC@assays@data$isTopAssay[,1:30] %>% rowSums() >=6
table(flag_top_2)
monkeyHSPC_top99_30 <- monkeyHSPC[flag_top_2, 1:30]
dim(monkeyHSPC_top99_30)
```

# Paper 81 barcodes (top10each from 30 samples)

- My selecting strategy to select `NK_CD56n_CD16p biased barcodes` by testing

 - [1] "(CelltypeNK_CD56n_CD16p)" v.s.                                      

 - [2] "(CelltypeB + CelltypeGr + CelltypeT + CelltypeNK_CD56p_CD16n)/4"

- Only keep the samples of timepoints at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m. (to be consistent with the paper's analysis)

## LFC approach

(i) >1% contribution to CD56−CD16+ cells and 

(ii) >10-fold abundance in contribution to CD56−CD16+ cells compared with T, B, Gr, and CD56+CD16− cells.

```{r}
list_BC_top10each_30_LFC <- calc_bias_LFC(top10each_30_test30)
union_LFC_top10each_30 <- unlist(list_BC_top10each_30_LFC) %>% unique()
```

```{r}
## check overlaps between months
venn0 <- eulerr::euler(list_BC_top10each_30_LFC)
plot(venn0, fills = c("white"), alpha = 0.5, labels = T, quantities = T)
```

## Test 30 samples

```{r checkcout df after contrast}
## For time points at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m, each collecting all cell types.
## These samples refer to the first 30 samples in the object, which is also in line with the samples in the table downloaded from the paper's materials.
dim(top10each_30)

##need to take Time into account
designs <- top10each_30[,1:30]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_30_test30 <- testBarcodeSignif(
  barbieQ = top10each_30[,1:30], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
# check design
# top10each_30_test30@elementMetadata$testingBarcode@metadata$design
top10each_30_test30@elementMetadata$testingBarcode@metadata$contrasts
dim(top10each_30_test30)
```

## Venn: heu vs test

```{r}
list_top10each_30 <- list(
    barbieQ = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    heuristics = union_LFC_top10each_30)

venn_top10each_30 <- eulerr::euler(
  list_top10each_30
)

plot(venn_top10each_30 , fills = c("hotpink", "slateblue"), alpha = 0.7, labels = T, quantities = T)


## save
png("../../output/WuC/venn_bbq_heu.png", width = 1500, height = 1000, res = 300)
# Plot
plot(venn_top10each_30 , fills = c("white", "white"), alpha = 0.7, labels = T, quantities = T)
# Close the device
dev.off()

```


## Boxplot FC

```{r}
FC_top10each_30 <- calc_bias_LFC(barbieQ = top10each_30[,1:30], get_FC = T) %>% as.data.frame()
rownames(FC_top10each_30) <- rownames(top10each_30)
## annotating BC by barbieQ seletion
FC_top10each_30$bbq <- top10each_30_test30@elementMetadata$testingBarcode$direction
## annotating BC by heuristic selection
FC_top10each_30$heu = 0
FC_top10each_30[union_LFC_top10each_30, "heu"] = 1
  ## adding BC name
FC_top10each_30 <- FC_top10each_30 %>% 
  mutate(Barcode = rownames(top10each_30)) %>% 
  ## summarizing two selecting methods overlaps
  mutate(identifiedBy = ifelse(bbq == 1, 
                             ifelse(heu == 1, "both", "barbieQ"), 
                             ifelse(heu == 1, "heuristics", "neither"))) %>% 
  dplyr::select(-c(heu, bbq))

## extract FC from testing
coff_FC <- top10each_30_test30@elementMetadata$testingBarcode$meanDiff
FC_top10each_30$testing_coeff <- coff_FC

long_FC_top10each_30 <- FC_top10each_30 %>% 
  mutate(Barcode = rownames(top10each_30)) %>% 
  pivot_longer(
      cols = -c(Barcode, identifiedBy, testing_coeff),     # keep Celltype and Months fixed
      names_to = "Month",
      values_to = "FC"
    )
```

```{r fig.width=10, fig.height=5}
## order BC in x axis by mean LFC
barcode_order <- long_FC_top10each_30 %>%
  group_by(Barcode) %>%
  summarise(mean_LFC = mean(log2(FC), na.rm = TRUE)) %>%
  arrange(mean_LFC) %>%
  pull(Barcode)

p_box_0 <- ggplot(long_FC_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = log2(FC), group = (Barcode), color = identifiedBy)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank()) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  labs(x = "Barocdes ordered by mean LFC across time",
       y = "Barcode proportion log2(FC) at each time") +
  scale_color_manual(values = c("both" = "tan3", "barbieQ" = "hotpink", "heuristics" = "royalblue", "neither" = "slategray"))

p_box_0

```

```{r}
## order BC in x axis by contrast coefficient
barcode_order <- long_FC_top10each_30 %>%
  group_by(Barcode) %>%
  summarise(mean_LFC = mean((testing_coeff), na.rm = TRUE)) %>%
  arrange(mean_LFC) %>%
  pull(Barcode)

long_FC_top10each_30$identifiedBy <- factor(
  long_FC_top10each_30$identifiedBy,
  levels = c("barbieQ", "heuristics", "both", "neither")
)

p_box_coef <- ggplot(long_FC_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = log2(FC), group = (Barcode), color = identifiedBy)) +
  geom_boxplot(varwidth = T) +
  # geom_jitter()+
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "top") +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "snow4") +
  labs(x = "",
       y = "Log2(FC) in proportion at each time") +
  scale_color_manual(
    values = c("both" = "slateblue", "barbieQ" = "hotpink", "heuristics" = "skyblue", "neither" = "slategray"),
    drop = FALSE)

p_box_coef_2 <- ggplot(long_FC_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = testing_coeff, group = (Barcode), color = identifiedBy)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "none") +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "snow4") +
  labs(x = "Barocdes ordered by barbieQ-estimated coefficient for the contrast",
       y = "Contrast coefficient") +
  scale_color_manual(values = c("both" = "slateblue", "barbieQ" = "hotpink", "heuristics" = "skyblue", "neither" = "slategray"))

p_box_coef / p_box_coef_2 + plot_layout(heights = c(3, 1))

ggsave("../../output/WuC/LFC_box.png", plot = p_box_coef / p_box_coef_2 + plot_layout(heights = c(3, 1)), 
       dpi = 300, width = 10, height = 6)
```


```{r}
ggplot(long_FC_top10each_30, aes(x = identifiedBy, y = log2(FC), group = (Barcode), color = identifiedBy)) +
  geom_boxplot(varwidth = F) +
  # scale_y_continuous(
  #   trans = "log2",
  #   breaks = trans_breaks("log2", function(x) 2^x),
  #   labels = trans_format("log2", math_format(2^.x))
  # ) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  theme_minimal()
```

# 1k BC (99%, n=6, from 53 samples)

```{r}
dim(monkeyHSPC_top99_30)
##need to take Time into account
designs <- monkeyHSPC_top99_30$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>%
  with(
    model.matrix(~0+ Celltype + Months)
)
# colnames(designs) <- gsub(":", ".", colnames(designs))
# Celltype + Celltype:Months didn't work well

top99n6_30_test <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top99_30, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top99n6_30_test)
# check design
# top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$design
top99n6_30_test@elementMetadata$testingBarcode@metadata$contrasts
```

## Validation

```{r}
plot_validation(top99n6_30_test)

ggsave(filename = "../../output/WuC/validation_top99n6_30.png", 
       plot = plot_validation(top99n6_30_test), 
       dpi = 300, width = 6, height = 5,)

## plot all 1k BC
top99n6_30_all <- top99n6_30_test
top99n6_30_all@elementMetadata$testingBarcode$direction <- 1
plot_validation(top99n6_30_all) + ylim(0, 1)

```
## Venn filter

```{r}
## signif. BC
filter_venn_list <- list(
  ## signif. BC
  top10each = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  barbieQ = rownames(top99n6_30_test)[top99n6_30_test@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  ## all BC in test
  top10each_all = rownames(top10each_30_test30),
  barbieQ_filter_all = rownames(top99n6_30_test))

lapply(filter_venn_list, length)

venn_filter <- eulerr::euler(filter_venn_list[1:2])

plot(venn_filter, fills = c("yellowgreen", "salmon"), alpha = 0.7, labels = T, quantities = T)

## save
png("../../output/WuC/venn_filter.png", width = 1500, height = 1000, res = 300)
# Plot
plot(
  venn_filter,
  fills = c("white", "white"),
  alpha = 0.5,
  labels = TRUE,
  quantities = TRUE
)
# Close the device
dev.off()



venn_filter <- eulerr::euler(filter_venn_list[1:2])

plot(venn_filter, fills = c("grey", "white"), alpha = 0.5, labels = T, quantities = T)
## total BC
venn_filter <- eulerr::euler(filter_venn_list)
plot(venn_filter, fills = c("grey","grey", "white", "white"), alpha = 0.5, labels = T, quantities = T)

venn_filter <- eulerr::euler(filter_venn_list[c(1,2,4)])
plot(venn_filter, fills = c("grey","grey", "white", "white"), alpha = 0.5, labels = T, quantities = T)
```

## Boxplot FC

```{r}
BC_filter_sig <- filter_venn_list[1:2] %>% unlist %>% unique()

FC_filter_sig <- calc_bias_LFC(barbieQ = monkeyHSPC[BC_filter_sig,1:30], get_FC = T) %>% as.data.frame()
Mean_filter_sig <- calc_bias_LFC(barbieQ = monkeyHSPC[BC_filter_sig,1:30], get_AMean = T) %>% as.data.frame()
rownames(FC_filter_sig) <- BC_filter_sig
## annotating BC by barbieQ seletion
FC_filter_sig$top10each = 0
FC_filter_sig$bbq = 0
## annotating BC by heuristic selection
FC_filter_sig[filter_venn_list[[1]], "top10each"] = 1
FC_filter_sig[filter_venn_list[[2]], "bbq"] = 1
  ## adding BC name
FC_filter_sig <- FC_filter_sig %>% 
  mutate(Barcode = BC_filter_sig) %>% 
  ## summarizing two selecting methods overlaps
  mutate(FilteringMethod = ifelse(top10each == 1, 
                                  ifelse(bbq == 1, "both", "top10each"),
                                  ifelse(bbq == 1, "barbieQ", "neither"))) %>% 
  dplyr::select(-c(top10each, bbq))

## extract FC from testing
# coff_FC <- top10each_30_test30@elementMetadata$testingBarcode$meanDiff
# FC_filter_sig$testing_coeff <- coff_FC

long_FC_filter_sig <- FC_filter_sig %>% 
  mutate(Barcode = BC_filter_sig) %>% 
  pivot_longer(
      cols = -c(Barcode, FilteringMethod),     # keep Celltype and Months fixed
      names_to = "Month",
      values_to = "FC"
    )
# Add barcode column to Mean_filter_sig
Mean_filter_sig <- Mean_filter_sig %>%
  mutate(Barcode = BC_filter_sig)

long_Mean <- Mean_filter_sig %>%
  pivot_longer(
    cols = -Barcode,
    names_to = "Month",
    values_to = "Mean"
  )

long_FC_filter_sig <- long_FC_filter_sig %>%
  left_join(long_Mean, by = c("Barcode", "Month"))

```

```{r fig.width=10, fig.height=5}

barcode_order <- long_FC_filter_sig %>%
  group_by(Barcode) %>%
  summarise(AMean = mean(Mean, na.rm = TRUE)) %>%
  arrange(AMean) %>%
  pull(Barcode)

barcode_order_df <- long_FC_filter_sig %>%
  group_by(Barcode, FilteringMethod) %>%
  summarise(AMean = mean(Mean, na.rm = TRUE)) %>%
  arrange(AMean)

p_AMean_0 <- ggplot()+
  geom_point(data = barcode_order_df,
             aes(x = factor(Barcode, levels = barcode_order),
                 y = log10(AMean),
                 color = FilteringMethod),
             size = 1.5, shape = 15,
             inherit.aes = FALSE) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "none") +
  labs(x = "Barocdes ordered by mean proportion across cell types across time (Amean)",
       y = "Log10 (Amean)") +
  scale_color_manual(values = c("both" = "royalblue", "barbieQ" = "salmon", "top10each" = "yellowgreen"))


p_box_0 <- ggplot(long_FC_filter_sig, aes(x = factor(Barcode, levels = barcode_order), y = log2(FC), group = (Barcode), color = FilteringMethod)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "top") +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  labs(x = "",
       y = "Log2(FC) in proportion at each time") +
  scale_color_manual(values = c("both" = "royalblue", "barbieQ" = "salmon", "top10each" = "yellowgreen"))

p_box_0 / p_AMean_0 + plot_layout(heights = c(3, 1))


ggsave("../../output/WuC/LFC_box_filter.png", plot = p_box_0 / p_AMean_0 + plot_layout(heights = c(3, 1)), 
       dpi = 300, width = 10, height = 6)

```



