---
title: "case_study"
author: "LF"
date: "2025-06-19"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(proxy) ## for dist
library(circlize)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_ZG66_allPhases_barbieQ.rda")
```


```{r}
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "early",]
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "late",]
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

# Function: Top10BC per sample

```{r filtering original}
tagTop10 <- function(col, n=10) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= n
  # summary(isTop)
  return(isTop)
}


```

# Function: Validation 

```{r function validation}
plot_validation <- function(myObject) {
  
  flag <- rowData(myObject)$testingBarcode$direction == 1
  myObject <- myObject[flag,]
  
  NK_clone_contributionU <- assays(myObject)$proportion %>% colSums()
  dat <- data.frame(
    contr = NK_clone_contributionU,
    Celltype = factor(myObject$sampleMetadata$Celltype,
                      levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p")),
    Months = myObject$sampleMetadata$Months)
  ## sort samples by celltype and months
  dat <- dat %>%
    arrange(Celltype, Months) %>% 
    mutate(id = row_number()) 
  ## add dummy rows for spacing between celltypes
  dat <- dat %>%
    group_by(Celltype) %>%
    mutate(dummy = n() + 1) %>%    # Add a dummy count for spacing
    ungroup() %>%
    bind_rows(dat %>%             # Add dummy rows for spacing
                group_by(Celltype) %>%
                slice_tail(n = 1) %>%
                mutate(id = as.numeric(id) + 0.5, contr = NA, Months = NA)) %>%
    arrange(as.numeric(id))        # Arrange with dummy rows in place
  
  celltype_colors <- c(
  "T" = "yellowgreen",
  "B" = "yellowgreen",
  "Gr" = "yellowgreen",
  "NK_CD56p_CD16n" = "yellowgreen",
  "NK_CD56n_CD16p" = "slateblue1" # "orchid"
)
  
  p <- ggplot(dat, aes(x = as.factor(id), y = contr, color = Celltype)) +
  geom_bar(stat = "identity", fill = NA) + 
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 0.5, size = 10),
        axis.title.x = element_text(size = 14), 
        axis.title.y = element_text(size = 14), 
        panel.grid.major.x = element_blank()) +
  scale_x_discrete(labels = ifelse(is.na(dat$Months), "", dat$Months)) + # Relabel with `Celltype`
  scale_color_manual(values = celltype_colors) +  # <- set bar outline color
  scale_fill_manual(values = celltype_colors) +   # <- set tile fill color
  scale_y_continuous(limits = c(-0.09, 0.8)) +
  labs(x = "Samples of cell types at different months", y = "Total prop. of selected barcodes") +  # Update axis labels
  geom_tile(
    data = dat,
    aes(x = ifelse(is.na(Months), NA,order(id)), y = -0.03, fill = Celltype),
    height = 0.03, width = 1
  ) +
  geom_text(
    data = dat %>%
      mutate(id = order(id)) %>%
      group_by(Celltype) %>%
      mutate(Celltype = recode(Celltype, "NK_CD56n_CD16p" = "NK.CD56-CD16+")) %>% 
      mutate(Celltype = recode(Celltype, "NK_CD56p_CD16n" = "NK.CD56+CD16-")) %>% 
      summarize(mid_point = mean(as.numeric(id))),
    aes(x = mid_point-0.5, y = -0.03, label = Celltype),
    size = 3, vjust = 0.5, color = "black"
  )
  
  return(p)
}
```

# Function: line validation

```{r}
plot_line_validation <- function(barcode_set_list, barbieQ) {
  ## Example for barcode set list
  # barcode_set_list <- list(
  #   test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  #   test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  #   heuristic = union_LFC_top10each_30
  # )
  
  ## check barcode set list and barbieQ contents
  print(names(barcode_set_list))
  print(dim(barbieQ))
  
  # Create named vector with default value "none"
  selectedBy <- setNames(rep("none", nrow(barbieQ)), rownames(barbieQ))
  
  # Assign sources
  for (source_name in names(barcode_set_list)) {
    matching_barcodes <- barcode_set_list[[source_name]]
    selectedBy[matching_barcodes] <- ifelse(
      selectedBy[matching_barcodes] == "none",
      source_name,
      paste(selectedBy[matching_barcodes], source_name, sep = ",")
    )
  }
  
  ## extract props. and sample conditions
  prop. <- assays(barbieQ)$proportion
  cond. <- barbieQ$sampleMetadata %>% as.data.frame()
  
  cond.$Celltype <- factor(cond.$Celltype,
                      levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p", 
                                 "NK_NKG2Ap_CD16p", "NK_NKG2Ap_CD16p_KIR3DL01n",
                                 "NK_NKG2Ap_CD16p_KIR3DL01p"))

  ## combine slots
  prop_df <- data.frame(cond., t(prop.))
  ## keep only early and mid samples
  prop_df <- prop_df %>% 
    # dplyr::filter(Phase %in% c("early", "mid")) %>% 
    select(-Phase)
  ## classify BC by selected by test or LFC
  
  ## picot to long data
  prop_long <- prop_df %>%
    pivot_longer(
      cols = -c(Celltype, Months),     # keep Celltype and Months fixed
      names_to = "Barcode",
      values_to = "Proportion"
    )
  ## assign to barcode selectedBy status
  prop_long$selectedBy <- selectedBy[as.character(prop_long$Barcode)]
  
  ## plot
  p <- ggplot(prop_long) +
    geom_line(aes(x = Months, y = Proportion, group = Barcode, color = selectedBy)) +
    facet_grid(selectedBy ~ Celltype, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "top")
  
  return(p)
  
}
```


# Function: Calc LFC

```{r}
calc_bias_LFC <- function(barbieQ) {
  ## extract props. and sample conditions
  prop. <- assays(barbieQ)$proportion
  cond. <- barbieQ$sampleMetadata %>% as.data.frame()
  ## group samples by NKnp or others
  cond. <- cond. %>% 
    mutate(isNK = ifelse(Celltype == "NK_CD56n_CD16p", "NKnp", "others"))
  cond.$Celltype %>% table()
  
  ## calc mean per group at each month
  group_labels <- paste(cond.$Months, cond.$isNK, sep = "__")
  group_means <- sapply(unique(group_labels), function(g) {
    idx <- which(group_labels == g)
    rowMeans(prop.[, idx, drop = FALSE])
  })
  
  ## calc group LFC at each month
  colnames(group_means)
  months <- gsub("(\\1)__.*", "\\1",colnames(group_means))
  
  LFC_months <- sapply(unique(months), function(m) {
    idx <- which(months == m)
    sub <- group_means[, idx, drop = FALSE]
    group <- gsub(".*__()", "\\1",colnames(sub))
    NK <- sub[, which(group == "NKnp"), drop = F]
    others <- sub[, which(group=="others"),drop = F]
    NK / others 
  })
  
  ## decide if each barcode LFC is high at each month
  is_LFC_high <- LFC_months > 10
  colnames(is_LFC_high)
  
  ## calc if each barcode NKnp  prop > 0.01 at each month
  NK_prop <- prop.[,cond.$isNK == "NKnp", drop=F]
  is_NKnp_high <- NK_prop > 0.01
  colnames(is_NKnp_high)
  
  ## barcodes meet both metric at each month
  select_months <- (is_LFC_high & is_NKnp_high)
  
  ## select barcodes at each month
  select_barcodes_months <- lapply(colnames(select_months), function(j) 
    rownames(prop.)[select_months[,j]])
  
  names(select_barcodes_months) <- paste0("heuristic at ", colnames(is_LFC_high), "m")
  
  return(select_barcodes_months)
}

```

# Data

```{r}
# first 30 samples are included in the paper table
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,1:30])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_30 <- monkeyHSPC
rowData(top10each_30)$isTopBarcode$isTop <- colTagVec
# assays(top10each_30)$isTopAssay <- colTagMat
top10each_30 <- top10each_30[colTagVec,]

```

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,flag_early_mid])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_early_mid <- monkeyHSPC
rowData(top10each_early_mid)$isTopBarcode$isTop <- colTagVec
# assays(top10each_early_mid)$isTopAssay <- colTagMat
top10each_early_mid <- top10each_early_mid[colTagVec,]

```

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
# tag top
monkeyHSPC_top99_early_mid_0 <- tagTopBarcodes(barbieQ = monkeyHSPC[,flag_early_mid], nSampleThreshold = 6)
flag_top <- monkeyHSPC_top99_early_mid_0@elementMetadata$isTopBarcode$isTop
table(flag_top)
monkeyHSPC_top99_early_mid <- monkeyHSPC_top99_early_mid_0[flag_top,]
```

subset NKnp samples of all 8 timepoints

Choose BC by top10each (suggested by paper): 39 BCs

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

colTags_NKnp <- lapply(as.data.frame(assays(NKnp)$proportion), 
                  function(col) tagTop10(col, n = 10))
colTagMat_NKnp <- do.call(cbind, colTags_NKnp)
colTagVec_NKnp <- rowSums(colTagMat_NKnp) >= 1
sum(colTagVec_NKnp) 
all(colSums(colTagMat_NKnp) == 15)

top10each_NKnp <- monkeyHSPC[colTagVec_NKnp,]

## same data used by paper
NKnp <- NKnp[colTagVec_NKnp,]
```


# Top10BC, 39BC from 8 NKnp samples

## HP clustering - paper analysis

### clustering approach


hierarchically clustering barcodes and extract the clusters

```{r}
NKnp_slot <- NKnp@assays@data$proportion
NKnp_hp <- Heatmap(log(NKnp_slot+1e-6), show_row_names = F, 
        column_order = order(NKnp$sampleMetadata$Months),
        col = colorRamp2(
    c(log(1+1e-6), log(0.01+1e-6), log(1e-6)),                     # log10 values for 1, 0.01, 0.000025
    c("red", "white", "blue")             # choose any colors you prefer
  ))

# NKnp_hp
```

```{r}
# 3. Extract row dendrogram and cut into 3 clusters
NKnp_row_dend <- row_dend(NKnp_hp)
NKnp_clusters <- cutree(as.hclust(NKnp_row_dend), k = 5)

# 5. Create row annotation based on clusters
row_ha <- rowAnnotation(
  NKnp_clusters = as.factor(NKnp_clusters),
  col = list(NKnp_clusters = structure(
    RColorBrewer::brewer.pal(5, "Set1"),
    names = as.character(1:5)
  )),
  show_annotation_name = TRUE
)

# 6. Plot heatmap with annotation
ht_with_anno <- Heatmap(
  log(NKnp_slot + 1e-6),
  show_row_names = FALSE, column_labels = NKnp$sampleMetadata$Months,
  column_order = order(NKnp$sampleMetadata$Months),
  # col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  left_annotation = row_ha,
  row_split = as.factor(NKnp_clusters)
)

ht_with_anno
```

### barcodetrackR source code

!!! this equals to the default clustering and algorithms in ComplexHeatmap

clustering_data <- SummarizedExperiment::assays(your_SE)[["logs"]]

log(NKnp_slot+1e-6)

clustering_data.dist <- proxy::dist(clustering_data, method = distance_method, p = minkowski_power)
hclustering <- hclust(clustering_data.dist, method = hclust_linkage)
clusters = cutree(hclustering, clusters)

```{r}
# Compute matrix and clustering
NKnp_slot <- NKnp@assays@data$proportion
log_mat <- log(NKnp_slot + 1e-6)

# Your existing clustering:
clustering_data.dist <- proxy::dist(log_mat, method = "euclidean", p = 2)
hclustering <- hclust(clustering_data.dist, method = "complete")
clusters <- cutree(hclustering, 5)
table(clusters)

# Make sure cluster vector is a factor in the same order as the matrix rows
row_split <- factor(clusters[rownames(log_mat)], levels = 1:5)

# Plot the heatmap with row_split
Heatmap(
  log_mat,
  show_row_names = FALSE,
  column_labels = NKnp$sampleMetadata$Months,
  column_order = order(NKnp$sampleMetadata$Months),
  # col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  row_split = row_split
)
```

## Test NKnp, early_mid, top10each_NKnp

```{r}
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")


##need to take Time into account
designs <- top10each_NKnp[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


top10each_NKnp_test_time_early_mid <- testBarcodeSignif(
  barbieQ = top10each_NKnp[,flag_early_mid], designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## Test NKnp, NKnp only, top10each_NKnp - Need to fix pkg here!!!

```{r eval=FALSE, include=FALSE}
# NKnp <- monkeyHSPC[,monkeyHSPC_top$sampleMetadata$Celltype == "NK_CD56n_CD16p"]
# 
# designs <- NKnp$sampleMetadata %>% as.data.frame() %>%
#   with(
#     model.matrix(~0+ Months)
# )
# 
# asinTrans_NKnp_testTime <- testBarcodeSignif(
#   barbieQ = NKnp, ,sampleMetadata = NKnp$sampleMetadata, sampleGroup = "Months",
#   method = "diffProp", transformation = "asin-sqrt"
# )
```

## Compare

### Venn

```{r}
cluster_selected_list <- lapply(1:5, function(i) names(NKnp_clusters)[NKnp_clusters == i])
names(cluster_selected_list) <- paste("cluster_", 1:5, sep = "_")

seletced_barbieQ_NKnp_up <- rowData(top10each_NKnp_test_time_early_mid)$testingBarcode$direction == 1
seletced_barbieQ_NKnp_down <- rowData(top10each_NKnp_test_time_early_mid)$testingBarcode$direction == -1

cluster_selected_list[["barbieQ_NKnp_up"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_up]
cluster_selected_list[["barbieQ_NKnp_down"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_down]

venn99 <- eulerr::euler(
  cluster_selected_list
)

plot(venn99, fills = c("white", "white","white","white","white","grey","grey"), alpha = 0.5, labels = T, quantities = T)


```

### Line plot

```{r, fig.width=10, fig.height=12}
plot_line_validation(cluster_selected_list[c(1,2,3,5,7)], top10each_NKnp_test_time_early_mid)
```

```{r, fig.width=10, fig.height=6}
plot_line_validation(cluster_selected_list[c(4,6)], top10each_NKnp_test_time_early_mid)
```
# More BC

## Test NKnp, early_mid, 132 BC from 53 samples of all TP all celltype

```{r}
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")

##need to take Time into account
designs <- top10each_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


top10each_early_mid_test_time_early_mid <- testBarcodeSignif(
  barbieQ = top10each_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```

## Test, 1.6k BC

```{r}
dim(monkeyHSPC_top99_early_mid)
##need to take Time into account
designs <- monkeyHSPC_top99_early_mid$sampleMetadata %>% as.data.frame() %>%
  with(
    model.matrix(~0+ Celltype + Celltype:Months)
)
colnames(designs)<- gsub(":", replacement = ".", colnames(designs))


top99n6_early_mid_test_time_early_mid <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top99_early_mid, designMatrix = designs,
  contrastFormula = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)
```


## Compare

### Venn

```{r}
# cluster_selected_list <- lapply(1:5, function(i) names(NKnp_clusters)[NKnp_clusters == i])
# names(cluster_selected_list) <- paste("cluster_", 1:5, sep = "_")

NKnp_up_132BC <- top10each_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_132BC <- top10each_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == -1

NKnp_up_39BC <- top10each_NKnp_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_39BC <- top10each_NKnp_test_time_early_mid@elementMetadata$testingBarcode$direction == -1

NKnp_up_1.6kBC <- top99n6_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == 1
NKnp_down_1.6kBC <- top99n6_early_mid_test_time_early_mid@elementMetadata$testingBarcode$direction == -1


cluster_selected_list[["barbieQ_NKnp_up"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_up]
cluster_selected_list[["barbieQ_NKnp_down"]] <- rownames(top10each_NKnp_test_time_early_mid)[seletced_barbieQ_NKnp_down]

list_test_NK_early_mid <- list(
    NKnp_up_132BC = rownames(top10each_early_mid_test_time_early_mid)[NKnp_up_132BC],
    NKnp_down_132BC = rownames(top10each_early_mid_test_time_early_mid)[NKnp_down_132BC],
    NKnp_up_39BC = rownames(top10each_NKnp_test_time_early_mid)[NKnp_up_39BC],
    NKnp_down_39BC = rownames(top10each_NKnp_test_time_early_mid)[NKnp_down_39BC],
    NKnp_up_1.6kBC = rownames(top99n6_early_mid_test_time_early_mid)[NKnp_up_1.6kBC],
    NKnp_down_1.6kBC = rownames(top99n6_early_mid_test_time_early_mid)[NKnp_down_1.6kBC]
  )

venn_test_NK_early_mid <- eulerr::euler(
  list_test_NK_early_mid[1:4]
)

plot(venn_test_NK_early_mid, fills = c("white", "white","grey","grey"), alpha = 0.5, labels = T, quantities = T)


```
```{r}
venn_test_NK_early_mid <- eulerr::euler(
  list_test_NK_early_mid
)

plot(venn_test_NK_early_mid, fills = c("white", "white","grey","grey", "grey2", "grey2"), alpha = 0.5, labels = T, quantities = T)
```

## Line plot

```{r}
plot_line_validation(list_test_NK_early_mid[c(1,3)], top10each_early_mid[,flag_early_mid])
plot_line_validation(list_test_NK_early_mid[c(2,4)], top10each_early_mid[,flag_early_mid])
```

```{r, fig.width=10, fig.height=6}
plot_line_validation(list_test_NK_early_mid[c(1,3,5)], monkeyHSPC[flag_top,])
plot_line_validation(list_test_NK_early_mid[c(2,4,6)], monkeyHSPC[flag_top,])
```

