---
title: "case_study"
author: "LF"
date: "2025-06-18"
output: workflowr::wflow_html
output_dir: ../../docs/WuC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(circlize)
library(limma)

library(SummarizedExperiment)
library(S4Vectors)
library(barbieQ)
```

# Load Processed Data

```{r load rda}
getwd()
load("../../output/WuC/monkeyHSPC_ZG66_allPhases_barbieQ.rda")
```


```{r}
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "early",]
monkeyHSPC$sampleMetadata[monkeyHSPC$sampleMetadata$Phase == "late",]
```

 - Loading the barcodes selected by the paper for `NK_CD56n_CD16p biased` using the paper's strategy.

```{r load xlsx}
## load dataset provided by the paper corresponding to specific figures
paper_data <- read_excel("../../data/WuC/aat9781_table_s2.xlsx", sheet = "Figure 2 ZG66")
paper_data <- as.data.frame(paper_data)
rownames(paper_data) <- paper_data[[1]]
ZG66_Fig2 <- paper_data[,-1]

```

# Source FUN

```{r}
getwd()
source("FUN_validation.R")
```

# Data

Same 81 BC in paper Fig2

```{r}
# first 30 samples are included in the paper table
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,1:30])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_30 <- monkeyHSPC
rowData(top10each_30)$isTopBarcode$isTop <- colTagVec
# assays(top10each_30)$isTopAssay <- colTagMat
top10each_30 <- top10each_30[colTagVec,]

```

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
colTags <- lapply(as.data.frame(assays(monkeyHSPC[,flag_early_mid])$proportion), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
sum(colTagVec) 
sum(colTagVec) / nrow(monkeyHSPC)
dim(colTagMat)
all(colSums(colTagMat) == 10)

top10each_early_mid <- monkeyHSPC
rowData(top10each_early_mid)$isTopBarcode$isTop <- colTagVec
# assays(top10each_early_mid)$isTopAssay <- colTagMat
top10each_early_mid <- top10each_early_mid[colTagVec,]

```

subset NKnp samples of all 8 timepoints

Choose BC by top10each (suggested by paper): 39 BCs

Same 39 BC in paper Fig3

```{r}
NKnp <- monkeyHSPC[,monkeyHSPC$sampleMetadata$Celltype == "NK_CD56n_CD16p"]

colTags_NKnp <- lapply(as.data.frame(assays(NKnp)$proportion), 
                  function(col) tagTop10(col, n = 10))
colTagMat_NKnp <- do.call(cbind, colTags_NKnp)
colTagVec_NKnp <- rowSums(colTagMat_NKnp) >= 1
sum(colTagVec_NKnp) 
all(colSums(colTagMat_NKnp) == 15)

top10each_NKnp <- monkeyHSPC[colTagVec_NKnp,]

## same data used by paper
NKnp <- NKnp[colTagVec_NKnp,]
```

monkeyHSPC_top99_early_mid saved in the rda object - no need to run again.

```{r eval=FALSE, include=FALSE}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
# tag top
monkeyHSPC <- tagTopBarcodes(barbieQ = monkeyHSPC, activeSamples = flag_early_mid, nSampleThreshold = 6)
flag_top <- monkeyHSPC@elementMetadata$isTopBarcode$isTop
table(flag_top)
monkeyHSPC_top99_early_mid <- monkeyHSPC[flag_top,]

```

## Save to rda

```{r}

save(
  monkeyHSPC, monkeyHSPC_top99_early_mid, file = "../../output/WuC/cluster_pca_facet.pngmonkeyHSPC_ZG66_allPhases_barbieQ.rda"
)

```


## Paper table

Downloaded table from the original paper contains 81 barcodes and 30 samples. Barcodes selected using the "top10each" approach based on the 30 samples from the original ZG66 dataset, monkeyHSPC, perfectly overlap with the 81 barcodes in the downloaded table.

However, when samples of earlier or later months are included, more barcodes will be selected.

```{r overlap}
venn2 <- eulerr::euler(
  list(
    top10each = rownames(top10each), 
    top10each_early_mid = rownames(top10each_early_mid), 
    top10each_30 = rownames(top10each_30), 
    Paper_select = rownames(ZG66_Fig2)))

plot(venn2, fills = c("white", "grey"), alpha = 0.5, labels = TRUE, quantities = TRUE)
```

## Heatmap by Phase

Quite a few shared barcodes between early+mid phases and late phase where samples are mature NK subtypes.

```{r}
plotBarcodeHeatmap(top10each, splitSamples = T, sampleGroup = "Phase")
```

# Paper 81 barcodes (top10each from 30 samples)

- My selecting strategy to select `NK_CD56n_CD16p biased barcodes` by testing

 - [1] "(CelltypeNK_CD56n_CD16p)" v.s.                                      

 - [2] "(CelltypeB + CelltypeGr + CelltypeT + CelltypeNK_CD56p_CD16n)/4"

- Only keep the samples of timepoints at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m. (to be consistent with the paper's analysis)

## LFC approach

(i) >1% contribution to CD56−CD16+ cells and 

(ii) >10-fold abundance in contribution to CD56−CD16+ cells compared with T, B, Gr, and CD56+CD16− cells.

```{r}
list_BC_top10each_30_LFC <- calc_bias_LFC(top10each_30_test30)
```

### Venn: LFC at each month

```{r}
## check overlaps between months
venn0 <- eulerr::euler(list_BC_top10each_30_LFC)
plot(venn0, fills = c("white"), alpha = 0.5, labels = T, quantities = T)
```


## Test 30 samples

```{r checkcout df after contrast}
## For time points at 6.5, 12.0, 17.0, 27.0, 36.0, 42.0 m, each collecting all cell types.
## These samples refer to the first 30 samples in the object, which is also in line with the samples in the table downloaded from the paper's materials.
dim(top10each_30)

##need to take Time into account
designs <- top10each_30[,1:30]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_30_test30 <- testBarcodeSignif(
  barbieQ = top10each_30[,1:30], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
# check design
# top10each_30_test30@elementMetadata$testingBarcode@metadata$design
top10each_30_test30@elementMetadata$testingBarcode@metadata$contrasts
dim(top10each_30_test30)



## test Barcode differential occurrence between sample groups
## set up the targets (sample conditions)
# targets <- top10each_30$sampleMetadata %>% 
#   as.data.frame() %>% 
#   mutate(Group = ifelse(
#     Celltype == "NK_CD56n_CD16p",
#     "NK_CD56n_CD16p",
#     "B.Gr.T.NK_CD56p_CD16n"))
# 
# occTest <- testBarcodeSignif(
#   barbieQ = top10each_30, 
#   sampleMetadata = targets[,"Group", drop=FALSE], 
#   method = "diffOcc"
# )
```


## Test early+mid

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")
dim(top10each_30)

top10each_30$sampleMetadata %>% as.data.frame() %>% View()

##need to take Time into account
designs <- top10each_30[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

## find time points with only 1 sample
iso_times <- colSums(designs) == 1
## locate samples causing these time points
iso_samples <- rowSums(designs[, iso_times]) == 1

dim(designs)
dim(designs[!iso_samples,!iso_times])

dim(top10each_30[,flag_early_mid])
dim(top10each_30[,flag_early_mid][,!iso_samples])

top10each_30_test_ealy_mid <- testBarcodeSignif(
  barbieQ = top10each_30[,flag_early_mid][,!iso_samples], designMatrix = designs[!iso_samples,!iso_times],
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top10each_30_test_ealy_mid)
# check design
# top10each_30_test_ealy_mid@elementMetadata$testingBarcode@metadata$design %>% View()
top10each_30_test_ealy_mid@elementMetadata$testingBarcode@metadata$design %>% colSums()
top10each_30_test_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```

## Compare test vs. LFC

### Venn

```{r}
## check overlaps
compare_list <- list_BC_top10each_30_LFC 
compare_list[["test_30sample"]] <- rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"]
compare_list[["test_early_mid"]] <- rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"]

venn00 <- eulerr::euler(compare_list[1:7])

plot(venn00, fills = c("grey", "grey", "grey", "grey", "grey", "grey", "white"), alpha = 0.5, labels = T, quantities = T)
```


```{r}
union_LFC_top10each_30 <- unlist(list_BC_top10each_30_LFC) %>% unique()

list_top10each_30 <- list(
    test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    union_heuristic = union_LFC_top10each_30)

venn_top10each_30 <- eulerr::euler(
  list_top10each_30[c(1,3)]
)

plot(venn_top10each_30 , fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)


venn_top10each_30 <- eulerr::euler(
  list_top10each_30[c(1,2)]
)

plot(venn_top10each_30 , fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)
```

### Validation: LFC at each month

```{r}
plot_list_LFC_top10each_30 <- lapply(list_BC_top10each_30_LFC, function(m) plot_validation(top10each_30_test30[m,]))

plot_list_LFC_top10each_early_mid <- lapply(list_BC_top10each_30_LFC, function(m) plot_validation(top10each_30_test_ealy_mid[m,]))
```


```{r, fig.width=12, fig.height=8}
suppressWarnings(wrap_plots(plot_list_LFC_top10each_30, ncol = 3))
```

```{r, fig.width=18, fig.height=8}
suppressWarnings(wrap_plots(plot_list_LFC_top10each_early_mid, ncol = 3))

```


```{r}
top10each_30_LFC <- top10each_30[union_LFC_top10each_30, 1:30]
dim(top10each_30_LFC)
rowData(top10each_30_LFC)$testingBarcode$direction = 1

top10each_early_mid_LFC <- top10each_early_mid[union_LFC_top10each_30, flag_early_mid]
rowData(top10each_early_mid_LFC)$testingBarcode$direction = 1

top10each_30_allBC <- top10each_30[, 1:30]
dim(top10each_30_allBC)
rowData(top10each_30_allBC)$testingBarcode$direction = 1

plot_validation(top10each_30_LFC) 
plot_validation(top10each_early_mid_LFC)
plot_validation(top10each_30_allBC)

```

```{r, fig.width=12, fig.height=5}
plot_validation(top10each_30_LFC) + plot_validation(top10each_30_test30)

```

### Line plot

```{r, fig.width=8, fig.height=7}
plot_line_validation(
  barcode_set_list = list_top10each_30[c(1,3)], 
  barbieQ = top10each_30[,flag_early_mid])

```

#### by time


```{r, fig.width=15, fig.height=5}
p_by_Time <- plot_line_validation(
  barcode_set_list = list_top10each_30[c(1,3)], barbieQ = top10each_30[,flag_early_mid], get_long_data = T)

suppressMessages({
  ggplot(p_by_Time) +
  # geom_point(aes(x = Celltype, y = Proportion, group = Barcode, color = selectedBy)) +
  geom_line(aes(x = Celltype, y = Proportion, group = Barcode, color = selectedBy)) +
    facet_grid(selectedBy~Months, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "top", axis.text.x = element_text(angle = 45, hjust = 1))
})


```

```{r, fig.width=15, fig.height=15}
p_by_Time <- plot_line_validation(
  barcode_set_list = list_top10each_30[c(1,3)], barbieQ = top10each_30[,flag_early_mid], get_long_data = T)

suppressMessages({
  ggplot(p_by_Time %>% filter(selectedBy == "test_30sample")) +
    # geom_point(aes(x = Celltype, y = Proportion, group = Barcode, color = selectedBy)) +
  geom_line(aes(x = Celltype, y = Proportion, group = Barcode, color = selectedBy)) +
    facet_grid(Barcode~Months, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "top", axis.text.x = element_text(angle = 45, hjust = 1))
})

suppressMessages({
  ggplot(p_by_Time %>% filter(selectedBy == "test_30sample,union_heuristic")) +
    geom_point(aes(x = Celltype, y = Proportion, group = Barcode, color = selectedBy), size = 1) +
  geom_line(aes(x = Celltype, y = Proportion, group = Barcode, color = selectedBy)) +
    facet_grid(Barcode~Months, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "top", axis.text.x = element_text(angle = 45, hjust = 1))
})

```

### Boxplot FC

```{r}
FC_top10each_30 <- calc_bias_LFC(barbieQ = top10each_30[,1:30], get_FC = T) %>% as.data.frame()
rownames(FC_top10each_30) <- rownames(top10each_30)
## annotating BC by barbieQ seletion
FC_top10each_30$bbq <- top10each_30_test30@elementMetadata$testingBarcode$direction
FC_top10each_30$bbq2 <- top10each_30_test_ealy_mid@elementMetadata$testingBarcode$direction
## annotating BC by heuristic selection
FC_top10each_30$heu = 0
FC_top10each_30[union_LFC_top10each_30, "heu"] = 1
  ## adding BC name
FC_top10each_30 <- FC_top10each_30 %>% 
  mutate(Barcode = rownames(top10each_30)) %>% 
  ## summarizing two selecting methods overlaps
  # mutate(selectedBy = ifelse(bbq2 == 1, "bbQ_smp+", "")) %>% 
  # mutate(selectedBy = ifelse(bbq == 1, "bbq", ifelse(bbq2 == 1, "bbQ_smp+", ""))) %>% 
  mutate(selectedBy = ifelse(bbq == 1, "bbq", "")) %>% 
  mutate(selectedBy = paste(selectedBy, ifelse(heu == 1, "heu", ""), sep = "_")) %>% 
  select(-c(heu, bbq, -bbq2))

## extract FC from testing
coff_FC <- top10each_30_test30@elementMetadata$testingBarcode$meanDiff
FC_top10each_30$testing_coeff <- coff_FC

long_FC_top10each_30 <- FC_top10each_30 %>% 
  mutate(Barcode = rownames(top10each_30)) %>% 
  pivot_longer(
      cols = -c(Barcode, selectedBy, testing_coeff),     # keep Celltype and Months fixed
      names_to = "Month",
      values_to = "FC"
    )

# Calculate number of unique barcodes per selectedBy
barcode_counts <- long_FC_top10each_30 %>%
  group_by(selectedBy) %>%
  summarise(n_barcodes = n_distinct(Barcode))

ggplot(long_FC_top10each_30, aes(x = as.numeric(Month), y = log2(FC), group = paste(Month, selectedBy), color = selectedBy)) +
  geom_boxplot()

```

```{r fig.width=10, fig.height=5}

barcode_order <- long_FC_top10each_30 %>%
  group_by(Barcode) %>%
  summarise(mean_LFC = mean(log2(FC), na.rm = TRUE)) %>%
  arrange(mean_LFC) %>%
  pull(Barcode)

p_box_0 <- ggplot(long_FC_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = log2(FC), group = (Barcode), color = selectedBy)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank()) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  labs(x = "Barocdes ordered by mean LFC across time",
       y = "Barcode proportion log2(FC) at each time")

barcode_order <- long_FC_top10each_30 %>%
  group_by(Barcode) %>%
  summarise(mean_LFC = mean((testing_coeff), na.rm = TRUE)) %>%
  arrange(mean_LFC) %>%
  pull(Barcode)

p_box_coef <- ggplot(long_FC_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = log2(FC), group = (Barcode), color = selectedBy)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "top") +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  labs(x = "Barocdes ordered by barbieQ-estimated coefficient for the contrast",
       y = "Barcode proportion log2(FC) at each time")

p_box_coef_2 <- ggplot(long_FC_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = testing_coeff, group = (Barcode), color = selectedBy)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank()) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  labs(x = "Barocdes ordered by barbieQ-estimated coefficient for the contrast",
       y = "barbieQ modelled difference in contrast")

ggsave("../../output/WuC/LFC_box.png", plot = p_box_coef, 
       dpi = 300, width = 6, height = 4)

```

```{r}
ggplot(long_FC_top10each_30, aes(x = selectedBy, y = log2(FC), group = (Barcode), color = selectedBy)) +
  geom_boxplot(varwidth = F) +
  # scale_y_continuous(
  #   trans = "log2",
  #   breaks = trans_breaks("log2", function(x) 2^x),
  #   labels = trans_format("log2", math_format(2^.x))
  # ) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  theme_minimal()
```


```{r}
mean_top10each_30 <- calc_bias_LFC(barbieQ = top10each_30[,1:30], get_Mean = T) %>% as.data.frame()

meanNKnp_top10each_30 <- mean_top10each_30[,7:12]
meanNKnp_top10each_30$selectedBy <- FC_top10each_30$selectedBy

long_meanNKnp_top10each_30 <- meanNKnp_top10each_30 %>% 
  mutate(Barcode = rownames(top10each_30)) %>% 
  pivot_longer(
      cols = -c(Barcode, selectedBy),     # keep Celltype and Months fixed
      names_to = "Month",
      values_to = "NKnp_prop"
    )

barcode_order <- long_meanNKnp_top10each_30 %>%
  group_by(Barcode) %>%
  summarise(mean_NKnp_prop = mean(NKnp_prop, na.rm = TRUE)) %>%
  arrange(mean_NKnp_prop) %>%
  pull(Barcode)

ggplot(long_meanNKnp_top10each_30, aes(x = factor(Barcode, levels = barcode_order), y = NKnp_prop, group = (Barcode), color = selectedBy)) +
  geom_boxplot(varwidth = T) +
  theme(axis.text.x = element_blank())

ggplot(long_meanNKnp_top10each_30, aes(x = selectedBy, y = NKnp_prop, group = (Barcode), color = selectedBy)) +
  geom_boxplot(varwidth = F) +
  geom_abline(slope = 0, intercept = 0.01)
```


## Compare two tests

### Venn

```{r}
venn_top10each_30 <- eulerr::euler(
  list_top10each_30[c(1,2)]
)

plot(venn_top10each_30 , fills = c("white", "white","grey"), alpha = 0.5, labels = T, quantities = T)
```

### Validation

```{r, fig.width=15, fig.height=5}

plot_validation(top10each_30_test30) + 
plot_validation(top10each_30_test_ealy_mid)
```

### Line plot: test early+mid vs LFC

indicate each barcode is selected by test_30sample, test_early-mid, or LFC

```{r}
# Define barcode sets
barcode_list <- list(
  test_30sample = rownames(top10each_30_test30)[top10each_30_test30@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  test_early_mid = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  heuristic = union_LFC_top10each_30
)

dim(top10each_early_mid)
```


```{r}
plot_line_validation(
  barcode_set_list = list_top10each_30[1:2], 
  barbieQ = top10each_early_mid[,top10each_early_mid$sampleMetadata$Phase != "late"])
```

# More BC, using same 53 samples of all TP

## Test: 132 BC (top10each from 53 samples)

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- top10each_early_mid$sampleMetadata$Phase %in% c("early", "mid")
dim(top10each_early_mid)

##need to take Time into account
designs <- top10each_early_mid[,flag_early_mid]$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(
    model.matrix(~0+ Celltype + Months)
)

top10each_early_mid_test_ealy_mid <- testBarcodeSignif(
  barbieQ = top10each_early_mid[,flag_early_mid], designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top10each_early_mid_test_ealy_mid)
# check design
# top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$design
top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```

## Test bbQ BC (99%, n=6, from 53 samples)

```{r}
# include all early and mid phase samples (discussed in Fig3, S4 for time course)
flag_early_mid <- monkeyHSPC$sampleMetadata$Phase %in% c("early", "mid")

##need to take Time into account
designs <- monkeyHSPC_top99_early_mid$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>%
  with(
    model.matrix(~0+ Celltype + Months)
)
# colnames(designs) <- gsub(":", ".", colnames(designs))
# Celltype + Celltype:Months didn't work well

top99n6_early_mid_test_ealy_mid <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top99_early_mid, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)

dim(top99n6_early_mid_test_ealy_mid)
# check design
# top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$design
top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode@metadata$contrasts
```


## Compare

### Total prop

```{r}
plot_validation(top10each_30_test_ealy_mid)
plot_validation(top10each_early_mid_test_ealy_mid)
plot_validation(top99n6_early_mid_test_ealy_mid[,flag_early_mid])
```

### Venn

```{r}
list_test_early_mid <- list(
    top10each_30 = rownames(top10each_30_test_ealy_mid)[top10each_30_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    top10each_early_mid = rownames(top10each_early_mid_test_ealy_mid)[top10each_early_mid_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    top99n6 = rownames(top99n6_early_mid_test_ealy_mid)[top99n6_early_mid_test_ealy_mid@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"]
    ## to chech if the missing one is filtered out: no, it's included
    # all_top99n6 = rownames(top99n6_early_mid_test_ealy_mid)
    )
```


```{r}

venn_test_early_mid <- eulerr::euler(
  list_test_early_mid[c(1,2)]
)

plot(venn_test_early_mid, fills = c("grey", "white"), alpha = 0.5, labels = T, quantities = T)
```

```{r}

venn_test_early_mid <- eulerr::euler(
  list_test_early_mid
)

plot(venn_test_early_mid, fills = c("grey", "white","white"), alpha = 0.5, labels = T, quantities = T)
```

### Line plot

```{r}
plot_line_validation(list_test_early_mid[1:2], top10each_early_mid[,flag_early_mid])
```

```{r, fig.width=10, fig.height=5}
plot_line_validation(list_test_early_mid, monkeyHSPC_top99_early_mid[,flag_early_mid])
```


