---
title: "estimate_coeff"
author: "LF"
date: "2025-07-30"
output: workflowr::wflow_html
output_dir: ../../DEBRA
---

# Load Dependencies

```{r load library, message=FALSE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(plotROC)

library(limma)
library(edgeR)
library(car)

library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
library(bartools)
```

# Install barbieQ - devel

```{r install barbieQ}
# devtools::install_github("Oshlack/barbieQ", force = TRUE)

library(barbieQ)
```

# Load rds

```{r save rds}
## rda saved in "../../analysis/DEBRA/negsim_perturb_barbieQ.Rmd"
load("../../output/DEBRA/perturb_barbieQ.rda")
```

# Add 1 to data

```{r}
# perturb_top <- createBarbieQ(assay(perturb_top) +1, sampleMetadata = perturb_top$sampleMetadata)
```


# Designed v.s. estimated

## check equal mixture of pool1 and pool2 in baseline samples

- looking good.

```{r}
colnames(perturb)
props <- perturb_top@assays@data$proportion

pool1_vec <- props[,1]

pool2_vec <- props[,2]

## if equal mixture of pool1 and pool2; designed prop should be 1/2 of total props in original pool1 and pool2 
check_mix_df <- data.frame(ps = rowMeans(props[,1:2]), 
                           mixture_660 = props[,3])

ggplot(check_mix_df, aes(x = ps, y = mixture_660)) + 
  geom_point() + 
  coord_fixed() +
  geom_smooth()
```

## check zero in pool1 and pool2

```{r}
flag_no_zero <- props[,1] != 0 & props[,2] != 0
sum(flag_no_zero)
```

## check diff in pool1 and pool2

```{r}
flag_small_gap <- abs(props[,1] - props[,2]) < (props[,1] + props[,2])/10
sum(flag_small_gap)

flag_small_gap <- abs(props[,1] - props[,2]) < (props[,1] + props[,2])/5
sum(flag_small_gap)

flag_small_gap <- abs(props[,1] - props[,2]) < (props[,1] + props[,2])/2
sum(flag_small_gap)
```

## check any zeros in mixed samples
```{r}
any(props[,-c(1:2)] == 0)
```

## Designed differences

```{r}
## theory
a = 0.35
((1/2 + a) / (1+a)) - (1/2)
a = 0.27
(1/2 + a) / (1+a) - 1/2
a = 0.18
(1/2 + a) / (1+a) - 1/2

calc_theory <- function(a, trans = "none"){
  theory_pool1 = (1/2 + a) / (1+a)
  theory_pool2 = (1/2) / (1+a)
  
  trans_fun = ifelse(trans == "none", function(x) {x},
                     ifelse(trans == "asin-sqrt", function(x) {asin(sqrt(x))}, 
                            ifelse(trans == "logit", function(x) {log((x+1e-10)/(1-x+3e-7))}, NA))) ## {log((x+1e-10)/(1-x+1e-10))},or {logit(x)}
  
  
  
  # the ratio of a barcode in pool1 or pool2; ~= pool1+ pool2
  prop = pool1_vec + pool2_vec
  
  theory_FC1 = (trans_fun(theory_pool1 * pool1_vec )) - (trans_fun(1/2 * pool1_vec ))
  theory_FC2 = (trans_fun(theory_pool2 * pool2_vec )) - (trans_fun(1/2 * pool2_vec ))
  
  ## FC
  # theory_FC1 = (trans(theory_pool1) + 1e-10) / (trans(1/2) + 1e-10)
  # theory_FC2 = (trans(theory_pool2) + 1e-10) / (trans(1/2) + 1e-10)
  
  
  df <- data.frame(diff_1 = theory_FC1, diff_2 = theory_FC2)
  
  return(df)
}


calc_theory(a = 0.18)
calc_theory(a = 0.18, trans = "asin-sqrt")
calc_theory(a = 0.18, trans = "logit")

calc_theory(a = 0.27)
calc_theory(a = 0.27, trans = "asin-sqrt")
calc_theory(a = 0.27, trans = "logit")

calc_theory(a = 0.35)
calc_theory(a = 0.35, trans = "asin-sqrt")
calc_theory(a = 0.35, trans = "logit")

```

check out logit trans function

```{r}
# Set a value for u
u <- ((1/2 + a) / (1+a)) / (1/2)
log(((1/2 + a) / (1+a)) / (1/2))
log(((1/2 ) / (1+a)) / (1/2))


# Create a sequence of x values within a valid range
x_vals <- seq(0.00001, 0.1, length.out = 10000)  # avoid x = 1

# Compute y values
df <- data.frame(
  x = x_vals,
  y = log((u - x_vals) / (1 - x_vals))
)

# Plot
ggplot(df, aes(x = x, y = y)) +
  geom_line(color = "steelblue") +
  theme_minimal() +
  labs(title = paste0("y = log((u - x)/(1 - x)), u = ", u),
       x = "x", y = "y")
```


## Model perturbations

```{r}
colnames(perturb_top)
perturb_top$sampleMetadata %>% as.data.frame() %>% summary()

mixed <- perturb_top[, perturb_top$sampleMetadata$Subset %in% c("perturb", "null")]
mixed$sampleMetadata

mixed_target <- mixed$sampleMetadata %>% as.data.frame()
## factorize perturbation levels
mixed_target$Perturbation <- as.factor(mixed_target$Perturbation)
levels(mixed_target$Perturbation)
# mixed_target$Perturbation <- as.factor(mixed_target$Perturbation)
mixed_target$BarcodeSize <- as.factor(mixed_target$BarcodeSize)
mixed_design <- mixed_target %>%
    with(model.matrix(~0 + Perturbation + BarcodeSize))

```

## Apply tests: extract estimates

```{r}
apply_model <- function(a = 0.18, trans = "none", scale = 1) {
  
  mixed@assays@data$proportion <- mixed@assays@data$proportion * scale
  
  tester <- testBarcodeSignif(
    barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation", 
    contrastFormula = paste0("Perturbation", a, " - Perturbation0"), transformation = trans,
    designMatrix = mixed_design)
  
  veri_diff <- cbind(
    tester@elementMetadata$testingBarcode,
    calc_theory(a = a, trans = trans)) %>% as.data.frame()
  
  veri_diff <- rownames_to_column(veri_diff, var = "BC") 
  
  veri_diff$Contrast = paste0(a, " - 0")
  veri_diff$Method = trans
  
  veri_diff$uncertain_origin <- flag_small_gap
  veri_diff$noise_in_origin <- flag_no_zero
  
  return(veri_diff)
  
}
```


```{r}
es_de_none <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(a = i))
es_de_none_df <- do.call(rbind, es_de_none)

# es_de_none_scale <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(a = i, scale = 1e6))
# es_de_none_scale_df <- do.call(rbind, es_de_none_scale)

es_de_asin <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(a = i, trans = "asin-sqrt"))
es_de_asin_df <- do.call(rbind, es_de_asin)

es_de_logit <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(a = i, trans = "logit"))
es_de_logit_df <- do.call(rbind, es_de_logit)

es_de_df <- rbind(es_de_none_df, es_de_asin_df, es_de_logit_df)


```

check if BC of uncertain origins are picked out - no.

- good.

```{r}
es_de_df %>% 
  filter(uncertain_origin)
```


### Correlation: estimated and designed

#### winsorize: rm outlier

```{r}
## use robustHD::winsorise() to remove the outliers for cor(). especially those big correlated points.

regression_stats_none <- es_de_none_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(winsorize(diff_1 + diff_2), winsorize(meanDiff)),
    slope = coef(lm(winsorize(meanDiff) ~ I(winsorize(diff_1 + diff_2))))[2],
    interc = coef(lm(winsorize(meanDiff) ~ I(winsorize(diff_1 + diff_2))))[1],
    # res = resid(mod <- lm(meanDiff ~ I(diff_1 + diff_2))),
    # meanDiff = meanDiff,
    .groups = "drop"
  )
regression_stats_none$Method = "none"

regression_stats_asin <- es_de_asin_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(winsorize(diff_1 + diff_2), winsorize(meanDiff)),
    slope = coef(lm(winsorize(meanDiff) ~ I(winsorize(diff_1 + diff_2))))[2],
    interc = coef(lm(winsorize(meanDiff) ~ I(winsorize(diff_1 + diff_2))))[1],
    .groups = "drop"
  )
regression_stats_asin$Method = "asin-sqrt"

regression_stats_logit <- es_de_logit_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(winsorize(diff_1 + diff_2), winsorize(meanDiff)),
    slope = coef(lm(winsorize(meanDiff) ~ I(winsorize(diff_1 + diff_2))))[2],
    interc = coef(lm(winsorize(meanDiff) ~ I(winsorize(diff_1 + diff_2))))[1],
    .groups = "drop"
  )
regression_stats_logit$Method = "logit"

regression_stats <- rbind(regression_stats_asin, regression_stats_none, regression_stats_logit)

print(regression_stats_none)
print(regression_stats_asin)
print(regression_stats_logit)
```

#### no scaling

```{r}
## use robustHD::winsorise() to remove the outliers for cor(). especially those big correlated points.

## asin(sqrt(abs(x))) to rescale the none trans data
# re_scale <- function(x) {
#   asin(sqrt(abs(x)))
# }

# re_scale <- function(x) {
#   log10((abs(x) + 1e-7))
# }

re_scale <- function(x) {
  x
}

regression_stats_none <- es_de_none_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(
      re_scale((diff_1 + diff_2)), 
      re_scale((meanDiff))
      ),
    slope = coef(lm(
      re_scale((meanDiff)) ~ I(re_scale((diff_1 + diff_2)))
      ))[2],
    interc = coef(lm(
      re_scale((meanDiff)) ~ I(re_scale((diff_1 + diff_2)))
      ))[1],
    # res = resid(mod <- lm(meanDiff ~ I(diff_1 + diff_2))),
    # meanDiff = meanDiff,
    .groups = "drop"
  )
regression_stats_none$Method = "none"
regression_stats_none$x = -0.01
regression_stats_none$y = 0.01

# regression_stats_none_scale <- es_de_none_scale_df %>%
#   group_by(Contrast) %>%
#   summarise(
#     cor = cor(
#       re_scale((diff_1 + diff_2)), 
#       re_scale((meanDiff))
#       ),
#     slope = coef(lm(
#       re_scale((meanDiff)) ~ I(re_scale((diff_1 + diff_2)))
#       ))[2],
#     interc = coef(lm(
#       re_scale((meanDiff)) ~ I(re_scale((diff_1 + diff_2)))
#       ))[1],
#     # res = resid(mod <- lm(meanDiff ~ I(diff_1 + diff_2))),
#     # meanDiff = meanDiff,
#     .groups = "drop"
#   )
# regression_stats_none_scale$Method = "none+scale"

regression_stats_asin <- es_de_asin_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(re_scale(diff_1 + diff_2), re_scale(meanDiff)),
    slope = coef(lm(re_scale(meanDiff) ~ I(re_scale(diff_1 + diff_2))))[2],
    interc = coef(lm(re_scale(meanDiff) ~ I(re_scale(diff_1 + diff_2))))[1],
    .groups = "drop"
  )
regression_stats_asin$Method = "asin-sqrt"
regression_stats_asin$x = -0.001
regression_stats_asin$y = 0.001


regression_stats_logit <- es_de_logit_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(re_scale(diff_1 + diff_2), re_scale(meanDiff)),
    slope = coef(lm(re_scale(meanDiff) ~ I(re_scale(diff_1 + diff_2))))[2],
    interc = coef(lm(re_scale(meanDiff) ~ I(re_scale(diff_1 + diff_2))))[1],
    .groups = "drop"
  )
regression_stats_logit$Method = "logit"
regression_stats_logit$x = -0.05
regression_stats_logit$y = 1

regression_stats <- rbind(regression_stats_asin, regression_stats_none, regression_stats_logit)

# print(regression_stats_none)
# print(regression_stats_none_scale)
# print(regression_stats_asin)
# print(regression_stats_logit)
```


```{r}
es_de_none_df %>%
  filter(Contrast == "0.18 - 0") %>%
  summarise(
    De = winsorize(diff_1 + diff_2),
    Es = winsorize(meanDiff)) %>% 
  arrange(-De)

es_de_none_df %>%
  filter(Contrast == "0.18 - 0") %>%
  summarise(
    De = (diff_1 + diff_2),
    Es = (meanDiff)) %>% 
  arrange(-De)
```

### Plot coefficient

```{r, fig.width=8, fig.height=4}
p_es_de_asin <- ggplot(es_de_asin_df, aes(x = re_scale(diff_1 + diff_2), y = re_scale(meanDiff))) +
  geom_point(aes(color = P.Value < 0.05), alpha = 0.5) +
  # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
    geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
  theme_linedraw() +
  # coord_fixed() +
  facet_wrap(~paste0("asin-sqrt, Contrast: ", Contrast)) +
  geom_text(data = regression_stats_asin, x = -0.01, y = 0.01, 
            aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "top") +
  labs(color = "P-Value")


p_es_de_none <- ggplot(es_de_none_df, aes(x = re_scale(diff_1 + diff_2), y = re_scale(meanDiff))) +
  geom_point(aes(color = P.Value < 0.05), alpha = 0.5) +
  # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
  geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
  theme_linedraw()+
  # coord_fixed() +
  facet_wrap(~paste0("none, Contrast: ", Contrast)) +
    geom_text(data = regression_stats_none, x = -0.0015, y = 0.001, 
            aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
  labs( y = "Estimated differences")+
  scale_x_continuous(limits = c(-0.0025, 0.0025)) +
  scale_y_continuous(limits = c(-0.0025, 0.0025)) +
  theme(axis.title.x = element_blank(), legend.position = "none")

# p_es_de_none_scale <- ggplot(es_de_none_scale_df, aes(x = re_scale(diff_1 + diff_2)*1e6, y = re_scale(meanDiff))) +
#   geom_point(aes(color = P.Value < 0.05), alpha = 0.5) +
#   # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
#   geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
#   theme_linedraw() +
#   # coord_fixed() +
#   facet_wrap(~Contrast) +
#     geom_text(data = regression_stats_none, x = -0.001, y = 0.001, 
#             aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
#   labs(x = "Designed differences", y = "Estimated differences", title = "none") +
#   scale_x_continuous(limits = c(-2500, 2500)) +
#   scale_y_continuous(limits = c(-2500, 2500))

# identical(es_de_none_scale_df$tendencyTo, es_de_none_df$tendencyTo)

p_es_de_logit <- ggplot(es_de_logit_df, aes(x = re_scale(diff_1 + diff_2), y = re_scale(meanDiff))) +
  geom_point(aes(color = P.Value < 0.05), alpha = 0.5) +
  # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
  geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
  theme_linedraw()+
  facet_wrap(~paste0("logit, Contrast: ", Contrast)) +
    geom_text(data = regression_stats_logit, x = -0.15, y = 1, 
            aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
  labs(x = "Designed differences") +
  theme(axis.title.y = element_blank(), , legend.position = "none")
```


```{r, fig.width=8, fig.height=9}
p_es_de_asin / p_es_de_none / p_es_de_logit

ggsave("../../output/DEBRA/Es_de.png", dpi = 300, width = 8, height = 9, 
       p_es_de_asin / p_es_de_none / p_es_de_logit)
```

## TPR ~ AMean Quantiles

```{r}
num_quantiles <- 10

quantile_breaks <- quantile(es_de_none_df$Amean, probs = seq(0, 1, length.out = num_quantiles+1)) 
quantile_breaks 

es_de_bin <- es_de_df %>% 
  mutate(
    raw_Amean = rep(es_de_none_df$Amean, 3),
    bin = cut(raw_Amean, breaks = quantile_breaks, include.lowest = TRUE, right = TRUE))

es_de_bin_TPR <- es_de_bin %>% 
  group_by(bin, Method, Contrast) %>% 
  summarize(Frac_signif = sum(P.Value < 0.05) / n(), .groups = "drop")

TPR_quantile_es_de <- ggplot(es_de_bin_TPR, aes(x = bin, y = Frac_signif, group = paste0(Contrast, Method))) +
  geom_point(aes(shape = Contrast, color = Method)) +
  geom_line(aes(linetype = Contrast, color = Method)) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 75, vjust = 0.5, size = 10), legend.position = "top",
        aspect.ratio = 0.5) +
  labs(y = "Fraction of P-Value < 0.05", x = "Quantiles of barcode mean proportion (raw)")

TPR_quantile_es_de
```


```{r}
ggsave("../../output/DEBRA/TPR_quantile_es_de.png", dpi = 300, width = 6, height = 5, TPR_quantile_es_de)
```


## Residules

```{r}
# Create a sequence of x values within a valid range
x_vals <- seq(0.000001, 0.0001, length.out = 10000)  # avoid x = 1

# Compute y values
df <- data.frame(
  x = x_vals,
  y = asin(sqrt(x_vals))
)

# Plot
ggplot(df, aes(x = x, y = y)) +
  geom_line(color = "steelblue") +
  theme_minimal() +
  labs(title = paste0("asin(sqrt(x_vals)"),
       x = "x", y = "y")
```

### FUNC

```{r}
get_residules <- function(a = 0.18, trans = "none") {
  
  ## get props
  props <- mixed@assays@data$proportion
  
  ## set trans fun
  trans_fun = ifelse(trans == "none", function(x) {x},
                     ifelse(trans == "asin-sqrt", function(x) {asin(sqrt(x))}, 
                            ifelse(trans == "logit", function(x) {log((x+1e-10)/(1-x+3e-7))}, NA))) ## {log((x+1e-10)/(1-x+1e-10))},or {logit(x)}
  
  mydata <- trans_fun(props)
  
  myfit1 <- limma::lmFit(object = mydata, design = mixed_design)

  mycontrast <- limma::makeContrasts(contrasts = paste0("Perturbation", a, " - Perturbation0"), 
                                     levels = colnames(mixed_design))
  
  ## fit contrast
  myfit2 <- limma::contrasts.fit(fit = myfit1, contrasts = mycontrast)
  
  ## fit eBayes, moderated p.values obtained after applying empirical Bayes
  myfit3 <- limma::eBayes(myfit2)
  
  ## extract residuals
  resi <- limma::residuals.MArrayLM(object = myfit1, y = mydata)
  
  resi_df <- as.data.frame(as.table(resi))
  
  colnames(resi_df) <- c("BC", "Sample", "Residual")
  
  resi_df$Contrast <- paste0(a, " - 0")
  
  resi_df$Method <- trans

  return(resi_df)
  
}

resi_none <- lapply(c(0.18, 0.27, 0.35), function(i) get_residules(a = i, trans = "none"))
resi_asin <- lapply(c(0.18, 0.27, 0.35), function(i) get_residules(a = i, trans = "asin-sqrt"))
resi_logit <- lapply(c(0.18, 0.27, 0.35), function(i) get_residules(a = i, trans = "logit"))

resi_df <- rbind(do.call(rbind, resi_none), do.call(rbind, resi_asin), do.call(rbind, resi_logit))

```

### Distr, QQ

```{r}
p_resid_distr <- ggplot(resi_df %>% filter(Contrast == "0.35 - 0"),
                        aes(x = Residual, color = Method)) +
  geom_density() +
  labs(x = "Residual", y = "Density") +
  facet_wrap(~Method, scales = "free") +
  theme_linedraw() +
  theme(legend.position = "none")


p_resid_qq <- ggplot(resi_df %>% filter(Contrast == "0.35 - 0"), 
                     aes(sample = Residual, color = Method)) +
  stat_qq() +
  stat_qq_line(color = "black", linetype = "dashed") +
  labs(x = "Normal Quantiles", y = "Residule Quantiles") +
  facet_wrap(~Method, scales = "free") +
  theme_linedraw() +
  theme(legend.position = "none")
```

```{r, fig.width=9, fig.height=3}
p_resid_distr

p_resid_qq
```

```{r}
ggsave("../../output/DEBRA/resid_distribution.png", dpi = 300, width = 9, height = 3, p_resid_distr)
ggsave("../../output/DEBRA/resid_qq.png", dpi = 300, width = 9, height = 3, p_resid_qq)
```

## consistency in signif.

```{r}
asin_adjP_3con <- es_de_asin_df %>%
  dplyr::select(BC, Contrast, P.Value) %>%
  tidyr::pivot_wider(
    id_cols = BC,
    names_from = Contrast,
    values_from = P.Value
  )

euasin <- eulerr::euler(list(
  C0.18 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.18 - 0` < 0.05],
  C0.27 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.27 - 0` < 0.05],
  C0.35 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.35 - 0` < 0.05]
))

plot(euasin, quantities = TRUE)
```

```{r}
none_adjP_3con <- es_de_none_df %>%
  dplyr::select(BC, Contrast, P.Value) %>%
  tidyr::pivot_wider(
    id_cols = BC,
    names_from = Contrast,
    values_from = P.Value
  )

eunone <- eulerr::euler(list(
  C0.18 = rownames(none_adjP_3con)[none_adjP_3con$`0.18 - 0` < 0.05],
  C0.27 = rownames(none_adjP_3con)[none_adjP_3con$`0.27 - 0` < 0.05],
  C0.35 = rownames(none_adjP_3con)[none_adjP_3con$`0.35 - 0` < 0.05]
))

plot(eunone, quantities = TRUE)
plot(euasin, quantities = TRUE)
```

```{r}
eu <- eulerr::euler(list(
  asin_C0.18 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.18 - 0` < 0.05],
  asin_C0.27 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.27 - 0` < 0.05],
  asin_C0.35 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.35 - 0` < 0.05],
  none_C0.18 = rownames(none_adjP_3con)[none_adjP_3con$`0.18 - 0` < 0.05],
  none_C0.27 = rownames(none_adjP_3con)[none_adjP_3con$`0.27 - 0` < 0.05],
  none_C0.35 = rownames(none_adjP_3con)[none_adjP_3con$`0.35 - 0` < 0.05]
))

plot(eu, quantities = TRUE)
```

