---
title: "ROC_perturb_barbieQ"
author: "LF"
date: "2025-05-26"
output: workflowr::wflow_html
output_dir: ../../DEBRA
---

# Load Dependencies

```{r load library, message=FALSE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(plotROC)

library(limma)
library(edgeR)

library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
library(bartools)
```

# Install barbieQ - devel

```{r install barbieQ}
# devtools::install_github("Oshlack/barbieQ", force = TRUE)

library(barbieQ)
```

# Load rds

```{r save rds}
## rda saved in "../../analysis/DEBRA/negsim_perturb_barbieQ.Rmd"
load("../../output/DEBRA/perturb_barbieQ.rda")
```

# Check Top

```{r}
perturb@elementMetadata$isTopBarcode$isTop %>% sum()

perturb$sampleMetadata$Subset ## pool1 and pool2 in first 2 samples
(perturb@assays@data$isTopAssay[,1:2] %>% rowSums() >=1) %>% sum()
```


# Test barcode for Perturbation

## asin-sqrt

```{r asin}
colnames(perturb_top)
perturb_top$sampleMetadata %>% as.data.frame() %>% summary()

mixed <- perturb_top[, perturb_top$sampleMetadata$Subset %in% c("perturb", "null")]
mixed$sampleMetadata

mixed_target <- mixed$sampleMetadata %>% as.data.frame()
## calculate the theory difference
# a = as.numeric(mixed_target$Perturbation)
# mixed_target$Theo_perturb <- (1/2 + a) / (1+a) - 1/2
mixed_target$BarcodeSize <- as.factor(mixed_target$BarcodeSize)
mixed_design <- mixed_target %>%
    with(model.matrix(~0 + Perturbation + BarcodeSize))

mixed <- testBarcodeSignif(
  barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation",
  designMatrix = mixed_design)

mixed@elementMetadata$testingBarcode$meanDiff %>% summary()

bar_test <- rowData(mixed)$testingBarcode
# bar_test %>% as.data.frame() %>% View()
metadata(bar_test)$design
bar_test$direction %>% table()
bar_test$meanDiff %>% summary()

plotBarcodeMA(mixed)
```

## no Trans

```{r no trans}
mixed_noTrans <- testBarcodeSignif(
  barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation",
  designMatrix = mixed_design, transformation = "none")

plotBarcodeMA(mixed_noTrans) + scale_x_log10()
```

## logit

```{r logit}
mixed_logit <- testBarcodeSignif(
  barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation",
  designMatrix = mixed_design, transformation = "logit")

plotBarcodeMA(mixed_logit)
```

# Bin adj p.values

```{r}
stats_barcodes <- data.frame(
  Prop_asin = mixed@elementMetadata$testingBarcode$adj.P.Val,
  Prop_logit = mixed_logit@elementMetadata$testingBarcode$adj.P.Val,
  Prop_noTrans = mixed_noTrans@elementMetadata$testingBarcode$adj.P.Val,
  Amean_prop = mixed_noTrans@elementMetadata$testingBarcode$Amean
)

 stats_binned <- stats_barcodes %>%
    mutate(
      logAmean = log10(Amean_prop + 1e-7),
      bin = cut(logAmean, breaks = seq(-7, -1, by = 1), include.lowest = TRUE)
    ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    )

  # Compute medians per bin/method/N_Samples
  FPR_lines <- stats_binned %>%
    group_by(bin, method) %>%
    summarize(Frac_signif = sum(pvalue < 0.05) / n(), .groups = "drop")
    

  # Plot
p_bin_p <- ggplot(FPR_lines, aes(x = bin, y = Frac_signif, group = method, color = method)) +
    # geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8), outliers = F) +
    # geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
    geom_point(size = 1.8) +
    geom_line(linewidth = 0.8) +
    labs(
      x = "Log10(barcode mean prop.) bins",
      y = "Fraction of adj.P-Value < 0.05",
      fill = "Method",
      color = "Method"
    ) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top")

p_bin_p
```

```{r}
ggsave("../../output/DEBRA/TPR_binned_adjP.png", dpi = 300, width = 3, height = 3, p_bin_p)
```

# Bin p.values - unadjusted


```{r}
stats_barcodes <- data.frame(
  Prop_asin = mixed@elementMetadata$testingBarcode$P.Value,
  Prop_logit = mixed_logit@elementMetadata$testingBarcode$P.Value,
  Prop_noTrans = mixed_noTrans@elementMetadata$testingBarcode$P.Value,
  Amean_prop = mixed_noTrans@elementMetadata$testingBarcode$Amean
)

 stats_binned <- stats_barcodes %>%
    mutate(
      logAmean = log10(Amean_prop + 1e-7),
      bin = cut(logAmean, breaks = seq(-7, -1, by = 1), include.lowest = TRUE)
    ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    )

  # Compute medians per bin/method/N_Samples
  FPR_lines <- stats_binned %>%
    group_by(bin, method) %>%
    summarize(Frac_signif = sum(pvalue < 0.05) / n(), .groups = "drop")

  # Plot
p_bin_p <- ggplot(FPR_lines, aes(x = bin, y = Frac_signif, group = method, color = method)) +
    # geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8), outliers = F) +
    # geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
    geom_point(size = 1.8) +
    geom_line(linewidth = 0.8) +
    labs(
      x = "Log10(barcode mean prop.) bins",
      y = "Fraction of P-Value < 0.05",
      fill = "Method",
      color = "Method"
    ) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top")

p_bin_p
```

```{r}
ggsave("../../output/DEBRA/TPR_binned_unP.png", dpi = 300, width = 3, height = 3, p_bin_p)
```


# Bin mean prop.

```{r}
stats_binned <- stats_barcodes %>%
    select(Prop_noTrans, Amean_prop) %>% 
    mutate(
      logAmean = log10(Amean_prop + 1e-7),
      bin = cut(logAmean, breaks = seq(-7, -1, by = 1), include.lowest = TRUE)
    ) %>% 
    pivot_longer(
      cols = c(Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    )
  

n_barcode_lines <- stats_binned %>%
    group_by(bin) %>%
    summarize(n_barcode =  n(), .groups = "drop")
  
p_bin_n <- ggplot(n_barcode_lines, aes(x = bin, y = n_barcode)) +
    geom_col(color = "black", fill = "white") +
    labs(
      x = "Log10(barcode mean prop.) bins",
      y = "Num. barcodes per bin"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top")
  
p_bin_n
```

```{r}
ggsave("../../output/DEBRA/n_bar_binned_TP.png", dpi = 300, width = 3, height = 3, p_bin_n)
```

# Bin verified

```{r}
## 1. Create stats_binned with Origin and Verification info
stats_binned <- stats_barcodes %>%
  mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, -1, by = 1), include.lowest = TRUE),
    Origin = (total_prop > k) & (abs(change_prop) > m),
    Verified = rowData(mixed)$testingBarcode$P.Value < 0.05
  )

## 2. Calculate number of barcodes and verified ratios in each bin
bin_summary <- stats_binned %>%
  group_by(bin) %>%
  summarize(
    n_barcode = n(),
    n_origin = sum(Origin),
    n_verified = sum(Origin & Verified),
    verified_ratio = ifelse(n_origin > 0, n_verified / n_origin, NA),
    .groups = "drop"
  )

## 3. Plot
p_bin_verified  <- ggplot(bin_summary, aes(x = bin, y = n_barcode)) +
  geom_col(color = "black", fill = "white") +
  geom_text(aes(label = scales::percent(verified_ratio, accuracy = 1)),
            vjust = -0.5, size = 3, na.rm = TRUE) +
  labs(
    x = "Log10(barcode mean prop.) bins",
    y = "Num. barcodes per bin",
    title = "Verified Ratio per Bin"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_bin_verified 
```


# Barcode origin in Pool1, Pool2

```{r pool1, pool2}
## manually check testing results
## compare high perturb sample with low
colnames(perturb_top)
## validate using the premix pools
pool1 <- assays(perturb_top[,"P1"])$proportion[,1]
pool2 <- assays(perturb_top[,"P2"])$proportion[,1]
```

## Ground truth threshold

```{r}
pool1_vec <- as.vector(pool1)
pool2_vec <- as.vector(pool2)
total_prop = pool1_vec + pool2_vec
change_prop = pool1_vec - pool2_vec
summary(total_prop)
summary(change_prop)
summary(abs(change_prop))


## k for total prop; m for diff prop
k <- 0.0001
m <- k/2

## both pass considered true positive; everything else considered true negative
Origin <- (total_prop > k) & (abs(change_prop) > m)
table(Origin)
```

```{r reference pools}
ggplot(data.frame(pool1 = pool1, pool2 = pool2)) +
  geom_point(aes(x = log10(pool1 + 1e-7), y = log10(pool2 + 1e-7), color = Origin)) +
  # scale_x_log10() +
  # scale_y_log10() +
  theme(aspect.ratio = 1) 

df <- data.frame(
  LFC = log2(pool1+1e-7) -log2(pool2+1e-7),
  meanLog10_proportion = (log10(pool1+1e-7) + log10(pool2+1e-7))/2,
  Log10_mean_proportion = log10((pool1+pool2)/2+1e-7),
  Origin = ifelse(Origin, ifelse(change_prop > 0, "Pool1", "Pool2"), "Negative")
  )
p_truth <- ggplot(df) +
  geom_point(
    aes(x = Log10_mean_proportion, y = LFC, color = Origin),
    alpha = 0.8, shape = 1, size = 3, stroke = 1) +
  theme_classic() +
  labs(x = "Log10(mean proportion in Pool1 and Pool2)", y = "LFC of proportion between Pool1 and Pool2") +
  theme(aspect.ratio = 1) +
  scale_color_manual(values = c("Pool1" = "firebrick1", "Pool2" = "deepskyblue", "Negative" = "gold4"))

p_truth
```

```{r}
ggsave("../../output/DEBRA/truth_km_-5_-5.png", dpi = 300, width = 5, height = 4, p_truth)
```

## TPR 

### asin-sqrt

```{r}
N1 <- sum(Origin & change_prop > 0)
N2 <- sum(Origin & change_prop < 0)

n1 <- sum(rowData(mixed)$testingBarcode$direction == 1)
n2 <- sum(rowData(mixed)$testingBarcode$direction == -1)

n1 / N1
n2 / N2

sum(rowData(mixed)$testingBarcode$P.Value < 0.05) / sum(Origin)
sum(rowData(mixed)$testingBarcode$adj.P.Val < 0.05) / sum(Origin)
```

### noTrans

```{r}
N1 <- sum(Origin & change_prop > 0)
N2 <- sum(Origin & change_prop < 0)

n1 <- sum(rowData(mixed_noTrans)$testingBarcode$direction == 1)
n2 <- sum(rowData(mixed_noTrans)$testingBarcode$direction == -1)

n1 / N1
n2 / N2

sum(rowData(mixed_noTrans)$testingBarcode$P.Value < 0.05) / sum(Origin)
sum(rowData(mixed_noTrans)$testingBarcode$adj.P.Val < 0.05) / sum(Origin)
```


### logit

```{r}
N1 <- sum(Origin & change_prop > 0)
N2 <- sum(Origin & change_prop < 0)

n1 <- sum(rowData(mixed_logit)$testingBarcode$direction == 1)
n2 <- sum(rowData(mixed_logit)$testingBarcode$direction == -1)

n1 / N1
n2 / N2

sum(rowData(mixed_logit)$testingBarcode$P.Value < 0.05) / sum(Origin)
sum(rowData(mixed_logit)$testingBarcode$adj.P.Val < 0.05) / sum(Origin)
```

## Origin MA vs. test results

### vs. asin-sqrt results

```{r vs asin}
## color by asin-sqrt testing results
perturb.Trend <- rowData(mixed)$testingBarcode$tendencyTo
df <- data.frame(
  LFC = log2(pool1+1e-7) - log2(pool2+1e-7),
  meanLog10_proportion = (log10(pool1+1e-7) + log10(pool2+1e-7))/2
  )
p_validate <- ggplot(df) +
  geom_point(
    aes(x = meanLog10_proportion, 
        y = LFC, 
        color  = perturb.Trend),
    alpha = 0.6, shape = 1, size = 4, stroke = 2) +
  scale_color_manual(values = metadata(mixed)$factorColors$testingBarcode) +
  theme_classic() +
  labs(x = "Mean log10 barcode prop.", y = "LFC in barcode prop. between pool_1 and pool_2") +
  theme(aspect.ratio = 1, legend.position = "top")

p_validate
```

### vs. no Trans

```{r vs no trans}
## color by asin-sqrt testing results
perturb.Trend <- rowData(mixed_noTrans)$testingBarcode$tendencyTo
df <- data.frame(
  LFC = log2(pool1+1e-7) - log2(pool2+1e-7),
  meanLog10_proportion = (log10(pool1+1e-7) + log10(pool2+1e-7))/2
  )
p_validate <- ggplot(df) +
  geom_point(
    aes(x = meanLog10_proportion, 
        y = LFC, 
        color  = perturb.Trend),
    alpha = 0.6, shape = 1, size = 4, stroke = 2) +
  scale_color_manual(values = metadata(mixed_noTrans)$factorColors$testingBarcode) +
  theme_classic() +
  labs(x = "Mean log10 barcode prop.", y = "LFC in barcode prop. between pool_1 and pool_2") +
  theme(aspect.ratio = 1, legend.position = "top")

p_validate
```

# ROC

## roaming on alpha

- with fixed ground truth

- TP: total prop. > 0.0001, changes > 0.0005

- TN: total prop. > 0.0001, changes < 0.0005

```{r roc 1}
# Example input
pool1_vec <- as.vector(pool1)
pool2_vec <- as.vector(pool2)
barcode_mask <- (pool1_vec + pool2_vec) > k

asin_adj_p <- rowData(mixed)$testingBarcode$adj.P.Val[barcode_mask]
noTrans_adj_p <- rowData(mixed_noTrans)$testingBarcode$adj.P.Val[barcode_mask]
logit_adj_p <- rowData(mixed_logit)$testingBarcode$adj.P.Val[barcode_mask]

ground_truth <- as.integer((abs(pool1_vec - pool2_vec) > m))[barcode_mask]

sum(ground_truth)
length(ground_truth)

df <- rbind(
  data.frame(D = ground_truth, M = -asin_adj_p, model = "asin-sqrt"),
  data.frame(D = ground_truth, M = -noTrans_adj_p, model = "no trans."),
  data.frame(D = ground_truth, M = -logit_adj_p, model = "logit")
)

roc <- ggplot(df, aes(d = D, m = M, color = model)) +
  geom_roc(n.cuts = 0, labels = FALSE) +
  style_roc() +
  ggtitle("ROC Curve Comparison") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") 

auc_table <- auc_table <- calc_auc(roc)

roc + annotate(
  "text", x = 0.75, y = 0.25, size = 3,
  label = paste0("AUC (asin-sqrt) = ", 
                 round(auc_table$AUC[auc_table$model == "asin-sqrt"], 4),
                 "\nAUC (no trans.) = ", 
                 round(auc_table$AUC[auc_table$model == "no trans."], 4),
                 "\nAUC (logit) = ", 
                 round(auc_table$AUC[auc_table$model == "logit"], 4)
                 ))
```

- with fixed ground truth

- TP: total prop. > 0.001, changes > 0.0005

- TN: everything else

```{r roc 2}
# Example input
pool1_vec <- as.vector(pool1)
pool2_vec <- as.vector(pool2)
barcode_mask <- (pool1_vec + pool2_vec) > k

asin_adj_p <- rowData(mixed)$testingBarcode$adj.P.Val
noTrans_adj_p <- rowData(mixed_noTrans)$testingBarcode$adj.P.Val
logit_adj_p <- rowData(mixed_logit)$testingBarcode$adj.P.Val

ground_truth <- as.integer((abs(pool1_vec - pool2_vec) > m) & (pool1_vec + pool2_vec)/2 > k)

sum(ground_truth)
length(ground_truth)

df <- rbind(
  data.frame(D = ground_truth, M = -asin_adj_p, model = "asin-sqrt"),
  data.frame(D = ground_truth, M = -noTrans_adj_p, model = "no trans."),
  data.frame(D = ground_truth, M = -logit_adj_p, model = "logit")
)

roc <- ggplot(df, aes(d = D, m = M, color = model)) +
  geom_roc(n.cuts = 0, labels = FALSE) +
  style_roc() +
  labs(color = "Method") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1)

auc_table <- auc_table <- calc_auc(roc)

p_roc <- roc + annotate(
  "text", x = 0.75, y = 0.25, size = 3,
  label = paste0("AUC (asin-sqrt) = ", 
                 round(auc_table$AUC[auc_table$model == "asin-sqrt"], 4),
                 "\nAUC (no trans.) = ", 
                 round(auc_table$AUC[auc_table$model == "no trans."], 4),
                 "\nAUC (logit) = ", 
                 round(auc_table$AUC[auc_table$model == "logit"], 4)
                 ))

p_roc

ggsave("../../output/DEBRA/roc_km_-5_-5.png", dpi = 300, width = 5, height = 4, p_roc)
```

## roaming on K

```{r inverse}
k_seq <- 10^seq(-6, 0, length.out = 100)

pool1_vec <- as.vector(pool1)
pool2_vec <- as.vector(pool2)


gt_list <- lapply(k_seq, function(k) {
  barcode_mask <- (pool1_vec + pool2_vec) > k
  gt <- (abs(pool1_vec - pool2_vec) > 0.0001) & ((pool1_vec + pool2_vec) > k)
  gt <- gt[barcode_mask]
  
  tpr_asin <- sum(gt & (asin_adj_p[barcode_mask] < 0.05)) / sum(gt)
  fpr_asin <- 1 - (sum((!gt) & (asin_adj_p[barcode_mask] > 0.05)) / (sum(!gt)+0.0000001))
  
  tpr_noTrans <- sum(gt & (noTrans_adj_p[barcode_mask] < 0.05)) / sum(gt)
  fpr_noTransn <- 1 - (sum((!gt) & (noTrans_adj_p[barcode_mask] > 0.05)) / (sum(!gt)+0.0000001))
  
  roc_asin <- data.frame(FPR = fpr_asin, TPR = tpr_asin, method = "asin-sqrt", k = k)
  roc_noTrans <- data.frame(FPR = fpr_noTrans, TPR = tpr_noTrans, method = "no trans.", k = k)
  
  roc_df <- rbind(roc_asin, roc_noTrans)
  
  roc_df
})
roc_df <- do.call(rbind, gt_list)
# 
# # Example FPR and TPR values
# tpr_asin <- apply(gt_df, MARGIN = 2, function(j) sum(j & (asin_adj_p[barcode_mask] < 0.05)) / sum(j))
# fpr_asin <- apply(gt_df, MARGIN = 2, function(j) 1 - (sum((!j) & (asin_adj_p[barcode_mask] > 0.05)) / sum(!j)))
# 
# tpr_noTrans <- apply(gt_df, MARGIN = 2, function(j) sum(j & (noTrans_adj_p[barcode_mask] < 0.05)) / sum(j))
# fpr_noTrans <- apply(gt_df, MARGIN = 2, function(j) 1 - (sum((!j) & (noTrans_adj_p[barcode_mask] > 0.05)) / sum(!j)))
# 
# roc_asin <- data.frame(FPR = fpr_asin, TPR = tpr_asin, method = "asin-sqrt")
# 
# roc_noTrans <- data.frame(FPR = fpr_noTrans, TPR = tpr_noTrans, method = "no trans.")
# 
# roc_df <- rbind(roc_asin, roc_noTrans)

ggplot(roc_df, aes(x = FPR, y = TPR, color = method)) +
  geom_line() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey") +
  theme_minimal() +
  labs(title = "ROC Curve (from FPR and TPR)", x = "False Positive Rate", y = "True Positive Rate")

```

# Discoveries at K

```{r performence plot}
# Define the k sequence on a log scale
k_seq <- 10^seq(-6, 0, length.out = 100)

# Create vectors to store results
frac_n1_asin <- numeric(length(k_seq))
frac_n2_asin <- numeric(length(k_seq))
frac_n1_no <- numeric(length(k_seq))
frac_n2_no <- numeric(length(k_seq))

# Extract data
pool1_vec <- as.vector(pool1)
pool2_vec <- as.vector(pool2)
direction_asin <- rowData(mixed)$testingBarcode$direction
direction_no <- rowData(mixed_noTrans)$testingBarcode$direction

# Loop over each k
for (i in seq_along(k_seq)) {
  k <- k_seq[i]
  
  # Define N1 and N2 as "truth" based on threshold
  N1 <- sum((pool1_vec - pool2_vec > k/2) & ((pool1_vec + pool2_vec) > k))
  N2 <- sum((pool2_vec - pool1_vec > k/2) & ((pool1_vec + pool2_vec) > k))
  
  # Subset direction vector by same conditions
  barcode_mask <- (pool1_vec + pool2_vec) > k
  
  # asin-sqrt
  n1_asin <- sum(direction_asin[barcode_mask] == 1)
  n2_asin <- sum(direction_asin[barcode_mask] == -1)
  
  # no transform
  n1_no <- sum(direction_no[barcode_mask] == 1)
  n2_no <- sum(direction_no[barcode_mask] == -1)
  
  # Avoid division by 0
  frac_n1_asin[i] <- ifelse(N1 > 0, n1_asin / N1, NA)
  frac_n2_asin[i] <- ifelse(N2 > 0, n2_asin / N2, NA)
  frac_n1_no[i]   <- ifelse(N1 > 0, n1_no / N1, NA)
  frac_n2_no[i]   <- ifelse(N2 > 0, n2_no / N2, NA)
}

# Combine into data.frame for ggplot
plot_df <- tibble(
  k = k_seq,
  asin_pool1 = frac_n1_asin,
  asin_pool2 = frac_n2_asin,
  noTrans_pool1 = frac_n1_no,
  noTrans_pool2 = frac_n2_no
) %>%
  pivot_longer(-k, names_to = "method_pool", values_to = "fraction") %>%
  separate(method_pool, into = c("method", "pool"), sep = "_")

# Plot
p_frac <- ggplot(plot_df, aes(x = k, y = fraction, color = pool, linetype = method)) +
  geom_line(size = 1) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Threshold for total prop. (log scale)", y = "TPR", color = "Barcode origin") +
  theme_classic() +
  theme(legend.position = "top")
p_frac


```

# TPR 

```{r}
plot_TPR_over_k <- function(adj.P = FALSE) {
  # Define the k sequence on a log scale
k_seq <- 10^seq(-6, 0, length.out = 100)

# Create vectors to store results
frac_asin <- numeric(length(k_seq))
frac_logit <- numeric(length(k_seq))
frac_no <- numeric(length(k_seq))

# Extract data
if(adj.P){
  direction_asin <- rowData(mixed)$testingBarcode$adj.P.Val < 0.05
  direction_logit <- rowData(mixed_logit)$testingBarcode$adj.P.Val < 0.05
  direction_no <- rowData(mixed_noTrans)$testingBarcode$adj.P.Val < 0.05
} else {
  ## raw P-Value
  direction_asin <- rowData(mixed)$testingBarcode$P.Value < 0.05
  direction_logit <- rowData(mixed_logit)$testingBarcode$P.Value < 0.05
  direction_no <- rowData(mixed_noTrans)$testingBarcode$P.Value < 0.05
}


# Loop over each k
for (i in seq_along(k_seq)) {
  k <- k_seq[i]
  
  # Define N1 and N2 as "truth" based on threshold
  N <- ((total_prop > k) & (abs(change_prop) > m)) %>% sum()
  
  # Subset direction vector by same conditions
  barcode_mask <- (total_prop > k) & (abs(change_prop) > k/2)
  
  # asin-sqrt
  n_asin <- sum(direction_asin[barcode_mask])
  
  # logit
  n_logit <- sum(direction_logit[barcode_mask])
  
  # no transform
  n_no <- sum(direction_no[barcode_mask])
  
  # Avoid division by 0
  frac_asin[i] <- ifelse(N > 0, n_asin / N, NA)
  frac_logit[i] <- ifelse(N > 0, n_logit / N, NA)
  frac_no[i]   <- ifelse(N > 0, n_no / N, NA)
}

# Combine into data.frame for ggplot
plot_df <- tibble(
  k = k_seq,
  asin = frac_asin,
  logit = frac_logit,
  noTrans = frac_no
) %>%
  pivot_longer(-k, names_to = "method", values_to = "fraction")

# Plot
p_frac <- ggplot(plot_df, aes(x = k, y = fraction, color = method)) +
  geom_line(size = 1) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Threshold for total prop. (log scale)", y = "TPR", color = "Method") +
  theme_minimal() +
  theme(legend.position = "top")
p_frac
}

```

```{r}
plot_TPR_over_k(adj.P = FALSE)
plot_TPR_over_k(adj.P = TRUE)
```

```{r}
ggsave("../../output/DEBRA/TPR_over_k_rawP.png", dpi = 300, width = 5, height = 4, plot_TPR_over_k(adj.P = FALSE))
ggsave("../../output/DEBRA/TPR_over_k_adjP.png", dpi = 300, width = 5, height = 4, plot_TPR_over_k(adj.P = TRUE))
```

# Fold change

```{r}
asin_estimate <- mixed_noTrans@elementMetadata$testingBarcode$meanDiff
```


```{r}
## theory
a = 0.35
(1/2 + a) / (1+a) - 1/2
a = 0.27
(1/2 + a) / (1+a) - 1/2
a = 0.18
(1/2 + a) / (1+a) - 1/2

calc_theory <- function(a, trans = function(x) {x}){
  theory_pool1 = (1/2 + a) / (1+a)
  theory_pool2 = (1/2) / (1+a)
  
  # the ratio of a barcode in pool1 or pool2; ~= pool1+ pool2
  prop = pool1_vec + pool2_vec
  
  theory_FC1 = (trans(theory_pool1 * pool1_vec )) - (trans(1/2 * pool1_vec ))
  theory_FC2 = (trans(theory_pool2 * pool2_vec )) - (trans(1/2 * pool2_vec ))
  
  ## FC
  # theory_FC1 = (trans(theory_pool1) + 1e-10) / (trans(1/2) + 1e-10)
  # theory_FC2 = (trans(theory_pool2) + 1e-10) / (trans(1/2) + 1e-10)
  
  
  df <- data.frame(diff_1 = theory_FC1, diff_2 = theory_FC2)
  
  return(df)
}


calc_theory(a = 0.18)
calc_theory(a = 0.18, trans = function(x) {asin(sqrt(x))})
calc_theory(a = 0.18, trans = function(x) {log((x+1e-10)/(1-x+1e-10))})

calc_theory(a = 0.27)
calc_theory(a = 0.27, trans = function(x) {asin(sqrt(x))})
calc_theory(a = 0.27, trans = function(x) {log((x+1e-10)/(1-x+1e-10))})

calc_theory(a = 0.35)
calc_theory(a = 0.35, trans = function(x) {asin(sqrt(x))})
calc_theory(a = 0.35, trans = function(x) {log((x+1e-10)/(1-x+1e-10))})



```


```{r}
colnames(perturb_top)
perturb_top$sampleMetadata %>% as.data.frame() %>% summary()

mixed <- perturb_top[, perturb_top$sampleMetadata$Subset %in% c("perturb", "null")]
mixed$sampleMetadata

mixed_target <- mixed$sampleMetadata %>% as.data.frame()
mixed_target$Perturbation <- as.factor(mixed_target$Perturbation)
levels(mixed_target$Perturbation)
# mixed_target$Perturbation <- as.factor(mixed_target$Perturbation)
mixed_target$BarcodeSize <- as.factor(mixed_target$BarcodeSize)
mixed_design <- mixed_target %>%
    with(model.matrix(~0 + Perturbation + BarcodeSize))

mixed <- testBarcodeSignif(
  barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation", 
  contrastFormula = ("Perturbation0.27 - Perturbation0"),
  designMatrix = mixed_design)

mixed@elementMetadata$testingBarcode@metadata$contrasts
mixed@elementMetadata$testingBarcode$meanDiff

bar_test <- rowData(mixed)$testingBarcode
# bar_test %>% as.data.frame() %>% View()
metadata(bar_test)$design
bar_test$direction %>% table()
bar_test$meanDiff %>% summary()
```

```{r}
tester <- testBarcodeSignif(
  barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation", 
  contrastFormula = ("Perturbation0.27 - Perturbation0"), transformation = "asin-sqrt",
  designMatrix = mixed_design)

veri_diff <- cbind(
  tester@elementMetadata$testingBarcode,
  calc_theory(a = 0.27, trans = function(x) {asin(sqrt(x))})) %>% as.data.frame()

ggplot(veri_diff ) +
  geom_point(aes(x = diff_1 + diff_2, y = meanDiff, color = tendencyTo)) +
  coord_fixed()

cor(veri_diff$diff_1 + veri_diff$diff_2, veri_diff$meanDiff)

veri_sig <- veri_diff %>% 
  filter(tendencyTo != "n.s.") 

cor(veri_sig$diff_1 + veri_sig$diff_2, veri_sig$meanDiff)

y = veri_sig$diff_1 + veri_sig$diff_2
x = veri_sig$meanDiff

coef(lm(y ~x))

veri_diff %>% 
  filter(tendencyTo == "n.s.") %>% 
  with(cor(diff_1 + diff_2, meanDiff))

```

```{r}
sum(rowData(tester)$testingBarcode$P.Value < 0.05) / sum(Origin)
sum(rowData(tester)$testingBarcode$adj.P.Val < 0.05) / sum(Origin)
```


# Example

```{r fig.width=10, fig.height=5}
mixed <- perturb_top[, perturb_top$sampleMetadata$Subset %in% c("perturb", "null")]
mixed_dge <- DGEList(
  counts = assay(mixed), 
  group = mixed$sampleMetadata$Subset)
bartools::plotBarcodeHistogram(mixed_dge)

perturbed <- perturb_top[, perturb_top$sampleMetadata$Subset %in% c("perturb")]

perturbed_dge <- DGEList(
  counts = assay(perturbed[,order(perturbed$sampleMetadata$Perturbation)]), 
  group = perturbed$sampleMetadata$Subset[order(perturbed$sampleMetadata$Perturbation)])
bartools::plotBarcodeHistogram(perturbed_dge)
```

