---
title: "barbieQ_paper Supplementary 2: Type I error rate using `AML data`"
author: "Liyang Fei"
date: |
  Initiate: 2025-04-04 
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Links to Type I error rate assessment using other datasets in the barbieQ paper: 
 
 - [Mixture data](barbieQ_paper_Figure3.html)
 
 - [AML data]()
 
 - [HSPC xenograft data](barbieQ_paper_FigureS2_xenoHSPC.html)

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)
library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)
# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
# library(bartools)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/F3_simulation.R") ## for negative simulation
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Load preprocessed data

Links to [AML data preprocessing](barbieQ_paper_FigureS1_AML.html)

```{r save rds}
load("output/AML_barbieQ.rda")
```

# Samples level QC

## PCA

```{r}
pca_results <- (assays(AML_top)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp()

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  AML_top$sampleMetadata
)
```

```{r}
ggplot(pca_df, aes(x = PC1, y = PC2, color = Treatment)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Dose)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Timepoint)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```


## Plot propportion

```{r bartools}
## order samples
order_sample <- AML_top$sampleMetadata %>% with(order(Treatment, Dose, Timepoint))

## create bartools object
AML_dge <- DGEList(
  counts = assay(AML_top), 
  group = AML_top$sampleMetadata$Treatment)

## plot 
plotBarcodeHistogram(AML_dge, orderSamples = colnames(AML_top)[order_sample])
```

## Subet DMSO

```{r subset dmso}
DMSO <- AML_top[, AML_top$sampleMetadata$Treatment == "DMSO"]
```

```{r}
pca_results <- (assays(DMSO)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp()

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_dmso <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  DMSO$sampleMetadata
)

ggplot(pca_dmso, aes(x = PC1, y = PC2, color = Timepoint)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```

## FS2G: prop. DMSO

```{r, fig.width= 4, fig.height=5}
## order samples
order_sample <- DMSO$sampleMetadata %>% with(order(Timepoint))

## create bartools object
DMSO_dge <- DGEList(
  counts = assay(DMSO), 
  group = DMSO$sampleMetadata$Timepoint)

## plot 
plotBarcodeHistogram(DMSO_dge, orderSamples = colnames(DMSO)[order_sample]) + 
  labs(title = "AML: DMSO") -> p_prop_DMSO

p_prop_DMSO
```

# Type I error rate

## Negative simulation

**avoid running this chunk during rendering. `global_all_loops` was saved from the first run.**

```{r eval=FALSE, fig.height=8, fig.width=16, include=TRUE}
results_negsi <- suppressMessages({
  random_sampling(barbieQ = DMSO, loop_times = 100)
  })

## `random_sampling()` save the simulations to global environment as `global_all_loops`
save(global_all_loops, file = "output/AML_negative_simulation.rda")
```

- load results

```{r}
## loading random sampling results to avoid running it repeatedly
load("output/AML_negative_simulation.rda")

# extract results from the loops
end_sampling <- floor(ncol(DMSO)/2)

## extract FPR
all_FPR <- lapply(seq(3:end_sampling), function(n) {
  global_all_loops[[n]]$FPR
})
df_FPR_DMSO <- do.call(rbind, all_FPR)
```

- plot FPR

```{r}
df_FPR_DMSO$N_Samples <- as.factor(df_FPR_DMSO$N_Samples)

## Diff_Prop
## reshape the data to fit methods for testing
df_FPR_DMSO_long <- df_FPR_DMSO %>%
  pivot_longer(cols = starts_with("Prop_"),
               names_to = "Method",
               values_to = "Pval_Prop") %>%
  mutate(Method = factor(Method, levels = c("Prop_asin", "Prop_logit", "Prop_noTrans"),
                         labels = c("asin-sqrt", "logit", "no trans.")))

P_FPR_prop_DMSO <- ggplot(
  df_FPR_DMSO_long, aes(x = factor(N_Samples), y = Pval_Prop, color = Method)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       # title = "DMSO",
       subtitle = "Differential proportion test") +
  # facet_wrap(~Method, scales = "free_y") +
  theme_classic() +
  theme(legend.position = "top")

## Diff_Occ
P_FPR_occ_DMSO <- ggplot(
  df_FPR_DMSO, aes(x = factor(N_Samples), y = Occ_firth)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Differential occurrrence test" = "black")) +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       titlte = "",
       subtitle = "Differential occurrence test") +
  theme_classic() +
  theme(legend.position = "top")

```

## FS2H: FPR

```{r, fig.width=8, fig.height=5}
(P_FPR_prop_DMSO + theme(legend.position = "top")) + (P_FPR_occ_DMSO) -> P_FPR_DMSO

P_FPR_DMSO
```


# Figure S2-AML

```{r, fig.width=12, fig.height=5}
layout = "
GHH
"
fs2_aml <- (
    wrap_elements(p_prop_DMSO + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
  wrap_elements(P_FPR_DMSO + theme(legend.position = "top") + theme(plot.margin = unit(rep(0,4), "cm")))
) + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("G", "H"))) &
  theme(
    plot.tag = element_text(size = 20, face = "bold", family = "arial"),
    axis.title = element_text(size = 17),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

fs2_aml
```

```{r}
ggsave(
  filename = "output/fs2_aml.png",
  plot = fs2_aml,
  width = 12,
  height = 5,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [fs2_AML](../output/fs2_aml.png)