---
title: "barbieQ_paper_Figure2: Preprocessing `Monkey HSPC dataset`"
author: "Liyang Fei"
date: |
  Initiated: 2025-09-07  
  Rendered: `r format(Sys.Date(), '%Y-%m-%d')`
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Linked to Preprocessing analyses of all datasets in the barbieQ paper: 
 
 - [Monkey HSPC]()
 
 - [AML data]()
 
 - [Mouse HSPC]()
 
 - [Mixture data]()

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(data.table) # for data.table()
library(SummarizedExperiment)
library(S4Vectors)
library(ggplot2)
library(ggbreak)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
getwd()
```


# Clean up datasets {.tabset .tabset-fade .tabset-pills}

The data was downloaded from [barcodetrackRData](https://github.com/dunbarlabNIH/barcodetrackRData) repository on GitHub. The experiment was described in [Wu, Chuanfeng, et al. "Clonal expansion and compartmentalized maintenance of rhesus macaque NK cell subsets." Science Immunology (2018)](http://dx.doi.org/10.1126/sciimmunol.aat9781).

Load barcode count matrix. Check barcode sequence: pass.

```{r}
## data is separated by "\t". so we use `read.table()` for loading data.
count <- read.table("data/WuC/monkey_ZG66.txt", header = TRUE, sep = "\t", row.names = 1)

## rownames are barcode sequence, basically consist of ATGCs: "^[ATGC]+$"
## check patterns of rownames
lapply(rownames(count), function(x) grepl("^[ATGC]+$", x)) %>% unlist() %>% table()
## print untypical rownamse
rownames(count)[!lapply(rownames(count), function(x) grepl("^[ATGC]+$", x)) %>% unlist()]
## allow "N" in rowname patterns: "^[ATGCN]+$"
lapply(rownames(count), function(x) grepl("^[ATGCN]+$", x)) %>% unlist() %>% table()
```

Load sample meta information. Clean up experimental conditions.

```{r}
metadata <- read.table("data/WuC/monkey_ZG66_metadata.txt", header = 1, row.names = 1)

## update the variable names using the camelCase style.
colnames(metadata)[grepl("Cell_type", colnames(metadata))] <- "Celltype"
print(colnames(metadata))

## `Timepoint` and `Months` variables in the `metadata` are repetition, delete `Timepoint`.
identical(gsub("m$", "", metadata$Timepoint) %>%  as.numeric(), metadata$Months)
metadata$Timepoint <- NULL
print(colnames(metadata))

## all samples are from the "ZG66". delete `Monkey` variable.
metadata$Monkey %>% table()
metadata$Monkey <- NULL
print(colnames(metadata))

## categorize `Months` by `Phase`
## replace "Gr_1" and "Gr_2" by "Gr" - I don't know what exactly happened in the experiment regarding the unmatched cell type names.
metadata$Months
metadata$Celltype %>% table()
metadata <- metadata %>% 
  as.data.frame() %>% 
  select(Celltype, Months) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype)) %>% 
  ## to repeat the analyses in the original paper, we select same set of samples, which is the first 30 samples, basically collected from those time points
  mutate(inOriginalPaper = Months %in% c(6.5, 12, 17, 27, 36, 42))

## turn `Celltype` into factors and set up levels.
metadata$Celltype <- factor(metadata$Celltype, levels = c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p", "NK_NKG2Ap_CD16p", "NK_NKG2Ap_CD16p_KIR3DL01n", "NK_NKG2Ap_CD16p_KIR3DL01p"))

data.table(metadata)
```

# Create `barbieQ` object

Store count matrix and sample metadata into a `barbieQ` object.

```{r}
## set up the color mapping for the conditions
conditionColor0 <- list(
  Celltype = setNames(
    c("yellowgreen", "darkolivegreen", "goldenrod", "steelblue1", "slateblue1", "orchid", "lightpink", "violetred"), 
    c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p", "NK_NKG2Ap_CD16p", "NK_NKG2Ap_CD16p_KIR3DL01n", "NK_NKG2Ap_CD16p_KIR3DL01p")),
  Months = circlize::colorRamp2(
    breaks = c(6.5, 12, 17, 27, 36, 42, 68),
    colors = c("#CFCFFF", "#B2B2FF", "#8282FF", "#4D4FFF", "#1D1DFF", "#0000B2", "#00006A"))
)

## for this specific subset of samples - removing conditions not existing
conditionColor <- list(
    Celltype = setNames(c("yellowgreen", "darkolivegreen", "goldenrod", "steelblue1", "slateblue1"), 
                        c("T", "B", "Gr", "NK_CD56p_CD16n", "NK_CD56n_CD16p")),
    Months = circlize::colorRamp2(
      breaks = c(6.5, 12, 17, 27, 36, 42),
      colors = c("#CFCFFF", "#B2B2FF", "#8282FF", "#4D4FFF", "#1D1DFF", "#0000B2"))
)
```

```{r}
## create an object to store the barcode count matrix and sample metadata.
## only passing Celltype and Months to sample metadata
## to repeat the analyses in the original paper, we select same set of samples, which is the first 30 samples, basically samples of Months %in% c(6.5, 12, 17, 27, 36, 42)
monkeyHSPC_raw <- barbieQ::createBarbieQ(
  object = count, 
  sampleMetadata = metadata, 
  factorColors = conditionColor0)

## add monkey ID to object metadata.
monkeyHSPC_raw@metadata$Monkey <- list("Monkey" = "ZG66")
```


```{r}
## create an object to store the barcode count matrix and sample metadata.
## only passing Celltype and Months to sample metadata
## to repeat the analyses in the original paper, we select same set of samples, which is the first 30 samples, basically samples of Months %in% c(6.5, 12, 17, 27, 36, 42)
monkeyHSPC <- barbieQ::createBarbieQ(
  object = count[,1:30], 
  sampleMetadata = metadata[1:30, c("Celltype", "Months")], 
  factorColors = conditionColor)

## add monkey ID to object metadata.
monkeyHSPC@metadata$Monkey <- list("Monkey" = "ZG66")
```

# Cluster correlating barcodes {.tabset .tabset-fade .tabset-pills}

Select "top" barcodes in advance to save the time & space for running `plotBarcodePairCorrelation`.

```{r}
## check min group size as `nSampleThreshold` in `tagTopBarcodes`
table(monkeyHSPC$sampleMetadata$Celltype)
## select top barcodes
monkeyHSPC <- barbieQ::tagTopBarcodes(monkeyHSPC, nSampleThreshold = 6)
## print number of unselected / selected barcodes
monkeyHSPC@elementMetadata$isTopBarcode %>% table()
```

Compute barcode pair-wise correlation. Default to Pearson correlation.

```{r, fig.width=5, fig.height=4}
p_monkey_pair_cor <- plotBarcodePairCorrelation(barbieQ = monkeyHSPC[rowData(monkeyHSPC)$isTopBarcode$isTop,])
f2a <- p_monkey_pair_cor + theme(legend.position = "top")
f2a
```

Cluster barcodes by correlation.

```{r}
monkeyHSPC_top <- barbieQ::clusterCorrelatingBarcodes(barbieQ = monkeyHSPC[rowData(monkeyHSPC)$isTopBarcode$isTop,])
```

Inspect barcode clusters.

```{r}
p_cluster_list <- barbieQ::inspectCorrelatingBarcodes(monkeyHSPC_top)
```

`p_cluster` is not a classical ggplot object, instead, it's converted from a base R plot generated by `igraph`. Here's how to change the `geom_node_text` size for it.

```{r fig.width=4, fig.height=4}
## update the geom_node_text for `p_cluster`
## find the geom_node_text layer
layer_index <- which(sapply(p_cluster_list$p_cluster$layers, function(x) inherits(x$geom, "GeomTextRepel")))
## update the size for that layer
p_cluster_list$p_cluster$layers[[layer_index]]$aes_params$size <- 1.4  # new size
f2b <- p_cluster_list$p_cluster + theme(legend.position = "none")

f2b
```

```{r include=FALSE}
f2c <- (((p_cluster_list$p_cluster_size + theme(axis.text.x = element_blank(),
                                         axis.title.x = element_blank(),
                                         axis.ticks.x = element_blank(),
                                         axis.line.x = element_blank(),
                                         strip.text = element_blank(),
                                         plot.margin = unit(c(0, 0, 0, 0.1), "lines"))
   ) %>% print() %>% wrap_elements(clip = FALSE)) / ((p_cluster_list$p_cluster_prop + theme(plot.margin = unit(c(0,0.6,0,0.6), "line"))) %>% wrap_elements(clip = FALSE)) + plot_layout(guides = "collect"))
```

```{r fig.width=4, fig.height=8}
f2c
```



# Merge correlating barcodes

Use the barcode correlation information saved in `monkeyHSPC_top` to merge barcodes in the initial data `monkeyHSPC`.

Barcode proportion, CPM, rank, ... are re-calculated based on the selected barcodes in the new barbieQ object.

```{r}
monkeyHSPC_merged <- barbieQ::mergeCorrelatingBarcodes(barbieQ_clustered = monkeyHSPC_top, barbieQ_toMerge = monkeyHSPC)
dim(monkeyHSPC_merged)

monkeyHSPC_merged@metadata$predicted_barcode_clusters
monkeyHSPC_merged@metadata$removed_barcodes
```


# Filter low-count barcodes {.tabset .tabset-fade .tabset-pills}

Tag top barcodes.

```{r}
## check min group size as `nSampleThreshold` in `tagTopBarcodes`
table(monkeyHSPC_merged$sampleMetadata$Celltype)
## select top barcodes
monkeyHSPC_merged <- barbieQ::tagTopBarcodes(monkeyHSPC_merged, nSampleThreshold = 6)
## print number of unselected / selected barcodes
monkeyHSPC_merged@elementMetadata$isTopBarcode %>% table()
```

Visualize barcodes by "top" and "bottom".

```{r, fig.width=5, fig.height=4}
f2d <- barbieQ::plotBarcodePareto(barbieQ = monkeyHSPC_merged) +
  ylim(-8, 7) + 
  annotate("text", x = c(pi * 0.05), y = c(7), label = c("Relative proportion %"), color = "#999966", size = 3, angle = 0, fontface = "bold", hjust = 1)
f2e <- barbieQ::plotBarcodeSankey(barbieQ = monkeyHSPC_merged)

f2d
f2e
```


```{r}
## subset barcodes by top barcodes.
monkeyHSPC_merged_top <- monkeyHSPC_merged[monkeyHSPC_merged@elementMetadata$isTopBarcode$isTop,]
dim(monkeyHSPC_merged_top)
```

Visualize barcode heatmaps by pre- and post- filtering.

```{r}
p_ht_all <- barbieQ::plotBarcodeHeatmap(monkeyHSPC_merged)
## setting heatmap matrix size with accessors
p_ht_all@matrix_param$width <- grid::unit(7, "cm")
p_ht_all@matrix_param$height <- grid::unit(7, "cm")
```

```{r}
p_ht_top <- barbieQ::plotBarcodeHeatmap(monkeyHSPC_merged_top)

p_ht_top@matrix_param$width <- grid::unit(7, "cm")
p_ht_top@matrix_param$height <- grid::unit(7, "cm")

# barbieQ::plotBarcodeHeatmap(monkeyHSPC_merged_top, colorMapTo = "proportion")
# barbieQ::plotBarcodeHeatmap(monkeyHSPC_merged_top, colorMapTo = "logit proportion")
```

# Figure2

```{r, fig.width=13, fig.height=12}
layout = "
AAAAABBBBCCCC
AAAAABBBBCCCC
DDDDDEEEECCCC
DDDDDEEEECCCC
FFFFFGGGGGG##
FFFFFGGGGGG##
"
f2 <- (wrap_elements(f2a + theme(plot.margin = unit(c(0,0.6,0,0.6), "line"))) + 
    wrap_elements(f2b + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(f2c + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(f2d + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(f2e + theme(plot.margin = unit(rep(0,4), "cm"), legend.position = "top")) + 
    wrap_plots(list(p_ht_all %>% draw(heatmap_legend_side = "left", show_annotation_legend = FALSE) %>% grid.grabExpr())) +
    wrap_plots(list(p_ht_top %>% draw(heatmap_legend_side = "left") %>% grid.grabExpr())) 
  ) + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("A","B","C", "D", "E", "F", ""))) &
  theme(
    plot.tag = element_text(size = 24, face = "bold", family = "arial"),
    axis.title = element_text(size = 17),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

f2
```

```{r}
ggsave(
  filename = "output/f2.png",
  plot = f2,
  width = 13,
  height = 12,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 300
  )
```

Saving this figure in [f2](../output/f2.png)

# Save preprocessed `barbieQ` objects in .rda

```{r}
save(monkeyHSPC_merged, monkeyHSPC_merged_top, file = "output/monkeyHSPC_barbieQ.rda")
save(monkeyHSPC_raw, file = "output/monkeyHSPC_raw_barbieQ.rda")
```

