---
title: "barbieQ_paper Supplementary 1: Preprocessing `HSPC xenograft data`"
author: "Liyang Fei"
date: |
  Initiate: 2025-04-03 
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Links to preprocessing other datasets in the barbieQ paper: 
 
 - [Monkey HSPC](barbieQ_paper_Figure2.html)
 
 - [AML data](barbieQ_paper_FigureS1_AML.html)
 
 - [HSPC xenograft data]()
 
 - [Mixture data](barbieQ_paper_FigureS1_Mixture.html)

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)
library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Read data

Read every sheet in the excel.

8 mice: "C1"  "C2"  "C3"  "C10" "C15" "C22" "C21" "C23"

overlapping barcodes: hundreds

mouse-specific barcodes: 1-2k.

According to the original publication, 'library' in columns may refer to lentiviral level barcode library - before integrated into cell genome.

```{r read xl}
sheets <- readxl::excel_sheets("data/BelderbosME/mmc2.xlsx")

data_list <- purrr::map(sheets, ~readxl::read_excel("data/BelderbosME/mmc2.xlsx", sheet = .x))

# bind sheets
# library(dplyr)
# combined_data <- bind_rows(data_list, .id = "sheet_name")

names(data_list) <- sheets

C1_bar <- data_list$C1$ID
C2_bar <- data_list$C2$`['ID'`
C3_bar <- data_list$C3$`['ID'`
C10_bar <- data_list$C10$ID
C15_bar <- data_list$C15$...1
C22_bar <- data_list$C22$ID
C21_bar <- data_list$C21$ID
C23_bar <- data_list$C23$ID

bar_list <- list(
  C1 = C1_bar,
  C2 = C2_bar,
  C3 = C3_bar,
  C10 = C10_bar,
  C15 = C15_bar,
  C22 = C22_bar,
  C21 = C21_bar,
  C23 = C23_bar
)

ggVennDiagram::ggVennDiagram(x = bar_list)

```

# Clean up data

## C1

```{r clean up C1}
class(data_list$C1)

data_list$C1
data_list$C1 <- as.data.frame(data_list$C1)
class(data_list$C1)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C1[, 1]))
data_list$C1 <- data_list$C1[!is.na(data_list$C1[, 1]), ]
dim(data_list$C1)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C1) <- data_list$C1[,1]
data_list$C1 <- data_list$C1[,-1]

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C1 <- data_list$C1[, !colnames(data_list$C1) %in% del]

# delete processed "sum" columns
data_list$C1 <- data_list$C1[, !grepl("_Sum", colnames(data_list$C1))] 
data_list$C1 <- data_list$C1[, !grepl("_sum", colnames(data_list$C1))] 

# delete processed "shannon" rows
rownames(data_list$C1) [!grepl("^[ATGC]+$", rownames(data_list$C1))]
rownames(data_list$C1) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C1))]
data_list$C1 <- data_list$C1[grepl("^[ATGC\\\"]+$", rownames(data_list$C1)),]

dim(data_list$C1)

# Convert NA values to 0 in the matrix
data_list$C1[is.na(data_list$C1)] <- 0
!any(is.na(data_list$C1))

# Convert character data to numeric
C1_count_numeric <- apply(data_list$C1, 2, as.numeric)
rownames(C1_count_numeric) <- rownames(data_list$C1)

data_list$C1 <- C1_count_numeric

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C1))
xCondition <- gsub("(.*)\\_.*", "\\1", xCondition)
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk[0-9]{1,}).*", "\\1", colnames(data_list$C1))
Time[grepl("\\_", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub(".*\\_(BM)\\_.*", "\\1", colnames(data_list$C1))
Tissue <- gsub(".*\\_(Blood)\\_.*", "\\1", Tissue)
Tissue[grepl("\\_", Tissue)] <- "Blood"
Tissue

# extract celltype from colnames
Celltype <- gsub(".*\\_.*\\_([A-Z])", "\\1", colnames(data_list$C1))
Celltype[grepl("\\_", Celltype)] <- "Unsorted"
Celltype[grepl("T", Celltype)] <- "Tcell"
Celltype[grepl("B", Celltype)] <- "Bcell"
Celltype[grepl("G", Celltype)] <- "Grn"


colnames(data_list$C1)

C1_meta <- data.frame(
  Sample = colnames(data_list$C1),
  Donor = rep("C1", ncol(data_list$C1)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Celltype = Celltype
)

# View(data_list$C1)

```

## C2

```{r clean up C2}
# data_list$C2 %>% View
data_list$C2 <- as.data.frame(data_list$C2)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C2[, 1]))
data_list$C2 <- data_list$C2[!is.na(data_list$C2[, 1]), ]
dim(data_list$C2)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C2) <- data_list$C2[,1]
data_list$C2 <- data_list$C2[,-1]

colnames(data_list$C2)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C2 <- data_list$C2[, !colnames(data_list$C2) %in% del]

# delete processed "sum" columns
data_list$C2 <- data_list$C2[, !grepl("Sum", colnames(data_list$C2))] 
data_list$C2 <- data_list$C2[, !grepl("sum", colnames(data_list$C2))] 

# delete processed "shannon" rows
rownames(data_list$C2) [!grepl("^[ATGC]+$", rownames(data_list$C2))]
rownames(data_list$C2) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C2))]
data_list$C2 <- data_list$C2[grepl("^[ATGC\\\"]+$", rownames(data_list$C2)),]

dim(data_list$C2)

# Convert NA values to 0 in the matrix
data_list$C2[is.na(data_list$C2)] <- 0
!any(is.na(data_list$C2))

# Convert character data to numeric
C2_count_numeric <- apply(data_list$C2, 2, as.numeric)
rownames(C2_count_numeric) <- rownames(data_list$C2)

data_list$C2 <- C2_count_numeric

# View sample names
colnames(data_list$C2)

# fix "library" clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C2) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C2))

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C2))
xCondition <- gsub("(.*)\\_.*", "\\1", xCondition)
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk[0-9]{1,}).*", "\\1", colnames(data_list$C2))
Time[grepl("\\_", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub(".*\\_(BM).*", "\\1", colnames(data_list$C2))
Tissue <- gsub(".*\\_(Liver).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Spleen).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Blood).*", "\\1", Tissue)
Tissue[grepl("\\_", Tissue)] <- "Blood"
Tissue

# extract position, like BMLeft, BMSpine
Position <- vector("character", length(Tissue))
Position[grepl("Left", colnames(data_list$C2))] <- "Left"
Position[grepl("Front", colnames(data_list$C2))] <- "Front"
Position[grepl("Right", colnames(data_list$C2))] <- "Right"
Position[grepl("Spine", colnames(data_list$C2))] <- "Spine"

# extract celltype from colnames
Celltype <- gsub(".*\\_.*\\_([A-Z])", "\\1", colnames(data_list$C2))
Celltype[grepl("Tcell", Celltype)] <- "Tcell"
Celltype[grepl("Bcell", Celltype)] <- "Bcell"
Celltype[grepl("Grn", Celltype)] <- "Grn"
Celltype[grepl("Other", Celltype)] <- "Other"
Celltype[grepl("\\_", Celltype)] <- "Unsorted"
Celltype

colnames(data_list$C2)

C2_meta <- data.frame(
  Sample = colnames(data_list$C2),
  Donor = rep("C2", ncol(data_list$C2)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

# View(data_list$C2)
```

## C3

There are two GK_BM_T. One might be typo, originally GK_BM_G.

```{r clean up C3}
# data_list$C3 %>% View
data_list$C3 <- as.data.frame(data_list$C3)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C3[, 1]))
data_list$C3 <- data_list$C3[!is.na(data_list$C3[, 1]), ]
dim(data_list$C3)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C3) <- data_list$C3[,1]
data_list$C3 <- data_list$C3[,-1]

colnames(data_list$C3)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C3 <- data_list$C3[, !colnames(data_list$C3) %in% del]

# delete processed "sum" columns
data_list$C3 <- data_list$C3[, !grepl("Sum", colnames(data_list$C3))] 
data_list$C3 <- data_list$C3[, !grepl("sum", colnames(data_list$C3))] 

# delete processed "shannon" rows
rownames(data_list$C3) [!grepl("^[ATGC]+$", rownames(data_list$C3))]
rownames(data_list$C3) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C3))]
data_list$C3 <- data_list$C3[grepl("^[ATGC\\\"]+$", rownames(data_list$C3)),]

dim(data_list$C3)

# Convert NA values to 0 in the matrix
data_list$C3[is.na(data_list$C3)] <- 0
!any(is.na(data_list$C3))

# Convert character data to numeric
C3_count_numeric <- apply(data_list$C3, 2, as.numeric)
rownames(C3_count_numeric) <- rownames(data_list$C3)

data_list$C3 <- C3_count_numeric

# View sample names
colnames(data_list$C3)

# fix "library" clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C3) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C3))

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C3))
xCondition <- gsub("(.*)\\_.*", "\\1", xCondition)
xCondition[grepl("TK3BMCellsbeforetransplant", xCondition)] <- "time0"
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk[0-9]{1,}).*", "\\1", colnames(data_list$C3))
Time[grepl("\\_", Time)] <- "sac"
Time[grepl("TK3BMCellsbeforetransplant", Time)] <- "time0"
Time

# extract tissue from colnames
Tissue <- gsub(".*\\_(BM).*", "\\1", colnames(data_list$C3))
Tissue <- gsub(".*\\_(Liver).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Spleen).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Blood).*", "\\1", Tissue)
Tissue[grepl("\\_", Tissue)] <- "Blood"
Tissue[grepl("TK3BMCellsbeforetransplant", Tissue)] <- "time0"
Tissue

# extract celltype from colnames
Celltype <- gsub(".*\\_.*\\_([A-Z])", "\\1", colnames(data_list$C3))
Celltype[grepl("TK3BMCellsbeforetransplant", Celltype)] <- "time0"
Celltype[grepl("T", Celltype)] <- "Tcell"
Celltype[grepl("B", Celltype)] <- "Bcell"
Celltype[grepl("G", Celltype)] <- "Grn"
Celltype[grepl("\\_", Celltype)] <- "Unsorted"
Celltype

colnames(data_list$C3)

C3_meta <- data.frame(
  Sample = colnames(data_list$C3),
  Donor = rep("C3", ncol(data_list$C3)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Celltype = Celltype
)

# View(data_list$C3)
```

## C10

```{r clean up C10}
# data_list$C10 %>% View
data_list$C10 <- as.data.frame(data_list$C10)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C10[, 1]))
data_list$C10 <- data_list$C10[!is.na(data_list$C10[, 1]), ]
dim(data_list$C10)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C10) <- data_list$C10[,1]
data_list$C10 <- data_list$C10[,-1]

colnames(data_list$C10)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C10 <- data_list$C10[, !colnames(data_list$C10) %in% del]

# delete processed "sum" columns
data_list$C10 <- data_list$C10[, !grepl("Sum", colnames(data_list$C10))] 
data_list$C10 <- data_list$C10[, !grepl("sum", colnames(data_list$C10))] 

# delete processed "shannon" rows
rownames(data_list$C10) [!grepl("^[ATGC]+$", rownames(data_list$C10))]
rownames(data_list$C10) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C10))]
data_list$C10 <- data_list$C10[grepl("^[ATGC\\\"]+$", rownames(data_list$C10)),]

dim(data_list$C10)

# Convert NA values to 0 in the matrix
data_list$C10[is.na(data_list$C10)] <- 0
!any(is.na(data_list$C10))

# Convert character data to numeric
C10_count_numeric <- apply(data_list$C10, 2, as.numeric)
rownames(C10_count_numeric) <- rownames(data_list$C10)

data_list$C10 <- C10_count_numeric

# View sample names
colnames(data_list$C10)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C10) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C10))

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub(".*\\_.*\\_.*\\_(.*)", "\\1", colnames(data_list$C10))
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk\\_[0-9]{1,}).*", "\\1", colnames(data_list$C10))
Time <- gsub("\\_", "", Time)
Time

# extract tissue from colnames
Tissue <- gsub("(.*)\\_.*\\_.*\\_.*", "\\1", colnames(data_list$C10))
Tissue[grepl("Blood", Tissue)] <- "Blood"
Tissue[grepl("BM", Tissue)] <- "BM"
Tissue[grepl("Liver", Tissue)] <- "Liver"
Tissue[grepl("Spleen", Tissue)] <- "Spleen"
Tissue

# extract position
mix <- gsub("(.*)\\_.*\\_.*\\_.*", "\\1", colnames(data_list$C10))
Position <- vector("character", length(mix))
Position[grepl("BM", mix)] <- gsub("(.*)BM.*", "\\1", mix)[grepl("BM", mix)]
Position

# extract celltype from colnames
mix <- gsub("(.*)\\_.*\\_.*\\_.*", "\\1", colnames(data_list$C10))
Celltype <- vector("character", length(mix))
Celltype[grepl("Blood", mix)] <- gsub(".*Blood(.*)", "\\1", mix)[grepl("Blood", mix)]
Celltype[grepl("BM", mix)] <- gsub(".*BM(.*)", "\\1", mix)[grepl("BM", mix)]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype

colnames(data_list$C10)

C10_meta <- data.frame(
  Sample = colnames(data_list$C10),
  Donor = rep("C10", ncol(data_list$C10)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

# View(data_list$C10)
```

## C15

```{r clean up C15}
# data_list$C15 %>% View
data_list$C15 <- as.data.frame(data_list$C15)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C15[, 1]))
data_list$C15 <- data_list$C15[!is.na(data_list$C15[, 1]), ]
dim(data_list$C15)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C15) <- data_list$C15[,1]
data_list$C15 <- data_list$C15[,-1]

colnames(data_list$C15)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C15 <- data_list$C15[, !colnames(data_list$C15) %in% del]

# delete processed "sum" columns
data_list$C15 <- data_list$C15[, !grepl("Sum", colnames(data_list$C15))] 
data_list$C15 <- data_list$C15[, !grepl("sum", colnames(data_list$C15))] 

# delete processed "shannon" rows
rownames(data_list$C15) [!grepl("^[ATGC]+$", rownames(data_list$C15))]
rownames(data_list$C15) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C15))]
data_list$C15 <- data_list$C15[grepl("^[ATGC\\\"]+$", rownames(data_list$C15)),]

dim(data_list$C15)

# Convert NA values to 0 in the matrix
data_list$C15[is.na(data_list$C15)] <- 0
!any(is.na(data_list$C15))

# Convert character data to numeric
C15_count_numeric <- apply(data_list$C15, 2, as.numeric)
rownames(C15_count_numeric) <- rownames(data_list$C15)

data_list$C15 <- C15_count_numeric

# View sample names
colnames(data_list$C15)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C15) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C15))

# extract time from colnames
Time <- gsub(".*(wk[0-9]{1,}).*", "\\1", colnames(data_list$C15))
Time[grepl("\\_", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- vector("character", ncol(data_list$C15))
Tissue[grepl("BM", colnames(data_list$C15))] <- "BM"
Tissue[Tissue == ""] <- "Blood"
Tissue

# extract position
mix <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C15))
Position <- vector("character", ncol(data_list$C15))
Position[grepl("BM", mix)] <- gsub("BM(.*)", "\\1", mix)[grepl("BM", mix)]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C15))
Celltype[grepl("\\_",colnames(data_list$C15))] <- gsub(".*\\_(.*)", "\\1", colnames(data_list$C15))[grepl("\\_",colnames(data_list$C15))]

Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C15)

C15_meta <- data.frame(
  Sample = colnames(data_list$C15),
  Donor = rep("C15", ncol(data_list$C15)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

# View(data_list$C15)
```

## C21

```{r clean up C21}
# data_list$C21 %>% View
data_list$C21 <- as.data.frame(data_list$C21)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C21[, 1]))
data_list$C21 <- data_list$C21[!is.na(data_list$C21[, 1]), ]
dim(data_list$C21)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C21) <- data_list$C21[,1]
data_list$C21 <- data_list$C21[,-1]

colnames(data_list$C21)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C21 <- data_list$C21[, !colnames(data_list$C21) %in% del]

# delete processed "sum" columns
data_list$C21 <- data_list$C21[, !grepl("Sum", colnames(data_list$C21))] 
data_list$C21 <- data_list$C21[, !grepl("sum", colnames(data_list$C21))] 

# delete processed "shannon" rows
rownames(data_list$C21) [!grepl("^[ATGC]+$", rownames(data_list$C21))]
rownames(data_list$C21) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C21))]
data_list$C21 <- data_list$C21[grepl("^[ATGC\\\"]+$", rownames(data_list$C21)),]

dim(data_list$C21)

# Convert NA values to 0 in the matrix
data_list$C21[is.na(data_list$C21)] <- 0
!any(is.na(data_list$C21))

# Convert character data to numeric
C21_count_numeric <- apply(data_list$C21, 2, as.numeric)
rownames(C21_count_numeric) <- rownames(data_list$C21)

data_list$C21 <- C21_count_numeric

# View sample names
colnames(data_list$C21)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C21) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C21))

# extract time from colnames
Time <- gsub("(wk[0-9]{1,}).*", "\\1", colnames(data_list$C21))
Time[grepl("BM|Spleen|Liver", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub("(BM|Spleen|Liver).*", "\\1", colnames(data_list$C21))
Tissue[grepl("wk", Tissue)] <- "Blood"
Tissue

# extract position
Position <- vector("character", ncol(data_list$C21))
Position[grepl("BM", colnames(data_list$C21))] <- gsub("BM(.*)\\_.*", "\\1", colnames(data_list$C21))[grepl("BM", colnames(data_list$C21))]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C21))
Celltype[grepl("\\_", colnames(data_list$C21))] <- gsub(".*\\_(.*)", "\\1", colnames(data_list$C21))[grepl("\\_", colnames(data_list$C21))]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "unsorted"] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C21)

C21_meta <- data.frame(
  Sample = colnames(data_list$C21),
  Donor = rep("C21", ncol(data_list$C21)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

# View(data_list$C21)
```

## C22

```{r clean up C22}
# data_list$C22 %>% View
data_list$C22 <- as.data.frame(data_list$C22)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C22[, 1]))
data_list$C22 <- data_list$C22[!is.na(data_list$C22[, 1]), ]
dim(data_list$C22)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C22) <- data_list$C22[,1]
data_list$C22 <- data_list$C22[,-1]

colnames(data_list$C22)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C22 <- data_list$C22[, !colnames(data_list$C22) %in% del]

# delete processed "sum" columns
data_list$C22 <- data_list$C22[, !grepl("Sum", colnames(data_list$C22))] 
data_list$C22 <- data_list$C22[, !grepl("sum", colnames(data_list$C22))] 

# delete processed "shannon" rows
rownames(data_list$C22) [!grepl("^[ATGC]+$", rownames(data_list$C22))]
rownames(data_list$C22) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C22))]
data_list$C22 <- data_list$C22[grepl("^[ATGC\\\"]+$", rownames(data_list$C22)),]

dim(data_list$C22)

# Convert NA values to 0 in the matrix
data_list$C22[is.na(data_list$C22)] <- 0
!any(is.na(data_list$C22))

# Convert character data to numeric
C22_count_numeric <- apply(data_list$C22, 2, as.numeric)
rownames(C22_count_numeric) <- rownames(data_list$C22)

data_list$C22 <- C22_count_numeric

# View sample names
colnames(data_list$C22)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C22) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C22))

# extract time from colnames
Time <- gsub(".*\\_(wk\\_[0-9]{1,}).*", "\\1", colnames(data_list$C22))
Time[grepl("\\_", Time)] <- "sac"
Time[Time == "Spleen"] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub("(.*)\\_.*\\_.*", "\\1", colnames(data_list$C22))
Tissue <- gsub("(.*)\\_.*", "\\1", Tissue)
Tissue[grepl("wk", Tissue)] <- "Blood"
Tissue

# extract position
Position <- vector("character", ncol(data_list$C22))
Position[grepl("BM", colnames(data_list$C22))] <- gsub("BM\\_(.*)\\_.*", "\\1", colnames(data_list$C22))[grepl("BM", colnames(data_list$C22))]
Position[grepl("BM", Position)] <- gsub("BM\\_(.*)", "\\1", Position)[grepl("BM", Position)]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C22))
Celltype[grepl("BM\\_.*\\_.*", colnames(data_list$C22))] <- gsub("BM\\_.*\\_(.*)", "\\1", colnames(data_list$C22))[grepl("BM\\_.*\\_.*", colnames(data_list$C22))]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "unsorted"] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C22)

C22_meta <- data.frame(
  Sample = colnames(data_list$C22),
  Donor = rep("C22", ncol(data_list$C22)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

# View(data_list$C22)
```

## C23

```{r clean up C23}
# data_list$C23 %>% View
data_list$C23 <- as.data.frame(data_list$C23)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C23[, 1]))
data_list$C23 <- data_list$C23[!is.na(data_list$C23[, 1]), ]
dim(data_list$C23)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C23) <- data_list$C23[,1]
data_list$C23 <- data_list$C23[,-1]

colnames(data_list$C23)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C23 <- data_list$C23[, !colnames(data_list$C23) %in% del]

# delete processed "sum" columns
data_list$C23 <- data_list$C23[, !grepl("Sum", colnames(data_list$C23))] 
data_list$C23 <- data_list$C23[, !grepl("sum", colnames(data_list$C23))] 

# delete processed "shannon" rows
rownames(data_list$C23) [!grepl("^[ATGC]+$", rownames(data_list$C23))]
rownames(data_list$C23) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C23))]
data_list$C23 <- data_list$C23[grepl("^[ATGC\\\"]+$", rownames(data_list$C23)),]

dim(data_list$C23)

# Convert NA values to 0 in the matrix
data_list$C23[is.na(data_list$C23)] <- 0
!any(is.na(data_list$C23))

# Convert character data to numeric
C23_count_numeric <- apply(data_list$C23, 2, as.numeric)
rownames(C23_count_numeric) <- rownames(data_list$C23)

data_list$C23 <- C23_count_numeric

# View sample names
colnames(data_list$C23)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C23) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C23))

# extract time from colnames
Time <- gsub("(wk[0-9]{1,}).*", "\\1", colnames(data_list$C23))
Time[grepl("BM|Spleen|Liver|Sac", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub("(BM|Spleen|Liver).*", "\\1", colnames(data_list$C23))
Tissue[grepl("wk", Tissue)] <- "Blood"
Tissue[grepl("Sac", Tissue)] <- "sac"
Tissue

# extract position
Position <- vector("character", ncol(data_list$C23))
Position[grepl("BM\\_.*\\_.*", colnames(data_list$C23))] <- gsub("BM\\_(.*)\\_.*", "\\1", colnames(data_list$C23))[grepl("BM\\_.*\\_.*", colnames(data_list$C23))]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C23))
Celltype[grepl("\\_", colnames(data_list$C23))] <- gsub(".*\\_(.*)", "\\1", colnames(data_list$C23))[grepl("\\_", colnames(data_list$C23))]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "unsorted"] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C23)

C23_meta <- data.frame(
  Sample = colnames(data_list$C23),
  Donor = rep("C23", ncol(data_list$C23)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

# View(data_list$C23)
```

## Combine donors

```{r combine}
Metadata <- dplyr::bind_rows(C1_meta, C2_meta, C3_meta, C10_meta, C15_meta, C22_meta, C21_meta, C23_meta)
Metadata[is.na(Metadata)] <- ""

data_list <- lapply(data_list, function(mat)  as.data.frame(t(mat)))
CountData <- dplyr::bind_rows(data_list)
CountData <- t(CountData)
```

# Save to a barbieQ object

**Note that clones (barcodes) between different donors don't match.** 

**Clones from each donor were labelled separately, using the same barcode (lentiviral) library.**

**Same barcode across donors doesn't refer to the same clone!**

**For convenience, herein we combined samples from all donors together.**

```{r save-to-barbieQ}
## convert all NA into zero and continue
CountData %>% is.na() %>% any()
CountData[is.na(CountData)] <- 0
CountData %>% is.na() %>% any()

xenoHSPC <- createBarbieQ(object = CountData, sampleMetadata = Metadata)
```

# Tag top barcodes

```{r tag top}
dim(xenoHSPC)
## find minimum group size
targets <- as.data.frame(xenoHSPC$sampleMetadata) %>% 
  mutate(Group = paste0(Donor, Celltype))
targets$Group %>% table()

## select sample to consider, exclude TP0 when tagging
xenoHSPC <- barbieQ::tagTopBarcodes(barbieQ = xenoHSPC, 
  # activeSamples = xenoHSPC$sampleMetadata$Sample != "library",
  nSampleThreshold = 3
  )

rowData(xenoHSPC)$isTopBarcode$isTop %>% table()
```

According to the original paper, for each donor, barcode library contains around 800 unique barcodes.

Check out filtered "top" barcodes.

```{r}
n_uni_BC <- xenoHSPC@assays@data$isTopAssay %>% colSums()

n_uni_BC %>% sort(decreasing = T)
```


```{r vis tag}
barbieQ::plotBarcodePareto(barbieQ = xenoHSPC)
barbieQ::plotBarcodeSankey(barbieQ = xenoHSPC)
```

# Cluster barcodes

## FS1C: pairwise correlation

```{r bar cor, fig.width=4, fig.height=4}
plotBarcodePairCorrelation(
  barbieQ = xenoHSPC[rowData(xenoHSPC)$isTopBarcode$isTop,], 
  propThresh = 0.005, transformation = "none") + 
  theme(legend.position = "top") + 
  labs(title = "HSPC xenograft data") -> fs1c

fs1c
```

## FS1F: clusters barcodes

```{r cluster barcodes}
## cluster barcoes
xenoHSPC_top0 <- barbieQ::clusterCorrelatingBarcodes(
  barbieQ = xenoHSPC[rowData(xenoHSPC)$isTopBarcode$isTop,], 
  transformation = "none", propThresh = 0.005)

p_cluster_list <- barbieQ::inspectCorrelatingBarcodes(xenoHSPC_top0)

p_cluster_list

layer_index <- which(sapply(p_cluster_list$p_cluster$layers, function(x) inherits(x$geom, "GeomTextRepel")))
## update the size for that layer
p_cluster_list$p_cluster$layers[[layer_index]]$aes_params$size <- 1.4  # new size
fs1f <- p_cluster_list$p_cluster + theme(legend.position = "none")

fs1f
```

## merge clusters

```{r}
xenoHSPC_merged <- barbieQ::mergeCorrelatingBarcodes(barbieQ_clustered = xenoHSPC_top0, barbieQ_toMerge = xenoHSPC)
dim(xenoHSPC_merged)

xenoHSPC_merged@metadata$predicted_barcode_clusters
xenoHSPC_merged@metadata$removed_barcodes
```

# Tag top barcode after collapsing

## FS1H,K: inspect top/bottom

```{r tag after cor, fig.width=5, fig.height=4}
xenoHSPC_merged <- tagTopBarcodes(
  barbieQ = xenoHSPC_merged,
  # activeSamples = xenoHSPC$sampleMetadata$Sample != "library",
  nSampleThreshold = 3)

barbieQ::plotBarcodePareto(barbieQ = xenoHSPC_merged) +
  ylim(-8, 12) + 
  annotate("text", x = c(pi * 0.05), y = c(12), 
           label = c("Relative proportion %"), color = "#999966", 
           size = 3, angle = 0, fontface = "bold", hjust = 1) -> fs1h

fs1h
```

```{r, fig.width=3, fig.height=4}
barbieQ::plotBarcodeSankey(barbieQ = xenoHSPC_merged) +
  theme(legend.position = "top")  -> fs1k

fs1k
```

## Filter "top" barcodes

```{r}
xenoHSPC_top <- xenoHSPC_merged[rowData(xenoHSPC_merged)$isTopBarcode$isTop,]
```

## Heatmap

```{r echo=TRUE}
## pre-filtering
Heatmap(name = "asin-sqrt prop.",
  matrix = asin(sqrt(assays(xenoHSPC_merged)$proportion)), 
  row_title = paste0(nrow(xenoHSPC_merged), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)
```

```{r hp}
## post-filtering
Heatmap(name = "asin-sqrt prop.",
  matrix = asin(sqrt(assays(xenoHSPC_top)$proportion)), 
  row_title = paste0(nrow(xenoHSPC_top), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)
```

# Save rds

```{r save rds}
save(xenoHSPC, xenoHSPC_merged, xenoHSPC_top, file = "output/xenoHSPC_barbieQ.rda")

```

# Top10

```{r}
tagTop10 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 10
  # summary(isTop)
  return(isTop)
}

tester <- tagTop10(assays(xenoHSPC_merged)$CPM[,1])
summary(tester)

colTags <- lapply(as.data.frame(assays(xenoHSPC_merged)$CPM), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat[,xenoHSPC_merged$sampleMetadata$Sample != "library"]) >= 1
summary(colTagVec)

xenoHSPC_merged_top10 <- xenoHSPC_merged
rowData(xenoHSPC_merged_top10)$isTopBarcode <- DataFrame(isTop = colTagVec)
summary(rowData(xenoHSPC_merged_top10)$isTopBarcode$isTop)

barbieQ::plotBarcodePareto(barbieQ = xenoHSPC_merged_top10)
barbieQ::plotBarcodeSankey(barbieQ = xenoHSPC_merged_top10)
```

# FigureS1-HSPC xenograft

```{r, fig.width=4, fig.height=16}
layout = "
C
F
H
K
"
fs1_xeno <- (
  wrap_elements(fs1c + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
    wrap_elements(fs1f + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(fs1h + theme(plot.margin = unit(rep(0,4), "cm"))) +
    wrap_elements(fs1k + theme(plot.margin = unit(rep(0,4), "cm")))
  ) + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("C","F","H", "K"))) &
  theme(
    plot.tag = element_text(size = 20, face = "bold", family = "arial"),
    axis.title = element_text(size = 17),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

fs1_xeno
```


```{r}
ggsave(
  filename = "output/fs1_xeno.png",
  plot = fs1_xeno,
  width = 4,
  height = 16,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [fs1_xenoHSPC](../output/fs1_xeno.png)