---
title: "barbieQ_paper Supplementary 2: Type I error rate using `HSPC xenograft data`"
author: "Liyang Fei"
date: |
  Initiate: 2025-04-11 
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Links to Type I error rate assessment using other datasets in the barbieQ paper: 
 
 - [Mixture data](barbieQ_paper_Figure3.html)
 
 - [AML data](barbieQ_paper_FigureS2_AML.html)
 
 - [HSPC xenograft data]()
 
# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)
library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)
# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
# library(bartools)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/F3_simulation.R") ## for negative simulation
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Load preprocessed data 

Links to [HSPC xenograft data preprocessing](barbieQ_paper_FigureS1_xenoHSPC.html)

```{r save rds}
load("output/xenoHSPC_barbieQ.rda")
```

# Samples level QC

## PCA

"library" samples and "time0" samples differ from recipient samples.

```{r tsne}
pca_results <- (assays(xenoHSPC_top)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp( scale. = TRUE)

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  xenoHSPC_top$sampleMetadata
)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Tissue)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

```

Delete "library" and "time0" samples in subsequent analysis.

```{r}
xenoHSPC_recipient <- xenoHSPC_top[,!xenoHSPC_top$sampleMetadata$Tissue %in% c("library", "time0")]

dim(xenoHSPC_recipient)

xenoHSPC_recipient$sampleMetadata$Time <- factor(
  xenoHSPC_recipient$sampleMetadata$Time, 
  levels =  c("wk4", "wk6", "wk7", "wk9", "wk10", "wk11", "wk12", "wk14", "wk16", "wk18", "wk19", "wk20", "wk22", "wk24", "wk25", "wk27", "wk30", "wk33", "wk36", "wk37", "wk42", "wk43", "wk47", "sac") 
  )     

wk <- gsub("wk(\\d+)", "\\1", xenoHSPC_recipient$sampleMetadata$Time)
wk <- gsub("sac", 50, wk) ## assign "sac" pseudotime by wk50
wk <- as.numeric(wk)

xenoHSPC_recipient$sampleMetadata$Week <- wk
``` 

```{r}
pca_results <- (assays(xenoHSPC_recipient)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp()

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  xenoHSPC_recipient$sampleMetadata
)

```

ggplot PCA

```{r}
ggplot(pca_df, aes(x = PC1, y = PC2, color = Donor)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Week)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Tissue)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```


```{r, fig.width=10, fig.height=10}
ggplot(pca_df, aes(x = PC1, y = PC2, color = Week)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_linedraw() +
  facet_grid(Donor~Celltype)
```


## Plot propportion

```{r bartools, fig.width=30, fig.height=10}
## order samples
order_sample <- xenoHSPC_recipient$sampleMetadata %>% with(order(Donor, Tissue, Celltype, Time))

## assign unique names to samples
ps.counts <- assay(xenoHSPC_recipient)
colnames(ps.counts) <- make.unique(xenoHSPC_recipient$sampleMetadata$Donor)

## create bartools object
Donor_dge <- DGEList(
  counts = ps.counts, 
  group = xenoHSPC_recipient$sampleMetadata$Donor)

## plot 
plotBarcodeHistogram(Donor_dge, orderSamples = colnames(ps.counts)[order_sample])
```

## Heatmap

```{r xenoHSPC, fig.width=8, fig.height=6}
barbieQ::plotBarcodeHeatmap(
  xenoHSPC_recipient, sampleMetadata = xenoHSPC_recipient$sampleMetadata[,c(2,5,6,7,8)], splitSamples = T, sampleGroup = "Donor")
```

# Subset samples of indivisual donors {.tabset}

QC samples

## Donor C21

```{r C21}
## subet C21 Donor samples
C21 <- xenoHSPC_recipient[, xenoHSPC_recipient$sampleMetadata$Donor == "C21"]

## flag C21 samples of extremely low libsize
colSums(assay(C21)) %>% sort()
flag_low_C21 <- colSums(assay(C21)) < 100
## remove the low libsize samples
C21 <- C21[, !flag_low_C21]
colSums(assay(C21)) %>% sort()

## tag top barcodes within C21
C21$sampleMetadata %>% with(paste(Donor, Tissue, Celltype, Time)) %>% table()
C21 <- tagTopBarcodes(C21, nSampleThreshold = 1)
plotBarcodePareto(C21)
plotBarcodeSankey(C21)
## filter top barcodes
C21 <- C21[C21@elementMetadata$isTopBarcode$isTop,]

barbieQ::plotBarcodeHeatmap(C21, sampleMetadata = C21$sampleMetadata[,c(5,6,7,8)])
```

- PCA

```{r}
pca_results <- (assays(C21)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp()

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  C21$sampleMetadata
)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Week)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Tissue)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Celltype)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```

## Donor C22

```{r C22}
## subet C21 Donor samples
C22 <- xenoHSPC_recipient[, xenoHSPC_recipient$sampleMetadata$Donor == "C22"]

## flag C22 samples of extremely low libsize
colSums(assay(C22)) %>% sort()

## tag top barcodes within C22
C22$sampleMetadata %>% with(paste(Donor, Tissue, Celltype, Time)) %>% table()
C22 <- tagTopBarcodes(C22, nSampleThreshold = 1)
plotBarcodePareto(C22)
plotBarcodeSankey(C22)
## filter top barcodes
C22 <- C22[C22@elementMetadata$isTopBarcode$isTop,]

barbieQ::plotBarcodeHeatmap(C22, sampleMetadata = C22$sampleMetadata[,c(5,6,7,8)])
```

- PCA

```{r}
pca_results <- (assays(C22)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp()

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  C22$sampleMetadata
)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Week)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Tissue)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Celltype)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```

## Donor C23

```{r C23}
## subet C23 Donor samples
C23 <- xenoHSPC_recipient[, xenoHSPC_recipient$sampleMetadata$Donor == "C23"]

## flag C23 samples of extremely low libsize
colSums(assay(C23)) %>% sort()

## tag top barcodes within C23
C23$sampleMetadata %>% with(paste(Donor, Tissue, Celltype, Time)) %>% table()
C23 <- tagTopBarcodes(C23, nSampleThreshold = 1)
plotBarcodePareto(C23)
plotBarcodeSankey(C23)
## filter top barcodes
C23 <- C23[C23@elementMetadata$isTopBarcode$isTop,]

barbieQ::plotBarcodeHeatmap(C23, sampleMetadata = C23$sampleMetadata[,c(5,6,7,8)])
```

- PCA

```{r}
pca_results <- (assays(C23)$proportion) %>% sqrt() %>% asin() %>% 
  t() %>% 
  prcomp()

var_pct <- pca_results$sdev^2 / sum(pca_results$sdev^2) * 100
var_pct

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  C23$sampleMetadata
)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Week)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Tissue)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Celltype)) +
  geom_point(size = 3) +
  labs(title = "PCA on asin-sqrt prop.", 
       x = paste0("PC1 (", round(var_pct[1], 2), " %)"), 
       y = paste0("PC2 (", round(var_pct[2], 2), " %)")) +
  theme_classic() +
  theme(aspect.ratio = 1)
```

# Plot proportion {.tabset}

## Donor C21

```{r}
## order samples
C21_id <- C21$sampleMetadata %>% with(paste(Tissue, Celltype, Position, Time, sep = "."))
order_C21 <- order(C21$sampleMetadata %>% with(paste(Tissue, Celltype, Position, Week, sep = ".")))

## create bartools object
C21_count <- assay(C21)
colnames(C21_count) <- C21_id

C21_dge <- DGEList(
  counts = C21_count, 
  group = C21$sampleMetadata$Donor)

## plot 
plotBarcodeHistogram(C21_dge, orderSamples = C21_id[order_C21]) + 
  labs(title = "HSPC xenograft: Donor C21") -> p_prop_C21
```

## Donor C22

```{r}
## order samples
C22_id <- C22$sampleMetadata %>% with(paste(Tissue, Position, Celltype, Time, sep = "."))
order_C22 <- order(C22$sampleMetadata %>% with(paste(Tissue, Position, Celltype, Week, sep = ".")))

## create bartools object
C22_count <- assay(C22)
colnames(C22_count) <- C22_id

C22_dge <- DGEList(
  counts = C22_count, 
  group = C22$sampleMetadata$Donor)

## plot 
plotBarcodeHistogram(C22_dge, orderSamples = C22_id[order_C22]) + 
  labs(title = "HSPC xenograft: Donor C22") -> p_prop_C22
```

## Donor C23

```{r}
## order samples
C23_id <- C23$sampleMetadata %>% with(paste(Tissue, Position, Celltype, Time, sep = "."))
order_C23 <- order(C23$sampleMetadata %>% with(paste(Tissue, Position, Celltype, Week, sep = ".")))

## create bartools object
C23_count <- assay(C23)
colnames(C23_count) <- C23_id

C23_dge <- DGEList(
  counts = C23_count, 
  group = C23$sampleMetadata$Donor)

## plot 
plotBarcodeHistogram(C23_dge, orderSamples = C23_id[order_C23]) + 
  labs(title = "HSPC xenograft: Donor C23") -> p_prop_C23
```

## FS2A,C,E: prop. of three donors

```{r, fig.width= 4, fig.height=15}
p_prop_C21 / p_prop_C22 / p_prop_C23
```

# Type I error rate {.tablet}

## Donor C21

- Negative simulation

**avoid running this chunk during rendering. `global_all_loops` was saved from the first run.**

```{r eval=FALSE, fig.height=8, fig.width=16, include=TRUE}
results_negsi <- suppressMessages({
  random_sampling(barbieQ = C21, loop_times = 100)
  })

## `random_sampling()` save the simulations to global environment as `global_all_loops`
save(global_all_loops, file = "output/xenoC21_negative_simulation.rda")
```

- load results

```{r}
## loading random sampling results to avoid running it repeatedly
load("output/xenoC21_negative_simulation.rda")

# extract results from the loops
end_sampling <- floor(ncol(C21)/2)

## extract FPR
all_FPR <- lapply(seq(3:end_sampling), function(n) {
  global_all_loops[[n]]$FPR
})
df_FPR_C21 <- do.call(rbind, all_FPR)
```

- plot FPR

```{r}
df_FPR_C21$N_Samples <- as.factor(df_FPR_C21$N_Samples)

## Diff_Prop
## reshape the data to fit methods for testing
df_FPR_C21_long <- df_FPR_C21 %>%
  pivot_longer(cols = starts_with("Prop_"),
               names_to = "Method",
               values_to = "Pval_Prop") %>%
  mutate(Method = factor(Method, levels = c("Prop_asin", "Prop_logit", "Prop_noTrans"),
                         labels = c("asin-sqrt", "logit", "no trans.")))

P_FPR_prop_C21 <- ggplot(
  df_FPR_C21_long, aes(x = factor(N_Samples), y = Pval_Prop, color = Method)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       # title = "C21",
       subtitle = "Differential proportion test") +
  # facet_wrap(~Method, scales = "free_y") +
  theme_classic() +
  theme(legend.position = "top")

## Diff_Occ
P_FPR_occ_C21 <- ggplot(
  df_FPR_C21, aes(x = factor(N_Samples), y = Occ_firth)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Differential occurrrence test" = "black")) +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       titlte = "",
       subtitle = "Differential occurrence test") +
  theme_classic() +
  theme(legend.position = "top")

```

## Donor C22

- Negative simulation

**avoid running this chunk during rendering. `global_all_loops` was saved from the first run.**

```{r eval=FALSE, fig.height=8, fig.width=16, include=TRUE}
results_negsi <- suppressMessages({
  random_sampling(barbieQ = C22, loop_times = 100)
  })

## `random_sampling()` save the simulations to global environment as `global_all_loops`
save(global_all_loops, file = "output/xenoC22_negative_simulation.rda")
```

- load results

```{r}
## loading random sampling results to avoid running it repeatedly
load("output/xenoC22_negative_simulation.rda")

# extract results from the loops
end_sampling <- floor(ncol(C22)/2)

## extract FPR
all_FPR <- lapply(seq(3:end_sampling), function(n) {
  global_all_loops[[n]]$FPR
})
df_FPR_C22 <- do.call(rbind, all_FPR)
```

- plot FPR

```{r}
df_FPR_C22$N_Samples <- as.factor(df_FPR_C22$N_Samples)

## Diff_Prop
## reshape the data to fit methods for testing
df_FPR_C22_long <- df_FPR_C22 %>%
  pivot_longer(cols = starts_with("Prop_"),
               names_to = "Method",
               values_to = "Pval_Prop") %>%
  mutate(Method = factor(Method, levels = c("Prop_asin", "Prop_logit", "Prop_noTrans"),
                         labels = c("asin-sqrt", "logit", "no trans.")))

P_FPR_prop_C22 <- ggplot(
  df_FPR_C22_long, aes(x = factor(N_Samples), y = Pval_Prop, color = Method)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       # title = "C22",
       subtitle = "Differential proportion test") +
  # facet_wrap(~Method, scales = "free_y") +
  theme_classic() +
  theme(legend.position = "top")

## Diff_Occ
P_FPR_occ_C22 <- ggplot(
  df_FPR_C22, aes(x = factor(N_Samples), y = Occ_firth)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Differential occurrrence test" = "black")) +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       titlte = "",
       subtitle = "Differential occurrence test") +
  theme_classic() +
  theme(legend.position = "top")

```

## Donor C23

- Negative simulation

**avoid running this chunk during rendering. `global_all_loops` was saved from the first run.**

```{r eval=FALSE, fig.height=8, fig.width=16, include=TRUE}
results_negsi <- suppressMessages({
  random_sampling(barbieQ = C23, loop_times = 100)
  })

## `random_sampling()` save the simulations to global environment as `global_all_loops`
save(global_all_loops, file = "output/xenoC23_negative_simulation.rda")
```

- load results

```{r}
## loading random sampling results to avoid running it repeatedly
load("output/xenoC23_negative_simulation.rda")

# extract results from the loops
end_sampling <- floor(ncol(C23)/2)

## extract FPR
all_FPR <- lapply(seq(3:end_sampling), function(n) {
  global_all_loops[[n]]$FPR
})
df_FPR_C23 <- do.call(rbind, all_FPR)
```

- plot FPR

```{r}
df_FPR_C23$N_Samples <- as.factor(df_FPR_C23$N_Samples)

## Diff_Prop
## reshape the data to fit methods for testing
df_FPR_C23_long <- df_FPR_C23 %>%
  pivot_longer(cols = starts_with("Prop_"),
               names_to = "Method",
               values_to = "Pval_Prop") %>%
  mutate(Method = factor(Method, levels = c("Prop_asin", "Prop_logit", "Prop_noTrans"),
                         labels = c("asin-sqrt", "logit", "no trans.")))

P_FPR_prop_C23 <- ggplot(
  df_FPR_C23_long, aes(x = factor(N_Samples), y = Pval_Prop, color = Method)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       # title = "C23",
       subtitle = "Differential proportion test") +
  # facet_wrap(~Method, scales = "free_y") +
  theme_classic() +
  theme(legend.position = "top")

## Diff_Occ
P_FPR_occ_C23 <- ggplot(
  df_FPR_C23, aes(x = factor(N_Samples), y = Occ_firth)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Differential occurrrence test" = "black")) +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       titlte = "",
       subtitle = "Differential occurrence test") +
  theme_classic() +
  theme(legend.position = "top")

```

## FS2B,D,F: FPR of three donors

```{r, fig.width=8, fig.height=5}
(P_FPR_prop_C21 + theme(legend.position = "top")) + (P_FPR_occ_C21) -> P_FPR_C21

(P_FPR_prop_C22 + theme(legend.position = "top")) + (P_FPR_occ_C22) -> P_FPR_C22

(P_FPR_prop_C23 + theme(legend.position = "top")) + (P_FPR_occ_C23) -> P_FPR_C23

P_FPR_C21
P_FPR_C22
P_FPR_C23
```

# Figure S2-HSPC xenograft

```{r, fig.width=12, fig.height=15}
layout = "
ABB
CDD
EFF
"
fs2_xeno <- (
    wrap_elements(p_prop_C21 + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
  wrap_elements(P_FPR_C21 + theme(legend.position = "top") + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(p_prop_C22 + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
  wrap_elements(P_FPR_C22 + theme(legend.position = "top") + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(p_prop_C23 + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
  wrap_elements(P_FPR_C23 + theme(legend.position = "top") + theme(plot.margin = unit(rep(0,4), "cm")))
) + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("A", "B", "C", "D", "E", "F"))) &
  theme(
    plot.tag = element_text(size = 20, face = "bold", family = "arial"),
    axis.title = element_text(size = 17),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

fs2_xeno
```

```{r}
ggsave(
  filename = "output/fs2_xeno.png",
  plot = fs2_xeno,
  width = 12,
  height = 15,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [fs2_xenoHSPC](../output/fs2_xeno.png)