---
title: "barbieQ_paper_Figure2 & Supplementary1: Preprocessing."
author: "Liyang Fei"
date: |
  Initiated: 2025-09-07  
  Rendered: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(SummarizedExperiment)
library(S4Vectors)
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
```

# Install barbieQ

```{r load barbieQ}
## install the latest devel version of barbieQ
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

# Clean up datasets {.tabset .tabset-fade .tabset-pills}

## Monkey HSPC

```{r}
## data is separated by "\t", so use read.table()
count <- read.table("data/WuC/monkey_ZG66.txt", header = TRUE, sep = "\t", row.names = 1)

## rownames are barcode sequence, basically consist of ATGCs: "^[ATGC]+$"
# lapply(rownames(count), function(x) grepl("^[ATGC]+$", x)) %>% unlist() %>% table()

## read metadata of samples.
metadata <- read.table("data/WuC/monkey_ZG66_metadata.txt", header = 1, row.names = 1)

## update the variable names following camelCase style.
colnames(metadata)[grepl("Cell_type", colnames(metadata))] <- "Celltype"
# print(colnames(metadata))

## Timepoint and Months variables in the `metadata` are simply repetition, delete one
metadata$Timepoint <- NULL

## create an object to store the barcode count matrix and sample metadata.
monkeyHSPC <- barbieQ::createBarbieQ(object = count, sampleMetadata = metadata)
```

```{r}
## clean up sample metadata
updateSampleMetadata <- monkeyHSPC$sampleMetadata %>% 
  as.data.frame() %>% 
  select(Celltype, Months) %>% 
  mutate(Phase = ifelse(Months < 6, "early", ifelse(Months >=55, "late", "mid"))) %>% 
  mutate(Celltype = gsub("(Gr).*", "\\1", Celltype))
## update sample metadata
monkeyHSPC$sampleMetadata <- S4Vectors::DataFrame(updateSampleMetadata)
## add more info
metadata(monkeyHSPC) <- list("Monkey" = "ZG66")

## subset samples by excluding the late time points where samples were sorted into NK subtypes - we don;t need them here
monkeyHSPC <- monkeyHSPC[, monkeyHSPC$sampleMetadata$Phase != "late"]
```

## AML data

## Mouse HSPC

## Mixture data

# Identify correlating barcodes {.tabset .tabset-fade .tabset-pills}

## Monkey HSPC

Tag top barcodes in advance to save the time for running 

```{r}
## check min group size
table(monkeyHSPC$sampleMetadata$Celltype)

monkeyHSPC <- barbieQ::tagTopBarcodes(monkeyHSPC, nSampleThreshold = 6)
monkeyHSPC@elementMetadata$isTopBarcode %>% table()
```

```{r}
p_monkey_pair_cor <- plotBarcodePairCorrelation(barbieQ = monkeyHSPC[rowData(monkeyHSPC)$isTopBarcode$isTop,])
p_monkey_pair_cor
```


## AML data

## Mouse HSPC

## Mixture data

# Filter low-count barcodes {.tabset .tabset-fade .tabset-pills}

## Monkey HSPC

```{r}

```

## AML data

## Mouse HSPC

## Mixture data

# Figure 2
