---
title: "barbieQ_paper Supplementary 1: Preprocessing `Mixture data`"
author: "Liyang Fei"
date: |
  Initiate: 2025-09-07
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Links to preprocessing other datasets in the barbieQ paper: 
 
 - [Monkey HSPC](barbieQ_paper_Figure2.html)
 
 - [AML data](barbieQ_paper_FigureS1_AML.html)
 
 - [HSPC xenograft data](barbieQ_paper_FigureS1_xenoHSPC.html)
 
 - [Mixture data]()

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(data.table) # for data.table()
library(SummarizedExperiment)
library(S4Vectors)
library(grid)
library(ggplot2)
library(ggbreak)
library(ggnewscale) # ggnewscale::new_scale_fill() 
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(magick) # image_read()
library(eulerr)
library(edgeR)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Clean up data

```{r}
EV1 <- read.csv("data/DEBRA/msb199195-sup-0004-datasetev1.csv", row.names = 1)
```

Here, TechRep 1 and 2 only means two replicates under the same condition. They don't match between conditions.

The actual "block" should be `paste0(Subset, BarcodeSize, Perturbation)`.

```{r clean up data}
sampleName <- colnames(EV1)

## extract P1P2, nulls and perturbs
Subset <- gsub("m_null_.*", "perturb", sampleName)
Subset <- gsub("null.*", "null", Subset)
Subset[Subset=="P1"] <- "premix1"
Subset[Subset=="P2"] <- "premix2"

## extract cell size (Barcode size, i.e. total number of Barcodes in each sample)
BarcodeSize <- gsub(".*null_(\\d+)\\..*", "\\1", sampleName)
BarcodeSize <- gsub("(P\\d)", "", BarcodeSize)

## extract perturbation level in the perturb samples
Perturbation <- gsub(".*p(\\d+).*", "\\1", sampleName)
Perturbation <- gsub("null.*", "0", Perturbation)
Perturbation <- gsub("[A-Za-z]+.*", "", Perturbation)

## extract technical rep
TechRep <- gsub(".*\\.(\\d)", "\\1", sampleName)
TechRep <- gsub("P\\d", "1", TechRep)

## combine these into a data.frame
Design <- data.frame(
  Subset = factor(Subset, levels = c("premix1", "premix2", "null", "perturb")),
  BarcodeSize = as.numeric(BarcodeSize) * 1000,
  Perturbation = as.numeric(Perturbation) / 100,
  TechRep = factor(TechRep)
)

lapply(Design, table)
```

color mapping.

```{r}
color_mapping <- list(
  Subset = setNames(
    c("#FF5959", "#33AAFF", "yellowgreen", "slateblue"), 
    c("premix1", "premix2", "null", "perturb")), 
  BarcodeSize = setNames(
    c("#CFCFFF", "#B2B2FF", "#8282FF", "#4D4FFF", "#1D1DFF", "#0000B2"),
    c(20000, 40000, 80000, 160000, 330000, 660000)),
  Perturbation = setNames(
    c("#C3FF85", "#DBB8FF", "#A951FF", "#7500EA"),
    c(0, 0.18, 0.27, 0.35)),
  TechRep = setNames(c("grey25", "grey75"), c(1, 2))
)
```


# Save to barbieQ

- Save count to a barbieQ object

- Save (count+1) to a barbieQ object to address the drop-out issue.

```{r create barbieQ}
mixture <- createBarbieQ(object = EV1, sampleMetadata = Design, factorColors = color_mapping)
mixture_plus1 <- createBarbieQ(object = EV1+1, sampleMetadata = Design, factorColors = color_mapping)
```
# Tag top barcodes

Getting minimum group size for filtering parameter. 

- Set `nSampleThreshold = 8` for subsequent filtering.

```{r tag top}
## check min group size
mixture$sampleMetadata %>% as.data.frame() %>% with(paste0(Subset, Perturbation, BarcodeSize)) %>% table()
mixture$sampleMetadata %>% as.data.frame() %>% with(paste0(Subset, Perturbation)) %>% table()
```

Find appropriate `proportionThreshold`.

- In the two premix samples, solid barcodes should be only detected once.

- Minimize the number of barcodes being tagged as "top" in twice, which are considered "noisy barcodes" based on the experimental design, that barcodes shouldn't overlap between the two premix samples.

```{r}
## tag top barcodes, `proportionThreshold` defaults to 0.99
mixture <- tagTopBarcodes(mixture, nSampleThreshold = 8)
mixture_plus1 <- tagTopBarcodes(mixture_plus1, nSampleThreshold = 8)

## flag the two premix samples
flag_premix <- mixture$sampleMetadata$Subset %in% c("premix1", "premix2") ## two in total
## frequency of being tagged top for each barcode
mixture@assays@data$isTopAssay[,flag_premix] %>% rowSums() %>% table()
## if going ahead with the current `proportionThreshold` to select "top" barcodes
## frequency of being tagged top for each "selected" barcode
mixture@assays@data$isTopAssay[mixture@elementMetadata$isTopBarcode$isTop,flag_premix] %>% rowSums() %>% table()

## same applied to mixture_plus1
mixture_plus1@assays@data$isTopAssay[,flag_premix] %>% rowSums() %>% table()
mixture_plus1@assays@data$isTopAssay[mixture_plus1@elementMetadata$isTopBarcode$isTop,flag_premix] %>% rowSums() %>% table()
```

- Default `proportionThreshold=0.99` keeps too many noisy barcodes.

- Setting `proportionThreshold=0.95` keeps small amount of noisy barcodes, and gets close sets of selected barcodes between `mxiture` and `mixture_plus1` data.

- Set `proportionThreshold=0.95` for subsequent filtering.

```{r}
## tag top barcodes, `proportionThreshold` set to 0.97
mixture <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = 0.95)
mixture_plus1 <- tagTopBarcodes(mixture_plus1, nSampleThreshold = 8, proportionThreshold = 0.95)

## flag the two premix samples
flag_premix <- mixture$sampleMetadata$Subset %in% c("premix1", "premix2") ## two in total
## frequency of being tagged top for each barcode
mixture@assays@data$isTopAssay[,flag_premix] %>% rowSums() %>% table()
## if going ahead with the current `proportionThreshold` to select "top" barcodes
## frequency of being tagged top for each "selected" barcode
mixture@assays@data$isTopAssay[mixture@elementMetadata$isTopBarcode$isTop,flag_premix] %>% rowSums() %>% table()

## same applied to mixture_plus1
mixture_plus1@assays@data$isTopAssay[,flag_premix] %>% rowSums() %>% table()
mixture_plus1@assays@data$isTopAssay[mixture_plus1@elementMetadata$isTopBarcode$isTop,flag_premix] %>% rowSums() %>% table()
```

# Determine barcode occurrence

Zeros per BC.

- Total cell number for the "null" samples is known as `mixture_plus1$sampleMetadata$BarcodeSize`; for the "perturbed" samples, total cell number should be `(mixture_plus1$sampleMetadata$BarcodeSize)*(1+mixture_plus1$sampleMetadata$Perturbation)`. The "perturbation" for the null samples is 0.

- Raw counts bigger than (total barcode count / total cell number) are considered "present", otherwise "absent". Note that we've added 1 to the raw counts before.

```{r}
detection_level <- colSums(assay(mixture_plus1)) / ((mixture_plus1$sampleMetadata$BarcodeSize)*(1+mixture_plus1$sampleMetadata$Perturbation))

detection <- sweep(assay(mixture_plus1), 2, detection_level, FUN = "-")

## plot the number of absence for each barcode.
ggplot(data.frame(N_zero = rowSums(detection <= 0, na.rm = T))) +
  geom_freqpoly(aes(x = N_zero))
## plot the number of absence for each selected barcode by filtering.
ggplot(data.frame(N_zero = rowSums(detection[mixture_plus1@elementMetadata$isTopBarcode$isTop,] <= 0, na.rm = T))) +
  geom_freqpoly(aes(x = N_zero))


## update the "occurrence" assay based on the detection here.
mixture_plus1@assays@data$occurrence <- detection <= 0
```

Barcode detection in the two premix samples.

# Detect correlating barcodes

## FS1D: pairwise correlation

No correlating barcodes detected.Actually, 

```{r, fig.width=4, fig.height=4}
plotBarcodePairCorrelation(barbieQ = mixture_plus1[rowData(mixture_plus1)$isTopBarcode$isTop,], transformation = "none") + 
  theme(legend.position = "top") + 
  labs(title = "Mixture data") -> fs1d

fs1d 
```

- looking weird if using "asin-sqrt" transformation here.

```{r}
clusterCorrelatingBarcodes(barbieQ = mixture_plus1[rowData(mixture_plus1)$isTopBarcode$isTop,], transformation = "asin-sqrt") %>% 
  inspectCorrelatingBarcodes()
```

# Filter top barcodes

## FS1I,L: inspect top/bottom

```{r, fig.width=5, fig.height=4}
plotBarcodePareto(mixture_plus1) +
  ylim(-8, 7) + 
  annotate("text", x = c(pi * 0.05), y = c(7), 
           label = c("Relative proportion %"), color = "#999966", 
           size = 3, angle = 0, fontface = "bold", hjust = 1) -> fs1i

fs1i
```

```{r, fig.width=3, fig.height=4}
plotBarcodeSankey(mixture_plus1) + 
  theme(legend.position = "top") -> fs1l

fs1l
```

## Filter "top" barcodes

```{r}
mixture_plus1_top <- mixture_plus1[rowData(mixture_plus1)$isTopBarcode$isTop,]
```

## Heatmap

```{r echo=TRUE}
## pre-filtering
Heatmap(name = "asin-sqrt prop.",
  matrix = asin(sqrt(assays(mixture_plus1)$proportion)), 
  row_title = paste0(nrow(mixture_plus1), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)
```

```{r hp}
## post-filtering
Heatmap(name = "asin-sqrt prop.",
  matrix = asin(sqrt(assays(mixture_plus1_top)$proportion)), 
  row_title = paste0(nrow(mixture_plus1_top), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)
```

# Save to rda

```{r}
save(mixture_plus1, mixture_plus1_top, file = "output/mixture_barbieQ.rda")
```

# FigureS1-Mixture

```{r, fig.width=4, fig.height=16}
layout = "
DDDD
####
IIII
LLLL
"
fs1_mixture <- (
  wrap_elements(fs1d + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
    wrap_elements(fs1i + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(fs1l + theme(plot.margin = unit(rep(0,4), "cm")))
  ) + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("D"," ", "I","L"))) &
  theme(
    plot.tag = element_text(size = 20, face = "bold", family = "arial"),
    axis.title = element_text(size = 17),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

fs1_mixture
```


```{r}
ggsave(
  filename = "output/fs1_mixture.png",
  plot = fs1_mixture,
  width = 4,
  height = 16,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [fs1_mixture](../output/fs1_mixture.png)