---
title: "barbieQ_paper_Figure4: Case study of `monkey HSPC`"
author: "Liyang Fei"
date: |
  Initiate: 2025-09-07
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(data.table) # for data.table()
library(SummarizedExperiment)
library(S4Vectors)
library(grid)
library(ggplot2)
library(ggbreak)
library(ggnewscale) # ggnewscale::new_scale_fill() 
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(magick)
library(eulerr)
library(edgeR)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/F4_validation.R") ## functions for Figure4 validations
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Load preprocessed data

Loading `barbieQ` object preprocessed from [barbieQ_paper_Figure2]("barbieQ_paper_Figure2.html")

```{r}
load("output/monkeyHSPC_barbieQ.rda")
# load("output/monkeyHSPC_raw_barbieQ.rda")
```

# F4A: Experimental schematic

```{r}
schematic <- image_read("output/f4a_schematic.jpeg")
f4a <- rasterGrob(schematic, interpolate = FALSE) 
wrap_elements(full = f4a)
```

# F4B: Plot barcode proportion and sample annotation

`plotBarcodeHistogram` is customized from the `bartools` package.

```{r}
## get sample conditions and color mapping
metadata <- monkeyHSPC_merged_top$sampleMetadata %>% as.data.frame()
metadata$sample <- rownames(metadata)
metadata$y = 1.05 ## for ggplot tile position
color_mapping <- monkeyHSPC_merged_top@metadata$factorColors

## order samples
order_sample <- metadata %>% with(order(Celltype, Months))

monkey_dge <- DGEList(
  counts = assay(monkeyHSPC_merged_top), 
  group = metadata$Celltype)

## pull out ggplot data
p_stacked_prop <- plotBarcodeHistogram(monkey_dge, orderSamples = colnames(monkeyHSPC_merged_top)[order_sample])

f4b <- p_stacked_prop +
  ## label perturbations
  ggnewscale::new_scale_fill() +
  geom_tile(data = metadata, aes(x = sample,y = y, fill = Celltype), height = 0.04, inherit.aes = FALSE, color = "white") +
  scale_fill_manual(values = color_mapping$Celltype) +
  labs(fill = "Cell type") + 
  ## label libsize
  ggnewscale::new_scale_fill() +
  geom_tile(data = metadata, aes(x = sample,y = y+0.05, fill = as.factor(Months)), height = 0.04, inherit.aes = FALSE, color = "white") +
  scale_fill_manual(values = setNames(attr(color_mapping$Months, "colors"), attr(color_mapping$Months, "breaks"))) +
  labs(fill = "Month") +
  labs(x = "Sample", y = "Barcode proportion") +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2)) +
  theme(axis.text.x = element_blank(), panel.grid.major.x = element_blank())

f4b
```

# Compare filtering

## F4D: top10each vs. barbieQ filter

The original paper filter barcodes by tagging top 10 barcodes from each sample, and collectively keeping the tagged barcodes across all samples.
Here we compare this "top10each" (and "top20each") strategy with `barbieQ` filtering.

The `tagTopN` function is sourced from "F4_validation.R".

```{r, fig.width=12, fig.height=4}
## tagging top 10 in all samples
colTags <- lapply(as.data.frame(assays(monkeyHSPC_merged)$proportion), function(col) tagTopN(col, N = 10))
colTagMat <- do.call(cbind, colTags) %>% as.matrix()
colnames(colTagMat) <- colnames(monkeyHSPC_merged)
rownames(colTagMat) <- rownames(monkeyHSPC_merged)
## filtering by keeping barcodes tagged for at least once
colTagVec <- rowSums(colTagMat) >= 1
## ensuring we've tagged 10 barcodes within each sample
all(colSums(colTagMat) == 10)

## copying the barbieQ object and manually labeling the filtered barcodes based on the above strategy
monkeyHSPC_top10each <- monkeyHSPC_merged
rowData(monkeyHSPC_top10each)$isTopBarcode$isTop <- colTagVec
assays(monkeyHSPC_top10each)$isTopAssay <- colTagMat
p_sankey_top10 <- barbieQ::plotBarcodeSankey(monkeyHSPC_top10each)
p_sankey_barbieQ <- barbieQ::plotBarcodeSankey(monkeyHSPC_merged)

## repeat the process for top20 each
colTags2 <- lapply(as.data.frame(assays(monkeyHSPC_merged)$proportion), function(col) tagTopN(col, N = 20))
colTagMat2 <- do.call(cbind, colTags2) %>% as.matrix()
colnames(colTagMat2) <- colnames(monkeyHSPC_merged)
rownames(colTagMat2) <- rownames(monkeyHSPC_merged)
colTagVec2 <- rowSums(colTagMat2) >= 1
all(colSums(colTagMat2) == 20)

monkeyHSPC_top20each <- monkeyHSPC_merged
rowData(monkeyHSPC_top20each)$isTopBarcode$isTop <- colTagVec2
assays(monkeyHSPC_top20each)$isTopAssay <- colTagMat2
p_sankey_top20 <- barbieQ::plotBarcodeSankey(monkeyHSPC_top20each)


## plot sankeys
f4d <- p_sankey_barbieQ + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, vjust = 1), 
        strip.text.x = element_blank(), strip.text.y = element_blank()) + labs(title = "barbieQ") +
  p_sankey_top10 + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, vjust = 1),
        strip.text.x = element_blank(), strip.text.y = element_blank()) + labs(title = "top10each") + 
  p_sankey_top20 + 
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5, vjust = 1),
        strip.text.x = element_blank(), strip.text.y = element_blank()) + labs(title = "top20each")

f4d
```

## F4C: overlap

```{r, fig.width=6, fig.height=4}
list_top_BC <- list(
  barbieQ = rownames(monkeyHSPC_merged)[monkeyHSPC_merged@elementMetadata$isTopBarcode$isTop],
  top10each = rownames(monkeyHSPC_merged)[colTagVec],
  top20each = rownames(monkeyHSPC_merged)[colTagVec2]
  )

## plot Euler
# venn_filter <- eulerr::euler(list_top_BC) %>% 
#   plot(fills = c("pink", "skyblue", "steelblue"), 
#        alpha = 0.5, labels = list(font = 4, cex = 1), quantities = list(font = 1, cex = 1)) 
# 
# venn_filter %>% plot() %>% grid.grabExpr() %>% wrap_plots()

## plot ggVenn
# ggVennDiagram(list_top_BC, label = "count") + scale_fill_gradient(low="white", high = "skyblue")

## plot UpSet plot
f4c <- list_top_BC %>% 
  Venn() %>% 
  plot_upset(top.bar.numbers.size = 4, relative_width = 0.0) ## limit space for left bar plot

## getting theme settings from an example ggplot using barbieQ theme
tester <- ggplot(data.frame(X = 1:3, Y = 4:6)) +
  geom_point(aes(x = X, y = Y)) +
  theme_barbieQ_size()

tester_col <- ggplot(data.frame(X = 1:3, Y = 4:6)) +
  geom_col(aes(x = X, y = Y), width = 0.2)
tester_col@layers$geom_col$aes_params$width

mytittle <- tester@theme$axis.title
mytext <- tester@theme$axis.text

## modifying components in UpSet plots: three gplot objects
f4c[["plotlist"]][[1]]@theme$axis.title <- mytittle
f4c[["plotlist"]][[1]]@theme$axis.text <- mytext

f4c[["plotlist"]][[2]]@theme$axis.title <- mytittle
f4c[["plotlist"]][[2]]@theme$axis.text <- mytext

f4c[["plotlist"]][[2]]@layers$geom_col$aes_params$width <- 0.4

f4c[["plotlist"]][[3]]@theme$axis.title <- mytittle
f4c[["plotlist"]][[3]]@theme$axis.text <- mytext

f4c[["plotlist"]][[3]]@theme$axis.text@size <- 11
f4c[["plotlist"]][[3]]@theme$axis.text@angle <- 15

## removing the left bar plot
f4c[["plotlist"]][[3]] <- f4c[["plotlist"]][[3]] + theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank()
  )

## plot 4c
f4c
```

## F4E: inspect two sets: heatmap

```{r fig.width=7, fig.height=9}
barcodeArray <- monkeyHSPC_merged@elementMetadata$isTopBarcode$isTop | colTagVec
length(colTagVec)
length(monkeyHSPC_merged@elementMetadata$isTopBarcode$isTop)
table(barcodeArray)

monkeyHSPC_compare_filter <- monkeyHSPC_merged[barcodeArray,]

# Create a named vector indicating where each element belongs
region_labels <- data.frame(
  barcode = rownames(monkeyHSPC_compare_filter),
  region = "both")
region_labels[monkeyHSPC_merged@elementMetadata$isTopBarcode$isTop[barcodeArray] & !(colTagVec[barcodeArray]), "region"] <- "barbieQ only"
region_labels[!(monkeyHSPC_merged@elementMetadata$isTopBarcode$isTop[barcodeArray]) & colTagVec[barcodeArray], "region"] <- "top10each only"

region_labels$region %>% table()

ht_filter <- plotBarcodeHeatmap(
  monkeyHSPC_compare_filter, 
  barcodeAnnotation = rowAnnotation(
    filteredBy = region_labels$region,
    col = list(filteredBy = c("barbieQ only" = "hotpink", "both" = "slateblue", "top10each only" = "steelblue1")),
    annotation_name_side = "top", annotation_name_rot = 45, annotation_name_gp = grid::gpar(fontsize = 10)))

## ht_filter@heatmap_param[["calling_env"]][["sampleAnnotation"]]@anno_list[["Months"]]@show_legend <- FALSE
## ht_filter@heatmap_param[["calling_env"]][["groupAnnotation"]]@anno_list[["Celltype"]]@show_legend <- FALSE

f4e <- draw(ht_filter, row_split = region_labels$region, width = unit(6, "cm"), height = unit(15, "cm")) %>%
  grid.grabExpr() %>% list() %>% wrap_plots()

f4e <- Heatmap(
  monkeyHSPC_compare_filter@assays@data$proportion %>% sqrt() %>% asin(), name = "asin-sqrt prop.",
  cluster_rows = TRUE, cluster_columns = TRUE, show_row_names = FALSE, show_column_names = FALSE, 
  right_annotation = rowAnnotation(
    filteredBy = region_labels$region,
    col = list(filteredBy = c("barbieQ only" = "hotpink", "both" = "slateblue", "top10each only" = "steelblue1")),
    annotation_name_side = "top", annotation_name_rot = 45),
  row_split = region_labels$region, width = unit(6, "cm"), height = unit(15, "cm"), row_title = "Barcodes",
  top_annotation = HeatmapAnnotation(
    Celltype = monkeyHSPC_compare_filter$sampleMetadata$Celltype,
    Months = monkeyHSPC_compare_filter$sampleMetadata$Months,
    col = monkeyHSPC_compare_filter@metadata$factorColors,
    show_legend = FALSE, annotation_name_side = "left"
  ))

f4e
```

# Compare difference identification on top10each filtering approach

## barbieQ filter + barbieQ test

```{r}
monkeyHSPC_top10each@elementMetadata$isTopBarcode$isTop %>% table()
## subset the 79 barcodes filtered by top10each
monkeyHSPC_top10each_79 <- monkeyHSPC_top10each[monkeyHSPC_top10each@elementMetadata$isTopBarcode$isTop,]
##need to take Time into account
designs <- monkeyHSPC_top10each_79$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(model.matrix(~0+ Celltype + Months))

monkeyHSPC_top10each_79 <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top10each_79, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
# check design
monkeyHSPC_top10each_79@elementMetadata$testingBarcode@metadata$contrasts
monkeyHSPC_top10each_79@elementMetadata$testingBarcode@metadata$design %>% data.table(keep.rownames = T)
```

## barbieQ filter + LFC approach

(i) >1% contribution to CD56−CD16+ cells and 

(ii) >10-fold abundance in contribution to CD56−CD16+ cells compared with T, B, Gr, and CD56+CD16− cells.

`calc_bias_LFC` function sourced from "F4_validation.R".

```{r}
list_BC_top10each_LFC <- calc_bias_LFC(monkeyHSPC_top10each_79)
union_LFC_top10each <- unlist(list_BC_top10each_LFC) %>% unique()
```

```{r}
## check overlaps between months
venn_top10each_LFC <- eulerr::euler(list_BC_top10each_LFC)
plot(venn_top10each_LFC, fills = c("white"), alpha = 0.5, labels = T, quantities = T)
```

## F4G: overlap

```{r}
list_top10each_feature <- list(
    barbieQ = rownames(monkeyHSPC_top10each_79)[monkeyHSPC_top10each_79@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
    heuristics = union_LFC_top10each)

## plot Euler
# venn_top10each_feature <- eulerr::euler(list_top10each_feature) %>% 
#   plot(fills = c("pink","slateblue"),
#        alpha = 0.5, labels = list(font = 4, cex = 1), quantities = list(font = 1, cex = 1)) 
# 
# venn_top10each_feature %>% plot() %>% grid.grabExpr() %>% wrap_plots()

## plot UpSet plot
f4g <- list_top10each_feature %>% 
  Venn() %>% 
  plot_upset(top.bar.numbers.size = 4, relative_width = 0.0) ## limit left bar plot

## getting theme settings from an example ggplot using barbieQ theme
tester <- ggplot(data.frame(X = 1:3, Y = 4:6)) +
  geom_point(aes(x = X, y = Y)) +
  theme_barbieQ_size()

mytittle <- tester@theme$axis.title
mytext <- tester@theme$axis.text

## modifying components in UpSet plots: three gplot objects
f4g[["plotlist"]][[1]]@theme$axis.title <- mytittle
f4g[["plotlist"]][[1]]@theme$axis.text <- mytext

f4g[["plotlist"]][[2]]@theme$axis.title <- mytittle
f4g[["plotlist"]][[2]]@theme$axis.text <- mytext

f4g[["plotlist"]][[2]]@layers$geom_col$aes_params$width <- 0.4

f4g[["plotlist"]][[3]]@theme$axis.title <- mytittle
f4g[["plotlist"]][[3]]@theme$axis.text <- mytext

f4g[["plotlist"]][[3]]@theme$axis.text@size <- 11
f4g[["plotlist"]][[3]]@theme$axis.text@angle <- 15

## removing the left bar plot
f4g[["plotlist"]][[3]] <- f4g[["plotlist"]][[3]] + theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank()
  )

## plot 4g
f4g
```

## F4F: inspect two sets: boxplot

- Box plot to visualize the LFC of sets of selected barcodes.

```{r}
FC_top10each <- calc_bias_LFC(barbieQ = monkeyHSPC_top10each_79, get_FC = T) %>% as.data.frame()
rownames(FC_top10each) <- rownames(monkeyHSPC_top10each_79)
## annotating BC by barbieQ seletion
FC_top10each$bbq <- monkeyHSPC_top10each_79@elementMetadata$testingBarcode$direction
## annotating BC by heuristic selection
FC_top10each$heu = 0
FC_top10each[union_LFC_top10each, "heu"] = 1
  ## adding BC name
FC_top10each <- FC_top10each %>% 
  mutate(Barcode = rownames(monkeyHSPC_top10each_79)) %>% 
  ## summarizing two selecting methods overlaps
  mutate(identifiedBy = ifelse(bbq == 1, 
                             ifelse(heu == 1, "both", "barbieQ only"), 
                             ifelse(heu == 1, "heuristics only", "neither"))) %>% 
  dplyr::select(-c(heu, bbq))

## extract FC from testing
coff_FC <- monkeyHSPC_top10each_79@elementMetadata$testingBarcode$meanDiff
FC_top10each$testing_coeff <- coff_FC

long_FC_top10each <- FC_top10each %>% 
  mutate(Barcode = rownames(monkeyHSPC_top10each_79)) %>% 
  pivot_longer(
      cols = -c(Barcode, identifiedBy, testing_coeff),     # keep Celltype and Months fixed
      names_to = "Month",
      values_to = "FC"
    )
```

Draw box plot, barcodes ordered by mean LFC

```{r}
## order BC in x axis by contrast coefficient
barcode_order <- long_FC_top10each %>%
  group_by(Barcode) %>%
  summarise(mean_LFC = mean((testing_coeff), na.rm = TRUE)) %>%
  arrange(mean_LFC) %>%
  pull(Barcode)

long_FC_top10each$identifiedBy <- factor(
  long_FC_top10each$identifiedBy,
  levels = c("barbieQ only", "heuristics only", "both", "neither")
)

## remove level "heuristics only"
long_FC_top10each$identifiedBy <- factor(
  long_FC_top10each$identifiedBy,
  levels = c("barbieQ only",  "both", "neither")
)

p_box_coef <- ggplot(long_FC_top10each, aes(x = factor(Barcode, levels = barcode_order), y = log2(FC), group = (Barcode), color = identifiedBy)) +
  geom_boxplot(varwidth = T) +
  # geom_jitter()+
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "right", axis.title = element_text(size = 14)) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "snow4") +
  labs(x = "",
       y = "LFC",
       title = "Barocdes selected by top10each filtering") +
  scale_color_manual(
    values = c("both" = "slateblue", "barbieQ only" = "hotpink", "heuristics only" = "steelblue1", "neither" = "slategray"),
    drop = FALSE)

p_box_coef_2 <- ggplot(long_FC_top10each, aes(x = factor(Barcode, levels = barcode_order), y = testing_coeff, group = (Barcode), color = identifiedBy)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "none", axis.title = element_text(size = 14)) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "snow4") +
  labs(x = "Barocdes ordered by barbieQ-estimated size effect",
       y = "Size effect") +
  scale_color_manual(values = c("both" = "slateblue", "barbieQ only" = "hotpink", "heuristics only" = "steelblue1", "neither" = "slategray"))

f4f <- p_box_coef / p_box_coef_2 + plot_layout(heights = c(3, 1))
f4f
```

# Compare fitering with barbieQ testing

## top10each + barbieQ test

```{r}
##need to take Time into account
designs <- monkeyHSPC_merged_top$sampleMetadata %>% as.data.frame() %>%
  mutate(Months = as.factor(Months)) %>% 
  with(model.matrix(~0+ Celltype + Months))

monkeyHSPC_merged_top <- testBarcodeSignif(
  barbieQ = monkeyHSPC_merged_top, designMatrix = designs,
  contrastFormula = "(CelltypeNK_CD56n_CD16p) - (CelltypeB+CelltypeGr+CelltypeT+CelltypeNK_CD56p_CD16n)/4",
  method = "diffProp", transformation = "asin-sqrt"
)
# check design
monkeyHSPC_merged_top@elementMetadata$testingBarcode@metadata$contrasts
monkeyHSPC_merged_top@elementMetadata$testingBarcode@metadata$design %>% data.table(keep.rownames = T)
```

## F4H: stacked proportion

`plot_validation` sourced from "F4_validation.R"

```{r, fig.width=12, fig.height=5}
 f4h <- (plot_validation(monkeyHSPC_top10each_79) + 
    theme( plot.title = element_text(hjust = 0.5, vjust = 1)) + labs(title = "Based on top10each")) +
  (plot_validation(monkeyHSPC_merged_top) + 
     theme( plot.title = element_text(hjust = 0.5, vjust = 1)) + labs(title = "Based on barbieQ"))

f4h
```

## overlap

- compare sets of significant barcodes

```{r}
## signif. BC
filter_venn_signif_list <- list(
  ## signif. BC
  top10each = rownames(monkeyHSPC_top10each_79)[monkeyHSPC_top10each_79@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  barbieQ = rownames(monkeyHSPC_merged_top)[monkeyHSPC_merged_top@elementMetadata$testingBarcode$tendencyTo == "CelltypeNK_CD56n_CD16p"],
  ## all BC in test
  top10each_all = rownames(monkeyHSPC_top10each_79),
  barbieQ_filter_all = rownames(monkeyHSPC_merged_top))

venn_filter_signif <- eulerr::euler(filter_venn_signif_list[1:2]) %>% 
  plot(fills = c("skyblue", "pink"), 
       alpha = 0.5, labels = list(font = 4, cex = 1), quantities = list(font = 1, cex = 1))

venn_filter_signif  %>% plot() %>% grid.grabExpr() %>% wrap_plots()

```

## F4I: inspect two sets: boxplot

- box plot on two sets of significant barcodes based on two filtering appraoches

```{r}
BC_filter_sig <- filter_venn_signif_list[1:2] %>% unlist %>% unique()

FC_filter_sig <- calc_bias_LFC(barbieQ = monkeyHSPC_merged[BC_filter_sig,], get_FC = T) %>% 
  as.data.frame()
Mean_filter_sig <- calc_bias_LFC(barbieQ = monkeyHSPC_merged[BC_filter_sig,], get_AMean = T) %>%
  as.data.frame()
rownames(FC_filter_sig) <- BC_filter_sig
## annotating BC by barbieQ seletion
FC_filter_sig$top10each = 0
FC_filter_sig$bbq = 0
## annotating BC by heuristic selection, `filter_venn_signif_list` obtained from last chunk
FC_filter_sig[filter_venn_signif_list[[1]], "top10each"] = 1
FC_filter_sig[filter_venn_signif_list[[2]], "bbq"] = 1
  ## adding BC name
FC_filter_sig <- FC_filter_sig %>% 
  mutate(Barcode = BC_filter_sig) %>% 
  ## summarizing two selecting methods overlaps
  mutate(FilteringMethod = ifelse(top10each == 1, 
                                  ifelse(bbq == 1, "both", "top10each"),
                                  ifelse(bbq == 1, "barbieQ", "neither"))) %>% 
  dplyr::select(-c(top10each, bbq))

## extract FC from testing

long_FC_filter_sig <- FC_filter_sig %>% 
  mutate(Barcode = BC_filter_sig) %>% 
  pivot_longer(
      cols = -c(Barcode, FilteringMethod),     # keep Celltype and Months fixed
      names_to = "Month",
      values_to = "FC"
    )
# Add barcode column to Mean_filter_sig
Mean_filter_sig <- Mean_filter_sig %>%
  mutate(Barcode = BC_filter_sig)

long_Mean <- Mean_filter_sig %>%
  pivot_longer(
    cols = -Barcode,
    names_to = "Month",
    values_to = "Mean"
  )

long_FC_filter_sig <- long_FC_filter_sig %>%
  left_join(long_Mean, by = c("Barcode", "Month"))

```

```{r fig.width=10, fig.height=5}
barcode_order_filter_sig <- long_FC_filter_sig %>%
  group_by(Barcode) %>%
  summarise(AMean = mean(Mean, na.rm = TRUE)) %>%
  arrange(AMean) %>%
  pull(Barcode)

barcode_order_df_filter_sig <- long_FC_filter_sig %>%
  group_by(Barcode, FilteringMethod) %>%
  summarise(AMean = mean(Mean, na.rm = TRUE)) %>%
  arrange(AMean)

p_AMean_0 <- ggplot()+
  geom_point(data = barcode_order_df_filter_sig,
             aes(x = factor(Barcode, levels = barcode_order_filter_sig),
                 y = log10(AMean),
                 color = FilteringMethod),
             size = 1.5, shape = 15,
             inherit.aes = FALSE) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "none", axis.title = element_text(size = 14)) +
  labs(x = "Barocdes ordered by mean proportion across cell types across time",
       y = "Log10 prop.") +
  scale_color_manual(values = c("both" = "slateblue", "barbieQ" = "hotpink", "top10each" = "steelblue1"))


p_box_0 <- ggplot(long_FC_filter_sig, aes(x = factor(Barcode, levels = barcode_order_filter_sig), y = log2(FC), group = (Barcode), color = FilteringMethod)) +
  geom_boxplot(varwidth = T) +
  theme_classic() +
  theme(axis.text.x = element_blank(), legend.position = "right", axis.title = element_text(size = 14)) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, colour = "grey") +
  labs(x = "",
       y = "LFC",
       title = "Barcodes significant in barbieQ test, selected by either barbieQ or top10each filtering") +
  scale_color_manual(values = c("both" = "slateblue", "barbieQ" = "hotpink", "top10each" = "steelblue1"))

f4i <- p_box_0 / p_AMean_0 + plot_layout(heights = c(3, 1))
f4i
```

# F4J: Example difference over time

```{r eval=FALSE, include=FALSE}
flag_NK <- monkeyHSPC_raw$sampleMetadata$Celltype == "NK_CD56n_CD16p"
colTags <- lapply(
  as.data.frame(assays(monkeyHSPC_raw)$proportion[,flag_NK]), 
  function(col) tagTopN(col, N = 10)
  )
colTagMatNK <- do.call(cbind, colTags) %>% as.matrix()
colnames(colTagMatNK) <- colnames(monkeyHSPC_raw)[flag_NK]
rownames(colTagMatNK) <- rownames(monkeyHSPC_raw)
## filtering by keeping barcodes tagged for at least once
colTagVecNK <- rowSums(colTagMatNK) >= 1
sum(colTagVecNK)
## ensuring we've tagged 10 barcodes within each sample
all(colSums(colTagMatNK) == 10)

NKnp <- monkeyHSPC_raw[colTagVecNK,flag_NK]
dim(NKnp)
```

```{r}
designs <- monkeyHSPC_top10each_79$sampleMetadata %>% as.data.frame() %>%
  with(model.matrix(~0+ Celltype+ Celltype:Months))
designs <- designs[,c(1:5, 9:13)]
colnames(designs) <- gsub(":", ".", colnames(designs))

monkeyHSPC_top10each_79_NK <- testBarcodeSignif(
  barbieQ = monkeyHSPC_top10each_79, 
  sampleMetadata = designs, designMatrix = designs, sampleGroup = "CelltypeNK_CD56n_CD16p.Months",
  method = "diffProp", transformation = "asin-sqrt"
)

monkeyHSPC_top10each_79_NK@elementMetadata$testingBarcode$tendencyTo %>% table()
monkeyHSPC_top10each_79_NK@elementMetadata$testingBarcode@metadata$design %>% data.table()
monkeyHSPC_top10each_79_NK@elementMetadata$testingBarcode@metadata$contrasts
monkeyHSPC_top10each_79_NK@elementMetadata$testingBarcode@metadata$contrastGroups


NKnp_fig <- monkeyHSPC_top10each_79_NK[,monkeyHSPC_top10each_79_NK$sampleMetadata$Celltype == "NK_CD56n_CD16p"]
```


```{r eval=FALSE, include=FALSE}
designs <- NKnp$sampleMetadata %>% as.data.frame() %>%
  with(model.matrix(~0+ Months))

NKnp <- testBarcodeSignif(
  barbieQ = NKnp, sampleMetadata = NKnp$sampleMetadata[,2,drop=F], sampleGroup = "Months",
  method = "diffProp", transformation = "asin-sqrt", designMatrix = designs
)

NKnp@elementMetadata$testingBarcode@metadata$design
NKnp@elementMetadata$testingBarcode@metadata$contrasts
NKnp@elementMetadata$testingBarcode@metadata$contrastGroups

NKnp@elementMetadata$testingBarcode$tendencyTo %>% table()
# flag_sigNK <- NKnp@elementMetadata$testingBarcode$tendencyTo != "n.s."
NKnp_fig <- NKnp
```

```{r, fig.width=5, fig.height=9}
BC_kinetic <- NKnp_fig@elementMetadata$testingBarcode$tendencyTo

pch_fun <- ifelse(NKnp_fig@elementMetadata$testingBarcode$direction != 0, "*", NA)

kinetic_ha <- rowAnnotation(
  Kinetic = anno_simple(
    BC_kinetic,
    col = NKnp_fig@metadata$factorColors$testingBarcode,
    pch = pch_fun
    ),
  show_annotation_name = TRUE,
  annotation_name_side = "top",
  annotation_name_rot = 45
)

HP_anno_kinetic <- Heatmap(
  asin(sqrt(NKnp_fig@assays@data$proportion)), name = "asin-sqrt prop.",
  show_row_names = FALSE, column_labels = NKnp_fig$sampleMetadata$Months,
  column_title = "Months", column_names_side = "top", column_names_rot = 45,
  row_title = "Barcodes", row_km = 4,
  width = unit(4, "cm"), height = unit(12, "cm"),
  column_order = order(NKnp_fig$sampleMetadata$Months),
  # col = colorRamp2(c(0, -2, -4.602), c("red", "white", "blue")),
  # left_annotation = row_ha,
  right_annotation = kinetic_ha
)

lgd_sig = Legend(pch = "*", type = "points", labels = "signif.")

lgd_kinetic = Legend(labels = c("up", "down", "n.s."), title = "Kinetic", 
             legend_gp = gpar(fill = c("#FF5959", "#33AAFF", "#FFC000"))
             )
## heatmap_legend_side = "top", annotation_legend_side = "top"

f4j <- draw(HP_anno_kinetic, annotation_legend_list = list(lgd_kinetic, lgd_sig),
            annotation_legend_side = "bottom", heatmap_legend_side = "bottom") %>%
  grid.grabExpr() %>% list() %>% wrap_plots()

f4j
```

# Figure 4

```{r, fig.width=15, fig.height=21}
layout = "
AAAAAAAAAAAAA##
AAAAAAAAAAAAA##
BBBBBBCCCCEEEEE
BBBBBBCCCCEEEEE
BBBBBBCCCCEEEEE
DDDDDDDDDDEEEEE
DDDDDDDDDDEEEEE
DDDDDDDDDDEEEEE
FFFFFFFFFFGGGG#
FFFFFFFFFFGGGG#
FFFFFFFFFFGGGG#
HHHHHHHHHHJJJJ#
HHHHHHHHHHJJJJ#
HHHHHHHHHHJJJJ#
HHHHHHHHHHJJJJ#
IIIIIIIIIIJJJJ#
IIIIIIIIIIJJJJ#
IIIIIIIIIIJJJJ#
"
f4 <- (wrap_elements(full = f4a) + 
    wrap_elements(f4b + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(grid.grabExpr(print(f4c))) +
    wrap_elements(f4d + theme(plot.margin = unit(rep(0,4), "cm"))) +
    wrap_plots(list(f4e %>% draw() %>% grid.grabExpr())) +
    wrap_elements(f4f + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(grid.grabExpr(print(f4g))) + 
    wrap_elements(f4h + theme(plot.margin = unit(c(0,0,0,0), "cm"))) +
    wrap_elements(f4i + theme(plot.margin = unit(rep(0,4), "cm"))) +
    # wrap_elements(f4j + theme(plot.margin = unit(rep(0,4), "cm")))
  wrap_plots(list(HP_anno_kinetic %>% 
          draw(annotation_legend_list = list(lgd_kinetic, lgd_sig), 
               annotation_legend_side = "top",heatmap_legend_side = "top") %>% 
            grid.grabExpr()))
    )+ 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("A","B","C", "D", "E","F","G","H", "I", "J"))) &
  theme(
    plot.tag = element_text(size = 24, face = "bold", family = "arial"),
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

f4

```

```{r}
ggsave(
  filename = "output/f4.png",
  plot = f4,
  width = 15,
  height = 21,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [f4](../output/f4.png)



