---
title: "barbieQ_paper Supplementary 1: Preprocessing `AML data`"
author: "Liyang Fei"
date: |
  Initiate: 2025-04-03
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Links to preprocessing other datasets in the barbieQ paper: 
 
 - [Monkey HSPC](barbieQ_paper_Figure2.html)
 
 - [AML data]()
 
 - [HSPC xenograft data](barbieQ_paper_FigureS1_xenoHSPC.html)
 
 - [Mixture data](barbieQ_paper_FigureS1_Mixture.html)

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)
library(SummarizedExperiment)
library(SEtools) # for aggregate SE
library(S4Vectors)
library(devtools)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Read data

Sourced from: [(https://zenodo.org/badge/713176761.svg)](https://zenodo.org/doi/10.5281/zenodo.10896642)

```{r read counts}
dge <- readRDS("data/bartools_dose_escalation/dose2.rds")

# simplify barcode labels
rownames(dge) <- gsub("BFP_Barcode", "BC", rownames(dge))

```

# Save to a barbieQ obejct

```{r review metadata}
## check sample metadata
# dge$samples$Treatment %>% table()
# dge$samples %>% class()
# lapply(dge$samples, function (j) {table(j)})
```

```{r create barbieQ}
AML <- barbieQ::createBarbieQ(
  object = dge$counts, 
  sampleMetadata = dge$samples[, c("Treatment", "Strategy", "Technical_Replicate", "Dose", "Timepoint")])
metadata(AML)<- list(
  "Experiment" = "Dose_escalation_2",
  "Celltype" = "MLL-AF9",
  "Library" = "SPLINTR_BFP_V1"
  )
```

# Tag top barcodes

```{r tag top}
dim(AML)
colData(AML)$sampleMetadata
## find minimum group size
targets <- as.data.frame(AML$sampleMetadata) %>% 
  mutate(Group = paste0(Treatment, Strategy, Dose))
targets$Group %>% table()

## select sample to consider, exclude TP0 when tagging
AML <- barbieQ::tagTopBarcodes(
  barbieQ = AML, 
  activeSamples = AML$sampleMetadata$Treatment != "TP0", 
  nSampleThreshold = 2
  )

rowData(AML)$isTopBarcode$isTop %>% table()
```

```{r vis tag}
barbieQ::plotBarcodePareto(barbieQ = AML)
barbieQ::plotBarcodeSankey(barbieQ = AML)
```

# Cluster barcodes

## FS1B: pairwise correlation

```{r barcode correlation, fig.width=4, fig.height=4}
barbieQ::plotBarcodePairCorrelation(barbieQ = AML, propThresh = 0.005, transformation = "none") + 
  theme(legend.position = "top") + 
  labs(title = "AML data") -> fs1b

fs1b
```

## FS1E: clusters barcodes

```{r cluster barcodes}
## cluster barcoes
AML <- barbieQ::clusterCorrelatingBarcodes(barbieQ = AML, propThresh = 0.005, transformation = "none")

p_cluster_list <- barbieQ::inspectCorrelatingBarcodes(AML)

p_cluster_list

layer_index <- which(sapply(p_cluster_list$p_cluster$layers, function(x) inherits(x$geom, "GeomTextRepel")))
## update the size for that layer
p_cluster_list$p_cluster$layers[[layer_index]]$aes_params$size <- 1.4  # new size
fs1e <- p_cluster_list$p_cluster + theme(legend.position = "none")

fs1e
```

## merge clusters

```{r}
AML_merged <- barbieQ::mergeCorrelatingBarcodes(barbieQ_clustered = AML)
dim(AML_merged)

AML_merged@metadata$predicted_barcode_clusters
AML_merged@metadata$removed_barcodes
```

# Tag top barcode after collapsing

## FS1G,J: inspect top/bottom

```{r tag after cor, fig.width=5, fig.height=4}
AML_merged <- tagTopBarcodes(
  barbieQ = AML_merged,
  activeSamples = AML_merged$sampleMetadata$Treatment != "TP0", 
  nSampleThreshold = 2)

barbieQ::plotBarcodePareto(barbieQ = AML_merged) +
  ylim(-8, 17) + 
  annotate("text", x = c(pi * 0.05), y = c(17), 
           label = c("Relative proportion %"), color = "#999966", 
           size = 3, angle = 0, fontface = "bold", hjust = 1) -> fs1g

fs1g
```

```{r, fig.width=3, fig.height=4}
barbieQ::plotBarcodeSankey(barbieQ = AML_merged) +
  theme(legend.position = "top")  -> fs1j

fs1j
```

## Filter "top" barcodes

```{r}
AML_top <- AML_merged[rowData(AML_merged)$isTopBarcode$isTop,]
```

## Heatmap

```{r hp}
## pre-filtering
Heatmap(name = "asin-sqrt prop.",
  matrix = asin(sqrt(assays(AML_merged)$proportion)), 
  row_title = paste0(nrow(AML_merged), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)

## post-filtering
Heatmap(name = "asin-sqrt prop.",
  matrix = asin(sqrt(assays(AML_top)$proportion)), 
  row_title = paste0(nrow(AML_top), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)
```

# Save rds

```{r save rds}
save(AML, AML_merged, AML_top, file = "output/AML_barbieQ.rda")

```

# Top10

```{r}
tagTop10 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 10
  # summary(isTop)
  return(isTop)
}

tester <- tagTop10(assays(AML_merged)$CPM[,1])
summary(tester)

colTags <- lapply(as.data.frame(assays(AML_merged)$CPM), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat) >= 1
summary(colTagVec)

AML_merged_top10 <- AML_merged
rowData(AML_merged_top10)$isTopBarcode <- DataFrame(isTop = colTagVec)
summary(rowData(AML_merged_top10)$isTopBarcode$isTop)

barbieQ::plotBarcodePareto(barbieQ = AML_merged_top10)
barbieQ::plotBarcodeSankey(barbieQ = AML_merged_top10)
```

# Save FigureS1-AML

```{r, fig.width=4, fig.height=16}
layout = "
B
E
G
J
"
fs1_aml <- (
  wrap_elements(fs1b + theme(plot.margin = unit(c(0,0,0,0), "line"))) + 
    wrap_elements(fs1e + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(fs1g + theme(plot.margin = unit(rep(0,4), "cm"))) +
    wrap_elements(fs1j + theme(plot.margin = unit(rep(0,4), "cm")))
  ) + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = list(c("B","E","G", "J"))) &
  theme(
    plot.tag = element_text(size = 20, face = "bold", family = "arial"),
    axis.title = element_text(size = 17),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11))

fs1_aml
```

```{r}
ggsave(
  filename = "output/fs1_aml.png",
  plot = fs1_aml,
  width = 4,
  height = 16,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [fs1_AML](../output/fs1_aml.png)