---
title: "barbieQ_paper_Figure3_FigureS3: Type I error rate and power using `Mixture` data"
author: "Liyang Fei"
date: |
  Initiate: 2025-09-07
  
  Last update: `r format(Sys.Date(), '%Y-%m-%d')`
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library, message=FALSE, warning=TRUE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(data.table) # for data.table()
library(SummarizedExperiment)
library(S4Vectors)
library(grid)
library(ggplot2)
library(ggbreak)
library(ggnewscale) # ggnewscale::new_scale_fill() 
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(magick) # image_read()
library(eulerr)
library(edgeR)

source("analysis/plotBarcodeHistogram.R") ## accommodated from bartools::plotBarcodehistogram
source("analysis/F3_simulation.R") ## for negative simulation
source("analysis/ggplot_theme.R") ## setting ggplot theme
```

# Install `barbieQ`

Installing the latest devel version of `barbieQ` from GitHub.

```{r load barbieQ}
if (!requireNamespace("barbieQ", quietly = TRUE)) {
  remotes::install_github("Oshlack/barbieQ")
}

library(barbieQ)
```

Check the version of `barbieQ`.

```{r}
packageVersion("barbieQ")
```

# Set seeds

```{r}
set.seed(2025)
```

# Load data

```{r}
load("output/mixture_barbieQ.rda")
```

# F3A: Experimental schematic

```{r}
schematic <- image_read("output/f3a_schematic.png")
f3a <- rasterGrob(schematic, interpolate = FALSE) 
wrap_elements(full = f3a)
```

# F3B: Plot barcode proportion and sample annotation

`plotBarcodeHistogram`, sourced from "plotBarcodeHistogram.R", is customized from the `bartools` package.

```{r, fig.width=7, fig.height=4}
## pull out sample designs
design_mixed <- mixture_top$sampleMetadata %>% as.data.frame()
design_mixed$sample <- colnames(mixture_top)
str(design_mixed)

design_mixed$y = 1.05
color_mapping <- mixture_top@metadata$factorColors

## order samples
order_sample <- design_mixed %>% with(order(Subset, Perturbation, BarcodeSize))

mixed_dge <- DGEList(
  counts = assay(mixture_top), 
  group = mixture_top$sampleMetadata$Subset)

## pull out ggplot data
p_mixed <- plotBarcodeHistogram(mixed_dge, orderSamples = colnames(mixture_top)[order_sample])

p_mixed_illus <- p_mixed +
  ## label subset
  ggnewscale::new_scale_fill() +
  geom_tile(data = design_mixed, aes(x = sample,y = y, fill = Subset), height = 0.04, inherit.aes = FALSE, color = "white") +
  scale_fill_manual(values = color_mapping$Subset, labels = c(null = "baseline", perturb = "perturbed")) +
  labs(fill = "Subset") + 
  guides(fill = guide_legend(ncol = 2)) + 
  ## label perturbations
  ggnewscale::new_scale_fill() +
  geom_tile(data = design_mixed, aes(x = sample,y = y+0.05, fill = as.factor(Perturbation)), height = 0.04, inherit.aes = FALSE, color = "white") +
  scale_fill_manual(values = color_mapping$Perturbation) +
  labs(fill = "Perturbation level") + 
  guides(fill = guide_legend(ncol = 2)) + 
  ## label libsize
  ggnewscale::new_scale_fill() +
  geom_tile(data = design_mixed, aes(x = sample,y = y+0.10, fill = as.factor(BarcodeSize)), height = 0.04, inherit.aes = FALSE, color = "white") +
  scale_fill_manual(values = color_mapping$BarcodeSize)+
  labs(fill = "Num. of cells") + 
  guides(fill = guide_legend(ncol = 2)) +
  labs(x = "Sample", y = "Barcode proportion") +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2)) +
  theme(axis.text.x = element_blank(), panel.grid.major.x = element_blank())

f3b <- p_mixed_illus + 
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 11), aspect.ratio = 1,
        axis.title = element_text(size = 13), axis.text = element_text(size = 12)) + 
  guides(fill = guide_legend(ncol = 2))
f3b

```

# FPR

Using [mixture_top](barbieQ_paper_S1.html) data, where an offset (one) was added to the raw count, based on which proportion was calculated. However the "top" barcodes were filtered based on the raw count data.

Here we subset the 12 "null" samples for evaluating FPR of diff_prop, and diff_occ methods.

Three transformations with dif_prop were evaluated separately.

 - **avoid running this chunk during rendering. `global_all_loops` was saved from the first run.**

```{r eval=FALSE, fig.height=8, fig.width=16, include=TRUE}
nulls <- mixture_top[, mixture_top$sampleMetadata$Subset == "null"]

results_negsi <- suppressMessages({
  random_sampling(barbieQ = nulls, loop_times = 100)
  })

## `random_sampling()` save the simulations to global environment
## as `global_all_loops`
save(global_all_loops, file = "output/mixture_negative_simulation.rda")

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, perturb_null")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, perturb_null")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, perturb_null")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, perturb_null")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, perturb_null")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, perturb_null")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, perturb_null")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, perturb_null")

wrap_plots(results_negsi[1:8], ncol = 4)
```

## load results

```{r}
## loading random sampling results to avoid running it repeatedly
load("output/mixture_negative_simulation.rda")
```

```{r }
  # extract results from the loops
  end_sampling <- 6
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```

## F3C: Type I Error Rate

 - Diff_prop

```{r}
df_FPR$N_Samples <- as.factor(df_FPR$N_Samples)
  
  ## reshape the data to fit methods for testing
df_FPR_long <- df_FPR %>%
  pivot_longer(cols = starts_with("Prop_"),
               names_to = "Method",
               values_to = "Pval_Prop") %>%
  mutate(Method = factor(Method, levels = c("Prop_asin", "Prop_logit", "Prop_noTrans"),
                         labels = c("asin-sqrt", "logit", "no trans.")))

P_FPR_add1_prop <- ggplot(
  df_FPR_long, aes(x = factor(N_Samples), y = Pval_Prop, color = Method)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       title = "Type I error rate",
       subtitle = "Differential proportion test") +
  # facet_wrap(~Method, scales = "free_y") +
  theme_classic() +
  theme(legend.position = "top")
```

 - Diff_occ. when converting barcode presence / absence, setting the threshold as CPM>10 instead of the default 1, as an offset (one) was added to the raw count [preprocessing](barbieQ_paper_S1.html).

```{r}
P_FPR_occ <- ggplot(
  df_FPR, aes(x = factor(N_Samples), y = Occ_firth)) +
  geom_boxplot(outlier.shape = 1, position = position_dodge(width = 0.7)) +
  # geom_jitter(size = 1, position = position_dodge(width = 0.7)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Differential occurrrence test" = "black")) +
  # coord_cartesian(ylim = c(0, 0.2)) + 
  labs(
       y = "Fraction of P-Value < 0.05",
       x = "Number of samples per group",
       titlte = "",
       subtitle = "Differential occurrence test") +
  theme_classic() +
  theme(legend.position = "top")

```

```{r, fig.width=8, fig.height=5}
f3c <- (P_FPR_add1_prop + theme_barbieQ_size() + theme(legend.position = "top")) + 
  (P_FPR_occ + theme_barbieQ_size())

f3c
```

## F3H: Type I Error Rate as a function to N_BC

### Generate subsets of barcodes

```{r}
## with different filtering thresholds, how many "top" barcodes do we get
1:9/10
lapply(
  c(1:9/10, 0.95), 
  function(filter_level) {
    mixture <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = filter_level)
    paste0("filter_thresh=", filter_level, "; N_barcodes=", sum(rowData(mixture)$isTopBarcode$isTop)) %>% print()
    # plotBarcodeSankey(mixture)
    
    return(
      sum(rowData(mixture)$isTopBarcode$isTop)
    )
  }
  ) %>% unlist() %>% 
  setNames(nm = c(1:9/10, 0.95)) ->
  filter_to_N

filter_to_N
names(filter_to_N) %>% str()
```

### Update proportion

```{r}
## Compute medians per bin/method/N_Samples
method_labels <- c(
  "Prop_asin" = "asin-sqrt",
  "Prop_logit" = "logit",
  "Prop_noTrans" = "none"
 )

```

 - avoid running this chunk during rendering. `stats_barcodes_all_levels` was saved from the first run.

```{r eval=FALSE, include=TRUE}
stats_barcodes_all_levels <- lapply(
  c(1:9/10, 0.95),
  function(i) {
    stats_barcodes_i <- get_null_sample_by_filter(mixture, i, loop_times = 100)
    stats_barcodes_i$filter_level <- i
    stats_barcodes_i$N_BC <- filter_to_N[i %>% as.character()]
    return(stats_barcodes_i)
  }
)

stats_barcodes_all_levels <- do.call(rbind, stats_barcodes_all_levels)

save(stats_barcodes_all_levels, file = "output/fdr_by_Nbc.rda")
```


```{r}
load("output/fdr_by_Nbc.rda")

 FPR_box <- stats_barcodes_all_levels %>%
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>% 
    group_by(method, N_Samples, N_BC, Loop_N) %>%
    summarize(N_signif = sum(pvalue < 0.05), Frac_signif = sum(pvalue < 0.05) / n(), .groups = "drop") %>% 
    mutate(Method = recode(method, !!!method_labels))
  
  ggplot(FPR_box, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(N_BC, Method))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05", color = "Method", shape = "N_sample", linetype = "N_sample",
      title = "Type I Error Rate as a function to number of barcodes"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~N_Samples, ncol =4) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12)) + 
    scale_y_continuous(limits = c(0,0.2))
```

### Raw proportion


```{r}
## saving all p.vals from the N=1553 simulation
# df_P.Val
mixture <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = 0.95)
## get the subset of barcodes
lapply(c(1:9/10, 0.95), function(filter_level) {
  barbieQ <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = filter_level)
  tag <- rowData(barbieQ)$isTopBarcode$isTop
  tag <- tag[mixture@elementMetadata$isTopBarcode$isTop]
  return(tag)
}) -> BC_set_by_filter

BC_set_by_filter <- do.call(cbind, BC_set_by_filter)
colnames(BC_set_by_filter) <- filter_to_N[c(1:9/10, 0.95) %>% as.character()]
## check
colSums(BC_set_by_filter)

mean_prop <- rowMeans(mixture[mixture@elementMetadata$isTopBarcode$isTop,mixture$sampleMetadata$Subset=="null"]@assays@data$proportion)

## testing results obtained from raw proportin in stats_barcodes
df_P.Val_filter_tag <- cbind(
  stats_barcodes,  # testing results obtained from raw proportin
  mean_prop,
  BC_set_by_filter[
    rep(
      seq_len(nrow(BC_set_by_filter)),
      length.out = nrow(stats_barcodes)
      ),
    ]
  )
```

```{r}
FPR_box_raw <- df_P.Val_filter_tag %>%
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>% 
   pivot_longer(
     cols = as.character(filter_to_N[c(1:9/10, 0.95) %>% as.character()]),
     names_to = "N_BC",
     values_to = "selectedBy"
   ) %>% 
    group_by(method, N_Samples, Loop_N, N_BC) %>%
    summarize(
      N_signif = sum(pvalue < 0.05 & selectedBy), 
      Frac_signif = sum(pvalue < 0.05 & selectedBy) / sum(selectedBy), .groups = "drop") %>% 
    mutate(Method = recode(method, !!!method_labels))

FPR_box_raw$N_BC <- as.numeric(FPR_box_raw$N_BC)
  
ggplot(FPR_box_raw, aes(x= as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(Method, N_BC))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Valuev < 0.05", color = "Method", shape = "N_sample", linetype = "N_sample"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~N_Samples, ncol =4) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme_barbieQ_size() + 
    scale_y_continuous(limits = c(0,0.2)) ## y scale zoom in
```

```{r, fig.width=5, fig.height=7}
rbind(
  data.frame(FPR_box_raw, prop_type  = "Raw proportion"), 
  data.frame(FPR_box, prop_type  = "Updated proportion")
  ) %>% 
  mutate(prop_type = factor(prop_type, levels = c("Updated proportion", "Raw proportion"))) %>% 
  ## plot
  ggplot(aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(N_BC, Method))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05", color = "Method", shape = "N_sample", linetype = "N_sample",
      title = "Type I error rate to number of barcodes"
      ) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  facet_grid(prop_type~1, switch = "y") +
  theme_bw()+
  theme_barbieQ_size() +
  theme(panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5), 
        legend.position = "top",
        strip.placement = "outside", strip.text.y = element_text(size = 12),
        strip.text.x = element_blank(), strip.background.x = element_blank()) +
  scale_y_continuous(limits = c(0,0.2)) -> f3h
  ## scale_x_discrete(limits = rev) ## reverse x axis

f3h
  
```


## F3G: Density to N_Barcodes

### Updated proportion

```{r}
## get the subset of barcodes
lapply(c(1:9/10, 0.95), function(filter_level) {
  barbieQ <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = filter_level)
  tag <- rowData(barbieQ)$isTopBarcode$isTop
  
  prop_mat <- mixture[tag,mixture$sampleMetadata$Subset=="null"]@assays@data$proportion
  ## update prop
  prop_mat <- t(t(prop_mat) / colSums(prop_mat))
  
  df <- data.frame(
    prop = as.vector(prop_mat), 
    N_BC=filter_to_N[as.character(filter_level)]
    )
  return(df)
}) -> BC_set_prop_filter

BC_set_prop_filter <- do.call(rbind, BC_set_prop_filter)

ggplot(BC_set_prop_filter) +
  geom_density(aes(x = prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10(breaks = 10^seq(-6, -1, by = 1), labels = strip_zeros) +
  coord_cartesian(xlim = c(1e-6, 3e-1)) +
  labs(color = "Num. BC", x = "Barcode Proportion", y = "Density", subtitle = "with updated proportion") +
  theme_barbieQ() +
  theme(aspect.ratio = 0.9) -> f3g1
```


### Raw proportion

```{r}
## get the subset of barcodes
lapply(c(1:9/10, 0.95), function(filter_level) {
  barbieQ <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = filter_level)
  tag <- rowData(barbieQ)$isTopBarcode$isTop
  ## keep raw prop
  prop_mat <- mixture[tag,mixture$sampleMetadata$Subset=="null"]@assays@data$proportion
  df <- data.frame(
    prop = as.vector(prop_mat), 
    N_BC=filter_to_N[as.character(filter_level)]
    )
  return(df)
}) -> BC_set_prop_filter_raw

BC_set_prop_filter_raw <- do.call(rbind, BC_set_prop_filter_raw)

ggplot(BC_set_prop_filter_raw) +
  geom_density(aes(x = prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10(breaks = 10^seq(-6, -1, by = 1), labels = strip_zeros) +
  coord_cartesian(xlim = c(1e-6, 3e-1)) +
  theme_linedraw() +
  labs(color = "Num. BC", x = "Barcode Proportion", y = "Density", subtitle = "with raw proportion") +
  theme_barbieQ() +
  theme(aspect.ratio = 0.9) -> f3g2
```

```{r fig.width=4, fig.height=7}
rbind(
  data.frame(BC_set_prop_filter_raw, prop_type  = "Raw proportion"), 
  data.frame(BC_set_prop_filter, prop_type  = "Updated proportion")
  ) %>% 
  mutate(prop_type = factor(prop_type, levels = c("Updated proportion", "Raw proportion"))) %>% 
  ## pass to gplot
  ggplot() +
  geom_density(aes(x = prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10(breaks = 10^seq(-6, -1, by = 1), labels = strip_zeros) +
  coord_cartesian(xlim = c(1e-6, 3e-1)) +
  facet_grid(prop_type~1, switch = "y") +
  labs(color = "Num. BC", x = "Barcode Proportion", y = "Density", title = "Barcode distribution in datasets") +
  theme_bw() +
  theme_barbieQ_size() +
  theme(panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5), 
        legend.position = "top", legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size = 11), legend.text  = element_text(size = 10), 
        strip.placement = "outside", strip.text.y = element_text(size = 12),
        strip.text.x = element_blank(), strip.background.x = element_blank()) ->  ## disable column label
  f3g
  
f3g
```

## F3D: Quantile distribution

```{r, fig.width=4, fig.height=5}
data.frame(x = (stats_barcodes_all_levels %>% filter(filter_level==0.95))$Amean_prop) %>%
  mutate(quantile_bin = ntile(x, 10)) %>% 
  ggplot(aes(x = as.factor(quantile_bin), y = x, group = quantile_bin)) +
  geom_jitter(color = "grey", shape = ".") + 
  geom_boxplot() +
  scale_y_log10(breaks = 10^seq(-5, -1, by = 1), labels = strip_zeros) +
  theme_barbieQ() +
  labs(x = "Quantiles of barcode mean proportion", y = "Barcode mean proportion",
       title = "Barcode proportion distribution", subtitle = " ") +
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) -> f3d

f3d
```

## F3E: False positives in quantiles

- For diff_prop

```{r}
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.1), num_quantiles = 4)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.2), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.3), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.4), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.5), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.6), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.7), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.8), num_quantiles = 10)
# plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.9), num_quantiles = 10)

plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.95), num_quantiles = 10) + 
  theme_barbieQ() +
  labs(title = "Type I error rate in quanties") -> f3e
```

# TPR

## Check premix

- confirming an equal mixture of pool1 and pool2 in baseline samples (the "null" samples).

```{r}
colnames(mixture_top)

prop_raw <- mixture_top@assays@data$proportion

pool1_vec <- prop_raw[,1]
pool2_vec <- prop_raw[,2]

## if equal mixture of pool1 and pool2; designed prop should be 1/2 of total props in original pool1 and pool2 
check_mix_df <- data.frame(
  ps = rowMeans(prop_raw[,1:2]), 
  mixture_660 = prop_raw[,3]
  )

ggplot(check_mix_df, aes(x = ps, y = mixture_660)) + 
  geom_point() + 
  coord_fixed() +
  geom_smooth(method = "lm")
```

 - Confirming orthogonal between premix1 and premix2: barcodes mostly not overlapping between the two premix samples, but a few noise are observed in premix2.

```{r}
ggplot(data.frame(prop_raw) %>% sqrt() %>% asin()) + 
  geom_point(aes(x = P1, y = P2))

ggplot(data.frame(prop_raw) %>% sqrt() %>% asin(), aes(x = P1+P2, y = abs(P1-P2))) + 
  geom_point() +
  coord_fixed() +
  geom_smooth(method = "lm")

ggplot(data.frame(prop_raw), aes(x = P1+P2, y = abs(P1-P2))) + 
  geom_point() +
  coord_fixed() +
  geom_smooth(method = "lm")

```

## Designed differences

```{r}
## theory
a = 0.35
((1/2 + a) / (1+a)) - (1/2)
a = 0.27
(1/2 + a) / (1+a) - 1/2
a = 0.18
(1/2 + a) / (1+a) - 1/2

```

```{r eval=FALSE, include=FALSE}
calc_theory(a = 0.18)
calc_theory(a = 0.18, trans = "asin-sqrt")
calc_theory(a = 0.18, trans = "logit")

calc_theory(a = 0.27)
calc_theory(a = 0.27, trans = "asin-sqrt")
calc_theory(a = 0.27, trans = "logit")

calc_theory(a = 0.35)
calc_theory(a = 0.35, trans = "asin-sqrt")
calc_theory(a = 0.35, trans = "logit")
```

## Model perturbations

```{r}
mixture_top$sampleMetadata %>% as.data.frame() %>% summary()

mixed <- mixture_top[, mixture_top$sampleMetadata$Subset %in% c("perturb", "null")]
mixed$sampleMetadata %>% as.data.frame() %>%  data.table()

mixed_target <- mixed$sampleMetadata %>% as.data.frame()
## factorize perturbation levels and barcodeSize levels
mixed_target$Perturbation <- as.factor(mixed_target$Perturbation)
levels(mixed_target$Perturbation)
mixed_target$BarcodeSize <- as.factor(mixed_target$BarcodeSize)
levels(mixed_target$BarcodeSize)

mixed_design <- mixed_target %>%
    with(model.matrix(~0 + Perturbation + BarcodeSize))

myblock <- mixed_target %>% with(paste(BarcodeSize, Perturbation))
table(myblock)

```

## Apply tests: extract estimates


```{r}
myblock <- NULL

es_de_none <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(mixed, mixed_design, myblock, a = i))
es_de_none_df <- do.call(rbind, es_de_none)
es_de_none_df <- es_de_none_df %>% 
  mutate(Signif. = ifelse(P.Value<0.05, "signif.", "n.s.")) %>% 
  mutate(Signif. = factor(Signif., levels = c("signif.", "n.s.")))

es_de_asin <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(mixed, mixed_design, myblock, a = i, trans = "asin-sqrt"))
es_de_asin_df <- do.call(rbind, es_de_asin)
es_de_asin_df <- es_de_asin_df %>% 
  mutate(Signif. = ifelse(P.Value<0.05, "signif.", "n.s.")) %>% 
  mutate(Signif. = factor(Signif., levels = c("signif.", "n.s.")))

es_de_logit <- lapply(c(0.18, 0.27, 0.35), function(i) apply_model(mixed, mixed_design, myblock, a = i, trans = "logit"))
es_de_logit_df <- do.call(rbind, es_de_logit)
es_de_logit_df <- es_de_logit_df %>% 
  mutate(Signif. = ifelse(P.Value<0.05, "signif.", "n.s.")) %>% 
  mutate(Signif. = factor(Signif., levels = c("signif.", "n.s.")))

es_de_df <- rbind(es_de_none_df, es_de_asin_df, es_de_logit_df)

```

## FS3A: Validate estimates

```{r}
regression_stats_none <- es_de_none_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(diff, meanDiff),
    slope = coef(lm(meanDiff ~ I(diff)))[2],
    interc = coef(lm(meanDiff ~ I(diff)))[1],
    # res = resid(mod <- lm(meanDiff ~ I(diff))),
    # meanDiff = meanDiff,
    .groups = "drop"
  )
regression_stats_none$Method = "none"
regression_stats_none$x = -0.01
regression_stats_none$y = 0.01

regression_stats_asin <- es_de_asin_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(diff, meanDiff),
    slope = coef(lm((meanDiff) ~ I(diff)))[2],
    interc = coef(lm((meanDiff) ~ I(diff)))[1],
    .groups = "drop"
  )
regression_stats_asin$Method = "asin-sqrt"
regression_stats_asin$x = -0.001
regression_stats_asin$y = 0.001


regression_stats_logit <- es_de_logit_df %>%
  group_by(Contrast) %>%
  summarise(
    cor = cor(diff, (meanDiff)),
    slope = coef(lm((meanDiff) ~ I(diff)))[2],
    interc = coef(lm((meanDiff) ~ I(diff)))[1],
    .groups = "drop"
  )
regression_stats_logit$Method = "logit"
regression_stats_logit$x = -0.05
regression_stats_logit$y = 1

regression_stats <- rbind(regression_stats_asin, regression_stats_none, regression_stats_logit)
```

(facet_grid doesn't allow free x sclae for the same column. so plot three methods seperately.)

```{r, fig.width=8, fig.height=4}
p_es_de_asin <- ggplot(es_de_asin_df, aes(x = diff, y = (meanDiff))) +
  geom_point(aes(color = Signif.), alpha = 0.5) +
  # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
    geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
  theme_bw() +
  # coord_fixed() +
  facet_wrap(~paste0("Contrast: ", Contrast)) +
  geom_text(data = regression_stats_asin, x = -0.01, y = 0.01, 
            aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
  theme(axis.title.x = element_blank(), legend.position = "none") +
  labs(color = "Barcode", y = " \n \n (asin-sqrt)")


p_es_de_none <- ggplot(es_de_none_df, aes(x = diff, y = (meanDiff))) +
  geom_point(aes(color = Signif.), alpha = 0.5) +
  # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
  geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
  theme_linedraw()+
  # coord_fixed() +
  facet_wrap(~paste0("Contrast: ", Contrast)) +
    geom_text(data = regression_stats_none, x = -0.0015, y = 0.001, 
            aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
  labs(y = " \n \n (no transformation)", x = " \n Designed differences")+
  ## zoom in, removing a few barcodes with large proportion. 
  scale_x_continuous(limits = c(-0.003, 0.003)) +
  scale_y_continuous(limits = c(-0.003, 0.003)) +
  theme(legend.position = "none", strip.text = element_blank()) + # remove facet titles
  labs(color = "Barcode")

p_es_de_logit <- ggplot(es_de_logit_df, aes(x = diff, y = (meanDiff))) +
  geom_point(aes(color = Signif.), alpha = 0.5) +
  # stat_smooth(method = "lm", linetype = "dashed", colour = "black") +
  geom_abline(data = regression_stats_none, aes(intercept = interc, slope = slope), linetype = "dashed") +
  theme_linedraw()+
  facet_wrap(~paste0("Contrast: ", Contrast)) +
    geom_text(data = regression_stats_logit, x = -0.15, y = 1, 
            aes(label = paste0("r = ", round(cor, digits = 2), "\n", "slope = ", round(slope, digits = 2)))) + 
  labs(y = "Estimated differences \n \n (logit)", color = "Barcode") +
   ## zoom in, removing one barcode with large proportion. 
  # scale_x_continuous(limits = c(-0.3, 0.3)) +
  # scale_y_continuous(limits = c(-0.3, 0.3)) +
  theme(axis.title.x = element_blank(), legend.position = "right", strip.text = element_blank())  # remove facet titles
```

```{r, fig.width=9, fig.height=8}
fs3a <- (p_es_de_asin + theme(legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12)) ) / 
  (p_es_de_logit + theme(legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))) /
  (p_es_de_none + theme(legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))) 
fs3a
```
## F3I: TPR as a function to N_BC

### Updated proportion

```{r}
test_perturb <- function(mixed, mixed_target, mixed_design, myblock, trans = "none", a = 0.18) {
  tester <- testBarcodeSignif(
    barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation", 
    contrastFormula = paste0("Perturbation", a, " - Perturbation0"), transformation = trans,
    designMatrix = mixed_design, block = myblock)
  
  veri_diff <- as.data.frame(tester@elementMetadata$testingBarcode)
  
  veri_diff <- rownames_to_column(veri_diff, var = "BC") 
  veri_diff$Contrast = paste0(a, " - 0")
  veri_diff$Method = trans

  
  return(veri_diff)
}

```

```{r}
test_all_levels <- lapply(
  c(1:9/10, 0.95),
  function(i) {get_mixted_by_filter(barbieQ = mixture, filter_level = i)}
)

test_all_levels <- do.call(rbind, test_all_levels)

test_TPR <- test_all_levels %>% 
  group_by(Method, Contrast, N_BC) %>% 
  summarize(Frac_signif = sum(P.Value < 0.05) / n(), .groups = "drop")

```

```{r}
ggplot(test_TPR, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, linetype = Contrast, shape = Contrast)) +
  # geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
  geom_line(aes(group = paste0(Contrast, Method))) +
  geom_point() +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05",
      title = "True postive rate as a function to total barcode number"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))

```

### Raw proportion

```{r}
## test on raw proportion
test_0.95_raw <- get_mixted_by_filter(barbieQ = mixture, filter_level = 0.95, update_prop = FALSE)

test_all_levels_raw <- cbind(
  test_0.95_raw,  
  mean_prop,
  BC_set_by_filter[
    rep(
      seq_len(nrow(BC_set_by_filter)),
      length.out = nrow(test_0.95_raw)
      ), 
    ]
  )

test_TPR_raw <- test_all_levels_raw %>%
   select(-N_BC) %>% 
   pivot_longer(
     cols = as.character(filter_to_N[c(1:9/10, 0.95) %>% as.character()]),
     names_to = "N_BC",
     values_to = "selectedBy"
   ) %>% 
    group_by(Method, Contrast, N_BC) %>% 
    summarize(Frac_signif = sum(P.Value < 0.05 & selectedBy) / sum(selectedBy), .groups = "drop")

test_TPR_raw$N_BC <- as.numeric(test_TPR_raw$N_BC)

ggplot(test_TPR_raw, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(Method, Contrast))) +
  # geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
  geom_point(aes(shape = Contrast)) +
  geom_line(aes(linetype = Contrast)) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~Contrast, ncol =3) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))

```

```{r, fig.width=5, fig.height=7}
rbind(
  data.frame(test_TPR_raw, prop_type  = "Raw proportion"), 
  data.frame(test_TPR, prop_type  = "Updated proportion")
  ) %>% 
  mutate(prop_type = factor(prop_type, levels = c("Updated proportion", "Raw proportion"))) %>% 
  ## plot
  ggplot(aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(Method, Contrast))) +
  # geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
  geom_point(aes(shape = Contrast)) +
  geom_line(aes(linetype = Contrast)) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05", color = "Method", shape = "Contrast", linetype = "Contrast",
      title = "True positive rate to number of barcodes"
      ) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  facet_grid(prop_type~1, switch = "y") +
  theme_bw()+
  theme_barbieQ_size() +
  theme(panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5), 
        legend.position = "top",
        strip.placement = "outside", strip.text.y = element_text(size = 12),
        strip.text.x = element_blank(), strip.background.x = element_blank()) + guides(color = "none") -> f3i

f3i
```


## F3F: TPR in quantiles

```{r}
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.1), num_quantiles = 4)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.2), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.3), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.4), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.5), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.6), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.7), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.8), num_quantiles = 10)
# plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.9), num_quantiles = 10)

plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.95), num_quantiles = 10) +
  theme_barbieQ() +
  labs(title = "True positive rate in quanties") -> f3f
```


## FS3B: Residule Normality

```{r}
resi_none <- lapply(c(0.18, 0.27, 0.35), function(i) get_residules(mixed, mixed_design, myblock, a = i, trans = "none"))
resi_asin <- lapply(c(0.18, 0.27, 0.35), function(i) get_residules(mixed, mixed_design, myblock, a = i, trans = "asin-sqrt"))
resi_logit <- lapply(c(0.18, 0.27, 0.35), function(i) get_residules(mixed, mixed_design, myblock, a = i, trans = "logit"))

resi_df <- rbind(do.call(rbind, resi_none), do.call(rbind, resi_asin), do.call(rbind, resi_logit))

```

 - Distribution and QQ plot.

```{r}
p_resid_distr_35 <- ggplot(resi_df %>% filter(Contrast == "0.35 - 0"),
                        aes(x = Residual, color = Method)) +
  geom_density() +
  labs(x = "Residual", y = "Density") +
  facet_wrap(~Method, scales = "free", ncol = 1, strip.position = "left") +
  theme_bw() +
  theme(legend.position = "none", plot.title.position = "panel")


p_resid_qq_35 <- ggplot(resi_df %>% filter(Contrast == "0.35 - 0"), 
                     aes(sample = Residual, color = Method)) +
  stat_qq() +
  stat_qq_line(color = "black", linetype = "dashed") +
  labs(x = "Normal Quantiles", y = "\n Residual Quuantiles") +
  facet_wrap(~Method, scales = "free", ncol = 1, strip.position = "left") +
  theme_barbieQ() +
  theme(legend.position = "none", plot.title.position = "panel") ##strip.text = element_blank()
```

```{r, fig.width=4, fig.height=9}
(p_resid_distr_35 + 
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 11), strip.text = element_text(size = 12),
        axis.title = element_text(size = 13), axis.text = element_text(size = 10))
  )
```


```{r, fig.width=4, fig.height=9}
fs3b <-
(p_resid_qq_35 + 
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 11), 
        axis.title = element_text(size = 13), axis.text = element_text(size = 10)))

fs3b
```

```{r}
p_resid_distr_27 <- ggplot(resi_df %>% filter(Contrast == "0.27 - 0"), aes(x = Residual, color = Method)) +
  geom_density() +
  labs(x = "Residual", y = "Density") +
  facet_wrap(~Method, scales = "free", ncol = 1) +
  theme_bw() +
  theme(legend.position = "none", strip.text = element_blank())

p_resid_qq_27 <- ggplot(resi_df %>% filter(Contrast == "0.27 - 0"), aes(sample = Residual, color = Method)) +
  stat_qq() +
  stat_qq_line(color = "black", linetype = "dashed") +
  labs(x = "Normal Quantiles", y = "Residule Quantiles") +
  facet_wrap(~Method, scales = "free", ncol = 1) +
  theme_bw() +
  theme(legend.position = "none", strip.text = element_blank())

p_resid_distr_18 <- ggplot(resi_df %>% filter(Contrast == "0.18 - 0"), aes(x = Residual, color = Method)) +
  geom_density() +
  labs(x = "Residual", y = "Density") +
  facet_wrap(~Method, scales = "free", ncol = 1) +
  theme_bw() +
  theme(legend.position = "none", strip.text = element_blank())

p_resid_qq_18 <- ggplot(resi_df %>% filter(Contrast == "0.18 - 0"), aes(sample = Residual, color = Method)) +
  stat_qq() +
  stat_qq_line(color = "black", linetype = "dashed") +
  labs(x = "Normal Quantiles", y = "Residule Quantiles") +
  facet_wrap(~Method, scales = "free", ncol = 1) +
  theme_bw() +
  theme(legend.position = "none", strip.text = element_blank())
```

```{r, fig.width=8, fig.height=7}
p_resid_distr_35 + p_resid_distr_27 + p_resid_distr_18

p_resid_qq_35 + p_resid_qq_27 + p_resid_qq_18
```

## FS3C: Residule Variance

 - Residual standard deviation vs. mean to check homosdecasticity
 
```{r}
plot_residuals(mixed, mixed_design, myblock, a = 0.35, trans = "asin-sqrt")
plot_residuals(mixed, mixed_design, myblock, a = 0.35, trans = "logit")
plot_residuals(mixed, mixed_design, myblock, a = 0.35, trans = "none")
```

```{r, fig.width=4, fig.height=8}
## capture limma SA plot
make_sa_grob <- function(trans, a = 0.35) {
  grid::grid.grabExpr({
    gridGraphics::grid.echo(function() {
      limma::plotSA(
        fit = plot_residuals(mixed, mixed_design, myblock, a = a, trans = trans),
        xlab = "",
        ylab = ""
      )
    })
  })
}

sa_asin <- make_sa_grob("asin-sqrt")
sa_logit <- make_sa_grob("logit")
sa_none <- make_sa_grob("none")

fs3c_core <- (
  wrap_elements(full = sa_asin) /
  wrap_elements(full = sa_logit) /
  wrap_elements(full = sa_none)
)

ylab_plot <- ggplot() +
  annotate("text", x = 0.5, y = 0.5, label = "Residual SD", angle = 90, size = 4) +
  theme_void()

xlab_plot <- ggplot() +
  annotate("text", x = 0.5, y = 0.5, label = "Average transformed barcode proportion", size = 4) +
  theme_void()

fs3c <- (
  ylab_plot + fs3c_core +
    plot_layout(widths = c(0.1, 1))
) / xlab_plot +
  plot_layout(heights = c(1, 0.02))

fs3c

```

## Result consistency

```{r}
asin_adjP_3con <- es_de_asin_df %>%
  dplyr::select(BC, Contrast, P.Value) %>%
  tidyr::pivot_wider(
    id_cols = BC,
    names_from = Contrast,
    values_from = P.Value
  )
none_adjP_3con <- es_de_none_df %>%
  dplyr::select(BC, Contrast, P.Value) %>%
  tidyr::pivot_wider(
    id_cols = BC,
    names_from = Contrast,
    values_from = P.Value
  )

```

```{r}
eu <- eulerr::euler(list(
  asin_C0.18 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.18 - 0` < 0.05],
  asin_C0.27 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.27 - 0` < 0.05],
  asin_C0.35 = rownames(asin_adjP_3con)[asin_adjP_3con$`0.35 - 0` < 0.05],
  none_C0.18 = rownames(none_adjP_3con)[none_adjP_3con$`0.18 - 0` < 0.05],
  none_C0.27 = rownames(none_adjP_3con)[none_adjP_3con$`0.27 - 0` < 0.05],
  none_C0.35 = rownames(none_adjP_3con)[none_adjP_3con$`0.35 - 0` < 0.05]
))

plot(eu, quantities = TRUE)
```

# Figure 3

```{r, fig.width=15, fig.height=20}
layout = "
AAAAAAAAAAAAA##
AAAAAAAAAAAAA##
AAAAAAAAAAAAA##
BBBBBBBCCCCCCCC
BBBBBBBCCCCCCCC
BBBBBBBCCCCCCCC
BBBBBBBCCCCCCCC
DDDDEEEEEFFFFFF
DDDDEEEEEFFFFFF
DDDDEEEEEFFFFFF
DDDDEEEEEFFFFFF
DDDDEEEEEFFFFFF
GGGGHHHHHIIIIII
GGGGHHHHHIIIIII
GGGGHHHHHIIIIII
GGGGHHHHHIIIIII
GGGGHHHHHIIIIII
GGGGHHHHHIIIIII
GGGGHHHHHIIIIII
"
f3 <- ((wrap_elements(full = f3a) + theme(plot.margin = unit(c(0,0,0,0), "cm"))) + 
    wrap_elements(f3b + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(f3c + theme(plot.margin = unit(c(0,0,0,0), "cm"))) +
    wrap_elements(f3d + theme(legend.position = "top", plot.margin = unit(rep(0,4), "cm"))) +
    wrap_elements(f3e + theme(legend.position = "top", plot.margin = unit(rep(0,4), "cm"))) +
    wrap_elements(f3f + theme(legend.position = "top", plot.margin = unit(rep(0,4), "cm")) + guides(color = "none")) +
    wrap_elements(f3g + theme(legend.position = "top", plot.margin = unit(c(0,0,0,0), "cm"))) +
    wrap_elements(f3h + theme(legend.position = "top",plot.margin = unit(rep(0,4), "cm"))) +
    wrap_elements(f3i + theme(legend.position = "top",plot.margin = unit(rep(0,4), "cm"))) 
      
    )+ 
  plot_layout(design = layout) +
  ## space is okay but must be unique
  plot_annotation(tag_levels = list(c("A", "B","C", "D", "E", "F", "G", "H", "I"))) &
  theme(plot.tag = element_text(size = 24, face = "bold", family = "arial"))

f3
```

```{r}
ggsave(
  filename = "output/f3.png",
  plot = f3,
  width = 15,
  height = 20,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [f3](../output/f3.png)

# Figure S3

```{r, fig.width=9, fig.height=16}
layoutS4 = "
AAAAAAAAA
BBBBCCCC#
"
fs3 <- (
    wrap_elements(fs3a + theme(plot.margin = unit(rep(0,4), "cm"))) + 
    wrap_elements(fs3b + theme(plot.margin = unit(c(0,0,0,0.2), "cm"))) +
    (fs3c & theme(plot.margin = unit(c(0,0,1,0), "cm")))
    )+ 
  plot_layout(design = layoutS4) +
  ## space is okay but must be unique
  plot_annotation(tag_levels = list(c("A", "B", "C"))) &
  theme(plot.tag = element_text(size = 24, face = "bold", family = "arial"))

fs3
```

```{r}
ggsave(
  filename = "output/fs3.png",
  plot = fs3,
  width = 9,
  height = 16,
  units = "in", # for Rmd r chunk fig size, unit default to inch
  dpi = 350
  )
```

Saving this figure in [fs3](../output/fs3.png)