---
title: "FPR_to_NBC"
output: html_document
---

```{r}
load("output/mixture_barbieQ.rda")
load("output/mixture_negative_simulation.rda")
source("analysis/F3_simulation.R")
```

# Filtering levels

```{r}
## with different filtering thresholds, how many "top" barcodes do we get
1:9/10
lapply(
  c(1:9/10, 0.95), 
  function(filter_level) {
    mixture_plus1 <- tagTopBarcodes(mixture_plus1, nSampleThreshold = 8, proportionThreshold = filter_level)
    paste0("filter_thresh=", filter_level, "; N_barcodes=", sum(rowData(mixture_plus1)$isTopBarcode$isTop)) %>% print()
    # plotBarcodeSankey(mixture)
    
    return(
      sum(rowData(mixture_plus1)$isTopBarcode$isTop)
    )
  }
  ) %>% unlist() %>% 
  setNames(nm = c(1:9/10, 0.95)) ->
  filter_to_N

filter_to_N
names(filter_to_N) %>% str()
```

```{r}
lapply(
  c(1:9/10, 0.95), 
  function(filter_level) {
    mixture_plus1 <- tagTopBarcodes(mixture_plus1, nSampleThreshold = 8, proportionThreshold = filter_level)
    rownames(mixture)[rowData(mixture_plus1)$isTopBarcode$isTop]
  }
  ) -> BC_list_filter

names(BC_list_filter) <- c(1:9/10, 0.95)
eulerr::euler(BC_list_filter) %>% plot() 
```


# FPR

## Update proportion

```{r}
get_null_sample_by_filter <- function(barbieQ, filter_level = 0.95, loop_times = 3) {
  ## tag top BC from raw barbieQ
  barbieQ <- tagTopBarcodes(barbieQ, nSampleThreshold = 8, proportionThreshold = filter_level)
  ## select top BC from barbieQ_plus1
  barbieQ_top <- barbieQ[rowData(barbieQ)$isTopBarcode$isTop,]
  ## subset baseline "null" samples
  nulls <- barbieQ_top[, barbieQ_top$sampleMetadata$Subset == "null"]
  ## re-calculate proportions
  nulls@assays@data$proportion <- ((nulls@assays@data$proportion %>% t()) / colSums(nulls@assays@data$proportion)) %>% t()
  
  ## run random sampling loops
  result_loops <- suppressMessages({
    get_random_sampling_loops(barbieQ = nulls, loop_times = loop_times)
  })

  ## extract results from the loops
  end_sampling <- 6
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {result_loops[[n]]$FPR})
  df_FPR <- do.call(rbind, all_FPR)
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {result_loops[[n]]$P.Val})
  df_P.Val <- do.call(rbind, all_P.Val)
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {result_loops[[n]]$MA})
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {result_loops[[n]]$Group_Vec})
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
  
  return(stats_barcodes)
}

```

```{r}
## Helper to format in "×10^" notation
format_math <- function(x) {
  exp <- floor(log10(x))
  base <- x / 10^exp
  # Return expression object (not a string!)
  bquote(.(formatC(base, digits = 2, format = "f")) %*% 10^.(exp))
 }

## Compute medians per bin/method/N_Samples
method_labels <- c(
  "Prop_asin" = "asin-sqrt",
  "Prop_logit" = "logit",
  "Prop_noTrans" = "none"
 )

plot_fdr_on_quantile <- function(stats_barcodes_i, num_quantiles = 10) {
  
  quantile_breaks <- quantile(stats_barcodes_i$Amean_prop, probs = seq(0, 1, length.out = num_quantiles+1)) 

  bin_labels <- mapply(
  function(lo, hi) bquote("("*.(format_math(lo))*","*.(format_math(hi))*"]"),
  head(quantile_breaks, -1),
  tail(quantile_breaks, -1),
  SIMPLIFY = FALSE
)

  ## specify label here
  stats_binned <- stats_barcodes_i %>%
    mutate(bin = cut(Amean_prop, breaks = quantile_breaks, labels = bin_labels, include.lowest = TRUE, right = TRUE)) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    )

  FPR_lines <- stats_binned %>%
    group_by(bin, method, N_Samples, Loop_N) %>%
    summarize(Frac_signif = sum(pvalue < 0.05) / n(), .groups = "drop") %>% 
    mutate(Method = recode(method, !!!method_labels))
  
  p <- ggplot(FPR_lines, aes(x = bin, y = Frac_signif, color = Method, group = paste0(bin, Method))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8), outliers = T) +
    labs(
      x = "Quantiles of barcode mean proportion (raw)",
      y = "Fraction of P-Value < 0.05", color = "Method", shape = "N_sample", linetype = "N_sample"
      # caption = paste("Quantile mapping:\n", legend_text)
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    theme_linedraw() +
    scale_x_discrete(labels = scales::label_parse()) +
    theme(axis.text.x = element_text(angle = 75, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))

  return(p)
}

```


```{r}
stats_barcodes_all_levels <- lapply(
  c(1:9/10, 0.95),
  # c(1:3/10),
  function(i) {
    stats_barcodes_i <- get_null_sample_by_filter(mixture_plus1, i, loop_times = 100)
    stats_barcodes_i$filter_level <- i
    stats_barcodes_i$N_BC <- filter_to_N[i %>% as.character()]
    return(stats_barcodes_i)
  }
)

stats_barcodes_all_levels <- do.call(rbind, stats_barcodes_all_levels)

save(stats_barcodes_all_levels, file = "output/fdr_by_Nbc.rda")
```


```{r, fig.width=16, fig.height=5}
load("output/fdr_by_Nbc.rda")

 FPR_box <- stats_barcodes_all_levels %>%
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>% 
    group_by(method, N_Samples, N_BC, Loop_N) %>%
    summarize(N_signif = sum(pvalue < 0.05), Frac_signif = sum(pvalue < 0.05) / n(), .groups = "drop") %>% 
    mutate(Method = recode(method, !!!method_labels))
  
  p <- ggplot(FPR_box, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(N_BC, Method))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "False discovery rate (FDR)", color = "Method", shape = "N_sample", linetype = "N_sample",
      title = "FDR as a function to total barcode number"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~N_Samples, ncol =4) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))
  
  p
  
  p + 
    scale_y_continuous(limits = c(0,0.1)) + 
    labs(subtitle = "y scale zoom in") -> fdr_new
```

```{r}
ggplot(FPR_box, aes(x = as.factor(N_BC), y = N_signif, color = Method, group = paste0(N_BC, Method))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "False discovery number", color = "Method", shape = "N_sample", linetype = "N_sample",
      title = "FDR as a function to total barcode number"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~N_Samples, ncol =4) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))

ggplot(FPR_box, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(N_BC, Method))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "False discovery rate (FDR)", color = "Method", shape = "N_sample", linetype = "N_sample",
      title = "FDR as a function to total barcode number"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~N_Samples, ncol =4) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))
```


zoom in , removing outliers; split comparisons.

```{r, fig.width=15, fig.height=10}
ggplot(FPR_box, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(N_BC, Method, ~N_Samples))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05", color = "Method", shape = "N_sample", linetype = "N_sample"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    facet_wrap(~N_Samples, ncol =2) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12)) + 
    scale_y_continuous(limits = c(0,0.17)) + 
    labs(subtitle = "y scale zoom in")
```

Fraction of fdr > 0.05

```{r, fig.width=10, fig.height=7}
thresh = 0.05

FPR_box_high <- FPR_box %>% 
    group_by(Method, N_Samples, N_BC) %>%
    summarize(N_high_FDR = sum(Frac_signif > thresh))
  
ggplot(FPR_box_high, aes(x = as.factor(N_BC), y = N_high_FDR, color = Method, group = paste0(N_BC, Method))) +
    # geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    geom_point() +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = paste0("Num. simulations with FDR > ", thresh),
      title = "Number of high-FDR simulations as a function to total barcode number",
      subtitle = paste0("sum(FDR>", thresh, ") out of 100 simulations at each htoup size")
      ) +
    geom_hline(yintercept = thresh, linetype = "dashed") +
    facet_wrap(~N_Samples, ncol =2) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))
  
```


```{r}
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.1), num_quantiles = 4)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.2), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.3), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.4), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.5), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.6), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.7), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.8), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.9), num_quantiles = 10)
plot_fdr_on_quantile(stats_barcodes_all_levels %>% filter(filter_level==0.95), num_quantiles = 10)
```

## Raw proportion

Load data as "barbieQ_paper_Figure3.Rmd"

```{r}
## saving all p.vals from the N=2224 simulation
# df_P.Val
mixture_plus1 <- tagTopBarcodes(mixture_plus1, nSampleThreshold = 8, proportionThreshold = 0.95)
## get the subset of barcodes
lapply(c(1:9/10, 0.95), function(filter_level) {
  barbieQ <- tagTopBarcodes(mixture_plus1, nSampleThreshold = 8, proportionThreshold = filter_level)
  tag <- rowData(barbieQ)$isTopBarcode$isTop
  tag <- tag[mixture_plus1@elementMetadata$isTopBarcode$isTop]
  return(tag)
}) -> BC_set_by_filter

BC_set_by_filter <- do.call(cbind, BC_set_by_filter)
colnames(BC_set_by_filter) <- filter_to_N[c(1:9/10, 0.95) %>% as.character()]
## check
colSums(BC_set_by_filter)

mean_prop <- rowMeans(mixture_plus1[mixture_plus1@elementMetadata$isTopBarcode$isTop,mixture_plus1$sampleMetadata$Subset=="null"]@assays@data$proportion)

df_P.Val_0.95 <- stats_barcodes_all_levels %>% filter(filter_level==0.95)

df_P.Val_filter_tag <- cbind(
  df_P.Val_0.95,  
  mean_prop,
  BC_set_by_filter[rep(seq_len(nrow(BC_set_by_filter)),length.out = nrow(df_P.Val_0.95)), ])
```

```{r, fig.width=12}
FPR_box_raw <- df_P.Val_filter_tag %>%
  select(-N_BC) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>% 
   pivot_longer(
     cols = as.character(filter_to_N[c(1:9/10, 0.95) %>% as.character()]),
     names_to = "N_BC",
     values_to = "selectedBy"
   ) %>% 
    group_by(method, N_Samples, Loop_N, N_BC) %>%
    summarize(
      N_signif = sum(pvalue < 0.05 & selectedBy), 
      Frac_signif = sum(pvalue < 0.05 & selectedBy) / sum(selectedBy), .groups = "drop") %>% 
    mutate(Method = recode(method, !!!method_labels))

FPR_box_raw$N_BC <- as.numeric(FPR_box_raw$N_BC)
  
ggplot(FPR_box_raw, aes(x= as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(Method, N_BC))) +
    geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "False discovery rate (FDR)", color = "Method", shape = "N_sample", linetype = "N_sample",
      title = "FDR as a function to total barcode number"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~N_Samples, ncol =4) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12)) + 
    scale_y_continuous(limits = c(0,0.1)) + 
    labs(subtitle = "y scale zoom in") -> fdr_old

fdr_new +fdr_old
```


# TPR

## Update proportion

```{r}
test_perturb <- function(mixed, mixed_target, mixed_design, myblock, trans = "none", a = 0.18) {
  tester <- testBarcodeSignif(
    barbieQ = mixed, sampleMetadata = mixed_target, sampleGroup = "Perturbation", 
    contrastFormula = paste0("Perturbation", a, " - Perturbation0"), transformation = trans,
    designMatrix = mixed_design, block = myblock)
  
  veri_diff <- as.data.frame(tester@elementMetadata$testingBarcode)
  
  veri_diff <- rownames_to_column(veri_diff, var = "BC") 
  veri_diff$Contrast = paste0(a, " - 0")
  veri_diff$Method = trans

  
  return(veri_diff)
}

get_mixted_by_filter <- function(barbieQ, filter_level = 0.95) {
  ## tag top BC from raw barbieQ
  barbieQ <- tagTopBarcodes(barbieQ, nSampleThreshold = 8, proportionThreshold = filter_level)
  ## select top BC from barbieQ_plus1
  barbieQ_top <- barbieQ[rowData(barbieQ)$isTopBarcode$isTop,]
  ## subset baseline "null" samples and perturbed samples
  mixed <- barbieQ_top[, barbieQ_top$sampleMetadata$Subset %in% c("null", "perturb")]
  
  ## re-calculate proportions
  mixed@assays@data$proportion <- ((mixed@assays@data$proportion %>% t()) / colSums(mixed@assays@data$proportion)) %>% t()
  
  mixed_target <- mixed$sampleMetadata %>% as.data.frame()
  mixed_target$Perturbation <- as.factor(mixed_target$Perturbation)
  # levels(mixed_target$Perturbation)
  mixed_target$BarcodeSize <- as.factor(mixed_target$BarcodeSize)
  # levels(mixed_target$BarcodeSize)

  mixed_design <- mixed_target %>%
      with(model.matrix(~0 + Perturbation + BarcodeSize))
  myblock <- mixed_target %>% with(paste(BarcodeSize, Perturbation))
  
  es_de_none <- lapply(c(0.18, 0.27, 0.35), function(i) test_perturb(mixed, mixed_target, mixed_design, NULL, a = i, trans = "none"))
  
  es_de_asin <- lapply(c(0.18, 0.27, 0.35), function(i) test_perturb(mixed, mixed_target, mixed_design, NULL, a = i, trans = "asin-sqrt"))
  
  es_de_logit <- lapply(c(0.18, 0.27, 0.35), function(i) test_perturb(mixed, mixed_target, mixed_design, NULL, a = i, trans = "logit"))

  es_de_df <- rbind(do.call(rbind, es_de_none), do.call(rbind, es_de_asin), do.call(rbind, es_de_logit))
  
  mean_prop <- rowMeans(mixed@assays@data$proportion)
  
  es_de_df <- es_de_df %>% 
    mutate(raw_Amean = rep(mean_prop, 9)) %>% 
    mutate(Signif. = ifelse(P.Value<0.05, "signif.", "n.s.")) %>% 
    mutate(Signif. = factor(Signif., levels = c("signif.", "n.s.")))
  
  es_de_df$filter_level = filter_level
  es_de_df$N_BC <- filter_to_N[filter_level %>% as.character()]
  
  return(es_de_df)
}


```

```{r}
test_all_levels <- lapply(
  c(1:9/10, 0.95),
  function(i) {get_mixted_by_filter(barbieQ = mixture_plus1, filter_level = i)}
)

test_all_levels <- do.call(rbind, test_all_levels)


test_TPR <- test_all_levels %>% 
  group_by(Method, Contrast, N_BC) %>% 
  summarize(Frac_signif = sum(P.Value < 0.05) / n(), .groups = "drop")

```

```{r}
ggplot(test_TPR, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(N_BC, Method))) +
  geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05",
      title = "True positive rate as a function to total barcode number",
      subtitle = "Each box contains three points for different Contrasts"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~Method, ncol =3) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))

ggplot(test_TPR, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, linetype = Contrast, shape = Contrast)) +
  # geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
  geom_line(aes(group = paste0(Contrast, Method))) +
  geom_point() +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05",
      title = "True positive rate as a function to total barcode number"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))  -> tdr_new

```

```{r}
plot_tpr_on_quantile <- function(test_all_levels, num_quantiles = 10) {
  
  quantile_breaks <- quantile(test_all_levels$raw_Amean, probs = seq(0, 1, length.out = num_quantiles+1)) 

  bin_labels <- mapply(
  function(lo, hi) bquote("("*.(format_math(lo))*","*.(format_math(hi))*"]"),
  head(quantile_breaks, -1),
  tail(quantile_breaks, -1),
  SIMPLIFY = FALSE
)

  ## specify label here
  stats_binned <- test_all_levels %>%
    mutate(bin = cut(raw_Amean, breaks = quantile_breaks, labels = bin_labels, include.lowest = TRUE, right = TRUE))

  FPR_lines <- stats_binned %>%
    group_by(bin, Method, Contrast) %>%
    summarize(Frac_signif = sum(P.Value < 0.05) / n(), .groups = "drop")
  
  p <- ggplot(FPR_lines, aes(x = bin, y = Frac_signif, color = Method, linetype = Contrast, shape = Contrast)) +
    # geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8), outliers = T) +
    geom_line(aes(group = paste0(Contrast, Method))) +
    geom_point() +
    labs(
      x = "Quantiles of barcode mean proportion (raw)",
      y = "Fraction of P-Value < 0.05", color = "Method", shape = "Contrast", linetype = "Contrast"
      # caption = paste("Quantile mapping:\n", legend_text)
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    theme_linedraw() +
    scale_x_discrete(labels = scales::label_parse()) +
    theme(axis.text.x = element_text(angle = 75, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12))

  return(p)
}

```

```{r}
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.1), num_quantiles = 4)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.2), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.3), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.4), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.5), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.6), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.7), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.8), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.9), num_quantiles = 10)
plot_tpr_on_quantile(test_all_levels %>% filter(filter_level==0.95), num_quantiles = 10)
```


```{r}
performance <- left_join(
  FPR_box %>% 
  group_by(N_BC, Method) %>% 
  summarise(FPR = mean(Frac_signif)),
  
  test_TPR %>% 
  group_by(N_BC, Method) %>% 
  summarise(TPR = mean(Frac_signif))
)

ggplot(performance, aes(y = FPR, x = TPR, colour = Method)) +
  geom_point(aes(size = N_BC)) +
  geom_line()
```

## Raw proportion

```{r}

test_all_levels_raw <- cbind(
  test_all_levels %>% filter(filter_level == 0.95),  
  mean_prop,
  BC_set_by_filter[rep(seq_len(nrow(BC_set_by_filter)),length.out = nrow(test_all_levels %>% filter(filter_level == 0.95))), ])

test_TPR_raw <- test_all_levels_raw %>%
   select(-N_BC) %>% 
   pivot_longer(
     cols = as.character(filter_to_N[c(1:9/10, 0.95) %>% as.character()]),
     names_to = "N_BC",
     values_to = "selectedBy"
   ) %>% 
    group_by(Method, Contrast, N_BC) %>% 
    summarize(
      N_signif = sum(P.Value < 0.05 & selectedBy), 
      Frac_signif = sum(P.Value < 0.05 & selectedBy) / sum(selectedBy), .groups = "drop")

test_TPR_raw$N_BC <- as.numeric(test_TPR_raw$N_BC)

ggplot(test_TPR_raw, aes(x = as.factor(N_BC), y = Frac_signif, color = Method, group = paste0(Method, Contrast))) +
  # geom_boxplot(alpha = 0.6, position = position_dodge(), outliers = T) +
  geom_point(aes(shape = Contrast)) +
  geom_line(aes(linetype = Contrast)) +
    labs(
      x = paste0("Number of total barcodes after filtering"),
      y = "Fraction of P-Value < 0.05",
      title = "True positive rate as a function to total barcode number",
      subtitle = "Each box contains three points for different Contrasts"
      ) +
    geom_hline(yintercept = 0.05, linetype = "dashed") +
    # facet_wrap(~Contrast, ncol =3) +
    theme_linedraw()+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size = 10, hjust = 0.5), aspect.ratio = 0.5) + 
    theme(
        legend.title = element_text(size = 12), legend.text = element_text(size = 11),
        axis.title = element_text(size = 13), axis.text = element_text(size = 12)) -> tdr_old

```

```{r, fig.width=12, fig.height=6}
(fdr_old + fdr_new) / (tdr_old + tdr_new)
```

# Distribution of BC mean prop

## Update

```{r}
## get the subset of barcodes
lapply(c(1:9/10, 0.95), function(filter_level) {
  barbieQ <- tagTopBarcodes(mixture, nSampleThreshold = 8, proportionThreshold = filter_level)
  tag <- rowData(barbieQ)$isTopBarcode$isTop
  
  mean_prop <- mixture_plus1[tag,mixture_plus1$sampleMetadata$Subset=="null"]@assays@data$proportion %>% rowMeans()
  df <- data.frame(mean_prop, N_BC=filter_to_N[as.character(filter_level)])
  return(df)
}) -> BC_set_prop_filter

BC_set_prop_filter <- do.call(rbind, BC_set_prop_filter)

ggplot(BC_set_prop_filter) +
  geom_freqpoly(aes(x = mean_prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10() +
  theme_linedraw() -> freq_new

ggplot(BC_set_prop_filter) +
  geom_density(aes(x = mean_prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10() +
  theme_linedraw() -> dst_new

dst_new
freq_new
```


## Raw

```{r}
# BC_set_by_filter
# mean_prop

mean_prop_set_filter <- data.frame(
  mean_prop,
  BC_set_by_filter
) 
colnames(mean_prop_set_filter) <- c("mean_prop", colnames(BC_set_by_filter))
mean_prop_set_filter %>% 
  pivot_longer(
    cols = as.character(filter_to_N[as.character(c(1:9/10, 0.95))]), 
    names_to = "N_BC", values_to = "selectedBy") %>% 
  mutate(N_BC = as.numeric(N_BC)) %>% 
  filter(selectedBy) -> mean_long


ggplot(mean_long) +
  geom_freqpoly(aes(x = mean_prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10() +
  theme_linedraw() -> freq_old


ggplot(mean_long) +
  geom_density(aes(x = mean_prop, color = as.factor(N_BC), group = as.factor(N_BC))) +
  scale_x_log10() +
  theme_linedraw() -> dst_old
```

```{r}
(dst_old + dst_new) / (freq_old + freq_new)
```


# Mean-Variance between barcodes

"mean" reflects the total number of barcodes; "variance" reflects the variability between barcodes - some dominant, some minor - even between top barcodes. 

 - this sounds like the distribution is "memoryless", which is the key property of exponential distribution.
 
 - If barcode proportions within a sample follow exponential distribution, then mean should = sqrt(variance).
 
```{r}
get_MA_by_filter <- function(barbieQ, barbieQ_plus1, filter_level, logCount=FALSE, logProp=FALSE) {
  ## tag top BC from raw barbieQ
  barbieQ <- tagTopBarcodes(barbieQ, nSampleThreshold = 8, proportionThreshold = filter_level)
  ## select top BC from barbieQ_plus1
  barbieQ_plus1_top <- barbieQ_plus1[rowData(barbieQ)$isTopBarcode$isTop,]
  ## subset baseline "null" samples and perturbed samples
  mixed <- barbieQ_plus1_top[, barbieQ_plus1_top$sampleMetadata$Subset %in% c("null", "perturb")]
  
  ## re-calculate proportions
  mixed@assays@data$proportion <- ((mixed@assays@data$proportion %>% t()) / colSums(mixed@assays@data$proportion)) %>% t()
  
  mat <- mixed@assays@data$proportion
  
  if(logCount) {
    mat <- log(assay(mixed)+1)
  } else if(logProp){
    mat <- log(mat + 1e-10)
  }
  
  mean_var_df <- data.frame(
    mean = apply(mat, 2, mean, na.rm = TRUE),
    variance = apply(mat, 2, var, na.rm = TRUE)
  )
  mean_var_df$filter_level <- filter_level
  mean_var_df$N_BC <- filter_to_N[filter_level %>% as.character()]
  mean_var_df$Sample <- colnames(mixed)
  
  return(mean_var_df)
}
```

```{r}
## proportion MA at each filtering level
MA_by_filter_all <- lapply(c(1:9/10, 0.95), function(i) {
  get_MA_by_filter(mixture, mixture_plus1, i)
})

MA_by_filter_all <- do.call(rbind, MA_by_filter_all)

ggplot(MA_by_filter_all, aes(x = mean, y = sqrt(variance), group = N_BC)) +
  geom_boxplot() +
  # theme(aspect.ratio = 1)
  coord_fixed() +
  labs(title = "Mean-Vartiance between barcodes within a sample")

ggplot(MA_by_filter_all, aes(x = mean, y = (sqrt(variance)/mean), group = N_BC)) +
  geom_boxplot() +
  # theme(aspect.ratio = 1)
  # coord_fixed() +
  labs(title = "Mean-Vartiance between barcodes within a sample")

ggplot(MA_by_filter_all, aes(x = mean, y = sqrt(variance), group = N_BC)) +
  geom_boxplot() +
  scale_x_log10() +
  scale_y_log10() +
  # theme(aspect.ratio = 1)
  coord_fixed() +
  labs(title = "Mean-Vartiance between barcodes within a sample",
       subtitle = "both x and y axis log-scaled")
```

```{r}
## log count MA at each filtering level
MA_by_filter_all_logCount <- lapply(c(1:9/10, 0.95), function(i) {
  get_MA_by_filter(mixture, mixture_plus1, i, logCount = T)
})

MA_by_filter_all_logCount <- do.call(rbind, MA_by_filter_all_logCount)

ggplot(MA_by_filter_all_logCount, aes(x = mean, y = sqrt(variance), group = N_BC)) +
  geom_boxplot() +
  # theme(aspect.ratio = 1)
  coord_fixed() +
  labs(title = "Mean-Vartiance between log barcode count within a sample")

ggplot(MA_by_filter_all_logCount, aes(x = mean, y = sqrt(variance), group = N_BC)) +
  geom_boxplot() +
  scale_x_log10() +
  scale_y_log10() +
  # theme(aspect.ratio = 1)
  labs(title = "Mean-Vartiance between log barcode count within a sample",
       subtitle = "y axis log-scaled")
```


QQ plopt visual check

```{r}
plot_QQ_column <- function(x, sample_id) {
  # x <- your_data_vector

  # Estimate the rate parameter λ (lambda = 1 / mean)
  lambda_hat <- 1 / mean(x)
  
  # Compute theoretical vs observed quantiles
  n <- length(x)
  theoretical_q <- qexp(ppoints(n), rate = lambda_hat)
  observed_q <- sort(x)
  
  df <- data.frame(theoretical = theoretical_q, observed = observed_q)
  
  # Create ggplot QQ plot
  p <- ggplot(df, aes(x = theoretical, y = observed)) +
    geom_point(color = "steelblue", size = 1.5, alpha = 0.7) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
    labs(
      x = "Theoretical Quantiles (Exponential)",
      y = "Sample Quantiles",
      subtitle = sample_id
    ) +
    theme_minimal(base_size = 12)
  
  return(p)
}


plot_QQ_by_filter <- function(barbieQ, barbieQ_plus1, filter_level) {
  ## tag top BC from raw barbieQ
  barbieQ <- tagTopBarcodes(barbieQ, nSampleThreshold = 8, proportionThreshold = filter_level)
  ## select top BC from barbieQ_plus1
  barbieQ_plus1_top <- barbieQ_plus1[rowData(barbieQ)$isTopBarcode$isTop,]
  ## subset baseline "null" samples and perturbed samples
  mixed <- barbieQ_plus1_top[, barbieQ_plus1_top$sampleMetadata$Subset %in% c("null", "perturb")]
  
  ## re-calculate proportions
  mixed@assays@data$proportion <- ((mixed@assays@data$proportion %>% t()) / colSums(mixed@assays@data$proportion)) %>% t()
  
  mat <- mixed@assays@data$proportion
  plot_list <- lapply(colnames(mat), function(x) plot_QQ_column(mat[,x], x))
  plot_list[[1]] <- plot_list[[1]] + 
    labs(title = paste0("filter_level=", filter_level, ", N_barcode=", filter_to_N[filter_level %>% as.character()]))
  pp <- wrap_plots(plot_list, ncol=6)
  
  return(pp)
}


```

```{r, fig.width=20, fig.height=20}
qq_filter_list <- lapply(
  c(1:9/10, 0.95), 
  function(i) plot_QQ_by_filter(mixture, mixture_plus1, i))


(qq_filter_list)
```

