---
title: "mouseHSPC_preprocessing"
author: "LF"
date: "2025-04-03"
output: workflowr::wflow_html
output_dir: ../../docs/BelderbosME
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)

library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

```

# Install barbieQ - devel

```{r install barbieQ}
devtools::install_github("Oshlack/barbieQ")

library(barbieQ)
```

# Read data

Read every sheet in the excel.

8 mice: "C1"  "C2"  "C3"  "C10" "C15" "C22" "C21" "C23"

overlapping barcodes: hundreds

mouse-specific barcodes: 1-2k.

I think 'library' in columns refers to time cell pool.

```{r read xl}
sheets <- readxl::excel_sheets("../../data/BelderbosME/mmc2.xlsx")

data_list <- purrr::map(sheets, ~readxl::read_excel("../../data/BelderbosME/mmc2.xlsx", sheet = .x))

# bind sheets
# library(dplyr)
# combined_data <- bind_rows(data_list, .id = "sheet_name")

names(data_list) <- sheets

C1_bar <- data_list$C1$ID
C2_bar <- data_list$C2$`['ID'`
C3_bar <- data_list$C3$`['ID'`
C10_bar <- data_list$C10$ID
C15_bar <- data_list$C15$...1
C22_bar <- data_list$C22$ID
C21_bar <- data_list$C21$ID
C23_bar <- data_list$C23$ID

bar_list <- list(
  C1 = C1_bar,
  C2 = C2_bar,
  C3 = C3_bar,
  C10 = C10_bar,
  C15 = C15_bar,
  C22 = C22_bar,
  C21 = C21_bar,
  C23 = C23_bar
)

ggVennDiagram::ggVennDiagram(x = bar_list)

```

# Clean up data

## C1

```{r clean up C1}
class(data_list$C1)

data_list$C1
data_list$C1 <- as.data.frame(data_list$C1)
class(data_list$C1)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C1[, 1]))
data_list$C1 <- data_list$C1[!is.na(data_list$C1[, 1]), ]
dim(data_list$C1)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C1) <- data_list$C1[,1]
data_list$C1 <- data_list$C1[,-1]

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C1 <- data_list$C1[, !colnames(data_list$C1) %in% del]

# delete processed "sum" columns
data_list$C1 <- data_list$C1[, !grepl("_Sum", colnames(data_list$C1))] 
data_list$C1 <- data_list$C1[, !grepl("_sum", colnames(data_list$C1))] 

# delete processed "shannon" rows
rownames(data_list$C1) [!grepl("^[ATGC]+$", rownames(data_list$C1))]
rownames(data_list$C1) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C1))]
data_list$C1 <- data_list$C1[grepl("^[ATGC\\\"]+$", rownames(data_list$C1)),]

dim(data_list$C1)

# Convert NA values to 0 in the matrix
data_list$C1[is.na(data_list$C1)] <- 0
!any(is.na(data_list$C1))

# Convert character data to numeric
C1_count_numeric <- apply(data_list$C1, 2, as.numeric)
rownames(C1_count_numeric) <- rownames(data_list$C1)

data_list$C1 <- C1_count_numeric

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C1))
xCondition <- gsub("(.*)\\_.*", "\\1", xCondition)
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk[0-9]{1,}).*", "\\1", colnames(data_list$C1))
Time[grepl("\\_", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub(".*\\_(BM)\\_.*", "\\1", colnames(data_list$C1))
Tissue <- gsub(".*\\_(Blood)\\_.*", "\\1", Tissue)
Tissue[grepl("\\_", Tissue)] <- "Blood"
Tissue

# extract celltype from colnames
Celltype <- gsub(".*\\_.*\\_([A-Z])", "\\1", colnames(data_list$C1))
Celltype[grepl("\\_", Celltype)] <- "Unsorted"
Celltype[grepl("T", Celltype)] <- "Tcell"
Celltype[grepl("B", Celltype)] <- "Bcell"
Celltype[grepl("G", Celltype)] <- "Grn"


colnames(data_list$C1)

C1_meta <- data.frame(
  Sample = colnames(data_list$C1),
  Mouse = rep("C1", ncol(data_list$C1)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Celltype = Celltype
)

View(data_list$C1)

```

## C2

```{r clean up C2}
data_list$C2 %>% View
data_list$C2 <- as.data.frame(data_list$C2)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C2[, 1]))
data_list$C2 <- data_list$C2[!is.na(data_list$C2[, 1]), ]
dim(data_list$C2)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C2) <- data_list$C2[,1]
data_list$C2 <- data_list$C2[,-1]

colnames(data_list$C2)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C2 <- data_list$C2[, !colnames(data_list$C2) %in% del]

# delete processed "sum" columns
data_list$C2 <- data_list$C2[, !grepl("Sum", colnames(data_list$C2))] 
data_list$C2 <- data_list$C2[, !grepl("sum", colnames(data_list$C2))] 

# delete processed "shannon" rows
rownames(data_list$C2) [!grepl("^[ATGC]+$", rownames(data_list$C2))]
rownames(data_list$C2) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C2))]
data_list$C2 <- data_list$C2[grepl("^[ATGC\\\"]+$", rownames(data_list$C2)),]

dim(data_list$C2)

# Convert NA values to 0 in the matrix
data_list$C2[is.na(data_list$C2)] <- 0
!any(is.na(data_list$C2))

# Convert character data to numeric
C2_count_numeric <- apply(data_list$C2, 2, as.numeric)
rownames(C2_count_numeric) <- rownames(data_list$C2)

data_list$C2 <- C2_count_numeric

# View sample names
colnames(data_list$C2)

# fix "library" clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C2) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C2))

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C2))
xCondition <- gsub("(.*)\\_.*", "\\1", xCondition)
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk[0-9]{1,}).*", "\\1", colnames(data_list$C2))
Time[grepl("\\_", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub(".*\\_(BM).*", "\\1", colnames(data_list$C2))
Tissue <- gsub(".*\\_(Liver).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Spleen).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Blood).*", "\\1", Tissue)
Tissue[grepl("\\_", Tissue)] <- "Blood"
Tissue

# extract position, like BMLeft, BMSpine
Position <- vector("character", length(Tissue))
Position[grepl("Left", colnames(data_list$C2))] <- "Left"
Position[grepl("Front", colnames(data_list$C2))] <- "Front"
Position[grepl("Right", colnames(data_list$C2))] <- "Right"
Position[grepl("Spine", colnames(data_list$C2))] <- "Spine"

# extract celltype from colnames
Celltype <- gsub(".*\\_.*\\_([A-Z])", "\\1", colnames(data_list$C2))
Celltype[grepl("Tcell", Celltype)] <- "Tcell"
Celltype[grepl("Bcell", Celltype)] <- "Bcell"
Celltype[grepl("Grn", Celltype)] <- "Grn"
Celltype[grepl("Other", Celltype)] <- "Other"
Celltype[grepl("\\_", Celltype)] <- "Unsorted"
Celltype

colnames(data_list$C2)

C2_meta <- data.frame(
  Sample = colnames(data_list$C2),
  Mouse = rep("C2", ncol(data_list$C2)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

View(data_list$C2)
```

## C3

There are two GK_BM_T. One might be typo, originally GK_BM_G.

```{r clean up C3}
data_list$C3 %>% View
data_list$C3 <- as.data.frame(data_list$C3)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C3[, 1]))
data_list$C3 <- data_list$C3[!is.na(data_list$C3[, 1]), ]
dim(data_list$C3)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C3) <- data_list$C3[,1]
data_list$C3 <- data_list$C3[,-1]

colnames(data_list$C3)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C3 <- data_list$C3[, !colnames(data_list$C3) %in% del]

# delete processed "sum" columns
data_list$C3 <- data_list$C3[, !grepl("Sum", colnames(data_list$C3))] 
data_list$C3 <- data_list$C3[, !grepl("sum", colnames(data_list$C3))] 

# delete processed "shannon" rows
rownames(data_list$C3) [!grepl("^[ATGC]+$", rownames(data_list$C3))]
rownames(data_list$C3) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C3))]
data_list$C3 <- data_list$C3[grepl("^[ATGC\\\"]+$", rownames(data_list$C3)),]

dim(data_list$C3)

# Convert NA values to 0 in the matrix
data_list$C3[is.na(data_list$C3)] <- 0
!any(is.na(data_list$C3))

# Convert character data to numeric
C3_count_numeric <- apply(data_list$C3, 2, as.numeric)
rownames(C3_count_numeric) <- rownames(data_list$C3)

data_list$C3 <- C3_count_numeric

# View sample names
colnames(data_list$C3)

# fix "library" clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C3) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C3))

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C3))
xCondition <- gsub("(.*)\\_.*", "\\1", xCondition)
xCondition[grepl("TK3BMCellsbeforetransplant", xCondition)] <- "time0"
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk[0-9]{1,}).*", "\\1", colnames(data_list$C3))
Time[grepl("\\_", Time)] <- "sac"
Time[grepl("TK3BMCellsbeforetransplant", Time)] <- "time0"
Time

# extract tissue from colnames
Tissue <- gsub(".*\\_(BM).*", "\\1", colnames(data_list$C3))
Tissue <- gsub(".*\\_(Liver).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Spleen).*", "\\1", Tissue)
Tissue <- gsub(".*\\_(Blood).*", "\\1", Tissue)
Tissue[grepl("\\_", Tissue)] <- "Blood"
Tissue[grepl("TK3BMCellsbeforetransplant", Tissue)] <- "time0"
Tissue

# extract celltype from colnames
Celltype <- gsub(".*\\_.*\\_([A-Z])", "\\1", colnames(data_list$C3))
Celltype[grepl("TK3BMCellsbeforetransplant", Celltype)] <- "time0"
Celltype[grepl("T", Celltype)] <- "Tcell"
Celltype[grepl("B", Celltype)] <- "Bcell"
Celltype[grepl("G", Celltype)] <- "Grn"
Celltype[grepl("\\_", Celltype)] <- "Unsorted"
Celltype

colnames(data_list$C3)

C3_meta <- data.frame(
  Sample = colnames(data_list$C3),
  Mouse = rep("C3", ncol(data_list$C3)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Celltype = Celltype
)

View(data_list$C3)
```

## C10

```{r clean up C10}
data_list$C10 %>% View
data_list$C10 <- as.data.frame(data_list$C10)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C10[, 1]))
data_list$C10 <- data_list$C10[!is.na(data_list$C10[, 1]), ]
dim(data_list$C10)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C10) <- data_list$C10[,1]
data_list$C10 <- data_list$C10[,-1]

colnames(data_list$C10)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C10 <- data_list$C10[, !colnames(data_list$C10) %in% del]

# delete processed "sum" columns
data_list$C10 <- data_list$C10[, !grepl("Sum", colnames(data_list$C10))] 
data_list$C10 <- data_list$C10[, !grepl("sum", colnames(data_list$C10))] 

# delete processed "shannon" rows
rownames(data_list$C10) [!grepl("^[ATGC]+$", rownames(data_list$C10))]
rownames(data_list$C10) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C10))]
data_list$C10 <- data_list$C10[grepl("^[ATGC\\\"]+$", rownames(data_list$C10)),]

dim(data_list$C10)

# Convert NA values to 0 in the matrix
data_list$C10[is.na(data_list$C10)] <- 0
!any(is.na(data_list$C10))

# Convert character data to numeric
C10_count_numeric <- apply(data_list$C10, 2, as.numeric)
rownames(C10_count_numeric) <- rownames(data_list$C10)

data_list$C10 <- C10_count_numeric

# View sample names
colnames(data_list$C10)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C10) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C10))

# extract condition from colnames. I don;t know what GK / L means.
xCondition <- gsub(".*\\_.*\\_.*\\_(.*)", "\\1", colnames(data_list$C10))
xCondition

# extract time from colnames
Time <- gsub(".*\\_(wk\\_[0-9]{1,}).*", "\\1", colnames(data_list$C10))
Time <- gsub("\\_", "", Time)
Time

# extract tissue from colnames
Tissue <- gsub("(.*)\\_.*\\_.*\\_.*", "\\1", colnames(data_list$C10))
Tissue[grepl("Blood", Tissue)] <- "Blood"
Tissue[grepl("BM", Tissue)] <- "BM"
Tissue[grepl("Liver", Tissue)] <- "Liver"
Tissue[grepl("Spleen", Tissue)] <- "Spleen"
Tissue

# extract position
mix <- gsub("(.*)\\_.*\\_.*\\_.*", "\\1", colnames(data_list$C10))
Position <- vector("character", length(mix))
Position[grepl("BM", mix)] <- gsub("(.*)BM.*", "\\1", mix)[grepl("BM", mix)]
Position

# extract celltype from colnames
mix <- gsub("(.*)\\_.*\\_.*\\_.*", "\\1", colnames(data_list$C10))
Celltype <- vector("character", length(mix))
Celltype[grepl("Blood", mix)] <- gsub(".*Blood(.*)", "\\1", mix)[grepl("Blood", mix)]
Celltype[grepl("BM", mix)] <- gsub(".*BM(.*)", "\\1", mix)[grepl("BM", mix)]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype

colnames(data_list$C10)

C10_meta <- data.frame(
  Sample = colnames(data_list$C10),
  Mouse = rep("C10", ncol(data_list$C10)),
  xCondition = xCondition,
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

View(data_list$C10)
```

## C15

```{r clean up C15}
data_list$C15 %>% View
data_list$C15 <- as.data.frame(data_list$C15)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C15[, 1]))
data_list$C15 <- data_list$C15[!is.na(data_list$C15[, 1]), ]
dim(data_list$C15)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C15) <- data_list$C15[,1]
data_list$C15 <- data_list$C15[,-1]

colnames(data_list$C15)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C15 <- data_list$C15[, !colnames(data_list$C15) %in% del]

# delete processed "sum" columns
data_list$C15 <- data_list$C15[, !grepl("Sum", colnames(data_list$C15))] 
data_list$C15 <- data_list$C15[, !grepl("sum", colnames(data_list$C15))] 

# delete processed "shannon" rows
rownames(data_list$C15) [!grepl("^[ATGC]+$", rownames(data_list$C15))]
rownames(data_list$C15) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C15))]
data_list$C15 <- data_list$C15[grepl("^[ATGC\\\"]+$", rownames(data_list$C15)),]

dim(data_list$C15)

# Convert NA values to 0 in the matrix
data_list$C15[is.na(data_list$C15)] <- 0
!any(is.na(data_list$C15))

# Convert character data to numeric
C15_count_numeric <- apply(data_list$C15, 2, as.numeric)
rownames(C15_count_numeric) <- rownames(data_list$C15)

data_list$C15 <- C15_count_numeric

# View sample names
colnames(data_list$C15)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C15) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C15))

# extract time from colnames
Time <- gsub(".*(wk[0-9]{1,}).*", "\\1", colnames(data_list$C15))
Time[grepl("\\_", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- vector("character", ncol(data_list$C15))
Tissue[grepl("BM", colnames(data_list$C15))] <- "BM"
Tissue[Tissue == ""] <- "Blood"
Tissue

# extract position
mix <- gsub("(.*)\\_.*", "\\1", colnames(data_list$C15))
Position <- vector("character", ncol(data_list$C15))
Position[grepl("BM", mix)] <- gsub("BM(.*)", "\\1", mix)[grepl("BM", mix)]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C15))
Celltype[grepl("\\_",colnames(data_list$C15))] <- gsub(".*\\_(.*)", "\\1", colnames(data_list$C15))[grepl("\\_",colnames(data_list$C15))]

Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C15)

C15_meta <- data.frame(
  Sample = colnames(data_list$C15),
  Mouse = rep("C15", ncol(data_list$C15)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

View(data_list$C15)
```

## C22

```{r clean up C22}
data_list$C22 %>% View
data_list$C22 <- as.data.frame(data_list$C22)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C22[, 1]))
data_list$C22 <- data_list$C22[!is.na(data_list$C22[, 1]), ]
dim(data_list$C22)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C22) <- data_list$C22[,1]
data_list$C22 <- data_list$C22[,-1]

colnames(data_list$C22)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C22 <- data_list$C22[, !colnames(data_list$C22) %in% del]

# delete processed "sum" columns
data_list$C22 <- data_list$C22[, !grepl("Sum", colnames(data_list$C22))] 
data_list$C22 <- data_list$C22[, !grepl("sum", colnames(data_list$C22))] 

# delete processed "shannon" rows
rownames(data_list$C22) [!grepl("^[ATGC]+$", rownames(data_list$C22))]
rownames(data_list$C22) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C22))]
data_list$C22 <- data_list$C22[grepl("^[ATGC\\\"]+$", rownames(data_list$C22)),]

dim(data_list$C22)

# Convert NA values to 0 in the matrix
data_list$C22[is.na(data_list$C22)] <- 0
!any(is.na(data_list$C22))

# Convert character data to numeric
C22_count_numeric <- apply(data_list$C22, 2, as.numeric)
rownames(C22_count_numeric) <- rownames(data_list$C22)

data_list$C22 <- C22_count_numeric

# View sample names
colnames(data_list$C22)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C22) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C22))

# extract time from colnames
Time <- gsub(".*\\_(wk\\_[0-9]{1,}).*", "\\1", colnames(data_list$C22))
Time[grepl("\\_", Time)] <- "sac"
Time[Time == "Spleen"] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub("(.*)\\_.*\\_.*", "\\1", colnames(data_list$C22))
Tissue <- gsub("(.*)\\_.*", "\\1", Tissue)
Tissue[grepl("wk", Tissue)] <- "Blood"
Tissue

# extract position
Position <- vector("character", ncol(data_list$C22))
Position[grepl("BM", colnames(data_list$C22))] <- gsub("BM\\_(.*)\\_.*", "\\1", colnames(data_list$C22))[grepl("BM", colnames(data_list$C22))]
Position[grepl("BM", Position)] <- gsub("BM\\_(.*)", "\\1", Position)[grepl("BM", Position)]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C22))
Celltype[grepl("BM\\_.*\\_.*", colnames(data_list$C22))] <- gsub("BM\\_.*\\_(.*)", "\\1", colnames(data_list$C22))[grepl("BM\\_.*\\_.*", colnames(data_list$C22))]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "unsorted"] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C22)

C22_meta <- data.frame(
  Sample = colnames(data_list$C22),
  Mouse = rep("C22", ncol(data_list$C22)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

View(data_list$C22)
```

## C21

```{r clean up C21}
data_list$C21 %>% View
data_list$C21 <- as.data.frame(data_list$C21)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C21[, 1]))
data_list$C21 <- data_list$C21[!is.na(data_list$C21[, 1]), ]
dim(data_list$C21)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C21) <- data_list$C21[,1]
data_list$C21 <- data_list$C21[,-1]

colnames(data_list$C21)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C21 <- data_list$C21[, !colnames(data_list$C21) %in% del]

# delete processed "sum" columns
data_list$C21 <- data_list$C21[, !grepl("Sum", colnames(data_list$C21))] 
data_list$C21 <- data_list$C21[, !grepl("sum", colnames(data_list$C21))] 

# delete processed "shannon" rows
rownames(data_list$C21) [!grepl("^[ATGC]+$", rownames(data_list$C21))]
rownames(data_list$C21) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C21))]
data_list$C21 <- data_list$C21[grepl("^[ATGC\\\"]+$", rownames(data_list$C21)),]

dim(data_list$C21)

# Convert NA values to 0 in the matrix
data_list$C21[is.na(data_list$C21)] <- 0
!any(is.na(data_list$C21))

# Convert character data to numeric
C21_count_numeric <- apply(data_list$C21, 2, as.numeric)
rownames(C21_count_numeric) <- rownames(data_list$C21)

data_list$C21 <- C21_count_numeric

# View sample names
colnames(data_list$C21)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C21) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C21))

# extract time from colnames
Time <- gsub("(wk[0-9]{1,}).*", "\\1", colnames(data_list$C21))
Time[grepl("BM|Spleen|Liver", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub("(BM|Spleen|Liver).*", "\\1", colnames(data_list$C21))
Tissue[grepl("wk", Tissue)] <- "Blood"
Tissue

# extract position
Position <- vector("character", ncol(data_list$C21))
Position[grepl("BM", colnames(data_list$C21))] <- gsub("BM(.*)\\_.*", "\\1", colnames(data_list$C21))[grepl("BM", colnames(data_list$C21))]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C21))
Celltype[grepl("\\_", colnames(data_list$C21))] <- gsub(".*\\_(.*)", "\\1", colnames(data_list$C21))[grepl("\\_", colnames(data_list$C21))]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "unsorted"] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C21)

C21_meta <- data.frame(
  Sample = colnames(data_list$C21),
  Mouse = rep("C21", ncol(data_list$C21)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

View(data_list$C21)
```

## C23

```{r clean up C23}
data_list$C23 %>% View
data_list$C23 <- as.data.frame(data_list$C23)

# Subset the dataframe to exclude rows where the first column (barcodeID) is NA
any(is.na(data_list$C23[, 1]))
data_list$C23 <- data_list$C23[!is.na(data_list$C23[, 1]), ]
dim(data_list$C23)

# set up the first column of each mat as rownames (barcode ID)
rownames(data_list$C23) <- data_list$C23[,1]
data_list$C23 <- data_list$C23[,-1]

colnames(data_list$C23)

# delete empty or processed columns
del <- c("...39", "...40", "Present", "Total", "Log_total")
data_list$C23 <- data_list$C23[, !colnames(data_list$C23) %in% del]

# delete processed "sum" columns
data_list$C23 <- data_list$C23[, !grepl("Sum", colnames(data_list$C23))] 
data_list$C23 <- data_list$C23[, !grepl("sum", colnames(data_list$C23))] 

# delete processed "shannon" rows
rownames(data_list$C23) [!grepl("^[ATGC]+$", rownames(data_list$C23))]
rownames(data_list$C23) [!grepl("^[ATGC\\\"]+$", rownames(data_list$C23))]
data_list$C23 <- data_list$C23[grepl("^[ATGC\\\"]+$", rownames(data_list$C23)),]

dim(data_list$C23)

# Convert NA values to 0 in the matrix
data_list$C23[is.na(data_list$C23)] <- 0
!any(is.na(data_list$C23))

# Convert character data to numeric
C23_count_numeric <- apply(data_list$C23, 2, as.numeric)
rownames(C23_count_numeric) <- rownames(data_list$C23)

data_list$C23 <- C23_count_numeric

# View sample names
colnames(data_list$C23)

# clean any character that is not a letter, digit, or underscore with an empty string, effectively removing unwanted symbols.
colnames(data_list$C23) <- gsub("[^a-zA-Z0-9_]", "", colnames(data_list$C23))

# extract time from colnames
Time <- gsub("(wk[0-9]{1,}).*", "\\1", colnames(data_list$C23))
Time[grepl("BM|Spleen|Liver|Sac", Time)] <- "sac"
Time

# extract tissue from colnames
Tissue <- gsub("(BM|Spleen|Liver).*", "\\1", colnames(data_list$C23))
Tissue[grepl("wk", Tissue)] <- "Blood"
Tissue[grepl("Sac", Tissue)] <- "sac"
Tissue

# extract position
Position <- vector("character", ncol(data_list$C23))
Position[grepl("BM\\_.*\\_.*", colnames(data_list$C23))] <- gsub("BM\\_(.*)\\_.*", "\\1", colnames(data_list$C23))[grepl("BM\\_.*\\_.*", colnames(data_list$C23))]
Position

# extract celltype from colnames
Celltype <- vector("character", ncol(data_list$C23))
Celltype[grepl("\\_", colnames(data_list$C23))] <- gsub(".*\\_(.*)", "\\1", colnames(data_list$C23))[grepl("\\_", colnames(data_list$C23))]
Celltype[Celltype == ""] <- "Unsorted"
Celltype[Celltype == "unsorted"] <- "Unsorted"
Celltype[Celltype == "B"] <- "Bcell"
Celltype[Celltype == "T"] <- "Tcell"
Celltype[Celltype == "G"] <- "Grn"
Celltype[Celltype == "O"] <- "Other"
Celltype

colnames(data_list$C23)

C23_meta <- data.frame(
  Sample = colnames(data_list$C23),
  Mouse = rep("C23", ncol(data_list$C23)),
  Time = Time,
  Tissue = Tissue,
  Position = Position,
  Celltype = Celltype
)

View(data_list$C23)
```

## Combine mice

```{r combine}
Metadata <- dplyr::bind_rows(C1_meta, C2_meta, C3_meta, C10_meta, C15_meta, C22_meta, C21_meta, C23_meta)
Metadata[is.na(Metadata)] <- ""

data_list <- lapply(data_list, function(mat)  as.data.frame(t(mat)))
CountData <- dplyr::bind_rows(data_list)
CountData <- t(CountData)
```

# Save to barbieQ

```{r save to barbieQ}
mouseHSPC <- createBarbieQ(object = CountData, sampleMetadata = Metadata)
```

# Tag top barcodes

```{r tag top}
dim(mouseHSPC)
## find minimum group size
targets <- as.data.frame(mouseHSPC$sampleMetadata) %>% 
  mutate(Group = paste0(Mouse, Tissue, Celltype))
targets$Group %>% table()

## select sample to consider, exclude TP0 when tagging
mouseHSPC <- barbieQ::tagTopBarcodes(barbieQ = mouseHSPC, 
  activeSamples = mouseHSPC$sampleMetadata$Sample != "library", 
  nSampleThreshold = 2
  )

rowData(mouseHSPC)$isTopBarcode$isTop %>% table()
```

```{r vis tag}
barbieQ::plotBarcodePareto(barbieQ = mouseHSPC)
barbieQ::plotBarcodeSankey(barbieQ = mouseHSPC)
```

# Barcode Correlation

```{r bar cor}
ggsave(
  "../../output/BelderbosME//bar_cor.png",
  plotBarcodePairCorrelation(
    barbieQ = mouseHSPC[rowData(mouseHSPC)$isTopBarcode$isTop,], 
    cpmThresh = 2^11)
  )

mouseHSPC_top_cluster <- clusterCorrelatingBarcodes(
  barbieQ = mouseHSPC[rowData(mouseHSPC)$isTopBarcode$isTop,], 
  cpmThresh = 2^11
  )

## check out clusters
identifiedCluster <- rowData(mouseHSPC_top_cluster)$barcodeCorrelatedCluster %>% as.data.frame()

N.single <- identifiedCluster$cluster %>% min()

## send the results back to unfiltered data
DF <- DataFrame(row.names = rownames(mouseHSPC))
DF[rownames(identifiedCluster),"cluster"] <- identifiedCluster
N.single.add <- is.na(DF$cluster) %>% sum()
DF$cluster[is.na(DF$cluster)] <- -seq_len(N.single.add) + N.single

rowData(mouseHSPC)$barcodeCorrelatedCluster <- DF
```

```{r collapse barcodes}
identifiedCluster <- rowData(mouseHSPC)$barcodeCorrelatedCluster %>% as.data.frame()

mouseHSPC_agg <- SEtools::aggSE(mouseHSPC, by = identifiedCluster$cluster, assayFun = "max")

row_sums <- rowSums(assay(mouseHSPC))

max_rows_per_cluster <- tapply(
  rownames(identifiedCluster),  # Row names
  identifiedCluster$cluster,  # Grouping variable
  function(rows) rows[which.max(row_sums[rows])]  # Get row with max sum per cluster
)

max_rows_df <- data.frame(cluster = names(max_rows_per_cluster), rowname = unlist(max_rows_per_cluster))

max_rows_df <-max_rows_df[rownames(mouseHSPC_agg), ]

rownames(mouseHSPC_agg) <- max_rows_df$rowname

## recreate barbieQ
mouseHSPC_agg <- createBarbieQ(
  object = assay(mouseHSPC_agg), sampleMetadata = mouseHSPC$sampleMetadata)
metadata(mouseHSPC_agg) <- metadata(mouseHSPC)
```

```{r vis cluster}
## extract clusters
identifiedCluster$cluster[identifiedCluster$cluster < 0] <- -1

plot3 <- ggplot(data = identifiedCluster) +
  geom_histogram(data = filter(identifiedCluster, cluster > 0), 
                 aes(x = cluster, fill = as.factor(cluster)), 
                 binwidth = 1, color = "white") +
  theme_classic() +
  scale_x_continuous(breaks = 1:14, labels = 1:14) +
  theme(aspect.ratio = 1, legend.position = "none")

ggsave("../../output/BelderbosME/cluster_hist.png", plot = plot3, width = 3, height = 3, dpi = 300)

identifiedCluster <- identifiedCluster %>% 
  mutate(logMeanCPM = log2(rowMeans(assays(mouseHSPC)$CPM)+1))

plot4 <- ggplot(data = identifiedCluster) +
  geom_point(aes(x = cluster, y = logMeanCPM),
             color = "grey", shape = 1, size = 4) +
  geom_point(data = filter(identifiedCluster, cluster > 0),
             aes(x = cluster, 
                 y = logMeanCPM,
                 color = factor(cluster)),
             shape = 1, size = 4, stroke = 1) +
  scale_x_continuous(breaks = 1:14, labels = 1:14) +
  theme_classic() +
  theme(aspect.ratio = 1, legend.position = "none")

ggsave("../../output/BelderbosME/cluster_mean.png", plot = plot4, width = 3, height = 3, dpi = 300)
```

# Tag top barcode after collapsing

```{r tag after cor}
mouseHSPC_agg <- tagTopBarcodes(
  barbieQ = mouseHSPC_agg, 
  activeSamples = mouseHSPC_agg$sampleMetadata$Sample != "library",
  nSampleThreshold = 2)

rowData(mouseHSPC_agg)$isTopBarcode$isTop %>% table()
(assays(mouseHSPC_agg)$isTopAssay %>% rowSums() >= 2) %>% table()

barbieQ::plotBarcodePareto(barbieQ = mouseHSPC_agg)
barbieQ::plotBarcodeSankey(barbieQ = mouseHSPC_agg)

ggsave("../../output/BelderbosME/pareto.png", 
       plot = barbieQ::plotBarcodePareto(barbieQ = mouseHSPC_agg), 
       width = 10, height = 7, dpi = 300)
ggsave("../../output/BelderbosME/sankey.png", 
       plot = barbieQ::plotBarcodeSankey(barbieQ = mouseHSPC_agg), 
       width = 5, height = 5, dpi = 300)

```

# Heatmap

```{r hp}
# png(filename = "../../output/BelderbosME/HP_unfilter.png", width = 2000, height = 2000, res = 300)
# Heatmap(name = "logCPM",
#   matrix = log2(assays(mouseHSPC_agg)$CPM + 1), 
#   row_title = paste0(nrow(mouseHSPC_agg), " Barcodes"), 
#   show_row_names = FALSE, show_column_names = FALSE)
# dev.off()

mouseHSPC_top <- mouseHSPC_agg[rowData(mouseHSPC_agg)$isTopBarcode$isTop,]

png(filename = "../../output/BelderbosME/HP_filtered.png", width = 2000, height = 2000, res = 300)
Heatmap(name = "logCPM",
  matrix = log2(assays(mouseHSPC_top)$CPM + 1), 
  row_title = paste0(nrow(mouseHSPC_top), " Barcodes"), 
  show_row_names = FALSE, show_column_names = FALSE)
dev.off()
```

# Save rds

```{r save rds}
save(
  mouseHSPC, mouseHSPC_agg, mouseHSPC_top,
  file = "../../output/BelderbosME/mouseHSPC_barbieQ.rda")

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")
```

# Top10

```{r}
tagTop10 <- function(col) {
  ranks <- rank(-col)
  # summary(ranks)
  isTop <- ranks <= 10
  # summary(isTop)
  return(isTop)
}

tester <- tagTop10(assays(mouseHSPC_agg)$CPM[,1])
summary(tester)

colTags <- lapply(as.data.frame(assays(mouseHSPC_agg)$CPM), 
                  function(col) tagTop10(col))
colTagMat <- do.call(cbind, colTags)
colTagVec <- rowSums(colTagMat[,mouseHSPC_agg$sampleMetadata$Sample != "library"]) >= 1
summary(colTagVec)

mouseHSPC_agg_top10 <- mouseHSPC_agg
rowData(mouseHSPC_agg_top10)$isTopBarcode <- DataFrame(isTop = colTagVec)
summary(rowData(mouseHSPC_agg_top10)$isTopBarcode$isTop)



ggsave("../../output/BelderbosME/pareto_top10.png", 
       plot = barbieQ::plotBarcodePareto(barbieQ = mouseHSPC_agg_top10), 
       width = 10, height = 7, dpi = 300)
ggsave("../../output/BelderbosME/sankey_top10.png", 
       plot = barbieQ::plotBarcodeSankey(barbieQ = mouseHSPC_agg_top10), 
       width = 5, height = 5, dpi = 300)
```

```{r check}
load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

View(mouseHSPC_top$sampleMetadata %>% as.data.frame())

colnames(assay(mouseHSPC_top))
```

