---
title: "Neg_sim"
author: "LF"
date: "2024-05-31"
output: html_document
---


# Load Library

```{r load library}
library(readxl)
library(purrr)
library(dplyr)
library(dbscan)
library(ggbiplot)
library(gridExtra)
library(ComplexHeatmap)
library(Rtsne)
library(dbscan)
library(rpart)
library(rpart.plot)
library(pROC)
library(sjPlot)
library(glmnet)
library(logistf)

# Define the directory containing the R files - Barbie 
directory <- "/Users/feiliyang/Desktop/Barbie/R"

# Get the list of R files in the directory
file_list <- list.files(directory, pattern = "\\.R$", full.names = TRUE)

# Loop through each file and load it
for (file in file_list) {
  source(file)
}
```

# Read Object

```{r read data}
Belderbos <- readRDS(file = "/Users/feiliyang/Desktop/published_analysis_tools/barcodetrackRData/datasets/BelderbosME_etal/Belderbos_Barbie.RDS")
```


# Trim Data

```{r trim data}
del_out <- Belderbos$metadata$Time %in% c("library", "time0")

# |Belderbos$metadata$Mouse == "C23"
Belderbos_hem <- trimColumn(Barbie = Belderbos, keep_columns = !del_out)

Belderbos_hem <- getTopBar(Barbie = Belderbos_hem, cut_off = 0.90, num_appearance = 2)

sum(Belderbos_hem$is_top)

Belderbos_hem <- trimRow(Barbie = Belderbos_hem, keep_rows = Belderbos_hem$is_top)
```

# PCA

```{r PCA, fig.width=9, fig.height=12}
plotPsmp <- function(Barbie, fac, df, show_legend = TRUE){
  
  if(fac %in% colnames(Barbie$metadata)){
    col_point <- Barbie$metadata[,fac]
    col_palette <- fac
    # col_scale <- Barbie$color_panel[[col_palette]]
  }
  
  if(fac == "DBSCAN"){
    col_point <- as.factor(paste0("Group ", df$DBSCAN))
    col_palette <- "DBSCAN"
    my_colors <- c("red3", "slateblue", "forestgreen", "orange", "plum", "grey",
                   "turquoise","olivedrab2", "grey27")
    col_scale <- my_colors[df$DBSCAN |> unique() |> seq() ]
  }
  
  ggbiplot(df, ellipse = F, choices = c(1,2), 
           groups = col_point, 
           varname.size = 0, var.axes = F) +
    theme(aspect.ratio = 1) +
    # geom_point(size = 2, 
    #            shape = 21,
    #            stroke = 1,
    #            fill = "white",
    #            alpha = 0.8,
    #            aes(color = col_point)) +
    # scale_color_manual(values = col_scale) +
    # fix style
    theme_classic() +
          labs(color = col_palette, title = paste0("Samples Colored by \"", col_palette, "\"")) +
          theme(aspect.ratio = 1,
          plot.title = element_text(size = 12),
          axis.title.x = element_text(size = 10),
          axis.title.y = element_text(size = 10),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          legend.title = element_text(size = 0),
          legend.text = element_text(size = 6),
          legend.key.height = unit(0.4, "cm"),   # Adjust the height of the legend keys
          legend.key.width = unit(0.4, "cm"),
          legend.spacing = unit(0.05, "cm"),
          legend.position = ifelse(show_legend, "top", "none"),
          legend.justification = "left")
}

plot_sample_pca <- function(Barbie, db_tune = 0.9, show_legend = TRUE) {
  smpt <- t(log2(1+ Barbie$CPM))
  smpP <- as.matrix(smpt)
  constant_columns <- apply(smpP, 2, function(col) sd(col, na.rm = TRUE) == 0)
  smpP_filtered <- smpP[, !constant_columns]


  p <- prcomp(smpP_filtered, center = TRUE,scale. = TRUE)
  
  tpsmp_dbs <- p$x[,1:2]
  tpd <- dbscan(tpsmp_dbs, eps = db_tune, minPts = 2)
  p$DBSCAN <- factor(tpd$cluster + 1)
  
  pcls <- lapply(c("Mouse", "xCondition", "Time", "Tissue", "Celltype", "Position", "DBSCAN"), function(x) {
      plotPsmp(Barbie = Barbie, fac = x, df = p, show_legend = show_legend)
    })
  do.call(grid.arrange, c(pcls, ncol = 3))
}

plot_sample_pca(Barbie = Belderbos_hem)

ComplexHeatmap::Heatmap(
  matrix = log2(Belderbos_hem$CPM + 1), 
  show_row_names = FALSE,
  top_annotation = HeatmapAnnotation(
    Mouse = Belderbos_hem$metadata$Mouse,
    Celltype = Belderbos_hem$metadata$Celltype,
    Time = Belderbos_hem$metadata$Time,
    Tissue = Belderbos_hem$metadata$Tissue,
    Position = Belderbos_hem$metadata$Position,
    xCondition = Belderbos_hem$metadata$xCondition,
    annotation_name_side = "left",
    annotation_name_gp = gpar(fontsize = 10))
  ) 
```
# tSNE

!! there are duplicated columns - identical values. check it.

```{r tsne}

# Get a logical vector indicating duplicated rows
duplicated_samp <- duplicated(t(Belderbos_hem$CPM))
unique_samp <- !duplicated_samp

#trim the obj
Belderbos_uniqueCol<- trimColumn(Barbie = Belderbos_hem, keep_columns = unique_samp)

tsne_results <- (Belderbos_uniqueCol$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  Rtsne::Rtsne()
tsne_coords <- tsne_results$Y
tsne_coords <- as.data.frame(tsne_coords)

ggplot(tsne_coords) +
  geom_point(aes(x = V1, y = V2, col = Belderbos_uniqueCol$metadata$Mouse)) +
  labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", col = "Mouse") +
  theme(aspect.ratio = 1)

lapply(
  Belderbos_uniqueCol$metadata %>% ncol() %>% seq(), 
  function(i) {
    col_fac <- colnames(Belderbos_uniqueCol$metadata)[i]
    ggplot(tsne_coords) +
    geom_point(aes(x = V1, y = V2, col = Belderbos_uniqueCol$metadata[[i]])) +
    labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = col_fac) +
      theme_classic() +
    theme(aspect.ratio = 1)
  }
  )



```
- Mouse C10 and C15,

- C2, C3 and C1

show high similarity.

- Mouse C21, C22, C23 show low similarity.

- But Mouse C21, C22, C23 show high homogeneity within a mouse.

in a mouse invivo cell transplantation experiment, a factor contains values like "GK", "L", and "R". what exactly can this factor likely to be?
ChatGPT
In a mouse in vivo cell transplantation experiment, factors like "GK", "L", and "R" could represent different groups or conditions that are part of the experimental design. Here are some possible interpretations:

Transplantation Sites:

"GK": Could stand for a specific location in the mouse where the cells are transplanted, possibly "Gastrocnemius" (a muscle in the lower leg).
"L": Likely represents the "Left" side of the mouse.
"R": Likely represents the "Right" side of the mouse.
Cell Types or Treatments:

"GK": Might indicate a specific type of cell or treatment used in the experiment, potentially "Germline Knockout" if dealing with genetic modifications.
"L": Could still mean "Left," or it could represent a specific treatment or cell line labeled with "L".
"R": Similarly, this might stand for "Right" or a treatment or cell line labeled with "R".
Genetic Modifications:

"GK": Could denote a genetic modification, such as "Gene Knockout".
"L" and "R": If they represent genetic markers, they might denote different alleles or loci.
Experimental Groups:

"GK": Might denote a specific experimental group with a unique treatment or condition.
"L" and "R": Likely represent different groups within the experiment, possibly indicating different treatment conditions or controls.
Given the context of in vivo cell transplantation, the most likely interpretation is that "L" and "R" indicate the sides (Left and Right) of the mouse where the transplantations are performed, and "GK" could represent a specific anatomical site or condition relevant to the experiment, such as a muscle group or a type of genetic modification.

To confirm the exact meaning of these factors, it would be necessary to refer to the specific experimental protocol or design used in the study.

# DBScan

```{r dbscan}
dbscan_result <- dbscan(tsne_coords, eps = 1.1, minPts = 2)

ggplot(
  data = data.frame(tsne_coords, cluster = factor(dbscan_result$cluster)), 
  aes(x = V1, y = V2, color = cluster)
  ) +
  geom_point(size = 1) +
  labs(title = "DBSCAN Clustering on t-SNE result", x = "t-SNE 1", y = "t-SNE 2") +
  theme(aspect.ratio = 1)
```

# Decision Tree

```{r decision tree}
meta <- Belderbos_uniqueCol$metadata

dt_data <- meta[,-1]
dt_data$cluster <- factor(dbscan_result$cluster)
dt_data$Time[dt_data$Time != "sac"] <- "live"

tree_model <- rpart(cluster ~ ., data = dt_data, method = "class")

rpart.plot(tree_model, type = 4, extra = 104)

```

Pseudo replicate groups:
1. Mouse == C21
2. Mouse == C22
3. Mouse == C23
4. Mouse == C10 | C15 & Celltype == Unsorted | Tcell
5. Mouse == C10 | C15 & Celltype == Bcell | Grn | Other
6. Mouse == C1 | C2 | C3 & Time == sac
7. Mouse == C1 & Time == live 
8. Mouse == C2 | C3 & Time == live & Celltype == Unsorted
9. Mouse == C2 | C3 & Time == live & Celltype == Bcell | Tcell | Grn

# Pseudo Replicate Groups

```{r pseudo rep}
meta <- Belderbos_uniqueCol$metadata

meta$pseudo_group[meta$Mouse == "C21"] <- "C21"
meta$pseudo_group[meta$Mouse == "C22"] <- "C22"
meta$pseudo_group[meta$Mouse == "C23"] <- "C23"
meta$pseudo_group[
  meta$Mouse %in% c("C10", "C15") 
  & meta$Celltype %in% c("Unsorted", "Tcell")] <- "C10.15_U.T"
meta$pseudo_group[
  meta$Mouse %in% c("C10", "C15") 
  & meta$Celltype %in% c("Bcell", "Grn", "Other")] <- "C10.15_B.G.O"
meta$pseudo_group[
  meta$Mouse %in% c("C1", "C2", "C3") 
  & meta$Time == "sac"] <- "C1.2.3_sac"
meta$pseudo_group[
  meta$Mouse == "C1" 
  & meta$Time != "sac"] <- "C1.live"
meta$pseudo_group[
  meta$Mouse %in% c("C2", "C3") 
  & meta$Time != "sac"
  & meta$Celltype == "Unsorted"] <- "C2.3_live_U"
meta$pseudo_group[
  meta$Mouse %in% c("C2", "C3") 
  & meta$Time != "sac"
  & meta$Celltype %in% c("Bcell", "Tcell", "Grn")] <- "C2.3_live_B.T.G"

ggplot(
  data = data.frame(tsne_coords, pseudo_group = meta$pseudo_group), 
  aes(x = V1, y = V2, color = pseudo_group)
  ) +
  geom_point(size = 1) +
  labs(title = "Pseudo-Groups on t-SNE result", x = "t-SNE 1", y = "t-SNE 2") +
  theme(aspect.ratio = 1)
```

# Sampling within Pseudo-Groups

## Random Sample Size

```{r sampling C22}
# for example C22
C22_bb <- trimColumn(
   Barbie = Belderbos_uniqueCol,
   keep_columns = meta$pseudo_group == "C22")

samples <- C22_bb$metadata %>% nrow() %>% seq()

# Initialize vectors to store the results
all_pvalues <- vector("list", 100)
all_FPRs <- numeric(100)

set.seed(42) # for reproducibility

for (i in 1:100) {
  # Randomly shuffle the samples and split them
  shuffled_samples <- sample(samples)
  split_point <- sample(2:(length(samples) - 2), 1)
  
  group1_samples <- shuffled_samples[1:split_point]
  group2_samples <- shuffled_samples[(split_point + 1):length(samples)]
  
  # Create a group vector
  group_vec <- vector("character", length = length(samples))
  names(group_vec) <- samples
  group_vec[group1_samples] <- "group1"
  group_vec[group2_samples] <- "group2"
  
  # Get the contingency table
  c_table_i <- GetContingencyTable(Barbie = C22_bb, Vector_customized = group_vec)
  
  # Apply the test and get bias group
  C22_bb <- GetOccBiasGroup(Barbie = C22_bb, contingency_table_ls = c_table_i)
  
  # kableExtra::kable(C22_bb$Bias_Occ)
  # 
  # PlotBiasVsRank(Barbie = C22_bb, passing_data = "contribution")
  # 
  # PlotCpmHP_0(Barbie = C22_bb, show_bias = TRUE, Vector_customized = group_vec, split_by = group_vec) 
  
  # Store p-values
  p.values <- C22_bb$Bias_Occ$pvalue
  all_pvalues[[i]] <- p.values
  
  # Calculate and store FDR
  num_unbiased <- sum(C22_bb$Bias_Occ$group == "Unbiased")
  num_biased <- sum(C22_bb$Bias_Occ$group != "Unbiased")
  
  FPR <- num_biased / (num_biased + num_unbiased)
  all_FPRs[i] <- FPR
}

plot(density(all_FPRs))
plot(density(unlist(all_pvalues)))

sim_pvalues <- unlist(all_pvalues)
ground_truth <- sample(c(0, 1), length(sim_pvalues), replace = TRUE)  

# Simulated ground truth
roc_curve <- roc(ground_truth, sim_pvalues)
plot(roc_curve, main = "ROC Curve")

# Calculate the AUC
auc_value <- auc(roc_curve)
print(paste("AUC:", auc_value))

# Plot the ROC curve
plot(roc_curve, main = paste("ROC Curve (AUC =", round(auc_value, 3), ")"), col = "#1c61b6", lwd = 2)
abline(a = 1, b = -1, lty = 2, col = "red")

```

## Fixed Sample Size

```{r sampling fixed size}

set.seed(42) # for reproducibility

# function for sampling once making two equal groups
Sample_Groups <- function(Barbie, size_group = 3) {
  
  pool <- Barbie$metadata %>% nrow() %>% seq()
  
  # First sampling without replacement
  group1_samples <- sample(pool, size = size_group, replace = FALSE)
  # Remove the first sample from the pool
  remaining_pool <- setdiff(pool, group1_samples)
  
  # Second sampling without replacement from the remaining pool
  group2_samples <- sample(remaining_pool, size = size_group, replace = FALSE)
  
  # Create a group vector
  group_vec <- vector("character", length = length(pool))
  group_vec[group1_samples] <- "group1"
  group_vec[group2_samples] <- "group2"
  
  Barbie$metadata$group <- group_vec
  
  # Hem data by group vector
  sample_bb <- trimColumn(
    Barbie = Barbie,
    keep_columns = group_vec %in% c("group1", "group2")
  )
  
  c_table_i <- GetContingencyTable(
    Barbie = sample_bb,
    Vector_customized = sample_bb$metadata$group)
  
  sample_bb <- GetOccBiasGroup(Barbie = sample_bb, contingency_table_ls = c_table_i)
  
  mytargets <- data.frame(sample_bb$metadata)
  
  # model design matrix
  mydesign <- mytargets %>%
    with(model.matrix(~0 + group))
  
  # set up block
  myblock <- NULL
  
  # apply test for Bias_Prop
  sample_bb <- getPropBiasGroup(Barbie = sample_bb, mydesign = mydesign, myblock = myblock, mytargets = mytargets)
  
  #logistic model
  sample_bb <- GetLogitOcc(Barbie = sample_bb, penalized = F)
  
  #penalized logistic madel
  sample_bb <- GetLogitOcc(Barbie = sample_bb, penalized = T)
  
  # PlotBiasVsRank(Barbie = sample_bb, passing_data = "contribution")
  # 
  # PlotCpmHP_0(Barbie = sample_bb, show_bias = TRUE, Vector_customized = sample_bb$metadata$group, split_by = sample_bb$metadata$group)
  # 
  # PlotBiasVsRank(Barbie = sample_bb, passing_data = "contribution", bias_group = sample_bb$Bias_Prop$group, bias_pvalue = sample_bb$Bias_Prop$pvalue)
  # 
  # PlotCpmHP_0(Barbie = sample_bb, show_bias = TRUE, Vector_customized = sample_bb$metadata$group, split_by = sample_bb$metadata$group, bias_group = sample_bb$Bias_Prop$group)
  
  # extract p.values and FPR
  pvalues_occ <- sample_bb$Bias_Occ$pvalue
  pvalues_prop <- sample_bb$Bias_Prop$pvalue
  pvalues_LR_occ_classic <- sample_bb$LR_Occ_classic$p.values
  pvalues_LR_occ_penalized <- sample_bb$LR_Occ_penalized$p.values
  
  FPR_occ <- sum(sample_bb$Bias_Occ$group != "Unbiased") / length(sample_bb$Bias_Occ$group)
  
  FPR_prop <- sum(sample_bb$Bias_Prop$group != "Unbiased") / length(sample_bb$Bias_Prop$group)
  
  FPR_LR_occ_classic <- sum(sample_bb$LR_Occ_classic$group != "Unbiased") / length(sample_bb$LR_Occ_classic$group)
  
  FPR_LR_occ_penalized <- sum(sample_bb$LR_Occ_penalized$group != "Unbiased") / length(sample_bb$LR_Occ_penalized$group)
  
  return(
    list(FPR_occ = FPR_occ,
         FPR_prop = FPR_prop,
         FPR_LR_occ_classic = FPR_LR_occ_classic,
         FPR_LR_occ_penalized = FPR_LR_occ_penalized,
         pvalues_occ = pvalues_occ,
         pvalues_prop = pvalues_prop,
         pvalues_LR_occ_classic = pvalues_LR_occ_classic,
         pvalues_LR_occ_penalized = pvalues_LR_occ_penalized,
         group_vec = group_vec)
  )

}

Loop_Samples <- function(Barbie, size_group = 3, loop_times = 100) {
  
  FPR_occ_list <- list(length = loop_times)
  FPR_prop_list <- list(length = loop_times)
  FPR_LR_occ_classic_list <- list(length = loop_times)
  FPR_LR_occ_penalized_list <- list(length = loop_times)
  
  pvalues_occ_list <- list(length = loop_times)
  pvalues_prop_list <- list(length = loop_times)
  pvalues_LR_occ_classic_list <- list(length = loop_times)
  pvalues_LR_occ_penalize_list <- list(length = loop_times)
  
  group_vec_list <- list(length = group_vec)
  
  for(i in 1:loop_times) {
    results_i <- Sample_Groups(
      Barbie = Barbie, size_group = size_group)
    
    FPR_occ_list[[i]] <- results_i$FPR_occ
    FPR_prop_list[[i]] <- results_i$FPR_prop
    FPR_LR_occ_classic_list[[i]] <- results_i$FPR_LR_occ_classic
    FPR_LR_occ_penalized_list[[i]] <- results_i$FPR_LR_occ_penalized
    
    pvalues_occ_list[[i]] <- results_i$pvalues_occ
    pvalues_prop_list[[i]] <- results_i$pvalues_prop
    pvalues_LR_occ_classic_list[[i]] <- results_i$pvalues_LR_occ_classic
    pvalues_LR_occ_penalize_list[[i]] <- results_i$ pvalues_LR_occ_penalize
    
    group_vec_list[[i]] <- results_i$group_vec
  }
  
  FPR <- data.frame(
    N_Samples = rep(size_group, loop_times),
    Occ = FPR_occ_list %>% unlist(),
    Prop = FPR_prop_list %>% unlist(),
    LR_occ_classic = FPR_LR_occ_classic_list %>% unlist(),
    LR_occ_penalized = FPR_LR_occ_penalized_list %>% unlist()
    )
  
  pvalues <- data.frame(
    N_Samples = rep(size_group, loop_times),
    Occ = pvalues_occ_list %>% unlist(),
    Prop = pvalues_prop_list %>% unlist(),
    LR_occ_classic = pvalues_LR_occ_classic_list %>% unlist(),
    LR_occ_penalized = pvalues_LR_occ_penalize_list %>% unlist()
    )
  
  return(list(
    FPR = FPR,
    pvalues = pvalues,
    group_vec_list = group_vec_list))

}

```

### C22

```{r random sampling C22}
C22_bb <- trimColumn(
  Barbie = Belderbos_uniqueCol,
  keep_columns = meta$pseudo_group == "C22")
# run the loop for each size of samples

random_sampling <- function(Barbie) {

  num_samples <- ncol(Barbie$CPM)
  end_sampling <- floor(num_samples / 2)
  
  # stop running loop if end_sampling <= 3
  if(num_samples < 3) {
    return("not enough samples!")
  }
  
  set.seed(42) # for reproducibility
  
  all_loops <- lapply(3:end_sampling, function(n) {
    Loop_Samples(
      Barbie = Barbie, size_group = n, loop_times = 100)
  })
  
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    all_loops[[n]]$FPR
  }) 
  
  df_FPR <- do.call(rbind, all_FPR)
  
  df_FPR$N_Samples <- as.factor(df_FPR$N_Samples)
  
  all_pvalue <- lapply(seq(3:end_sampling), function(n) {
    all_loops[[n]]$pvalues
  })
  
  df_pvalues <- do.call(rbind, all_pvalue)
  
  df_pvalues$N_Samples <- as.factor(df_pvalues$N_Samples)
  
  gFDRocc <- ggplot(as.data.frame(df_FPR),
    aes(x = N_Samples, y = Occ, fill = N_Samples)) +
    # geom_violin(width = 0.5) +  # Set width for better visibility +
    geom_boxplot(width = 0.5, outlier.shape = NA) +
    geom_jitter(width = 0.1, height = 0, size = 1) +  # Add jittered points
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
    labs(title = "Diff_Occ, Mouse C22",
         y = "Prop. barcodes with p.value < 0.05",
         x = "No. samples per group") +
    theme_classic()
  
  gFDRprop <- ggplot(as.data.frame(df_FPR),
    aes(x = N_Samples, y = Prop, fill = N_Samples)) +
    # geom_violin(width = 0.5) +  # Set width for better visibility
    geom_boxplot(width = 0.5, outlier.shape = NA) +
    geom_jitter(width = 0.1, height = 0, size = 1) +  # Add jittered points
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
    labs(title = "Diff_Prop, Mouse C22",
         y = "Prop. barcodes with p.value < 0.05",
         x = "No. samples per group") +
    theme_classic()
  
  
  gPocc <- ggplot(as.data.frame(df_pvalues),
    aes(x = N_Samples, y = Occ, fill = N_Samples)) +
    geom_violin(width = 0.5) +
    geom_jitter(width = 0.1, height = 0, size = 1) +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
    labs(title = "Diff_Occ, Mouse C22",
         y = "p.value")
  
  gPprop <- ggplot(as.data.frame(df_pvalues),
    aes(x = N_Samples, y = Prop, fill = N_Samples)) +
    geom_violin() +
      geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
    labs(title = "Diff_Prop, Mouse C22",
         y = "p.value")
  
  return(
    list(gFDRocc = gFDRocc,
         gFDRprop = gFDRprop,
         gPocc = gPocc,
         gPprop = gPprop))

}

results_negsi <- random_sampling(Barbie = C22_bb)

lapply(results_negsi, plot)

# Calculate the combos
for(i in 3:8) {
  print(choose(17,i) * choose(17-i,i))
}

```

### C22 t-SNE

```{r C22 tsne}
# Ensure you have the right number of samples
num_samples <- ncol(C22_bb$CPM)
perplexity <- min(30, (num_samples - 1) / 3)

tsne_results <- (C22_bb$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  Rtsne::Rtsne(perplexity = perplexity)
tsne_coords <- tsne_results$Y
tsne_coords <- as.data.frame(tsne_coords)

lapply(
  C22_bb$metadata %>% ncol() %>% seq(), 
  function(i) {
    col_fac <- colnames(C22_bb$metadata)[i]
    ggplot(tsne_coords) +
    geom_point(aes(x = V1, y = V2, col = C22_bb$metadata[[i]])) +
    labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = col_fac) +
      theme_classic() +
    theme(aspect.ratio = 1)
  }
  )

# color samples by random groups
flagN4 <- all_loops[[2]]$FPR$Occ > 0.05
group_vec_N4 <- all_loops[[2]]$group_vec_list[flagN4]

flagN8 <- all_loops[[6]]$FPR$Occ > 0.05
group_vec_N8 <- all_loops[[6]]$group_vec_list[flagN8]

flagN3 <- all_loops[[1]]$FPR$Prop > 0.05
group_vec_N3 <- all_loops[[1]]$group_vec_list[flagN3]

# function for plotting random groups
color_rd_group <- function(group_info){
  ggplot(tsne_coords) +
  geom_point(aes(x = V1, y = V2, col = group_info,
                 shape = group_info), size = 3) +
  labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", col = "Random Group", shape = "Random Group") +
  theme_classic() +
  theme(aspect.ratio = 1) +
  scale_color_manual(
    values = c("grey", "group1" = "red", "group2" = "blue")) +
  scale_shape_manual(
    values = c(16,17,17)
  )
}

# plot sig. simulations
lapply(list(group_vec_N4[[1]],
            group_vec_N4[[2]],
            group_vec_N8[[1]],
            group_vec_N3[[1]],
            group_vec_N3[[2]],
            group_vec_N3[[3]]),
       function(x) {
        color_rd_group(x)
      })

# Unsorted subset has 10 samples in total.
for(i in 3:5) {
  print(choose(10,i) * choose(10-i,i))
}
```

### C22 details

```{r C22 details, fig.width=10, fig.height=7}
# subset of sig. simulations
C22_bb$metadata$group <- group_vec_N3[[2]]
NegSim_gp <- trimObjectByMetadata(
  Barbie = C22_bb,
  condition = "group",
  specified = c("group1", "group2"),
  keep = TRUE
)

ConTb_N3_1<- GetContingencyTable(
  Barbie = NegSim_gp,
  Factor = "group"
)

NegSim_gp <- GetOccBiasGroup(
  Barbie = NegSim_gp,
  contingency_table_ls = ConTb_N3_1
)

mytargets <- data.frame(NegSim_gp$metadata)
  
# model design matrix
mydesign <- mytargets %>%
    with(model.matrix(~0 + group))
  
# set up block
myblock <- NULL
  
# apply test for Bias_Prop
NegSim_gp <- getPropBiasGroup(
  Barbie = NegSim_gp, 
  mydesign = mydesign, 
  myblock = myblock, 
  mytargets = mytargets)

row_ha <- NegSim_gp %>% with(
  rowAnnotation(
    Prop = Bias_Prop$group,
    Occ = Bias_Occ$group,
    col = list(
      Prop = color_panel$bias_group,
      Occ = color_panel$bias_group
      ),
    show_legend = TRUE,
    show_annotation_name = TRUE,
    annotation_name_side = "bottom",
    annotation_name_gp = gpar(fontsize = 10)
    )
)

PlotCpmHP_0(
  Barbie = NegSim_gp,
  Vector_customized = NegSim_gp$metadata$group,
  show_bias = TRUE,
  row_ha = row_ha
  ) 

pointer_N3_1 <- NegSim_gp$Bias_Occ$group != "Unbiased"

NegSim_gp$Bias_Prop$group[pointer_N3_1]

ConTb_N3_1[pointer_N3_1]



# Example data:
#            Response
# Treatment   Yes No
# Treatment1  3  0
# Treatment2  0  3

# Creating the table in R
contingency_table <- matrix(c(0, 3, 3, 0), nrow = 2, byrow = TRUE)

# Add row and column names (optional, but helps with readability)
dimnames(contingency_table) <- list(
  Treatment = c("Treatment1", "Treatment2"),
  Response = c("Yes", "No")
)

# Print the table
print(contingency_table)

# Perform Fisher's Exact Test
fisher.test(contingency_table)

chisq.test(contingency_table)
```

#### Logistic Regression

```{r C22 logistic, fig.width=7, fig.height=6}
# NegSim_gp 

Heatmap(
  # matrix = log2(NegSim_gp$CPM + 1),
  matrix = NegSim_gp$presence + 1 -1,
  name = "NegSim_size8",
  row_split = NegSim_gp$Bias_Occ$group,
  column_split = NegSim_gp$metadata$group,
  show_row_names = F,
  show_column_names = F,
  width = unit(4, "cm"),
  height = unit(4, "cm"))

# prepare df for logistic regression
# factors in columns, samples in rows

# selcte a barcode as outcome
NegSim_gp_un <- trimRow(
  Barbie = NegSim_gp,
  keep_rows = NegSim_gp$Bias_Occ$group != "Unbiased")

outcome <- NegSim_gp$presence + 1 - 1

# Create a function to fit the logistic model to a single row
fit_logistic_model <- function(row) {
  dat <- data.frame(
    Group = as.factor(NegSim_gp$metadata$group),
    Barcode_i = factor(row)
  )
  
  logiM8 <- glm(
    Barcode_i ~ Group,
    data = dat,
    family = "binomial"
  )
  
  plot_model(logiM8,
  type = "pred",
  terms = "Group"
) +
  labs(y = "Prob(Presence)")
  
  return(summary(logiM8))
}

firth_model <- function(row) {
  dat <- data.frame(
    Group = as.factor(NegSim_gp$metadata$group),
    Barcode_i = factor(row)
  )
  
  firthM8 <- logistf(Barcode_i ~ Group, data = dat)
  
  return(summary(firthM8))
  
}

# Apply the function to each row of outcome and store the results
results <- lapply(1:nrow(outcome), function(i) fit_logistic_model(outcome[i, ]))

# View the summary of the first model as an example
results[[1]]$coefficients[2,4]

p.values.logi <- lapply(results, function(x) 
  x$coefficients[2,4]) %>% unlist()

beta.logi <- lapply(results, function(x) 
  x$coefficients[2,1]) %>% unlist()

signif.logi <- ifelse(p.values.logi >= 0.05,
                      "Unbiased",
                      ifelse(beta.logi > 0,
                             "Group2", "Group1"))

# Apply the function to each row of outcome and store the results
results_firth <- lapply(1:nrow(outcome), function(i) firth_model(outcome[i, ]))

# View the summary of the first model as an example
rr <- results_firth[[1]]
rr$prob[2]

p.values.firth <- lapply(
  results_firth, function(x) 
  x$prob[2]) %>% unlist()

beta.firth <- lapply(
  results_firth, function(x) 
  x$coefficients[2]) %>% unlist()

signif.firth <- ifelse(p.values.firth >= 0.05,
                      "Unbiased",
                      ifelse(beta.firth > 0,
                             "Group2", "Group1"))

set.seed(123)
row_logi <- rowAnnotation(
  Prop = NegSim_gp$Bias_Prop$group,
  Fishers = NegSim_gp$Bias_Occ$group,
  Logi = signif.logi,
  Firth = signif.firth
  # p.value = anno_text(
  #   p.values.logi %>% round(3),
  #   gp = gpar(fontsize = 6)),
  # beta.logi = anno_text(
  #   beta.logi %>% round(3),
  #   gp = gpar(fontsize = 6))
)

Heatmap(
  matrix = log2(NegSim_gp$CPM + 1),
  # matrix = NegSim_gp$presence + 1 -1,
  name = "NegSim_size8",
  row_split = factor(
    NegSim_gp$Bias_Occ$group, 
    levels = c("group2", "Unbiased")),
  cluster_row_slices = F,
  column_split = NegSim_gp$metadata$group,
  show_row_names = F,
  show_column_names = F,
  width = unit(4, "cm"),
  height = unit(10, "cm")
  # right_annotation = row_logi
  ) + 
  
  Heatmap(
  # matrix = log2(NegSim_gp$CPM + 1),
  matrix = NegSim_gp$presence + 1 -1,
  name = "NegSim_size8",
  row_split = factor(
    NegSim_gp$Bias_Occ$group, 
    levels = c("group2", "Unbiased")),
  cluster_row_slices = F,
  column_split = NegSim_gp$metadata$group,
  show_row_names = F,
  show_column_names = F,
  width = unit(4, "cm"),
  height = unit(10, "cm"),
  right_annotation = row_logi
  )

# find the extreme situations
flag_ex <- beta.logi > 20

presence <- outcome[flag_ex,][1,] + 1 - 1

# peak the extreme model
fit_logistic_model(presence)

  dat <- data.frame(
    Barcode_i = factor(presence),
    Group = as.factor(NegSim_gp$metadata$group)
  )
  
  logiM8 <- glm(
    Barcode_i ~ Group,
    data = dat,
    family = "binomial"
  )
  
plot_model(logiM8,
  type = "pred",
  terms = "Group"
) +
  labs(y = "Prob(Presence)")

table(dat)

# we get insignificant coefficient of group2 because of perfect separation, that all outcomes for group1 are 0 and all 1 are in gruop2.

# To solve this, we can use 1) Lasso Regularization 2) Firth's Penalized Likelihood.

# 1)

# Prepare data for glmnet
x <- model.matrix(Barcode_i ~ Group, data = dat)
y <- dat$Barcode_i

# Fit logistic regression with lasso regularization
fit <- glmnet(x, y, family = "binomial", alpha = 1)

# Print coefficients at lambda that minimizes cross-validated error
coef(fit, s = "lambda.min")


# 2)

# Fit logistic regression using Firth's penalized likelihood
fit_firth <- logistf(Barcode_i ~ Group, data = dat)

# Print summary of the model
summary(fit_firth)
```

#### Test

```{r logi test, fig.width=7, fig.height=6}
source("/Users/feiliyang/Desktop/Barbie/R/Get_LR_Occ.R")

NegSim_gp <-  GetLogitOcc(Barbie = NegSim_gp, penalized = F)
NegSim_gp <- GetLogitOcc(Barbie = NegSim_gp, penalized = T)

set.seed(123)
row_logi <- rowAnnotation(
  Prop = NegSim_gp$Bias_Prop$group,
  Fishers = NegSim_gp$Bias_Occ$group,
  Logi = NegSim_gp$LR_Occ_classic$group,
  Firth = NegSim_gp$LR_Occ_penalized$group
  # p.value = anno_text(
  #   p.values.logi %>% round(3),
  #   gp = gpar(fontsize = 6)),
  # beta.logi = anno_text(
  #   beta.logi %>% round(3),
  #   gp = gpar(fontsize = 6))
)

Heatmap(
  matrix = log2(NegSim_gp$CPM + 1),
  # matrix = NegSim_gp$presence + 1 -1,
  name = "NegSim_size8",
  row_split = factor(
    NegSim_gp$Bias_Occ$group, 
    levels = c("group2", "Unbiased")),
  cluster_row_slices = F,
  column_split = NegSim_gp$metadata$group,
  show_row_names = F,
  show_column_names = F,
  width = unit(4, "cm"),
  height = unit(10, "cm")
  # right_annotation = row_logi
  ) + 
  
  Heatmap(
  # matrix = log2(NegSim_gp$CPM + 1),
  matrix = NegSim_gp$presence + 1 -1,
  name = "NegSim_size8",
  row_split = factor(
    NegSim_gp$Bias_Occ$group, 
    levels = c("group2", "Unbiased")),
  cluster_row_slices = F,
  column_split = NegSim_gp$metadata$group,
  show_row_names = F,
  show_column_names = F,
  width = unit(4, "cm"),
  height = unit(10, "cm"),
  right_annotation = row_logi
  )

```


### C22 Unsorted

```{r C22 unsorted}
C22_Un <- trimColumn(
  Barbie = C22_bb,
  keep_columns = C22_bb$metadata$Celltype == "Unsorted")

random_sampling(Barbie = C22_Un)
```


### C21

```{r random sampling C21}
C21_bb <- trimColumn(
  Barbie = Belderbos_uniqueCol,
  keep_columns = meta$pseudo_group == "C21")

C21_bb$metadata %>% nrow()

# run the loop for each size of samples

all_loops <- lapply(3:11, function(n) {
  Loop_Samples(
    Barbie = C21_bb, size_group = n, loop_times = 100)
})

all_FPR <- lapply(seq(3:11), function(n) {
  all_loops[[n]]$FPR
}) 

df_FPR <- do.call(rbind, all_FPR)

df_FPR$N_Samples <- as.factor(df_FPR$N_Samples)

all_pvalue <- lapply(seq(3:11), function(n) {
  all_loops[[n]]$pvalues
})

df_pvalues <- do.call(rbind, all_pvalue)

df_pvalues$N_Samples <- as.factor(df_pvalues$N_Samples)

ggplot(as.data.frame(df_FPR[201:300,]),
  aes(x = N_Samples, y = Occ, fill = N_Samples)) +
  geom_violin() +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  labs(title = "Diff_Occ, Mouse C22",
       y = "FPR")

ggplot(as.data.frame(df_FPR),
  aes(x = N_Samples, y = Prop, fill = N_Samples)) +
  geom_violin() +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  labs(title = "Diff_Prop, Mouse C22",
       y = "FPR")


ggplot(as.data.frame(df_pvalues),
  aes(x = N_Samples, y = Occ, fill = N_Samples)) +
  geom_violin() +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  labs(title = "Diff_Occ, Mouse C22",
       y = "p.value")

ggplot(as.data.frame(df_pvalues),
  aes(x = N_Samples, y = Prop, fill = N_Samples)) +
  geom_violin() +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  labs(title = "Diff_Prop, Mouse C22",
       y = "p.value")


```


# Test Occ between Mouse

```{r test mice}
# subset: C23, C22, or C21?
bb_1 <- trimObjectByMetadata(
  Barbie = Belderbos_uniqueCol,
  condition = "Mouse",
  specified = c("C10", "C15"),
  keep = TRUE
)

bb_1 <- trimObjectByMetadata(
  Barbie = bb_1,
  condition = "Celltype",
  specified = c("Unsorted", "Tcell"),
  keep = TRUE
)

c_table_1 <- GetContingencyTable(
  Barbie = bb_1,
  Vector_customized = bb_1$metadata$Mouse
)

# lapply(c_table_1[1:5], function(x) {knitr::kable(x)})

# Apply test, and get Bias group
bb_1 <- GetOccBiasGroup(Barbie = bb_1, contingency_table_ls = c_table_1)

kableExtra::kable(bb_1$Bias_Occ)

PlotBiasVsRank(Barbie = bb_1, passing_data = "contribution")

PlotCpmHP_0(Barbie = bb_1, show_bias = TRUE, Vector_customized = bb_1$metadata$Mouse, split_by = bb_1$metadata$Mouse)
```

# Test Prop between Mouse

```{r test prop between mouse}
bb_1 <- trimObjectByMetadata(
  Barbie = Belderbos_uniqueCol,
  condition = "Mouse",
  specified = c("C21", "C22"),
  keep = TRUE
)

# make your own targets and design matrix
# targets
grouped_time <- bb_1$metadata$Time
grouped_time[bb_1$metadata$Time != "sac"] <- "live"

mytargets <- data.frame(
  bb_1$metadata, 
  grouped_time = grouped_time
  )

# model design matrix
mydesign <- mytargets %>%
  with(model.matrix(~0 + Mouse + Tissue + grouped_time + Celltype))

# set up block
myblock <- mytargets %>%
    with(paste(Mouse, Tissue, grouped_time, Celltype))

# apply test for Bias_Prop
bb_1 <- getPropBiasGroup(Barbie = bb_1, mydesign = mydesign, myblock = myblock)

PlotBiasVsRank(
  Barbie = bb_1, 
  passing_data = "contribution", 
  bias_group = bb_1$Bias_Prop$group,
  bias_pvalue = bb_1$Bias_Prop$pvalue
  )

PlotCpmHP_0(
  Barbie = bb_1, 
  show_bias = TRUE, 
  Vector_customized = bb_1$metadata$Mouse, 
  split_by = bb_1$metadata$Mouse,
  bias_group = bb_1$Bias_Prop$group) 
```


# Homogenous Sets

```{r homogenous, fig.height=15, fig.width=15}
meta <- Belderbos_uniqueCol$metadata
View(meta)
kableExtra::kable(meta)

# group wkxx before sac together
grouped_time <- meta$Time
grouped_time[grouped_time != "sac"] <- "live"

# group other tissues against BM
grouped_tissue <- meta$Tissue
grouped_tissue[grouped_tissue != "BM"] <- "non_BM"

meta <- meta %>% 
  mutate(combined = paste(grouped_tissue, Celltype, grouped_time, xCondition, sep = "."))

table(meta$combined)

homo <- duplicated(meta$combined, fromLast = TRUE) | duplicated(meta$combined)
homo_cond <- unique(meta$combined[homo])

table(meta$combined[homo])

tsne_coords <- as.data.frame(tsne_coords)

# use facet wrap to highlight each group of duplicated conditions.
tsne_df <- data.frame(
  tsne_coords, 
  combined = meta$combined,
  Mouse.Time.Tissue = paste(meta$Mouse, meta$Time, meta$Tissue, sep = "."),
  grouped_time = grouped_time,
  grouped_tissue = grouped_tissue)

ggplot() +
  geom_point(data = tsne_coords, aes(x = V1, y = V2), color = "grey") +
  geom_point(data = tsne_df[homo,], aes(x = V1, y = V2, color = combined)) +
  geom_text(data = tsne_df[homo,], aes(x = V1+0.5, y = V2+0.5, label = Mouse.Time.Tissue), size = 2, position = position_jitter(width = 0.3, height = 0.3)) +
  facet_wrap(~ combined) +
  labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = "combined") +
  theme(aspect.ratio = 1, legend.position = "none")
```

```{r homogenous dif position, fig.height=15, fig.width=15}
meta <- Belderbos_uniqueCol$metadata
View(meta)
kableExtra::kable(meta)

# group wkxx before sac together
grouped_time <- meta$Time
grouped_time[grouped_time != "sac"] <- "live"

# group other tissues against BM
grouped_tissue <- meta$Tissue
grouped_tissue[grouped_tissue != "BM"] <- "non_BM"

# group mice
grouped_mouse <- meta$Mouse
grouped_mouse[grouped_mouse %in% c("C1", "C2", "C3")] <- "Pseudo1"
grouped_mouse[grouped_mouse %in% c("C10", "C15")] <- "Pseudo2"

meta <- meta %>% 
  mutate(combined = paste(grouped_mouse, Tissue, Celltype, grouped_time, xCondition, sep = "."))

table(meta$combined)

homo <- duplicated(meta$combined, fromLast = TRUE) | duplicated(meta$combined)
homo_cond <- unique(meta$combined[homo])

table(meta$combined[homo])

BM <- meta$Tissue == "BM"

tsne_coords <- as.data.frame(tsne_coords)

# use facet wrap to highlight each group of duplicated conditions.
tsne_df <- data.frame(
  tsne_coords, 
  combined = meta$combined,
  Mouse.Position = paste(meta$Mouse, meta$Position, sep = "."),
  grouped_time = grouped_time,
  grouped_tissue = grouped_tissue)

ggplot() +
  geom_point(data = tsne_coords, aes(x = V1, y = V2), color = "grey") +
  geom_point(data = tsne_df[homo,], aes(x = V1, y = V2, color = combined)) +
  geom_text(data = tsne_df[homo,], aes(x = V1+0.5, y = V2+0.5, label = Mouse.Position), size = 2, position = position_jitter(width = 0.3, height = 0.3)) +
  facet_wrap(~ combined) +
  labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = "combined") +
  theme(aspect.ratio = 1, legend.position = "none")
```
