---
title: "neg_sim_mouse_HSOC_barbieQ"
author: "LF"
date: "2025-05-04"
output: workflowr::wflow_html
output_dir: ../../BelderbosME
---

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)

library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
library(bartools)
```

# Install barbieQ - devel

```{r install barbieQ}
# devtools::install_github("Oshlack/barbieQ", force = TRUE)
BiocManager::install("barbieQ")

library(barbieQ)
```

# Load rds

```{r save rds}
# rds saved in "../../analysis/BelderbosME/mouseHSPC_preprocessing.Rmd"

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

assay(mouseHSPC) %>% colSums() %>% sort()
assay(mouseHSPC_top) %>% colSums() %>% sort()

(colSums(assay(mouseHSPC)) - colSums(assay(mouseHSPC_top))) %>% sort()

```

## t-SNE

```{r tsne}
# Set an appropriate perplexity value (e.g., num_samples / 3)
# set.seed(12)
# perplexity_value <- min(30, floor(ncol(mouseHSPC_top) / 3))
# 
# tsne_results <- (assays(mouseHSPC_top)$CPM + 1) %>% 
#   t() %>% 
#   log2() %>% 
#   Rtsne::Rtsne(perplexit = perplexity_value)
# tsne_coords <- tsne_results$Y
# tsne_coords <- as.data.frame(tsne_coords)
# 
# ggplot(tsne_coords) +
#   geom_point(aes(x = V1, y = V2, col = mouseHSPC_top$sampleMetadata$Mouse)) +
#   labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", col = "Mouse") +
#   theme_classic() +
#   theme(aspect.ratio = 1)
# 
# lapply(
#   mouseHSPC_top$sampleMetadata %>% ncol() %>% seq(), 
#   function(i) {
#     col_fac <- colnames(mouseHSPC_top$sampleMetadata)[i]
#     ggplot(tsne_coords) +
#      geom_point(aes(x = V1, y = V2, col = mouseHSPC_top$sampleMetadata[[i]])) +
#      labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = col_fac) +
#       theme_classic() +
#      theme(aspect.ratio = 1)
#   }
#   )

pca_results <- (assays(mouseHSPC_top)$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  prcomp( scale. = TRUE)

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  Mouse = mouseHSPC_top$sampleMetadata$Mouse
)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Mouse)) +
  geom_point(size = 3) +
  labs(title = "PCA on logCPM", x = "PC1", y = "PC2") +
  theme_classic() +
  theme(aspect.ratio = 1)

mouseHSPC_top$sampleMetadata %>% as.data.frame() %>%  View()
assay(mouseHSPC_top) %>% colnames()
```

## bartools

```{r bartools, fig.width=30, fig.height=10}
mouse_ids <- mouseHSPC_top$sampleMetadata$Mouse

mouse_levels <- c("C22", setdiff(unique(mouse_ids), "C22"))

ordered_indices <- order(factor(mouse_ids, levels = mouse_levels))


mouseHSPC_top$sampleMetadata <- mouseHSPC_top$sampleMetadata[ordered_indices, ]

ps.counts <- assay(mouseHSPC_top)
ps.counts <- ps.counts[, ordered_indices]
colnames(ps.counts) <- make.unique(mouseHSPC_top$sampleMetadata$Mouse)
mouse_dge <- DGEList(
  counts = ps.counts, 
  group = mouseHSPC_top$sampleMetadata$Mouse)

bartools::plotBarcodeHistogram(mouse_dge)
```

# mouseHSPC

```{r mouseHSPC, fig.width=8, fig.height=6}
barbieQ::plotBarcodeHeatmap(
  mouseHSPC_top, sampleMetadata = mouseHSPC_top$sampleMetadata[,-1])

barbieQ::plotBarcodeHeatmap(
  mouseHSPC_top, sampleMetadata = mouseHSPC_top$sampleMetadata[,-1], splitSamples = T, sampleGroup = "Mouse")
```

### C21

```{r C21}
C21 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C21"]

barbieQ::plotBarcodeHeatmap(
  C21, sampleMetadata = C21$sampleMetadata[,-1])

## flag C21 samples of exremely low libsize
flag_low_C21 <- colSums(assay(C21)) < 1000
## remove the low libsize samples
C21 <- C21[, !flag_low_C21]
colSums(assay(C21))

barbieQ::plotBarcodeHeatmap(
  C21, sampleMetadata = C21$sampleMetadata[,-1])

assay(C21) %>% colSums() %>% sort()

## tag top barcoes
C21 <- tagTopBarcodes(C21)
rowData(C21)$isTopBarcode %>% table()

barbieQ::plotBarcodeProportion(C21, absoluteProportion = T)
```

```{r, fig.width=10}
C21 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C21"]
## flag C21 samples of exremely low libsize
flag_low_C21 <- colSums(assay(C21)) < 1000
## remove the low libsize samples
C21 <- C21[, !flag_low_C21]

C21_dge <- DGEList(
  counts = assay(C21), 
  group = C21$sampleMetadata$Mouse)

bartools::plotBarcodeHistogram(C21_dge)
ggsave("../../output/BelderbosME/bartool_C21.png", dpi = 300,width = 8, height = 6, bartools::plotBarcodeHistogram(C21_dge))
```


### C22

```{r C22}
C22 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C22"]
## flag C21 samples of exremely low libsize
flag_low_C22 <- colSums(assay(C22)) < 1000
## remove the low libsize samples
C22 <- C22[, !flag_low_C22]

## check libsize
assay(C22) %>% colSums()
## tag top barcodes
C22 <- tagTopBarcodes(C22)
rowData(C22)$isTopBarcode %>% table()

barbieQ::plotBarcodeHeatmap(
  C22, sampleMetadata = C22$sampleMetadata[,-1])

barbieQ::plotBarcodeProportion(C22, absoluteProportion = F)
rowMeans(assays(C22)$proportion) %>% summary()
rowMeans(assays(C22)$proportion) %>% sort(decreasing = T) 
rowMeans(assays(mouseHSPC_top)$proportion) %>% summary()


colSums(assay(C22))
```

```{r}
C22 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C22"]
## flag C22 samples of exremely low libsize
flag_low_C22 <- colSums(assay(C22)) < 1000
## remove the low libsize samples
C22 <- C22[, !flag_low_C22]
C22_dge <- DGEList(
  counts = assay(C22), 
  group = C22$sampleMetadata$Mouse)

bartools::plotBarcodeHistogram(C22_dge)
ggsave("../../output/BelderbosME/bartool_C22.png", dpi = 300,width = 8, height = 6, bartools::plotBarcodeHistogram(C22_dge))
```


### C23

```{r C23}
C23 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C23"]

## check libsize
assay(C23) %>% colSums()
## tag top barcoes
C23 <- tagTopBarcodes(C23)
rowData(C23)$isTopBarcode %>% table()

barbieQ::plotBarcodeHeatmap(
  C23, sampleMetadata = C23$sampleMetadata[,-1])

barbieQ::plotBarcodeProportion(C23, absoluteProportion = F)
rowMeans(assays(C23)$proportion) %>% summary()
rowMeans(assays(C23)$proportion) %>% sort(decreasing = T) 
rowMeans(assays(mouseHSPC_top)$proportion) %>% summary()

barbieQ::plotBarcodeProportion(C23, absoluteProportion = T)
```

```{r}
C23 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C23"]
## flag C23 samples of exremely low libsize
flag_low_C23 <- colSums(assay(C23)) < 1000
## remove the low libsize samples
C23 <- C23[, !flag_low_C23]
C23_dge <- DGEList(
  counts = assay(C23), 
  group = C23$sampleMetadata$Mouse)

bartools::plotBarcodeHistogram(C23_dge)
ggsave("../../output/BelderbosME/bartool_C23.png", dpi = 300,width = 8, height = 6, bartools::plotBarcodeHistogram(C23_dge))
```

# Subet C21 Mouse

## NegSim

```{r negsim dmso, fig.width=7, fig.height=5}
source("../../analysis/BelderbosME/NegSim-mouseHSPC.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

C21 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C21"]

## flag C21 samples of exremely low libsize
flag_low_C21 <- colSums(assay(C21)) < 1000
## remove the low libsize samples
C21 <- C21[, !flag_low_C21]

set.seed(2025)

results_negsi <- random_sampling(barbieQ = C21, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/BelderbosME/all_loops_C21.rda")

lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, C21")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, C21")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, C21")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, C21")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, C21")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, C21")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, C21")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, C21")

ggsave(
  "../../output/BelderbosME/test_C21.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 40, height = 15,
  units = "cm"
  )

```

## Inspect NegSim

### load results

```{r extract results from random sampling}
load("../../output/BelderbosME/all_loops_C21.rda")

  # extract results from the loops
  end_sampling <- 11
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```


### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


## flag out the FPR-high simulation
flag_highFPR <- which(df_FPR$Prop_asin > 0.05)
SimID_highFPR <- df_FPR[flag_highFPR,]$SimulationID

# df_Group_Vec[SimID_highFPR,]
```

### P.Val vs. Prop

```{r plot p.val}
stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
# stats_barcodes$Gmean_prop <- (2^stats_barcodes$Gmean_prop-0.5) / 1e6

ggplot(stats_barcodes) +
  geom_point(aes(y = -log10(Prop_asin), x = log10(Amean_prop + 1e-7)), color = "skyblue", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_logit), x = log10(Amean_prop + 1e-7)), color = "yellow", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_noTrans), x = log10(Amean_prop + 1e-7)), color = "pink", alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05))


```

### Binned P.Val vs. Prop

#### Function

```{r bin Amean}
plotP.Val_binnedAmean <- function(stats_barcodes) {
  stats_binned <- stats_barcodes %>%
    mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, 0, by = 1), include.lowest = TRUE)
  ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>%
    mutate(neglog10_p = -log10(pvalue + 1e-6))

ggplot(stats_binned, aes(x = bin, y = neglog10_p, fill = method)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ N_Samples, nrow = 1) +
  labs(x = "log10(barcode mean prop) bins", y = "-log10(p-value)", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r bin Amean, fig.width=27, fig.height=4}
plotP.Val_binnedAmean(stats_barcodes)
```


### High FPR simulation

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.08)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.3)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.05)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
# df_Group_Vec[flag_highFPR_asin,]
# df_Group_Vec[flag_highFPR_logit,]
# df_Group_Vec[flag_highFPR_noTrans,]

```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)

# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_logit))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_noTrans))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_asin < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_logit < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_noTrans < 0.05)))
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- C21
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

#### heatmap

```{r, fig.width=10, fig.height=6}
Heatmeap_list <- lapply(rownames(groups_highFPR), plotSimHeatmap)

# Set the resolution and output file for PNG
png("../../output/BelderbosME/heatmap_simulation_C21.png", width = 45, height = 30, units = "in", res = 300)

# Set layout: 2 rows × 2 columns
n_rows <- 6
n_cols <- 5

grid.newpage()
pushViewport(viewport(layout = grid.layout(n_rows, n_cols)))

vplayout <- function(row, col) viewport(layout.pos.row = row, layout.pos.col = col)

for (i in seq_along(Heatmeap_list)) {
  row <- ceiling(i / n_cols)
  col <- (i - 1) %% n_cols + 1
  pushViewport(vplayout(row, col))
  draw(Heatmeap_list[[i]], newpage = FALSE)  # This renders the heatmap in the current viewport
  upViewport()
}

# Close the PNG device
dev.off()
```

# Subet C23 Mouse

## NegSim

```{r negsim dmso, fig.width=7, fig.height=5}
source("../../analysis/BelderbosME/NegSim-mouseHSPC.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

C23 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C23"]

set.seed(2025)

results_negsi <- random_sampling(barbieQ = C23, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/BelderbosME/all_loops_C23.rda")

lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, C23")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, C23")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, C23")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, C23")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, C23")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, C23")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, C23")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, C23")

ggsave(
  "../../output/BelderbosME/test_C23.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 40, height = 15,
  units = "cm"
  )

```

## Inspect NegSim

### load results

```{r extract results from random sampling}
load("../../output/BelderbosME/all_loops_C23.rda")

  # extract results from the loops
  end_sampling <- 6
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```


### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


## flag out the FPR-high simulation
flag_highFPR <- which(df_FPR$Prop_asin > 0.05)
SimID_highFPR <- df_FPR[flag_highFPR,]$SimulationID

# df_Group_Vec[SimID_highFPR,]
```

### P.Val vs. Prop

```{r plot p.val}
stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
# stats_barcodes$Gmean_prop <- (2^stats_barcodes$Gmean_prop-0.5) / 1e6

ggplot(stats_barcodes) +
  geom_point(aes(y = -log10(Prop_asin), x = log10(Amean_prop + 1e-7)), color = "skyblue", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_logit), x = log10(Amean_prop + 1e-7)), color = "yellow", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_noTrans), x = log10(Amean_prop + 1e-7)), color = "pink", alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05))


```

### Binned P.Val vs. Prop

#### Function

```{r bin Amean}
plotP.Val_binnedAmean <- function(stats_barcodes) {
  stats_binned <- stats_barcodes %>%
    mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, 0, by = 1), include.lowest = TRUE)
  ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>%
    mutate(neglog10_p = -log10(pvalue + 1e-6))

ggplot(stats_binned, aes(x = bin, y = neglog10_p, fill = method)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ N_Samples, nrow = 1) +
  labs(x = "log10(barcode mean prop.) bins", y = "-log10(p-value)", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r bin Amean, fig.width=12, fig.height=4}
plotP.Val_binnedAmean(stats_barcodes)
```


### High FPR simulation

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.08)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.2)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.05)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
# df_Group_Vec[flag_highFPR_asin,]
# df_Group_Vec[flag_highFPR_logit,]
# df_Group_Vec[flag_highFPR_noTrans,]

```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)

# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_logit))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# ggplot(stats_barcodes_SimHigh) +
#   geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_noTrans))) +
#   coord_fixed() +
#   geom_vline(xintercept = -log10(0.05)) +
#   geom_hline(yintercept = -log10(0.05))
# 
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_asin < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_logit < 0.05)))
# 
# ggplot(as.data.frame(stats_barcodes_SimHigh)) +
#   geom_point(aes(x = Gmean_prop, y = LFC_prop, 
#                  color = as.factor(Prop_noTrans < 0.05)))
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- C23
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

#### heatmap

```{r, fig.width=10, fig.height=6}
Heatmeap_list <- lapply(rownames(groups_highFPR), plotSimHeatmap)

# Set the resolution and output file for PNG
png("../../output/BelderbosME/heatmap_simulation_C23.png", width = 45, height = 30, units = "in", res = 300)

# Set layout: 2 rows × 2 columns
n_rows <- 6
n_cols <- 5

grid.newpage()
pushViewport(viewport(layout = grid.layout(n_rows, n_cols)))

vplayout <- function(row, col) viewport(layout.pos.row = row, layout.pos.col = col)

for (i in seq_along(Heatmeap_list)) {
  row <- ceiling(i / n_cols)
  col <- (i - 1) %% n_cols + 1
  pushViewport(vplayout(row, col))
  draw(Heatmeap_list[[i]], newpage = FALSE)  # This renders the heatmap in the current viewport
  upViewport()
}

# Close the PNG device
dev.off()
```

