---
title: "neg_sim_mouse_HSOC_barbieQ"
author: "LF"
date: "2025-04-11"
output: workflowr::wflow_html
output_dir: ../../BelderbosME
---

# Load Dependencies

```{r load library}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr) # for pivot_longer
library(tibble) # for rownames_to _column
library(knitr) # for kable()
library(ggplot2)
library(patchwork)
library(scales)
library(ggVennDiagram)
library(ComplexHeatmap)
library(limma)
library(edgeR)

library(SummarizedExperiment)
library(SEtools)
library(S4Vectors)
library(devtools)

# devtools::install_github("DaneVass/bartools", dependencies = TRUE, force = TRUE)
library(bartools)
```

# Install barbieQ - devel

```{r install barbieQ}
# devtools::install_github("Oshlack/barbieQ", force = TRUE)
BiocManager::install("barbieQ")

library(barbieQ)
```

# Load rds

```{r save rds}
# rds saved in "../../analysis/BelderbosME/mouseHSPC_preprocessing.Rmd"

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

assay(mouseHSPC) %>% colSums() %>% sort()
assay(mouseHSPC_top) %>% colSums() %>% sort()

(colSums(assay(mouseHSPC)) - colSums(assay(mouseHSPC_top))) %>% sort()

```

## t-SNE

```{r tsne}
# Set an appropriate perplexity value (e.g., num_samples / 3)
# set.seed(12)
# perplexity_value <- min(30, floor(ncol(mouseHSPC_top) / 3))
# 
# tsne_results <- (assays(mouseHSPC_top)$CPM + 1) %>% 
#   t() %>% 
#   log2() %>% 
#   Rtsne::Rtsne(perplexit = perplexity_value)
# tsne_coords <- tsne_results$Y
# tsne_coords <- as.data.frame(tsne_coords)
# 
# ggplot(tsne_coords) +
#   geom_point(aes(x = V1, y = V2, col = mouseHSPC_top$sampleMetadata$Mouse)) +
#   labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", col = "Mouse") +
#   theme_classic() +
#   theme(aspect.ratio = 1)
# 
# lapply(
#   mouseHSPC_top$sampleMetadata %>% ncol() %>% seq(), 
#   function(i) {
#     col_fac <- colnames(mouseHSPC_top$sampleMetadata)[i]
#     ggplot(tsne_coords) +
#      geom_point(aes(x = V1, y = V2, col = mouseHSPC_top$sampleMetadata[[i]])) +
#      labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2", color = col_fac) +
#       theme_classic() +
#      theme(aspect.ratio = 1)
#   }
#   )

pca_results <- (assays(mouseHSPC_top)$CPM + 1) %>% 
  t() %>% 
  log2() %>% 
  prcomp( scale. = TRUE)

pca_df <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  Mouse = mouseHSPC_top$sampleMetadata$Mouse
)

ggplot(pca_df, aes(x = PC1, y = PC2, color = Mouse)) +
  geom_point(size = 3) +
  labs(title = "PCA on logCPM", x = "PC1", y = "PC2") +
  theme_classic() +
  theme(aspect.ratio = 1)

mouseHSPC_top$sampleMetadata %>% as.data.frame() %>%  View()
assay(mouseHSPC_top) %>% colnames()
```

## bartools

```{r bartools, fig.width=30, fig.height=10}
mouse_ids <- mouseHSPC_top$sampleMetadata$Mouse

mouse_levels <- c("C22", setdiff(unique(mouse_ids), "C22"))

ordered_indices <- order(factor(mouse_ids, levels = mouse_levels))

ps.counts <- ps.counts[, ordered_indices]

mouseHSPC_top$sampleMetadata <- mouseHSPC_top$sampleMetadata[ordered_indices, ]

ps.counts <- assay(mouseHSPC_top)
colnames(ps.counts) <- make.unique(mouseHSPC_top$sampleMetadata$Mouse)
mouse_dge <- DGEList(
  counts = ps.counts, 
  group = mouseHSPC_top$sampleMetadata$Mouse)

bartools::plotBarcodeHistogram(mouse_dge)
```

# mouseHSPC

```{r mouseHSPC, fig.width=8, fig.height=6}
barbieQ::plotBarcodeHeatmap(
  mouseHSPC_top, sampleMetadata = mouseHSPC_top$sampleMetadata[,-1])

barbieQ::plotBarcodeHeatmap(
  mouseHSPC_top, sampleMetadata = mouseHSPC_top$sampleMetadata[,-1], splitSamples = T, sampleGroup = "Mouse")
```

### C21

```{r C21}
C21 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C21"]

barbieQ::plotBarcodeHeatmap(
  C21, sampleMetadata = C21$sampleMetadata[,-1])

## flag C21 samples of exremely low libsize
flag_low_C21 <- colSums(assay(C21)) < 1000
## remove the low libsize samples
C21 <- C21[, !flag_low_C21]
colSums(assay(C21))

barbieQ::plotBarcodeHeatmap(
  C21, sampleMetadata = C21$sampleMetadata[,-1])

assay(C21) %>% colSums() %>% sort()

## tag top barcoes
C21 <- tagTopBarcodes(C21)
rowData(C21)$isTopBarcode %>% table()

```

### C22

```{r C22}
C22 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C22"]

## check libsize
assay(C22) %>% colSums()
## tag top barcodes
C22 <- tagTopBarcodes(C22)
rowData(C22)$isTopBarcode %>% table()

barbieQ::plotBarcodeHeatmap(
  C22, sampleMetadata = C22$sampleMetadata[,-1])

barbieQ::plotBarcodeProportion(C22, absoluteProportion = F)
rowMeans(assays(C22)$proportion) %>% summary()
rowMeans(assays(C22)$proportion) %>% sort(decreasing = T) 
rowMeans(assays(mouseHSPC_top)$proportion) %>% summary()



```

### C23

```{r C23}
C23 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C23"]

## check libsize
assay(C23) %>% colSums()
## tag top barcoes
C23 <- tagTopBarcodes(C23)
rowData(C23)$isTopBarcode %>% table()

barbieQ::plotBarcodeHeatmap(
  C23, sampleMetadata = C23$sampleMetadata[,-1])

barbieQ::plotBarcodeProportion(C23, absoluteProportion = F)
rowMeans(assays(C23)$proportion) %>% summary()
rowMeans(assays(C23)$proportion) %>% sort(decreasing = T) 
rowMeans(assays(mouseHSPC_top)$proportion) %>% summary()



```

## C21 - C22

```{r C21 C22, fig.width=10, fig.height=10}

## flag C21 samples of exremely low libsize
flag_low_mouseHSPC_top <- colSums(assay(mouseHSPC_top)) < 1000
## remove the low libsize samples
mouseHSPC_top_lib <- mouseHSPC_top[, !flag_low_mouseHSPC_top]

myds <- mouseHSPC_top_lib$sampleMetadata %>%
    with(model.matrix(~0 + Mouse))

mouseHSPC_top_lib_occ <- testBarcodeSignif(
  barbieQ = mouseHSPC_top_lib, designMatrix = myds, method = "diffOcc", contrastFormula = "MouseC22 - MouseC21")
plotSignifBarcodeProportion(mouseHSPC_top_lib_occ)
plotSignifBarcodeHeatmap(mouseHSPC_top_lib_occ)

ggsave("../../output/BelderbosME/C21_C22_diff_occ_sig.prop.png", width = 8, height = 8, 
       plotSignifBarcodeProportion(mouseHSPC_top_lib_occ))

png(filename = "../../output/BelderbosME/C21_C22_diff_occ_ht.png", res = 300,
    width = 2000, height = 2000)
plotSignifBarcodeHeatmap(mouseHSPC_top_lib_occ)
dev.off()


mouseHSPC_top_lib_prop <- testBarcodeSignif(
  barbieQ = mouseHSPC_top_lib, designMatrix = myds, contrastFormula = "MouseC22 - MouseC21")
plotSignifBarcodeProportion(mouseHSPC_top_lib_prop)

bar_test <- rowData(mouseHSPC_top_lib)$testingBarcode
bar_test %>% as.data.frame() %>% View()
metadata(bar_test)$design
bar_test$direction %>% table()

plotBarcodeMA(mouseHSPC_top_lib)
plotBarcodePValue(mouseHSPC_top_lib)
plotSignifBarcodeHeatmap(mouseHSPC_top_lib)
plotSignifBarcodeProportion(mouseHSPC_top_lib)


```


# Subet C22 Mouse

```{r subset dmso}
C22 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C22"]

myds <- mouseHSPC_top$sampleMetadata %>%
    with(model.matrix(~0 + Mouse))

mouseHSPC_top <- testBarcodeSignif(
  barbieQ = mouseHSPC_top, designMatrix = myds, method = "diffOcc", contrastFormula = "MouseC22 - MouseC21")

mouseHSPC_top <- testBarcodeSignif(
  barbieQ = mouseHSPC_top, designMatrix = myds, contrastFormula = "MouseC22 - MouseC21")

bar_test <- rowData(mouseHSPC_top)$testingBarcode
bar_test %>% as.data.frame() %>% View()
metadata(bar_test)$design
bar_test$direction %>% table()

plotBarcodeMA(mouseHSPC_top)
plotBarcodePValue(mouseHSPC_top)
plotSignifBarcodeHeatmap(mouseHSPC_top)
plotSignifBarcodeProportion(mouseHSPC_top)


```




## NegSim

```{r negsim dmso, fig.width=7, fig.height=5}
source("../../analysis/BelderbosME/NegSim-mouseHSPC.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

C22 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C22"]
C22$sampleMetadata$Mouse %>% table()

set.seed(2025)

results_negsi <- random_sampling(barbieQ = C22, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/BelderbosME/all_loops.rda")

lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, C22")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, C22")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, C22")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, C22")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, C22")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, C22")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, C22")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, C22")

ggsave(
  "../../output/BelderbosME/test.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 40, height = 15,
  units = "cm"
  )

```

## Inspect NegSim

### load results

```{r extract results from random sampling}
load("../../output/BelderbosME/all_loops.rda")

  # extract results from the loops
  end_sampling <- 8
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```


### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


## flag out the FPR-high simulation
flag_highFPR <- which(df_FPR$Prop_asin > 0.05)
SimID_highFPR <- df_FPR[flag_highFPR,]$SimulationID

# df_Group_Vec[SimID_highFPR,]
```

### P.Val vs. Prop

```{r plot p.val}
stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)
# stats_barcodes$Gmean_prop <- (2^stats_barcodes$Gmean_prop-0.5) / 1e6

ggplot(stats_barcodes) +
  geom_point(aes(y = -log10(Prop_asin), x = log10(Amean_prop + 1e-7)), color = "skyblue", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_logit), x = log10(Amean_prop + 1e-7)), color = "yellow", alpha = 0.6) +
  geom_point(aes(y = -log10(Prop_noTrans), x = log10(Amean_prop + 1e-7)), color = "pink", alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05))


```

### Binned P.Val vs. Prop

#### Function

```{r bin Amean}
plotP.Val_binnedAmean <- function(stats_barcodes) {
  stats_binned <- stats_barcodes %>%
    mutate(
    logAmean = log10(Amean_prop + 1e-7),
    bin = cut(logAmean, breaks = seq(-7, 0, by = 1), include.lowest = TRUE)
  ) %>% 
    pivot_longer(
      cols = c(Prop_asin, Prop_logit, Prop_noTrans),
      names_to = "method",
      values_to = "pvalue"
    ) %>%
    mutate(neglog10_p = -log10(pvalue + 1e-6))

ggplot(stats_binned, aes(x = bin, y = neglog10_p, fill = method)) +
  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ N_Samples, nrow = 1) +
  labs(x = "log10(barcode mean prop.) bins", y = "-log10(p-value)", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r bin Amean, fig.width=18, fig.height=4}
plotP.Val_binnedAmean(stats_barcodes)
```

```{r extreme p.val}
which(-log10(stats_barcodes_SimHigh$Prop_noTrans) > 200)
stats_barcodes_SimHigh[1131,]
```


### High FPR simulation

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.05)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.2)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.05)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
df_Group_Vec[flag_highFPR_asin,]
df_Group_Vec[flag_highFPR_logit,]
df_Group_Vec[flag_highFPR_noTrans,]

```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)

ggplot(stats_barcodes_SimHigh) +
  geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_logit))) +
  coord_fixed() +
  geom_vline(xintercept = -log10(0.05)) +
  geom_hline(yintercept = -log10(0.05))

ggplot(stats_barcodes_SimHigh) +
  geom_point(aes(x = -log10(Prop_asin), y = -log10(Prop_noTrans))) +
  coord_fixed() +
  geom_vline(xintercept = -log10(0.05)) +
  geom_hline(yintercept = -log10(0.05))


ggplot(as.data.frame(stats_barcodes_SimHigh)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(Prop_asin < 0.05)))

ggplot(as.data.frame(stats_barcodes_SimHigh)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(Prop_logit < 0.05)))

ggplot(as.data.frame(stats_barcodes_SimHigh)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(Prop_noTrans < 0.05)))
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- C22
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

#### heatmap

```{r, fig.width=30, fig.height=15}
Heatmeap_list <- lapply(rownames(groups_highFPR), plotSimHeatmap)

# Set the resolution and output file for PNG
png("../../output/BelderbosME/heatmap_simulation.png", width = 45, height = 30, units = "in", res = 300)

# Set layout: 2 rows × 2 columns
n_rows <- 6
n_cols <- 5

grid.newpage()
pushViewport(viewport(layout = grid.layout(n_rows, n_cols)))

vplayout <- function(row, col) viewport(layout.pos.row = row, layout.pos.col = col)

for (i in seq_along(Heatmeap_list)) {
  row <- ceiling(i / n_cols)
  col <- (i - 1) %% n_cols + 1
  pushViewport(vplayout(row, col))
  draw(Heatmeap_list[[i]], newpage = FALSE)  # This renders the heatmap in the current viewport
  upViewport()
}

# Close the PNG device
dev.off()
```

#### tsne / pca

```{r pca high FPR simulation}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimPCA <- function(sim) {
  ## the first dim is "sample_3_loop_1"
  # sim <- "sample_3_loop_1"
  ## add simulation groups to sample metadata
  simData <- C22
  simData$sampleMetadata <- cbind(simData$sampleMetadata, t(groups_highFPR[sim,, drop = F]))
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
  sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit< 0.05 | check_barcodes$Prop_noTrans < 0.05)
  
  pca_results <- prcomp(t(assays(simData)$proportion))
  ggplot(as.data.frame(pca_results$x)) +
    geom_point(aes(x = PC1, y = PC2, shape = groups_highFPR[sim,], color = simData$sampleMetadata$Celltype))
  
  # tsne_results <- Rtsne(t(assays(simData)$proportion), perplexity = 3)
  # ggplot(as.data.frame(tsne_results$Y)) +
  #   geom_point(aes(x = V1, y = V2, color = groups_highFPR[sim,]))
  # 
  # umap_results <- umap(t(assays(simData)$proportion), 
  #                      n_neighbors = 3, min_dist = 0.1, metric = "euclidean")
  # ggplot(as.data.frame(umap_results)) +
  #   geom_point(aes(x = V1, y = V2, color = groups_highFPR[sim,]))
}

```

```{r, fig.width=10, fig.height=6}
PCA_list <- lapply(rownames(groups_highFPR), plotSimPCA)

C22 <- tagTopBarcodes(C22)
rowData(C22)$isTopBarcode$isTop %>% summary()
C22 <- C22[rowData(C22)$isTopBarcode$isTop,]

plotBarcodeHeatmap(C22)

set.seed(2025)
plotBarcodeHeatmap(mouseHSPC_top, sampleMetadata = mouseHSPC_top$sampleMetadata[,-1])

```

```{r, fig.width=8, fig.height=6}
C21 <- mouseHSPC_top[,mouseHSPC_top$sampleMetadata$Mouse == "C21"]
C21 <- tagTopBarcodes(C21)
C21 <- C21[rowData(C21)$isTopBarcode$isTop,]
plotBarcodeHeatmap(C21)
```

```{r, fig.width=8, fig.height=6}
C21 <- mouseHSPC_top[,mouseHSPC_top$sampleMetadata$Mouse == "C21"]
C21 <- tagTopBarcodes(C21)
C21 <- C21[rowData(C21)$isTopBarcode$isTop,]
plotBarcodeHeatmap(C21)
```

## NegSim C22 top

```{r c22 top}
source("../../analysis/BelderbosME/NegSim-mouseHSPC.R")

library(SummarizedExperiment)
library(magrittr)
library(barbieQ)
library(ggplot2)
library(patchwork)

load("../../output/BelderbosME/mouseHSPC_barbieQ.rda")

C22 <- mouseHSPC_top[, mouseHSPC_top$sampleMetadata$Mouse == "C22"]

# filter bottom barcodes in C22
C22 <- tagTopBarcodes(C22)
rowData(C22)$isTopBarcode %>% table()
C22_top <- C22[rowData(C22)$isTopBarcode$isTop,]

set.seed(2025)

results_negsi <- random_sampling(barbieQ = C22_top, loop_times = 100)

## saved to global
save(global_all_loops, file = "../../output/BelderbosME/all_loops_top.rda")

lapply(results_negsi, plot)

results_negsi[[1]] <- results_negsi[[1]] + ggtitle("Prop_asin, C22_top")
results_negsi[[2]] <- results_negsi[[2]] + ggtitle("Prop_logit, C22_top")
results_negsi[[3]] <- results_negsi[[3]] + ggtitle("Prop_noTrans, C22_top")
results_negsi[[4]] <- results_negsi[[4]] + ggtitle("Occ_firth, C22_top")
results_negsi[[5]] <- results_negsi[[5]] + ggtitle("Prop_asin, C22_top")
results_negsi[[6]] <- results_negsi[[6]] + ggtitle("Prop_logit, C22_top")
results_negsi[[7]] <- results_negsi[[7]] + ggtitle("Prop_noTrans, C22_top")
results_negsi[[8]] <- results_negsi[[8]] + ggtitle("Occ_firth, C22_top")

ggsave(
  "../../output/BelderbosME/test_top.png",
  wrap_plots(results_negsi[1:8], ncol = 4),
  width = 40, height = 15,
  units = "cm"
  )

```

### load results

```{r extract results from random sampling}
load("../../output/BelderbosME/all_loops_top.rda")

  # extract results from the loops
  end_sampling <- 8
  
  ## extract FPR
  all_FPR <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$FPR
  })
  df_FPR <- do.call(rbind, all_FPR)
  
  ## extract P.Val
  all_P.Val <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$P.Val
  })
  df_P.Val <- do.call(rbind, all_P.Val)
  
  ## extract MA
  all_MA <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$MA
  })
  df_MA <- do.call(rbind, all_MA)
  
  all_Group_Vec <- lapply(seq(3:end_sampling), function(n) {
    global_all_loops[[n]]$Group_Vec
  })
  df_Group_Vec <- do.call(rbind, all_Group_Vec)
  
  stats_barcodes <- cbind(df_P.Val[,1:4], df_MA)

```

### high logit

#### locate the high FPRs

```{r high asin}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.2)
df_FPR[flag_highFPR_asin,"SimulationID"]

flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.3)
df_FPR[flag_highFPR_logit,"SimulationID"]

flag_highFPR_noTrans <- which(df_FPR$Prop_noTrans > 0.2)
df_FPR[flag_highFPR_noTrans, "SimulationID"]


## simulation groups
df_Group_Vec[flag_highFPR_asin,]
df_Group_Vec[flag_highFPR_logit,]
df_Group_Vec[flag_highFPR_noTrans,]


flag_highFPR_asin
flag_highFPR_logit
flag_highFPR_noTrans

## check logit high simulations
df_FPR %>% filter(Prop_asin > 0.18) %>% View()
df_FPR %>% filter(Prop_logit > 0.3) %>% View()
df_FPR %>% filter(Prop_logit > 0.18) %>% View()
```

#### union of the high FPR

```{r high union}
flag_highFPR_union <- union(union(flag_highFPR_asin, flag_highFPR_logit), flag_highFPR_noTrans)

SimID_highFPR <- df_FPR[flag_highFPR_union, "SimulationID"]

stats_barcodes_SimHigh <- stats_barcodes %>% filter(SimulationID %in% SimID_highFPR)
```

#### Function

```{r heatmap of high sim}
groups_highFPR <- df_Group_Vec[flag_highFPR_union,]

plotSimHeatmap <- function(sim) {
  ## the first dim is "sample_3_loop_39"
  # sim <- "sample_3_loop_39"
  ## add simulation groups to sample metadata
  simData <- C22_top
  simData$sampleMetadata <- cbind(simData$sampleMetadata[,-1], t(groups_highFPR[sim,, drop = F]))
  ## locate the barcode pvalues of the chosen simulation
  check_barcodes <- stats_barcodes_SimHigh %>% 
    filter(SimulationID == sim) 
  ## locate the significant barcodes in either test
#   sig_barcodes <- which(check_barcodes$Prop_asin < 0.05 | check_barcodes$Prop_logit < 0.05 | check_barcodes$Prop_noTrans < 0.05)
#   
#   sig_simData <- simData[sig_barcodes,]
#   barbieQ::plotBarcodeHeatmap(barbieQ = sig_simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
#     asin = check_barcodes[sig_barcodes,"Prop_asin"] < 0.05,
#     logit = check_barcodes[sig_barcodes,"Prop_logit"] < 0.05,
#     noTrans = check_barcodes[sig_barcodes,"Prop_noTrans"] < 0.05,
#     col = list(
#       asin = c(`TRUE` = "red", `FALSE` = "grey"),
#       logit = c(`TRUE` = "red", `FALSE` = "grey"),
#       noTrans = c(`TRUE` = "red", `FALSE` = "grey")
#     )
#     ))
# }

  barbieQ::plotBarcodeHeatmap(barbieQ = simData, splitSamples = T, sampleGroup = sim, barcodeAnnotation = rowAnnotation(
    asin = check_barcodes[,"Prop_asin"] < 0.05,
    logit = check_barcodes[,"Prop_logit"] < 0.05,
    noTrans = check_barcodes[,"Prop_noTrans"] < 0.05,
    col = list(
      asin = c(`TRUE` = "red", `FALSE` = "grey"),
      logit = c(`TRUE` = "red", `FALSE` = "grey"),
      noTrans = c(`TRUE` = "red", `FALSE` = "grey")
    )
    ))
}

```

```{r high logit HP}
flag_highFPR_logit <- which(df_FPR$Prop_logit > 0.30)
lapply(df_FPR$SimulationID[flag_highFPR_logit], plotSimHeatmap)
```

```{r high asin HP}
flag_highFPR_asin <- which(df_FPR$Prop_asin > 0.18)
lapply(df_FPR$SimulationID[flag_highFPR_asin], plotSimHeatmap)
```

### MA plot

```{r inspect random sampling}
## LFC and Gmean works better than linear diff and Amean
## the x axis is actually log10 Gmean proportion
ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_asin < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop,  
                 color = as.factor(df_P.Val$Prop_logit < 0.05)))

ggplot(as.data.frame(df_MA)) +
  geom_point(aes(x = Gmean_prop, y = LFC_prop, 
                 color = as.factor(df_P.Val$Prop_noTrans < 0.05)))


# df_Group_Vec[SimID_highFPR,]
```

### Binned P.Val

```{r bin Amean, fig.width=10, fig.height=8}
plotP.Val_binnedAmean(stats_barcodes)
```
